### 2024-05-02 paper/paper.tex
Commit paper, style files, and directory layout

\documentclass[draft]{agujournal2019}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage{soul}
\linenumbers

\draftfalse

\journalname{JGR: Oceans}

\begin{document}

\title{Dispersion of short waves riding on longer waves}

\authors{M. Curcic\affil{1,2}}

\affiliation{1}{
  Rosenstiel School of Marine, Atmospheric, and Earth Science,
  University of Miami,
  Miami, FL
}
\affiliation{2}{
  Frost Institute for Data Science and Computing,
  University of Miami,
  Miami, FL
}

\correspondingauthor{Milan Curcic}{mcurcic@miami.edu}

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

\begin{abstract}
Abstract goes here
\end{abstract}

\section*{Plain Language Summary}
Plain language summary goes here

\section{Introduction}

Short ocean surface waves are modulated by longer swell.
Following the result by \citeA{unna1941white,unna1942waves,unna1947sea},
\citeA{longuet1960changes} introduced an analytical solution for hydrodynamic
modulation of short waves by an underlying longer wave:

\begin{equation}
\label{eq:tau}
\widetilde{k} = k (1 + a_L k_L)
\end{equation}
where $\widetilde{k}$ and $k$ are the modulated and unmodulated wavenumbers of
short waves, respectively, and $a_L$ and $k_L$ are amplitude and wavenumber
of the long wave.

This result has been influential on the development of remote sensing of ocean
surface waves and currents \cite{keller1975microwave,hara1994hydrodynamic}
and has been used even recently \cite{peureux2021unsteady} to simulate
unsteady growth of short-waves due to longer waves.

In this study, we extend the \citeA{longuet1960changes} theory (hereafter L-HS)
to a higher-order solution following the linear wave theory, and present revised
phase-averaged estimates of short-wave properties in presence of long waves.

\section{Governing equations}

Paragraph: Describe the conditions, assumptions, and limitations.

\section{Analytical solutions}

\subsection{Derivation of the original L-HS solution}

The L-HS solution can be most readily obtained by starting from the dispersion
relationship for short waves:

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency,
$g$ is the gravitational acceleration,
and $U$ is the mean advective current in the direction of the wave propagation.
For simplicity, we will limit the derivation to gravity waves and ignore
surface tension effects, however, the solution can be as well derived for
gravity-capillary waves by carrying the surface tension terms.

Next, take the conservation of wave crests:

\begin{equation}
\label{eq:wave_crests1}
\dfrac{\partial k}{\partial t}
+ \dfrac{\partial \omega}{\partial x}
= 0
\end{equation}
where $\partial/\partial t$ and $\partial/\partial x$ are partial derivatives
in time and space, respectively.
Insert (\ref{eq:dispersion}) into (\ref{eq:wave_crests1}) to get:

\begin{equation}
\label{eq:wave_crests2}
\dfrac{\partial k}{\partial t}
+ C_g \dfrac{\partial k}{\partial x}
+ U \dfrac{\partial k}{\partial x}
+ k \dfrac{\partial U}{\partial x}
= 0
\end{equation}
where $C_g$ is the group velocity of short waves.
At this step, we assume that $g$ does not vary in space and time.
We will see later that to a second order $g$ of short waves varies in presence
of long waves.
If we follow the short waves in a Lagrangian frame, we can re-write
(\ref{eq:wave_crests2}) as:

\begin{equation}
\label{eq:wave_crests3}
\dfrac{d k}{dt}
+ k \dfrac{\partial U}{\partial x}
= 0
\end{equation}

\begin{equation}
\label{eq:wave_crests4}
\dfrac{d k}{k}
= - \dfrac{\partial U}{\partial x} dt
\end{equation}
which states that the relative contraction of short waves is largest where the
convergence of the ambient current $U$ is the largest.
For short waves riding on longer waves, the mean ambient current is the
horizontal near-surface orbital velocity of the long wave.
Rather than at the long-wave surface, L-HS took the ambient current to be the
long-wave orbital velocity at the mean water level, $z = 0$:

\begin{equation}
\label{eq:U_L1}
U = a_L \omega_L \cos{\psi}
\end{equation}
where $\psi$ is the long-wave phase such that the crest is located
at $\psi = 0$.
Insert (\ref{eq:U_L1}) into (\ref{eq:wave_crests4}) to obtain:

\begin{equation}
\label{eq:wave_crests5}
\dfrac{d k}{k}
= a_L k_L \omega_L \sin{\psi} dt
\end{equation}
To calculate the long-wave modulated short-wave wavenumber, integrate
(\ref{eq:wave_crests5}) from the initial time at which the short-waves
are unperturbed (at $z=0$, or $\psi=\pi/2$), to time $t$:

\begin{equation}
\label{eq:wave_crests6}
\int_k^{\ \widetilde k(t)} \dfrac{d k'}{k'}
= \int_0^t a_L k_L \omega_L \sin{\psi} dt'
\end{equation}
which yields:

\begin{equation}
\label{eq:k_short1}
\ \widetilde k(\psi) = k e^{a_L k_L \cos{\psi}}
\end{equation}
Approximating (\ref{eq:k_short1}) to the first-order Taylor expansion,
we obtain the original L-HS solution:

\begin{equation}
\label{eq:k_short2}
\ \widetilde k(\psi) = k (1 + a_L k_L \cos{\psi})
\end{equation}
The intuitive interpretation of (\ref{eq:k_short2}) is that as the short waves
contract on the front face of the long wave, they reach maximum contraction
(highest $\ \widetilde k$) at the crest of the long wave.
As they pass to the rear face of the long wave, they get elongated and reach
minimum contraction (lowest $\ \widetilde k$) in the trough of the long wave.

\subsection{Extending the L-HS solution}

L-HS made two approximations in their solution:
First, they evaluated the advective velocity of short waves at $z = 0$
rather than the surface, $z = \eta_L$, where $\eta_L$ is the long-wave
elevation; and second, they evaluated the solution to the first order
of the Taylor expansion series.
Both of these approximations cause an underestimate of the short-wave
modulation magnitude.

To extend the L-HS solution, evaluate the advective velocity at the surface
of the long-wave:

\begin{equation}
\label{eq:U_L2}
U = U_L(z=\eta) = a_L \omega_L e^{a_L k_L \cos{\psi}} \cos{\psi}
\end{equation}
and insert it into \ref{eq:wave_crests4} to obtain:

\begin{equation}
\label{eq:wave_crests7}
\dfrac{dk}{k} = a_L k_L \omega_L e^{a_L k_L \cos{\psi}} \sin{\psi} dt
\end{equation}
Integrate (\ref{eq:wave_crests7}) to the same limits as we did with
(\ref{eq:wave_crests6}) to obtain:

\begin{equation}
\label{eq:k_short3}
\widetilde{k}(\psi) = k e^{a_L k_L \cos{\psi} e^{a_L k_L \cos{\psi}}}
\end{equation}

We compare the solutions from (\ref{eq:k_short1}), (\ref{eq:k_short2}),
and (\ref{eq:k_short3}) for a moderately steep long wave with steepness
$a_L k_L = 0.2$ in Fig. (\ref{fig:wavenumber_modulation}).
The original L-HS solution is a cosine wave with an amplitude equal
to the long-wave steepness.
The other two solutions both introduce an increase in wave modulation
on the crest and a decrease in the trough.
At the crest, the relative modulation predicted by the exact L-HS solution
is 10.7\% larger than that predicted by the 1st order approximate L-HS
solution, while the solution that evaluates the long-wave orbital velocity
at the water surface rather than $z = 0$ is 38.4\% larger than that of the
1st order approximate L-HS solution.
Also, since the less-approximate solutions decrease the modulation in the
troughs and increase in on the crests, the average wavenumber modulation
is no longer 0, but 1\% in the case of the exact L-HS solution and 3.1\% in the
case of the extended solution.

\begin{figure}[h]
\label{fig:wavenumber_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_wavenumber_modulation_by_phase.png}
\caption{
  Comparison of relative short-wave modulation as function of the
  long-wave phase, for a moderately steep long wave of $a_L k_L = 0.2$.
  The dotted line shows the 1st-order approximated solution by L-HS,
  the dashed line shows the exact solution by L-HS,
  and the solid line shows the extended solution introduced in this paper.
}
\end{figure}

\section{Apparent gravity}

Long waves also induce orbital acceleration under their surface, so the short
waves experience an apparent gravity:

\begin{equation}
\label{eq:apparent_gravity1}
\widetilde{g} = g + \dfrac{\partial w_L}{\partial t}
\end{equation}
The positive sign in front of $\partial w_L/\partial t$ comes from our
coordinate-system convention where $w$ is defined positive-upward and $g$ is
defined positive-downward, and the fact that $\widetilde{g}$ is the apparent
gravity in the reference frame of the short waves.
Taking the long-wave vertical acceleration at the surface as:

\begin{equation}
\label{eq:w_L}
w_L = - a_L \omega_L e^{a_L k_L \cos{\psi}} \sin{\psi}
\end{equation}
we arrive at:

\begin{equation}
\label{eq:apparent_gravity2}
\widetilde{g}(\psi) = g \left[
  1 - a_L k_L e^{a_L k_L \cos{\psi}} \left(
    \cos{\psi} + a_L k_L \sin^2{\psi}
  \right)
\right]
\end{equation}
The first-order term in $a_L k_L$ is in phase with the long-wave elevation,
so the apparent gravity of the short waves is increased at the long-wave crests
and decreased in the troughs.
The second-order term in $a_L k_L$ is positive-definite and phase-shifted by
$\pi/2$ relative to the crests.
The second-order term thus increases the apparent gravity on the front and rear
faces of the long wave (Fig. \ref{fig:apparent_gravity}).

\begin{figure}[h]
\label{fig:apparent_gravity}
\centering
\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_by_phase.png}
\caption{
  Relative change in gravity of short waves due to a longer wave with steepness $a_L k_L = 0.2$,
  as a function of the long-wave phase and evaluated to the first (dashed line) and second (solid line) order in $a_L k_L$.
}
\end{figure}

\begin{figure}[h]
\label{fig:gravity_modulation_vs_ak}
\centering
\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_by_ak.png}
\caption{
  TODO
}
\end{figure}

With the new long-wave modulated expressions for the short-wave wavenumber
(Eq. \ref{eq:k_short3}) and gravity (Eq. \ref{eq:apparent_gravity2}), we can
evaluate the long-wave modulated frequency as simply
$\widetilde{\omega}(\psi) = \sqrt{\widetilde{g} \widetilde{k}}$, which we compare
for four different values of $a_L k_L$ in Fig. \ref{fig:frequency_modulation}.

\begin{figure}[h]
\label{fig:frequency_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_frequency_modulation.png}
\caption{
  Relative change in short-wave frequency due to the presence of longer
  waves, for four different values of long-wave steepness.
}
\end{figure}

\section{Lagrangian averaged short-wave properties}

Stokes drift is a residual Lagrangian drift that occurs due to the wave orbits
not being closed under a propagating wave
\cite{stokes1847,kenyon1969stokes,van2018stokes}:

\begin{equation}
\label{}
u_s = a^2 \omega k e^{2kz}
\end{equation}
for a monochromatic wave in deep water.
Stokes drift has been successfully used to predict material transport on the
ocean surface \cite{rohrs2012observation,curcic2016hurricane}.
It can be derived from linear wave theory by integrating the wave orbital
velocity following material orbits \cite{phillips1966dynamics}:

\begin{equation}
\label{eq:lagrangian_average}
u_s = \overline{
  u(\psi) + 
  \zeta \dfrac{\partial u}{\partial x} +
  \xi \dfrac{\partial w}{\partial z}
}
\end{equation}
where $\zeta$ and $\xi$ are horizontal and vertical excursions of the orbits,
respectively.

To obtain an average modulation of the short-wave properties, we apply the same
Lagrangian integration approach used to obtain Stokes drift.

\begin{equation}
\label{eq:stokes_wavenumber}
\widetilde{k_s} = \overline{
  \widetilde{k}(\psi) +
  \zeta \dfrac{\partial \widetilde{k}}{\partial x} +
  \xi \dfrac{\partial \widetilde{k}}{\partial z}
}
= k \left( 1 + a_L^2 k_L^2 \right)
\end{equation}

\begin{equation}
\label{eq:stokes_gravity}
\widetilde{g_s} = \overline{
  \widetilde{g}(\psi) +
  \zeta \dfrac{\partial \widetilde{g}}{\partial x} +
  \xi \dfrac{\partial \widetilde{g}}{\partial z}
}
= g \left( 1 + \dfrac{3}{2} a_L^2 k_L^2 \right)
\end{equation}

\section{Numerical solutions}

\section{Conclusions}

\acknowledgments
We thank anonymous reviewers for their help improving this paper.

\bibliography{references.bib}

\end{document}

### 2024-05-08 paper/scripts/modulation.py
Make rules for scripts and figures

@@ -1,0 +1,56 @@
+import numpy as np
+
+
+def gravity(
+    x: float, t: float, a: float, k: float, omega: float, g0: float, order: int = 3
+) -> float:
+    assert order in [0, 1, 2, 3]
+    phi = phase(x, t, k, omega)
+    term1 = -a * k * np.cos(phi)
+    term2 = -a * k * np.cos(phi) * np.exp(a * k * np.cos(phi))
+    term3 = -((a * k * np.sin(phi)) ** 2) * np.exp(a * k * np.cos(phi))
+    if order == 0:
+        return g0 * np.ones_like(x)
+    elif order == 1:
+        return g0 * (1 + term1)
+    elif order == 2:
+        return g0 * (1 + term2)
+    elif order == 3:
+        return g0 * (1 + term2 + term3)
+    else:
+        raise ValueError("Invalid order")
+
+
+def phase(
+    x: float | np.ndarray,
+    t: float | np.ndarray,
+    k: float | np.ndarray,
+    omega: float | np.ndarray,
+) -> float | np.ndarray:
+    return k * x - omega * t
+
+
+def wavenumber(
+    frequency: float | np.ndarray,
+    water_depth: float,
+    grav: float = 9.8,
+    surface_tension: float = 0.074,
+    water_density: float = 1e3,
+    num_iterations: int = 100,
+) -> float | np.ndarray:
+    """Solve the dispersion relationship for wavenumber using the
+    Newton-Raphson method."""
+    frequency_nondim = 2 * np.pi * np.sqrt(water_depth / grav) * frequency
+    k = frequency_nondim**2
+    surface_tension_nondim = surface_tension / (grav * water_density * water_depth**2)
+    count = 0
+    while count < num_iterations:
+        t = np.tanh(k)
+        dk = -(frequency_nondim**2 - k * t * (1 + surface_tension_nondim * k**2)) / (
+            3 * surface_tension_nondim * k**2 * t
+            + t
+            + k * (1 + surface_tension_nondim * k**2) * (1 - t**2)
+        )
+        k -= dk
+        count += 1
+    return k / water_depth

### 2024-05-08 paper/scripts/plot_frequency_modulation.py

@@ -1,0 +1,166 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib
+
+
+matplotlib.rc("font", size=16)
+
+
+def plot_fig_frequency_modulation():
+    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
+    g = 9.8
+
+    fig = plt.figure(figsize=(8, 6))
+    ax = fig.add_subplot(111)
+
+    for ak_L in [0.1, 0.2, 0.3, 0.4]:
+        term1 = -ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase))
+        term2 = -((ak_L * np.sin(phase)) ** 2) * np.exp(ak_L * np.cos(phase))
+        g_rel = 1 + term1 + term2
+        k_rel = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
+        omega_rel = np.sqrt(g_rel * k_rel)
+        plt.plot(
+            phase, omega_rel, marker="", linestyle="-", label=r"$a_L k_L=%.1f$" % ak_L
+        )
+
+    plt.legend()
+    plt.xlabel("Phase [rad]")
+    plt.ylabel(r"$\widetilde \omega/\omega$")
+    plt.grid()
+    plt.plot([0, 2 * np.pi], [1, 1], "k--")
+    plt.xlim(0, 2 * np.pi)
+    plt.xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
+    ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
+    plt.savefig("../figures/fig_frequency_modulation.png", dpi=200)
+    plt.close(fig)
+
+
+def plot_fig_phase_speed_modulation():
+    phase = np.arange(0, 2 * np.pi + 1e-3, 1e-3)
+
+    g0 = 9.8
+    k0 = 10
+    omega0 = np.sqrt(g0 * k0)
+    cp0 = omega0 / k0
+
+    k_L = 1
+    omega_L = np.sqrt(g0 * k_L)
+
+    fig = plt.figure(figsize=(8, 6))
+    ax = fig.add_subplot(111)
+
+    colors = ["tab:blue", "tab:orange", "tab:green", "tab:red"]
+
+    for n, ak_L in enumerate([0.1, 0.2, 0.3, 0.4]):
+        a_L = ak_L / k_L
+        eta = a_L * np.cos(phase)
[…]

### 2024-05-08 paper/scripts/plot_gravity_modulation.py

@@ -1,0 +1,108 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib
+
+matplotlib.rc("font", size=16)
+
+
+def g_eulerian(g, psi, a, k, order=3):
+    assert order in [1, 2, 3]
+    eta = a * np.cos(psi)
+    if order == 1:
+        res = g * (1 - a * k * np.cos(psi))
+    elif order == 2:
+        res = g * (1 - a * k * np.cos(psi) * np.exp(k * eta))
+    else:
+        res = g * (
+            1 - a * k * np.exp(k * eta) * (np.cos(psi) + a * k * np.sin(psi) ** 2)
+        )
+    return res
+
+
+def g_lagrangian(g, psi, a, k, order=4):
+    assert order in [3, 4]
+    eta = a * np.cos(psi)
+    g_E = g_eulerian(g, psi, a, k, order=3)
+    if order == 3:
+        res = g_E - a**2 * k**2 * np.exp(2 * k * eta)
+    if order == 4:
+        res = (
+            g_E
+            - a**2 * k**2 * np.exp(2 * k * eta)
+            - a**3
+            * k**3
+            * np.exp(2 * k * eta)
+            * (np.sin(psi) ** 2 * np.cos(psi) - np.sin(2 * psi) * np.sin(psi))
+        )
+    return res
+
+
+def plot_fig_apparent_gravity():
+    """Plot the apparent gravity figure."""
+
+    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
+    a_L = 0.2
+    k_L = 1
+
+    g = 9.8
+
+    g_eulerian_1 = g_eulerian(g, phase, a_L, k_L, order=1)
+    g_eulerian_2 = g_eulerian(g, phase, a_L, k_L, order=2)
+    g_eulerian_3 = g_eulerian(g, phase, a_L, k_L, order=3)
+    g_lagrangian_3 = g_lagrangian(g, phase, a_L, k_L, order=3)
+    # g_lagrangian_4 = g_lagrangian(g, phase, a_L, k_L, order=4) # Essentially the same as 3rd order; omit
+
+    fig = plt.figure(figsize=(8, 6))
+    ax = fig.add_subplot(111)
[…]

### 2024-05-08 paper/scripts/plot_wavenumber_modulation.py

@@ -1,0 +1,65 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from modulation import wavenumber
+import matplotlib
+
+matplotlib.rc("font", size=16)
+
+
+def plot_fig_wavenumber_modulation():
+    """Plot the wavenumber modulation comparison figure."""
+
+    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
+    k_L = 1
+    a_L = 0.1
+    ak_L = a_L * k_L
+
+    solution1 = 1 + ak_L * np.cos(phase)
+    solution2 = np.exp(ak_L * np.cos(phase))
+    solution3 = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
+
+    fig = plt.figure(figsize=(8, 6))
+    ax = fig.add_subplot(111)
+    ax.plot(phase, solution1, "k-", lw=2, label="L-H&S")
+    ax.plot(phase, solution3, "r-", lw=2, label="This paper")
+    ax.legend()
+    ax.set_xlabel("Phase [rad]")
+    ax.set_ylabel(r"$\widetilde k/k$")
+    ax.grid()
+    ax.plot([0, 2 * np.pi], [1, 1], "k--")
+    ax.set_xlim(0, 2 * np.pi)
+    ax.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
+    ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
+    ax.set_title(r"$\varepsilon_L = 0.1$")
+    plt.savefig("../figures/fig_wavenumber_modulation_by_phase.png", dpi=200)
+    plt.close(fig)
+
+    # Evaluate wa enumber modulation as function of ak
+    ak_range = np.arange(0, 0.41, 0.01)
+    k1_mean, k1_max = [], []
+    k2_mean, k2_max = [], []
+    for ak in ak_range:
+        k1 = 1 + ak * np.cos(phase)
+        k2 = np.exp(ak * np.cos(phase) * np.exp(ak * np.cos(phase)))
+        k1_mean.append(np.mean(k1))
+        k1_max.append(np.max(k1))
+        k2_mean.append(np.mean(k2))
+        k2_max.append(np.max(k2))
+
+    fig = plt.figure(figsize=(8, 6))
+    ax = fig.add_subplot(111)
[…]

### 2024-05-08 paper/paper.tex
WIP governing equations

@@
 
-\authors{M. Curcic\affil{1,2}}
+\authors{Milan Curcic\affil{1,2}}
 
@@
 \begin{keypoints}
-\item enter point 1 here
-\item enter point 2 here
-\item enter point 3 here
+\item We revisit analytical and numerical solutions for hydrodynamic modulation of short waves by longer waves from linear wave theory.
+\item We extend the original solution by \citeA{longuet1960changes} to a higher-order solution.
+\item In contrast to earlier studies, we find that there exist stable numerical solutions for modulation by realistic long-wave groups.
 \end{keypoints}
@@
 \begin{abstract}
-Abstract goes here
+Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
+a well-known ocean surface process that pertains to remote sensing, the
+interpretation of in situ wave measurements, and numerical wave forecasting.
+In this paper we revisit the early analytical work by \citeA{longuet1960changes},
+as well as the more recent numerical simulations by \citeA{peureux2021unsteady},
+to re-evaluate the changes in short-wave properties due to the presence of longer waves.
+We present two distinct improvements to the current state of the art:
+First, we derive a higher-order, non-linear solution for the short-wave wavenumber
+and action density as function of the long-wave phase and steepness.
+This solution suggests an increase in short-wave wavenumber by 18 and 38\% for
+long-wave steepenesses of 0.1 and 0.2, respectively, compared to the original
+linear solution.
+It is also in close agreement with the full numerical solutions up to the
+long-wave steepness of 0.25.
+Second, we find that the full numerical solutions, which were previously found
+to be unstable for homogeneous initial conditions, are in fact stable for more
+realistic long-wave groups.
+The key physical interpretation from these results is that the short waves can
+indefinitely grow and steepen only in the presence of inifinite long-wave trains,
+an unlikely occurence in the ocean.
 \end{abstract}
@@
 \section*{Plain Language Summary}
-Plain language summary goes here
-
+The ocean surface is typically covered by waves of many wavelengths.
+Longer waves bunch up and dilate the shorter waves, and this process is
+important to consider for both the measurements and the forecasting of ocean
+waves.
+Although there have been analytical and numerical solutions for how the short
+waves should change in the presence of longer waves, they have been simplified
+and incomplete.
+This paper revisits and extends these solutions to propose a more accurate
+description of how the short waves should change in the presence of longer
+waves.
 \section{Introduction}
@@
 
-Paragraph: Describe the conditions, assumptions, and limitations.
-
-\section{Analytical solutions}
-
-\subsection{Derivation of the original L-HS solution}
-
-The L-HS solution can be most readily obtained by starting from the dispersion
-relationship for short waves:
+In this paper we consider inviscid, irrotational, and incompressible flow.
+In these conditions the linear (sine-like), deep-water, surface gravity waves
+obey the dispersion relation:
 
@@
 \end{equation}
-where $\omega$ is the angular frequency,
-$g$ is the gravitational acceleration,
-and $U$ is the mean advective current in the direction of the wave propagation.
-For simplicity, we will limit the derivation to gravity waves and ignore
-surface tension effects, however, the solution can be as well derived for
-gravity-capillary waves by carrying the surface tension terms.
+where $\omega$ is the angular frequency, $g$ is the gravitational acceleration,
+$k$ is the wavenumber, and $U$ is the mean advective current in the direction
+of the wave propagation.
+A current in the direction of the waves thus increases its absolution frequency,
+wavenumber remaining unchanged.
+An important limitation here is that $U$ must be slow-varying on the scales of
+the wave period.
+In the context of short waves riding on long waves, the mean advective current
+is the horizontal near-surface orbital velocity of the long wave.
 
-Next, take the conservation of wave crests:
+The second important relation is the conservation of wave crests:
 
 \begin{equation}
-\label{eq:wave_crests1}
+\label{eq:wave_crests}
 \dfrac{\partial k}{\partial t}
@@
 \end{equation}
-where $\partial/\partial t$ and $\partial/\partial x$ are partial derivatives
-in time and space, respectively.
-Insert (\ref{eq:dispersion}) into (\ref{eq:wave_crests1}) to get:
+where $t$ and $x$ are time and space, respectively.
+This relation states that the wavenumber must change when the frequency varies
+in space.
+
+Equations (\ref{eq:dispersion}) and (\ref{eq:wave_crests}) only describe the
+evolution of the wavenumber.
+To describe the evolution of the wave amplitude, we use the conservation of wave
+action:
 
 \begin{equation}
-\label{eq:wave_crests2}
-\dfrac{\partial k}{\partial t}
-+ C_g \dfrac{\partial k}{\partial x}
-+ U \dfrac{\partial k}{\partial x}
-+ k \dfrac{\partial U}{\partial x}
+\label{eq:wave_action}
+\dfrac{\partial N}{\partial t}
++ \dfrac{\partial}{\partial x} \left[\left(C_g + U\right)N\right]
 = 0
 \end{equation}
-where $C_g$ is the group velocity of short waves.
-At this step, we assume that $g$ does not vary in space and time.
-We will see later that to a second order $g$ of short waves varies in presence
-of long waves.
-If we follow the short waves in a Lagrangian frame, we can re-write
-(\ref{eq:wave_crests2}) as:
+where $N$ is the action density, and $C_g = \partial \omega / \partial k$ is the
+group speed.
+In deep water, the group speed is half the phase speed, so $C_g = 1/2\sqrt{g/k}$.
+
+As the short waves ride on the surface of the long wave, defined by the elevation
+$\eta_L = a_L cos{\psi}$, where $\psi = k_L x - \omega_L t$ is the long-wave phase,
+they move in an accelerated reference frame and will thus experience effective
+gravitational acceleration that changes in both space and time.
+Inserting equation (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:
 
 \begin{equation}
-\label{eq:wave_crests3}
-\dfrac{d k}{dt}
+\label{eq:wavenumber}
+\dfrac{\partial k}{\partial t}
++ \left(C_g + U\right) \dfrac{\partial k}{\partial x}
 + k \dfrac{\partial U}{\partial x}
++ \dfrac{1}{2} \sqrt{\dfrac{k}{g}} \dfrac{\partial g}{\partial x}
 = 0
 \end{equation}
+
+Similarly, we expand the equation \ref{eq:wave_action} to obtain:
+
+\begin{equation}
+\label{eq:wave_action2}
+\dfrac{\partial N}{\partial t}
++ \left(C_g + U\right) \dfrac{\partial N}{\partial x}
++ N \dfrac{\partial U}{\partial x}
++ N \dfrac{\partial C_g}{\partial x}
+= 0
+\end{equation}
+
+\section{Analytical solutions}
+
+\subsection{Derivation of the original L-HS solution}
 

### 2024-05-09 paper/paper.tex
WIP analytical solutions

@@
 waves.
+
 \section{Introduction}
@@
 \citeA{longuet1960changes} introduced an analytical solution for hydrodynamic
-modulation of short waves by an underlying longer wave:
+modulation of short waves at the crest of an underlying longer wave:
 
@@
 short waves, respectively, and $a_L$ and $k_L$ are amplitude and wavenumber
-of the long wave.
-
+of the long wave, respectively.
 This result has been influential on the development of remote sensing of ocean
@@
 unsteady growth of short-waves due to longer waves.
+
+TODO extend with background.
 
@@
 To describe the evolution of the wave amplitude, we use the conservation of wave
-action:
+action in absence of sources (e.g. due to wind) and sinks (e.g. due to whitecapping):
 
@@
 
-As the short waves ride on the surface of the long wave, defined by the elevation
-$\eta_L = a_L cos{\psi}$, where $\psi = k_L x - \omega_L t$ is the long-wave phase,
+As the short waves ride on the surface of the long wave,
 they move in an accelerated reference frame and will thus experience effective
 gravitational acceleration that changes in both space and time.
-Inserting equation (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:
+Inserting Eq. (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:
 
@@
 \end{equation}
+From left to right, the terms in this equation represent the change in wavenumber
+due to the propagation and advection by ambient current $U$, the convergence of
+the ambient current, and the spatial inhomogeneity of the gravitational acceleration.
+In a non-accelerated reference frame (i.e. in absence of longer waves), the last
+term is zero.
+However, if gravitational acceleration varies in space, this term is necessary
+to conserve the wave crests.
 
-Similarly, we expand the equation \ref{eq:wave_action} to obtain:
+Similar to Eq. \ref{eq:wavenumber}, we expand Eq. \ref{eq:wave_action} to obtain:
 
@@
 \end{equation}
+This equation is similar to Eq. \ref{eq:wavenumber}, except for the last term
+which here represents the inhomogeneity of the group speed.
+As \citeA{peureux2021unsteady} explained, this term is responsible for unstable
+growth of wave action in infinite long-wave trains.
+Eqs. \ref{eq:wavenumber} and \ref{eq:wave_action2} will be the governing equations
+that we solve numerically in section \ref{section:numerical_solutions}.
+
+For the ambient forcing, we consider a single, monochromatic long wave
+defined by the elevation $\eta_L = a_L cos{\psi}$,
+where $a_L$ is the long-wave amplitude, $\psi = k_L x - \omega_L t$ is the
+long-wave phase, and $\omega_L$ its angular frequency.
+At this time we define the long-wave steepness $\varepsilon_L = a_L k_L$ which
+we will use thoughout the rest of the paper as a scaling parameter.
+The long wave induces horizontal and vertical orbital velocities whose values
+at the surface $\eta_L$ are:
+
+\begin{equation}
+\label{eq:U_L}
+U = a_L \omega_L e^{\varepsilon_L \cos{\psi}} \cos{\psi}
+\end{equation}
+
+\begin{equation}
+\label{eq:w_L}
+W = a_L \omega_L e^{\varepsilon_L \cos{\psi}} \sin{\psi}
+\end{equation}
+
+Eqs. \ref{eq:dispersion}-\ref{eq:w_L} is the full system of equations that we
+exploit in the remainder of this paper.
 
 \section{Analytical solutions}
+\label{section:analytical_solutions}
 
-\subsection{Derivation of the original L-HS solution}
+\subsection{Wavenumber and amplitude modulation}
+
+Solving analytically for the wavenumber requires dropping spatial gradients of
+$k$ in Eq. \ref{eq:wavenumber} and assuming homogeneous $g$:
 
 \begin{equation}
-\label{eq:wave_crests4}
-\dfrac{d k}{k}
-= - \dfrac{\partial U}{\partial x} dt
+\label{eq:wavenumber2}
+\dfrac{\partial k}{\partial t}
+= - k \dfrac{\partial U}{\partial x}
 \end{equation}
-which states that the relative contraction of short waves is largest where the
-convergence of the ambient current $U$ is the largest.
-For short waves riding on longer waves, the mean ambient current is the
-horizontal near-surface orbital velocity of the long wave.
-Rather than at the long-wave surface, L-HS took the ambient current to be the
-long-wave orbital velocity at the mean water level, $z = 0$:
+Although \citeA{peureux2021unsteady} pointed out that the solution by L-HS
+requires assuming homogeneous group speed of the short-wave field, in fact,
+it also requires assuming no horizontal propagation and advection of the short
+waves.
+Combining Eqs. \ref{eq:U_L} and \ref{eq:wavenumber2}, and integrating it in time
+yields:
 
 \begin{equation}
-\label{eq:U_L1}
-U = a_L \omega_L \cos{\psi}
+\label{eq:k_short_exact}
+\widetilde{k}(\psi) = k e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
 \end{equation}
-where $\psi$ is the long-wave phase such that the crest is located
-at $\psi = 0$.
-Insert (\ref{eq:U_L1}) into (\ref{eq:wave_crests4}) to obtain:
+
+Notice that the Taylor expansion of Eq. \ref{eq:k_short_exact} to the first order
+recovers the original solution by L-HS:
 
 \begin{equation}
-\label{eq:wave_crests5}
-\dfrac{d k}{k}
-= a_L k_L \omega_L \sin{\psi} dt
+\label{eq:k_short_lhs}
+\widetilde{k}(\psi) = k (1 + \varepsilon_L \cos{\psi})
 \end{equation}
-To calculate the long-wave modulated short-wave wavenumber, integrate
-(\ref{eq:wave_crests5}) from the initial time at which the short-waves
-are unperturbed (at $z=0$, or $\psi=\pi/2$), to time $t$:
 
-\begin{equation}
-\label{eq:wave_crests6}
-\int_k^{\ \widetilde k(t)} \dfrac{d k'}{k'}
-= \int_0^t a_L k_L \omega_L \sin{\psi} dt'
-\end{equation}
-which yields:
-
-\begin{equation}
-\label{eq:k_short1}
-\ \widetilde k(\psi) = k e^{a_L k_L \cos{\psi}}
-\end{equation}
-Approximating (\ref{eq:k_short1}) to the first-order Taylor expansion,
-we obtain the original L-HS solution:
-
-\begin{equation}
-\label{eq:k_short2}
-\ \widetilde k(\psi) = k (1 + a_L k_L \cos{\psi})
-\end{equation}
-The intuitive interpretation of (\ref{eq:k_short2}) is that as the short waves
-contract on the front face of the long wave, they reach maximum contraction
-(highest $\ \widetilde k$) at the crest of the long wave.
-As they pass to the rear face of the long wave, they get elongated and reach
-minimum contraction (lowest $\ \widetilde k$) in the trough of the long wave.
-
-\subsection{Extending the L-HS solution}
-
-L-HS made two approximations in their solution:
-First, they evaluated the advective velocity of short waves at $z = 0$
-rather than the surface, $z = \eta_L$, where $\eta_L$ is the long-wave
-elevation; and second, they evaluated the solution to the first order
+The L-HS solution requires two approximations in their solution:
+First, evaluating the long-wave velocity at $z = 0$
+rather than $z = \eta_L$; and second, evaluating the solution to the first order
 of the Taylor expansion series.
 Both of these approximations cause an underestimate of the short-wave
-modulation magnitude.
-
-To extend the L-HS solution, evaluate the advective velocity at the surface
-of the long-wave:
-
-\begin{equation}
-\label{eq:U_L2}
-U = U_L(z=\eta) = a_L \omega_L e^{a_L k_L \cos{\psi}} \cos{\psi}
-\end{equation}
-and insert it into \ref{eq:wave_crests4} to obtain:
-
-\begin{equation}
-\label{eq:wave_crests7}
-\dfrac{dk}{k} = a_L k_L \omega_L e^{a_L k_L \cos{\psi}} \sin{\psi} dt
-\end{equation}
-Integrate (\ref{eq:wave_crests7}) to the same limits as we did with
-(\ref{eq:wave_crests6}) to obtain:
-
-\begin{equation}
-\label{eq:k_short3}
-\widetilde{k}(\psi) = k e^{a_L k_L \cos{\psi} e^{a_L k_L \cos{\psi}}}
-\end{equation}
-
-We compare the solutions from (\ref{eq:k_short1}), (\ref{eq:k_short2}),
-and (\ref{eq:k_short3}) for a moderately steep long wave with steepness
-$a_L k_L = 0.2$ in Fig. (\ref{fig:wavenumber_modulation}).
+modulation magnitude (Fig. \ref{fig:wavenumber_modulation}).
 The original L-HS solution is a cosine wave with an amplitude equal
@@
 1st order approximate L-HS solution.
-Also, since the less-approximate solutions decrease the modulation in the
-troughs and increase in on the crests, the average wavenumber modulation
-is no longer 0, but 1\% in the case of the exact L-HS solution and 3.1\% in the
-case of the extended solution.
 
@@
   Comparison of relative short-wave modulation as function of the
-  long-wave phase, for a moderately steep long wave of $a_L k_L = 0.2$.
+  long-wave phase, for a moderately steep long wave of $a_L k_L = 0.1$.
   The dotted line shows the 1st-order approximated solution by L-HS,
@@
 
-\section{Apparent gravity}
+\subsection{Apparent gravity}
 
@@
 
-\section{Lagrangian averaged short-wave properties}
+\subsection{Lagrangian averaged short-wave properties}
 
@@
 \section{Numerical solutions}
+\label{section:numerical_solutions}
+
+\section{Scale analysis}
+\label{section:scale_analysis}
+
+Now that we showed that the analytical solution is consistent with full numerical
+solutions for $\varepsilon_L < 0.25$, we can use analytical solutions for scale
+analysis to quantify the relative contribution of tendencies in the full equation set.
+
+TODO
 
@@
 \acknowledgments
-We thank anonymous reviewers for their help improving this paper.
+Thanks to Peisen Tan, Nathan Laxague, and Fabrice Ardhuin for providing helpful
+feedback on the early drafts of this work.
 

### 2024-05-10 paper/scripts/plot_gravity_modulation.py
Update wavenumber and gravity modulation plots

@@ -52,23 +52,29 @@ def plot_fig_apparent_gravity():
     g_lagrangian_3 = g_lagrangian(g, phase, a_L, k_L, order=3)
     # g_lagrangian_4 = g_lagrangian(g, phase, a_L, k_L, order=4) # Essentially the same as 3rd order; omit
 
-    fig = plt.figure(figsize=(8, 6))
-    ax = fig.add_subplot(111)
-    ax.plot(phase, g_eulerian_1 / g, label=r"Eulerian, 1$^{st}$ order", lw=2)
-    ax.plot(phase, g_eulerian_2 / g, label=r"Eulerian, 2$^{nd}$ order", lw=2)
-    ax.plot(phase, g_eulerian_3 / g, label=r"Eulerian, 3$^{rd}$ order", lw=2)
-    ax.plot(phase, g_lagrangian_3 / g, label=r"Lagrangian, 3$^{rd}$ order", lw=2)
-    ax.legend()
-    ax.set_xlabel("Phase [rad]")
-    ax.set_ylabel(r"$\widetilde g/g$")
-    ax.grid()
-    ax.plot([0, 2 * np.pi], [1, 1], "k--")
-    ax.set_xlim(0, 2 * np.pi)
-    ax.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
-    ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
-    ax.set_title(r"$\varepsilon_L = 0.2$")
-    plt.savefig("../figures/fig_gravity_modulation_by_phase.png", dpi=200)
-    plt.close(fig)
+    fig = plt.figure(figsize=(14, 6))
+    ax1 = fig.add_subplot(121)
+    ax2 = fig.add_subplot(122)
+
+    ax1.text(
+        0, 1.02, "(a)", ha="left", va="bottom", transform=ax1.transAxes, fontsize=20
+    )
+    ax2.text(
+        0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
+    )
+
+    ax1.plot(phase, g_eulerian_1 / g, label=r"Eulerian, 1$^{st}$ order", lw=2)
[…]

### 2024-05-10 paper/scripts/plot_wavenumber_modulation.py

@@ -1,6 +1,5 @@
 import matplotlib.pyplot as plt
 import numpy as np
-from modulation import wavenumber
 import matplotlib
 
 matplotlib.rc("font", size=16)
@@ -11,28 +10,33 @@ def plot_fig_wavenumber_modulation():
 
     phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
     k_L = 1
-    a_L = 0.1
+    a_L = 0.2
     ak_L = a_L * k_L
 
     solution1 = 1 + ak_L * np.cos(phase)
-    solution2 = np.exp(ak_L * np.cos(phase))
     solution3 = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
 
-    fig = plt.figure(figsize=(8, 6))
-    ax = fig.add_subplot(111)
-    ax.plot(phase, solution1, "k-", lw=2, label="L-H&S")
-    ax.plot(phase, solution3, "r-", lw=2, label="This paper")
-    ax.legend()
-    ax.set_xlabel("Phase [rad]")
-    ax.set_ylabel(r"$\widetilde k/k$")
-    ax.grid()
-    ax.plot([0, 2 * np.pi], [1, 1], "k--")
-    ax.set_xlim(0, 2 * np.pi)
-    ax.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
-    ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
-    ax.set_title(r"$\varepsilon_L = 0.1$")
-    plt.savefig("../figures/fig_wavenumber_modulation_by_phase.png", dpi=200)
-    plt.close(fig)
+    fig = plt.figure(figsize=(14, 6))
+    ax1 = fig.add_subplot(121)
+    ax2 = fig.add_subplot(122)
+
+    ax1.text(
+        0, 1.02, "(a)", ha="left", va="bottom", transform=ax1.transAxes, fontsize=20
+    )
+    ax2.text(
+        0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
+    )
+
+    ax1.plot(phase, solution1, "k-", lw=2, label="L-H&S")
+    ax1.plot(phase, solution3, "r-", lw=2, label="This paper")
+    ax1.legend()
[…]

### 2024-05-10 paper/paper.tex
Add references

@@
 An important limitation here is that $U$ must be slow-varying on the scales of
-the wave period.
+the wave period \cite{bretherton1968wavetrains}.
 In the context of short waves riding on long waves, the mean advective current
@@
 they move in an accelerated reference frame and will thus experience effective
-gravitational acceleration that changes in both space and time.
+gravitational acceleration that changes in both space and time \cite{longuet1986eulerian,longuet1987propagation}.
 Inserting Eq. (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:
@@
 \subsection{Wavenumber and amplitude modulation}
+\label{subsection:analytical_solutions}
 
@@
 \centering
-\includegraphics[width=\textwidth]{figures/fig_wavenumber_modulation_by_phase.png}
+\includegraphics[width=\textwidth]{figures/fig_wavenumber_modulation_2panel.png}
 \caption{
   Comparison of relative short-wave modulation as function of the
-  long-wave phase, for a moderately steep long wave of $a_L k_L = 0.1$.
+  long-wave phase, for a moderately steep long wave of $\varepsilon = 0.1$.
   The dotted line shows the 1st-order approximated solution by L-HS,
@@
 
-\subsection{Apparent gravity}
+\subsection{Gravity modulation}
+\label{subsection:gravity_modulation}
 
-Long waves also induce orbital acceleration under their surface, so the short
-waves experience an apparent gravity:
+Long waves induce vertical orbital accelerations, so the short
+waves that ride on their surface move in an accelerated reference frame and
+experience effective gravitational acceleration that is lower at the crests and
+higher in the troughs \cite{longuet1986eulerian,longuet1987propagation}.
+A linearized (sine-like) wave has a modulation in gravity acceleration at its
+surface:
 
 \begin{equation}
-\label{eq:apparent_gravity1}
-\widetilde{g} = g + \dfrac{\partial w_L}{\partial t}
-\end{equation}
-The positive sign in front of $\partial w_L/\partial t$ comes from our
-coordinate-system convention where $w$ is defined positive-upward and $g$ is
-defined positive-downward, and the fact that $\widetilde{g}$ is the apparent
-gravity in the reference frame of the short waves.
-Taking the long-wave vertical acceleration at the surface as:
-
-\begin{equation}
-\label{eq:w_L}
-w_L = - a_L \omega_L e^{a_L k_L \cos{\psi}} \sin{\psi}
-\end{equation}
-we arrive at:
-
-\begin{equation}
-\label{eq:apparent_gravity2}
+\label{eq:effective_gravity}
 \widetilde{g}(\psi) = g \left[
-  1 - a_L k_L e^{a_L k_L \cos{\psi}} \left(
-    \cos{\psi} + a_L k_L \sin^2{\psi}
+  1 - \varepsilon_L e^{\varepsilon_L \cos{\psi}} \left(
+    \cos{\psi} + \varepsilon_L \sin^2{\psi}
   \right)
@@
 \begin{figure}[h]
-\label{fig:apparent_gravity}
+\label{fig:gravity_modulation}
 \centering
-\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_by_phase.png}
+\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_2panel.png}
 \caption{
@@
   as a function of the long-wave phase and evaluated to the first (dashed line) and second (solid line) order in $a_L k_L$.
-}
-\end{figure}
-
-\begin{figure}[h]
-\label{fig:gravity_modulation_vs_ak}
-\centering
-\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_by_ak.png}
-\caption{
-  TODO
 }

### 2024-05-14 paper/paper.tex
Fix sign in 3rd order grav. accel.

@@
 Equations (\ref{eq:dispersion}) and (\ref{eq:wave_crests}) only describe the
-evolution of the wavenumber.
+evolution of the short-wave wavenumber.
 To describe the evolution of the wave amplitude, we use the conservation of wave
@@
 
-For the ambient forcing, we consider a single, monochromatic long wave
-defined by the elevation $\eta_L = a_L cos{\psi}$,
+For the ambient forcing, we consider a train of monochromatic long wave
+defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
 where $a_L$ is the long-wave amplitude, $\psi = k_L x - \omega_L t$ is the
 long-wave phase, and $\omega_L$ its angular frequency.
+
 At this time we define the long-wave steepness $\varepsilon_L = a_L k_L$ which

### 2024-05-14 paper/scripts/modulation.py

@@ -8,7 +8,7 @@ def gravity(
     phi = phase(x, t, k, omega)
     term1 = -a * k * np.cos(phi)
     term2 = -a * k * np.cos(phi) * np.exp(a * k * np.cos(phi))
-    term3 = -((a * k * np.sin(phi)) ** 2) * np.exp(a * k * np.cos(phi))
+    term3 = ((a * k * np.sin(phi)) ** 2) * np.exp(a * k * np.cos(phi))
     if order == 0:
         return g0 * np.ones_like(x)
     elif order == 1:

### 2024-05-14 paper/scripts/plot_gravity_modulation.py

@@ -14,7 +14,7 @@ def g_eulerian(g, psi, a, k, order=3):
         res = g * (1 - a * k * np.cos(psi) * np.exp(k * eta))
     else:
         res = g * (
-            1 - a * k * np.exp(k * eta) * (np.cos(psi) + a * k * np.sin(psi) ** 2)
+            1 - a * k * np.exp(k * eta) * (np.cos(psi) - a * k * np.sin(psi) ** 2)
         )
     return res
 
@@ -46,11 +46,8 @@ def plot_fig_apparent_gravity():
 
     g = 9.8
 
-    g_eulerian_1 = g_eulerian(g, phase, a_L, k_L, order=1)
-    g_eulerian_2 = g_eulerian(g, phase, a_L, k_L, order=2)
-    g_eulerian_3 = g_eulerian(g, phase, a_L, k_L, order=3)
-    g_lagrangian_3 = g_lagrangian(g, phase, a_L, k_L, order=3)
-    # g_lagrangian_4 = g_lagrangian(g, phase, a_L, k_L, order=4) # Essentially the same as 3rd order; omit
+    g_E1 = g_eulerian(g, phase, a_L, k_L, order=1)
+    g_E3 = g_eulerian(g, phase, a_L, k_L, order=3)
 
     fig = plt.figure(figsize=(14, 6))
     ax1 = fig.add_subplot(121)
@@ -63,10 +60,8 @@ def plot_fig_apparent_gravity():
         0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
     )
 
-    ax1.plot(phase, g_eulerian_1 / g, label=r"Eulerian, 1$^{st}$ order", lw=2)
-    ax1.plot(phase, g_eulerian_2 / g, label=r"Eulerian, 2$^{nd}$ order", lw=2)
-    ax1.plot(phase, g_eulerian_3 / g, label=r"Eulerian, 3$^{rd}$ order", lw=2)
-    ax1.plot(phase, g_lagrangian_3 / g, label=r"Lagrangian, 3$^{rd}$ order", lw=2)
+    ax1.plot(phase, g_E1 / g, "k-", label=r"1$^{st}$ order", lw=2)
+    ax1.plot(phase, g_E3 / g, "r-", label=r"3$^{rd}$ order", lw=2)
     ax1.legend()
     ax1.set_xlabel("Phase (rad.)")
[…]

### 2024-05-14 paper/scripts/plot_wavenumber_modulation.py

@@ -40,20 +40,24 @@ def plot_fig_wavenumber_modulation():
 
     # Evaluate wa enumber modulation as function of ak
     ak_range = np.arange(0, 0.41, 0.01)
-    k1_mean, k1_max = [], []
-    k2_mean, k2_max = [], []
+    k1_mean, k1_max, k1_min = [], [], []
+    k2_mean, k2_max, k2_min = [], [], []
     for ak in ak_range:
         k1 = 1 + ak * np.cos(phase)
         k2 = np.exp(ak * np.cos(phase) * np.exp(ak * np.cos(phase)))
         k1_mean.append(np.mean(k1))
         k1_max.append(np.max(k1))
+        k1_min.append(np.min(k1))
         k2_mean.append(np.mean(k2))
         k2_max.append(np.max(k2))
+        k2_min.append(np.min(k2))
 
-    ax2.plot(ak_range, k1_max, "k-", label="L-H&S, max", lw=2)
+    ax2.plot(ak_range, k1_max, "k-", label="L-H&S, max.", lw=2)
     ax2.plot(ak_range, k1_mean, "k--", label="L-H&S, mean", lw=2)
-    ax2.plot(ak_range, k2_max, "r-", label="This paper, max", lw=2)
+    ax2.plot(ak_range, k1_min, "k:", label="L-H&S, min.", lw=2)
+    ax2.plot(ak_range, k2_max, "r-", label="This paper, max.", lw=2)
     ax2.plot(ak_range, k2_mean, "r--", label="This paper, mean", lw=2)
+    ax2.plot(ak_range, k2_min, "r:", label="This paper, min.", lw=2)
     ax2.legend(loc="upper left")
     ax2.grid()
     ax2.set_xlabel(r"$\varepsilon_L$")

### 2024-05-17 paper/scripts/plot_frequency_modulation.py
Frequency modulation 2-panel

@@ -10,29 +10,70 @@ def plot_fig_frequency_modulation():
     phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
     g = 9.8
 
-    fig = plt.figure(figsize=(8, 6))
-    ax = fig.add_subplot(111)
+    ak_L = 0.2
 
-    for ak_L in [0.1, 0.2, 0.3, 0.4]:
+    g1 = 1 - ak_L * np.cos(phase)
+    k1 = 1 + ak_L * np.cos(phase)
+    omega_rel1 = np.sqrt(g1 * k1)
+
+    term1 = -ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase))
+    term2 = ((ak_L * np.sin(phase)) ** 2) * np.exp(ak_L * np.cos(phase))
+    g2 = 1 + term1 + term2
+    k2 = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
+    omega_rel2 = np.sqrt(g2 * k2)
+
+    fig = plt.figure(figsize=(14, 6))
+    ax1 = fig.add_subplot(121)
+    ax2 = fig.add_subplot(122)
+
+    ax1.plot(phase, omega_rel1, "k-", label=r"1$^{st}$ order", lw=2)
+    ax1.plot(phase, omega_rel2, "r-", label=r"3$^{rd}$ order", lw=2)
+    ax1.legend()
+    ax1.set_xlabel("Phase [rad]")
+    ax1.set_ylabel(r"$\widetilde \omega/\omega$")
+    ax1.grid()
+    ax1.plot([0, 2 * np.pi], [1, 1], "k--")
+    ax1.set_xlim(0, 2 * np.pi)
+    ax1.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
+    ax1.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
+
+    # Evaluate frequency modulation as function of ak
+    ak_range = np.arange(0, 0.41, 0.01)
+    omega1_mean, omega1_max, omega1_min = [], [], []
+    omega2_mean, omega2_max, omega2_min = [], [], []
+    for ak_L in ak_range:
+        g1 = 1 - ak_L * np.cos(phase)
+        k1 = 1 + ak_L * np.cos(phase)
+        omega1 = np.sqrt(g1 * k1)
+
[…]

### 2024-05-22 paper/scripts/plot_model_infinite.py
Script to plot modulation 3-panel using the model from the 2wave library

@@ -1,0 +1,61 @@
+import matplotlib.pyplot as plt
+from model import WaveModulationModel
+import numpy as np
+
+
+def plot_modulation_3panel(model, contour_levels, title):
+    ds = model.to_xarray()
+
+    num_periods = 10
+    time = ds.time / model.T_long
+    ak = ds.amplitude * ds.wavenumber
+
+    fig = plt.figure(figsize=(8, 6))
+    axes = fig.subplots(1, 3, sharey=True)
+
+    axes[0].contourf(ds.space, time, ds.wave_action, contour_levels)
+    axes[0].contour(ds.space, time, ds.wave_action, [1], colors="k", linewidths=0.5)
+    axes[0].set_ylabel("Time (long-wave periods)")
+    axes[0].set_title(r"$N$")
+
+    axes[1].contourf(ds.space, time, ds.wavenumber / model.k_short, contour_levels)
+    axes[1].contour(
+        ds.space, time, ds.wavenumber / model.k_short, [1], colors="k", linewidths=0.5
+    )
+    axes[1].set_title(r"$k$")
+
+    cm = axes[2].contourf(
+        ds.space, time, ak / model.k_short / model.a_short, contour_levels
+    )
+    axes[2].contour(
+        ds.space,
+        time,
+        ak / model.k_short / model.a_short,
+        [1],
+        colors="k",
+        linewidths=0.5,
+    )
+    axes[2].set_title(r"$ak$")
+
+    for ax in axes:
+        ax.set_yticks(range(0, num_periods + 1, 1))
+        ax.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
+        ax.set_xticklabels([r"$0$", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
+        ax.set_xlabel("Phase")
+
+    fig.subplots_adjust(left=0.1, right=0.9, bottom=0.0, top=0.9)
+    cbar = fig.colorbar(
+        cm, ax=axes, orientation="horizontal", pad=0.15, shrink=0.7, label="Modulation"
+    )
+
+    plt.savefig(
+        f"../figures/fig_modulation_3panel_{title.lower().replace(' ', '_')}.png",
+        dpi=200,
+    )
+    plt.close(fig)
+
+
+m = WaveModulationModel()
+m.run()
+
+plot_modulation_3panel(m, np.arange(0.5, 2.2, 0.1), "Infinite wave train")

### 2024-05-23 paper/paper.tex
Frequency modulation 2-panel
Start working on scale analysis

@@
 
+TODO amplitude modulation;
+
 \subsection{Gravity modulation}
@@
 \label{eq:effective_gravity}
-\widetilde{g}(\psi) = g \left[
+\widetilde{g}(\psi) = g + \dfrac{\partial W}{\partial t} = g \left[
   1 - \varepsilon_L e^{\varepsilon_L \cos{\psi}} \left(
-    \cos{\psi} + \varepsilon_L \sin^2{\psi}
+    \cos{\psi} - \varepsilon_L \sin^2{\psi}
   \right)
@@
 \centering
-\includegraphics[width=\textwidth]{figures/fig_frequency_modulation.png}
+\includegraphics[width=\textwidth]{figures/fig_frequency_modulation_2panel.png}
 \caption{
@@
 \end{figure}
+
+\subsection{Frequency and group speed modulation}
 
@@
 
+We now proceed to integrate the full numerical solutions for the short-wave
+wavenumber and action density in presence of long waves.
+Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
+2nd order central finite difference in space and integrated in time using the
+4th order Runge-Kutta method \cite{butcher1996runge}.
+Although the 2nd order central finite difference is not appropriate for many
+numerical problems, in our case it is sufficient because it is conservative and
+the fields that we compute derivatives of are smooth and continuous.
+The space is divided into 100 grid points.
+
+\subsection{Unstable growth in infinite long-wave trains}
+\label{subsection:unstable_growth}
+
+\begin{figure}[h]
+\label{fig:frequency_modulation}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_infinite_wave_train.png}
+\caption{
+  TODO
+}
+\end{figure}
+
+\begin{figure}[h]
+\label{fig:frequency_modulation}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_linear_ramp.png}
+\caption{
+  TODO
+}
+\end{figure}
+
+\subsection{Short-wave modulation by long-wave groups}
+\label{subsection:wave_groups}
+
+\begin{figure}[h]
+\label{fig:frequency_modulation}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_wave_group.png}
+\caption{
+  TODO
+}
+\end{figure}
+
 \section{Scale analysis}

### 2024-05-23 paper/scripts/plot_frequency_modulation.py

@@ -26,6 +26,13 @@ def plot_fig_frequency_modulation():
     ax1 = fig.add_subplot(121)
     ax2 = fig.add_subplot(122)
 
+    ax1.text(
+        0, 1.02, "(a)", ha="left", va="bottom", transform=ax1.transAxes, fontsize=20
+    )
+    ax2.text(
+        0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
+    )
+
     ax1.plot(phase, omega_rel1, "k-", label=r"1$^{st}$ order", lw=2)
     ax1.plot(phase, omega_rel2, "r-", label=r"3$^{rd}$ order", lw=2)
     ax1.legend()

### 2024-05-23 paper/scripts/plot_numerical_simulations.py

@@ -56,6 +56,11 @@ def plot_modulation_3panel(model, contour_levels, title):
 
 
 m = WaveModulationModel()
-m.run()
+m.run(ramp_type=None)
+plot_modulation_3panel(m, np.arange(0.5, 2.3, 0.1), "Infinite wave train")
 
-plot_modulation_3panel(m, np.arange(0.5, 2.2, 0.1), "Infinite wave train")
+m.run(ramp_type="linear")
+plot_modulation_3panel(m, np.arange(0.8, 1.24, 0.04), "Linear ramp")
+
+m.run(ramp_type="groups")
+plot_modulation_3panel(m, np.arange(0.8, 1.24, 0.04), "Wave group")

### 2024-05-23 paper/scripts/scale_analysis.py

@@ -1,0 +1,40 @@
+import matplotlib.pyplot as plt
+from model import diff, WaveModulationModel
+import numpy as np
+
+phase = np.linspace(0, 2 * np.pi, 100, endpoint=False)
+a_long = 0.1
+k_long = 1
+ak = a_long * k_long
+x = phase / k_long
+dx = x[1] - x[0]
+
+k0 = 10
+g0 = 9.8
+
+omega_long = np.sqrt(g0 * k_long)
+U = a_long * omega_long * np.cos(phase) * np.exp(ak * np.cos(phase))
+
+k = k0 * np.exp(ak * np.cos(phase) * np.exp(ak * np.cos(phase)))
+g = g0 * (
+    1 - ak * np.exp(ak * np.cos(phase)) * (np.cos(phase) + ak * np.sin(phase) ** 2)
+)
+omega = np.sqrt(g * k)
+Cg = 0.5 * omega / k
+
+propagation = -Cg * diff(k) / dx
+advection = -U * diff(k) / dx
+convergence = -k * diff(U) / dx
+inhomogeneity = -0.5 * np.sqrt(k / g) * diff(g) / dx
+
+
+m = WaveModulationModel(a_long=a_long, k_long=k_long)
+m.run(ramp_type="groups")
+ds = m.to_xarray()
+
+fig = plt.figure(figsize=(8, 6))
+ax = fig.add_subplot(111)
+ax.plot(phase, k0 * (1 + ak * np.cos(phase)))
+ax.plot(phase, k)
+ax.plot(phase, ds.wavenumber[1004, :])
+ax.grid()

### 2024-06-07 paper/paper.tex
Updates on numerical results

@@
 TODO amplitude modulation;
+The modulation of short-wave amplitude can be derived in the same way by
+dropping the non-linear terms in the wave action balance equation
+(Eq. \ref{eq:wave_action2}) and integrating wave action in time:
+
+\begin{equation}
+\label{eq:wave_action_modulation}
+\widetilde{N}(\psi) = N e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
+\end{equation}
+
+Since wave action is variance divided by the intrinsic frequency $\sigma$,
+and variance scales with the square of the amplitude, the modulation of
+amplitude is proportional to the square root of the modulation of wave action.
+
+\begin{equation}
+\label{eq:wave_amplitude_modulation}
+\widetilde{a}(\psi) = a \sqrt{\dfrac{\sigma}{\widetilde{\sigma}} \dfrac{\widetilde{N}}{N}}
+\end{equation}
+To determine $\widetilde{\sigma}$, we will need to evaluate the modulation of
+the gravitational acceleration at the surface of the long wave.
 
@@
 higher in the troughs \cite{longuet1986eulerian,longuet1987propagation}.
-A linearized (sine-like) wave has a modulation in gravity acceleration at its
-surface:
+A linear wave causes a modulation in gravity acceleration at its surface:
 
 \begin{equation}
-\label{eq:effective_gravity}
+\label{eq:gravity_modulation_linear}
 \widetilde{g}(\psi) = g + \dfrac{\partial W}{\partial t} = g \left[
@@
 The second-order term thus increases the apparent gravity on the front and rear
-faces of the long wave (Fig. \ref{fig:apparent_gravity}).
+faces of the long wave (Fig. \ref{fig:gravity_modulation}).
 
@@
 
+We can also derive the short-wave gravitational acceleration on the surface of
+a non-linear Stokes wave with the elevation (to the third order in $\varepsilon_L$):
+
+\begin{equation}
+\label{eq:eta_stokes}
+\eta_{St} = a_L \left[
+  \cos{\psi} +
+  \dfrac{1}{2} \varepsilon_L \cos{2\psi} +
+  \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
+\right]
+\end{equation}
+
+The corresponding gravitational acceleration at the surface of the Stokes wave
+is:
+
+\begin{equation}
+\label{eq:gravity_modulation_stokes}
+\widetilde{g}(\psi) =
+g \left\{
+  1 - \varepsilon_L e^{k \eta_{St}}
+  \left[ \cos{\psi} -
+    \varepsilon_L \sin{\psi} \left(
+      \sin{\psi}
+      + \varepsilon_L \sin{2\psi}
+      - \dfrac{1}{16} \varepsilon_L^2 \sin{\psi}
+      + \dfrac{9}{8} \varepsilon_L^2 \sin{3\psi}
+    \right)
+  \right]
+\right\}
+\end{equation}
+
 With the new long-wave modulated expressions for the short-wave wavenumber
-(Eq. \ref{eq:k_short3}) and gravity (Eq. \ref{eq:apparent_gravity2}), we can
-evaluate the long-wave modulated frequency as simply
-$\widetilde{\omega}(\psi) = \sqrt{\widetilde{g} \widetilde{k}}$, which we compare
-for four different values of $a_L k_L$ in Fig. \ref{fig:frequency_modulation}.
+(Eq. \ref{eq:k_short_exact}) and gravity
+(Eqs. \ref{eq:gravity_modulation_linear}-\ref{eq:gravity_modulation_stokes}),
+we can evaluate the long-wave modulated frequency as
+$\widetilde{\omega}(\psi) = \sqrt{\widetilde{g} \widetilde{k}}$.
+We compare the modulated intrinsic frequency for short waves on the surface of
+linear and Stokes waves with $\varepsilon_L=2$ as function of phase, and their
+phase-integral statistics in Fig. \ref{fig:frequency_modulation}.
 
@@
 \caption{
-  Relative change in short-wave frequency due to the presence of longer
-  waves, for four different values of long-wave steepness.
+  (a) Relative change in the short-wave intrinsic frequency on the surface of a
+  longer linear (black) and Stokes (red) wave of steepness $\varepsilon_L = 0.2$
+  as function of long-wave phase;
+  (b) Long-wave phase maximum (solid), mean (dashed), and minimum (dotted)
+  relative change in the short-wave intrinsic frequency as function of long-wave
+  steepness in the case of a linear (black) and Stokes (red) long wave.
 }
 \end{figure}
-
-\subsection{Frequency and group speed modulation}
 
@@
 We now proceed to integrate the full numerical solutions for the short-wave
-wavenumber and action density in presence of long waves.
+wavenumber and action density in presence of long waves, and in absence of
+sources or sinks of wave energy.
 Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
-2nd order central finite difference in space and integrated in time using the
-4th order Runge-Kutta method \cite{butcher1996runge}.
-Although the 2nd order central finite difference is not appropriate for many
+2$^{nd}$  order central finite difference in space and integrated in time using the
+4$^{th}$ order Runge-Kutta method \cite{butcher1996runge}.
+The space is divided into 100 grid points.
+This is effectively the same equation set and numerical configuration as that of
+\citeA{peureux2021unsteady}, except that, like \citeA{longuet1960changes}, their
+long-wave orbital velocities are evaluated at $z=0$ instead of $z=\eta_L$.
+The key difference is that evaluating orbital velocities at $z=0$ neglects the
+Stokes drift of short-wave groups.
+Although not significantly consequential for the modulation results here, there
+is no good reason to exclude it given that its implementation is trivial.
+
+Another difference is the choice of the numerical scheme for spatial differences,
+which in their case is the MUSCL4 scheme by Kurganov and Tadmor (2000 TODO).
+Although the 2$^{nd}$ order central finite difference is not appropriate for many
 numerical problems, in our case it is sufficient because it is conservative and
 the fields that we compute derivatives of are smooth and continuous.
-The space is divided into 100 grid points.
+We show in the next section that our numerical results in infinite long-wave
+trains are qualitatively equivalent to those of \citeA{peureux2021unsteady}.
+
+Finally, the numerical equations here are integrated in a fixed reference frame,
+rather than one that moves with the long wave phase speed.
+Either approach produces qualitatively equivalent results, but we find that
+a fixed reference frame allows for more intuitive interpretation of the results.
+However, this difference is important to keep in mind when comparing the results
+of \citeA{peureux2021unsteady} and those presented here--in their Fig. 1 the
+long-wave phase is fixed and the short waves are moving leftward
+with the speed of $C_{pL} - C_g - U$ (neglecting any group speed inhomogeneity),
+whereas in the figures that follow, the long wave is moving rightward with its
+phase speed $C_{pL}$ and the short waves (that is, their energy) are moving
+rightward with the speed of $C_g + U$ (neglecting any group wave inhomogeneity).
 
@@
 
+The key result of \citeA{peureux2021unsteady} is that of unsteady steepening of
+short waves when a homogeneous field of short waves is forced by an infinite
+long-wave train.
+The steepening is driven by the progressive increase in the action of short
+waves (as opposed to their wavenumber), and is centered on the group of short
+waves that begins its journey in the trough of the long wave.
+This is because the unmodulated short waves that begin in the trough enter into
+the area where the surface velocities are converging.
+Inversely, the short waves that begin around the crest of the long-wave will
+enter into the area of surface divergence and will thus have their wave action
+and wavenumber, decreased.
+We reproduce this behavior in Fig. \ref{fig:modulation_3panel_infinite}, which
+shows the evolution of the short-wave action, wavenumber, and steepness
+modulation (here defined as the change of a quantity relative to its initial value).
+After 10 long-wave periods, the short-wave action modulation $\widetilde{N}/N$
+is approximately doubled, which almost exactly corresponds to the unsteady growth
+reported by \citeA{peureux2021unsteady} in their Fig. 1.
+Another important property of this solution is that the short-wave group that
+experiences peak amplification (dampening) are not locked-in to the long-wave
+crests (troughs), as they are in the L-HS solution.
+
 \begin{figure}[h]
-\label{fig:frequency_modulation}
+\label{fig:modulation_3panel_infinite}
 \centering
@@
 \caption{
-  TODO
+  Change in the short-wave action (left), wavenumber (middle), and steepness (right)
+  relative to their initial values as function of initial long-wave phase and time.
+  $k_L = 1$, $k = 10$, $\varepsilon_L = 0.1$.
 }
@@
 
+Why does this unsteady growth of short-wave action occur?
+As \citeA{peureux2021unsteady} explain, "the additional change in advection
+velocity due to the modulation of the short wave group speed gives an additional
+amplification."
+That is, it is the inhomogeneity term $N \dfrac{\partial C_g}{\partial x}$ that
+is responsible for the instability.
+However, the instability only occurs if the short waves are initialized as a
+uniform field.
+If they are initialized instead as a periodic function of the long-wave phase
+such that their wavenumber (and optionally, action) is higher on the crests
+and lower in the troughs, akin to the L-HS solution, the instability does not
+occur.
+
 \begin{figure}[h]
-\label{fig:frequency_modulation}
+\label{fig:modulation_3panel_ramp}
 \centering
@@
 \caption{
-  TODO
+  Same as Fig. \ref{fig:modulation_3panel_infinite} but with a linear ramp
+  applied to $a_L$ during the initial five long-wave periods.
+}
+\end{figure}
+
+Why is it, then, that the solution is so different depending on the short-wave
+initial conditions?
+\citeA{peureux2021unsteady} did not provide an explanation for this behavior.
+At first, it may seem that there is something special about the uniform initial
+conditions that makes its solutions grow unsteadily and indefinitely.
+\citeA{peureux2021unsteady} describe this scenario as "...the sudden appearance
+of a long wave perturbation in the middle of a homogeneous short wave field".
+The sudden appearance of long waves may be a clue.
+The exact numerical simulation as shown in Fig. \ref{fig:modulation_3panel_infinite},
+except for a linear ramp applied to $a_L$ (and consequently, $U$ and $W$ as well)
+for the first five periods reveals that the unsteady growh no longer occurs
+(Fig. \ref{fig:modulation_3panel_ramp}).
+
+\begin{figure}[h]
+  \label{fig:modulation_3panel_ramp}
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig_modulation_ramp_timeseries.png}
+  \caption{
+    Maximum short-wave steepness modulation as function of time in case of
+    infinite long-wave train case (blue) and the same case but with a linear
+    ramp during the initial five long-wave periods (orange). The dashed black
+    line corresponds to the short-wave steepness of 0.4 at which most waves are
+    expected to break.
   }
@@
 \begin{figure}[h]
-\label{fig:frequency_modulation}
+\label{fig:modulation_3panel_groups}
 \centering
@@
 \caption{
-  TODO
+  Same as Fig. \ref{fig:modulation_3panel_infinite} but for a long-wave group,
+  causing the long-wave amplitude to gradually increase and peak at $a_L = 0.1$
+  after five long-wave periods, and then conversely decay back to a calm sea state.
 }

### 2024-06-07 paper/scripts/scale_analysis.py

@@ -27,7 +27,6 @@ advection = -U * diff(k) / dx
 convergence = -k * diff(U) / dx
 inhomogeneity = -0.5 * np.sqrt(k / g) * diff(g) / dx
 
-
 m = WaveModulationModel(a_long=a_long, k_long=k_long)
 m.run(ramp_type="groups")
 ds = m.to_xarray()

### 2024-06-20 paper/paper.tex
Small additions

@@
 
-\title{Dispersion of short waves riding on longer waves}
+%\title{Dispersion of short waves riding on longer waves}
+\title{Revisiting the linear theory of hydrodynamic wave modulation}
 
@@
 
-TODO extend with background.
+Working through the theory of \citeA{longuet1960changes,longuet1987propagation}
+and the numerical simulation of \citeA{peureux2021unsteady}, I found several
+gaps in the current understanding of the hydrodynamic modulation of short waves
+based on linear wave theory that I thought were worth revisiting.
 
@@
 exploit in the remainder of this paper.
+Note that here we derive the solutions from a different starting set of
+equations than that of \citeA{longuet1960changes}.
+Rather than starting from the conservation of crests and wave action, as we do
+here, they derived their solution from the linear superposition of the velocity
+potential of two waves.
+It appears the starting from the conservation of crests and wave action allows
+for a simpler and more intuitive derivation.
 
@@
 at the water surface rather than $z = 0$ is 38.4\% larger than that of the
-1st order approximate L-HS solution.
+1$^{st}$ order approximate L-HS solution.
 
@@
 Since wave action is variance divided by the intrinsic frequency $\sigma$,
-and variance scales with the square of the amplitude, the modulation of
-amplitude is proportional to the square root of the modulation of wave action.
+and variance scales with $ga^2$, the modulation of the amplitude is proportional
+to the square root of the modulation of wave action.
 
@@
 \label{eq:wave_amplitude_modulation}
-\widetilde{a}(\psi) = a \sqrt{\dfrac{\sigma}{\widetilde{\sigma}} \dfrac{\widetilde{N}}{N}}
+\widetilde{a}(\psi) = a \sqrt{
+  \dfrac{g}{\widetilde{g}}
+  \dfrac{\sigma}{\widetilde{\sigma}}
+  \dfrac{\widetilde{N}}{N}}
 \end{equation}

### 2024-06-21 paper/paper.tex
Working on amplitude modulation

@@
 %\title{Dispersion of short waves riding on longer waves}
-\title{Revisiting the linear theory of hydrodynamic wave modulation}
+\title{Revisiting the hydrodynamic wave modulation from linear wave theory}
 
@@
 
-TODO amplitude modulation;
 The modulation of short-wave amplitude can be derived in the same way by
@@
   \dfrac{\widetilde{N}}{N}}
+=
+  \left( \dfrac{\widetilde{g}}{g} \right)^{0.75}
+  \left( \dfrac{\widetilde{k}}{k} \right)^{0.25}
+  \left( \dfrac{\widetilde{N}}{N} \right)^{0.5}
 \end{equation}
+To the 1$^{st}$ order in $\varepsilon_L$, the amplitude modulation is then:
+
+\begin{equation}
+\label{eq:wave_amplitude_modulation_order1}
+\dfrac{\widetilde{a}}{a} = 
+  \left( \dfrac{\widetilde{g}}{g} \right)^{0.75}
+  \left( 1 + \varepsilon_L \right)^{0.75}
+\end{equation}
+
 To determine $\widetilde{\sigma}$, we will need to evaluate the modulation of

### 2024-06-21 paper/paper.tex
Fix amplitude calculation in the paper, again

@@
 
-Since wave action is variance divided by the intrinsic frequency $\sigma$,
-and variance scales with $ga^2$, the modulation of the amplitude is proportional
-to the square root of the modulation of wave action.
+Since wave action is energy divided by the intrinsic frequency $\sigma$,
+and energy scales with $ga^2$, the modulation of the amplitude is readily
+related to the modulations of gravity, wavenumber, and action:
 
@@
 \label{eq:wave_amplitude_modulation}
-\widetilde{a}(\psi) = a \sqrt{
+\dfrac{\widetilde{a}}{a} = \sqrt{
   \dfrac{g}{\widetilde{g}}
-  \dfrac{\sigma}{\widetilde{\sigma}}
+  \dfrac{\widetilde{\sigma}}{\sigma}
   \dfrac{\widetilde{N}}{N}}
 =
-  \left( \dfrac{\widetilde{g}}{g} \right)^{0.75}
+  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
   \left( \dfrac{\widetilde{k}}{k} \right)^{0.25}
@@
 \end{equation}
-To the 1$^{st}$ order in $\varepsilon_L$, the amplitude modulation is then:
+To the 1$^{st}$ order in $\varepsilon_L$, and using Eqs. \ref{eq:k_short_exact}
+and \ref{eq:wave_action_modulation}, we get:
 
@@
 \dfrac{\widetilde{a}}{a} = 
-  \left( \dfrac{\widetilde{g}}{g} \right)^{0.75}
+  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
   \left( 1 + \varepsilon_L \right)^{0.75}
@@
 
-To determine $\widetilde{\sigma}$, we will need to evaluate the modulation of
+To determine $\widetilde{\g}$, we will need to evaluate the modulation of
 the gravitational acceleration at the surface of the long wave.

### 2024-06-21 paper/scripts/plot_numerical_simulations.py
Fix syntax and color range in modulation 3-panels

@@ -60,7 +60,7 @@ m.run(ramp_type=None)
 plot_modulation_3panel(m, np.arange(0.5, 2.3, 0.1), "Infinite wave train")
 
 m.run(ramp_type="linear")
-plot_modulation_3panel(m, np.arange(0.8, 1.24, 0.04), "Linear ramp")
+plot_modulation_3panel(m, np.arange(0.75, 1.3, 0.05), "Linear ramp")
 
 m.run(ramp_type="groups")
-plot_modulation_3panel(m, np.arange(0.8, 1.24, 0.04), "Wave group")
+plot_modulation_3panel(m, np.arange(0.75, 1.3, 0.05), "Wave group")

### 2024-06-24 paper/paper.tex
Fix syntax and color range in modulation 3-panels
Comparing analytical and numerical solutions

@@
 
-\subsection{Wavenumber and amplitude modulation}
-\label{subsection:analytical_solutions}
+\subsection{Wavenumber modulation}
+\label{subsection:wavenumber_modulation}
 
@@
 
-The modulation of short-wave amplitude can be derived in the same way by
-dropping the non-linear terms in the wave action balance equation
-(Eq. \ref{eq:wave_action2}) and integrating wave action in time:
-
-\begin{equation}
-\label{eq:wave_action_modulation}
-\widetilde{N}(\psi) = N e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
-\end{equation}
-
-Since wave action is energy divided by the intrinsic frequency $\sigma$,
-and energy scales with $ga^2$, the modulation of the amplitude is readily
-related to the modulations of gravity, wavenumber, and action:
-
-\begin{equation}
-\label{eq:wave_amplitude_modulation}
-\dfrac{\widetilde{a}}{a} = \sqrt{
-  \dfrac{g}{\widetilde{g}}
-  \dfrac{\widetilde{\sigma}}{\sigma}
-  \dfrac{\widetilde{N}}{N}}
-=
-  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
-  \left( \dfrac{\widetilde{k}}{k} \right)^{0.25}
-  \left( \dfrac{\widetilde{N}}{N} \right)^{0.5}
-\end{equation}
-To the 1$^{st}$ order in $\varepsilon_L$, and using Eqs. \ref{eq:k_short_exact}
-and \ref{eq:wave_action_modulation}, we get:
-
-\begin{equation}
-\label{eq:wave_amplitude_modulation_order1}
-\dfrac{\widetilde{a}}{a} = 
-  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
-  \left( 1 + \varepsilon_L \right)^{0.75}
-\end{equation}
-
-To determine $\widetilde{\g}$, we will need to evaluate the modulation of
-the gravitational acceleration at the surface of the long wave.
-
 \subsection{Gravity modulation}
@@
 \end{equation}
-
 The corresponding gravitational acceleration at the surface of the Stokes wave
-is:
+is, to the 4$^{th}$ order:
 
@@
 
-\subsection{Lagrangian averaged short-wave properties}
+\subsection{Amplitude and steepness modulation}
+\label{subsection:amplitude_modulation}
 
-Stokes drift is a residual Lagrangian drift that occurs due to the wave orbits
-not being closed under a propagating wave
-\cite{stokes1847,kenyon1969stokes,van2018stokes}:
+The modulation of short-wave amplitude can be derived in the same way by
+dropping the non-linear terms in the wave action balance equation
+(Eq. \ref{eq:wave_action2}) and integrating wave action in time:
 
 \begin{equation}
-\label{}
-u_s = a^2 \omega k e^{2kz}
+\label{eq:wave_action_modulation}
+\widetilde{N}(\psi) = N e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
 \end{equation}
-for a monochromatic wave in deep water.
-Stokes drift has been successfully used to predict material transport on the
-ocean surface \cite{rohrs2012observation,curcic2016hurricane}.
-It can be derived from linear wave theory by integrating the wave orbital
-velocity following material orbits \cite{phillips1966dynamics}:
+
+Since wave action is energy divided by the intrinsic frequency $\sigma$,
+and energy scales with $ga^2$, the modulation of the amplitude is readily
+related to the modulations of gravity, wavenumber, and action:
 
 \begin{equation}
-\label{eq:lagrangian_average}
-u_s = \overline{
-  u(\psi) + 
-  \zeta \dfrac{\partial u}{\partial x} +
-  \xi \dfrac{\partial w}{\partial z}
-}
+\label{eq:wave_amplitude_modulation}
+\dfrac{\widetilde{a}}{a} = \sqrt{
+  \dfrac{g}{\widetilde{g}}
+  \dfrac{\widetilde{\sigma}}{\sigma}
+  \dfrac{\widetilde{N}}{N}}
+=
+  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
+  \left( \dfrac{\widetilde{k}}{k} \right)^{0.25}
+  \left( \dfrac{\widetilde{N}}{N} \right)^{0.5}
 \end{equation}
-where $\zeta$ and $\xi$ are horizontal and vertical excursions of the orbits,
-respectively.
-
-To obtain an average modulation of the short-wave properties, we apply the same
-Lagrangian integration approach used to obtain Stokes drift.
+To the 1$^{st}$ order in $\varepsilon_L$, and using Eqs. \ref{eq:k_short_exact}
+and \ref{eq:wave_action_modulation}, we get:
 
 \begin{equation}
-\label{eq:stokes_wavenumber}
-\widetilde{k_s} = \overline{
-  \widetilde{k}(\psi) +
-  \zeta \dfrac{\partial \widetilde{k}}{\partial x} +
-  \xi \dfrac{\partial \widetilde{k}}{\partial z}
-}
-= k \left( 1 + a_L^2 k_L^2 \right)
-\end{equation}
-
-\begin{equation}
-\label{eq:stokes_gravity}
-\widetilde{g_s} = \overline{
-  \widetilde{g}(\psi) +
-  \zeta \dfrac{\partial \widetilde{g}}{\partial x} +
-  \xi \dfrac{\partial \widetilde{g}}{\partial z}
-}
-= g \left( 1 + \dfrac{3}{2} a_L^2 k_L^2 \right)
+\label{eq:wave_amplitude_modulation_order1}
+\dfrac{\widetilde{a}}{a} = 
+  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
+  \left( 1 + \varepsilon_L \right)^{0.75}
 \end{equation}

### 2024-06-24 paper/scripts/plot_numerical_vs_analytical_solutions.py

@@ -1,0 +1,105 @@
+import matplotlib.pyplot as plt
+from model import WaveModulationModel
+import numpy as np
+import matplotlib
+
+matplotlib.rc("font", size=16)
+
+
+def g_modulation(psi, a, k):
+    eta = a * np.cos(psi)
+    res = 1 - a * k * np.exp(k * eta) * (np.cos(psi) - a * k * np.sin(psi) ** 2)
+    return res
+
+
+k_long = 1
+
+ak_range = np.arange(0, 0.41, 0.01)
+solutions = []
+for ak in ak_range:
+    a_long = ak / k_long
+    m = WaveModulationModel(a_long=a_long, k_long=k_long)
+    m.run(ramp_type="groups")
+    solutions.append(m)
+
+k_modulation_numerical = np.array(
+    [np.max(m.to_xarray().wavenumber) / m.k_short for m in solutions]
+)
+a_modulation_numerical = np.array(
+    [np.max(m.to_xarray().amplitude) / m.a_short for m in solutions]
+)
+ak_modulation_numerical = np.array(
+    [
+        np.max(m.to_xarray().amplitude * m.to_xarray().wavenumber)
+        / m.a_short
+        / m.k_short
+        for m in solutions
+    ]
+)
+
+phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
+a_modulation_analytical = []
+k_modulation_analytical = []
+ak_modulation_analytical = []
+for ak in ak_range:
+    a_long = ak / k_long
+    kmod = np.exp(ak * np.cos(phase) * np.exp(ak * np.cos(phase)))
+    Nmod = kmod
+    gmod = g_modulation(phase, a_long, k_long)
+    amod = gmod ** (-0.25) * kmod**0.25 * Nmod**0.5
+    a_modulation_analytical.append(np.max(amod))
+    k_modulation_analytical.append(np.max(kmod))
+    ak_modulation_analytical.append(np.max(amod * kmod))
+a_modulation_analytical = np.array(a_modulation_analytical)
+k_modulation_analytical = np.array(k_modulation_analytical)
+ak_modulation_analytical = np.array(ak_modulation_analytical)
+
+
+fig = plt.figure(figsize=(8, 6))
+ax = fig.add_subplot(111)
+
+ax.plot(ak_range, (1 + ak_range), "k-", lw=2, label="Amplitude, L-HS 1960")
[…]

### 2024-06-25 paper/scripts/plot_amplitude_modulation.py
Updates on amplitude modulation

@@ -1,0 +1,99 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib
+
+matplotlib.rc("font", size=16)
+
+
+def g_eulerian(g, psi, a, k, order=3):
+    assert order in [1, 2, 3]
+    eta = a * np.cos(psi)
+    if order == 1:
+        res = g * (1 - a * k * np.cos(psi))
+    elif order == 2:
+        res = g * (1 - a * k * np.cos(psi) * np.exp(k * eta))
+    else:
+        res = g * (
+            1 - a * k * np.exp(k * eta) * (np.cos(psi) - a * k * np.sin(psi) ** 2)
+        )
+    return res
+
+
+def plot_amplitude_modulation():
+    """Plot the apparent gravity figure."""
+
+    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
+    a_L = 0.1
+    k_L = 1
+    ak = a_L * k_L
+
+    g = 9.8
+
+    g_modulation = g_eulerian(g, phase, a_L, k_L, order=3) / g
+    k_modulation = np.exp(ak * np.cos(phase) * np.exp(ak * np.cos(phase)))
+    N_modulation = k_modulation
+    a_modulation = g_modulation ** (-0.25) * k_modulation**0.25 * N_modulation**0.5
+
+    a_modulation_lhs = 1 + ak * np.cos(phase)
+
+    fig = plt.figure(figsize=(14, 6))
+    ax1 = fig.add_subplot(121)
+    ax2 = fig.add_subplot(122)
+
+    ax1.text(
+        0, 1.02, "(a)", ha="left", va="bottom", transform=ax1.transAxes, fontsize=20
+    )
+    ax2.text(
+        0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
+    )
+
+    ax1.plot(phase, a_modulation_lhs, "k-", label=r"L-HS 1960", lw=2)
+    ax1.plot(phase, a_modulation, "r-", label=r"This paper", lw=2)
+    ax1.legend()
+    ax1.set_xlabel("Phase (rad.)")
+    ax1.set_ylabel(r"$a'/a$")
+    ax1.grid()
+    ax1.plot([0, 2 * np.pi], [1, 1], "k--")
+    ax1.set_xlim(0, 2 * np.pi)
[…]

### 2024-06-25 paper/scripts/plot_numerical_vs_analytical_solutions.py

@@ -14,26 +14,48 @@ def g_modulation(psi, a, k):
 
 k_long = 1
 
-ak_range = np.arange(0, 0.41, 0.01)
-solutions = []
+ak_range = np.arange(0, 0.45, 0.05)
+solutions1 = []
+solutions2 = []
 for ak in ak_range:
     a_long = ak / k_long
+
     m = WaveModulationModel(a_long=a_long, k_long=k_long)
     m.run(ramp_type="groups")
-    solutions.append(m)
+    solutions1.append(m)
 
-k_modulation_numerical = np.array(
-    [np.max(m.to_xarray().wavenumber) / m.k_short for m in solutions]
+    m = WaveModulationModel(a_long=a_long, k_long=k_long)
+    m.run(wave_type="stokes", ramp_type="groups")
+    solutions2.append(m)
+
+
+k_modulation_numerical_linear = np.array(
+    [np.max(m.to_xarray().wavenumber) / m.k_short for m in solutions1]
 )
-a_modulation_numerical = np.array(
-    [np.max(m.to_xarray().amplitude) / m.a_short for m in solutions]
+a_modulation_numerical_linear = np.array(
+    [np.max(m.to_xarray().amplitude) / m.a_short for m in solutions1]
 )
-ak_modulation_numerical = np.array(
+k_modulation_numerical_stokes = np.array(
+    [np.max(m.to_xarray().wavenumber) / m.k_short for m in solutions2]
+)
+a_modulation_numerical_stokes = np.array(
+    [np.max(m.to_xarray().amplitude) / m.a_short for m in solutions2]
+)
+ak_modulation_numerical_linear = np.array(
     [
         np.max(m.to_xarray().amplitude * m.to_xarray().wavenumber)
         / m.a_short
         / m.k_short
-        for m in solutions
+        for m in solutions1
+    ]
+)
+
+ak_modulation_numerical_stokes = np.array(
+    [
+        np.max(m.to_xarray().amplitude * m.to_xarray().wavenumber)
+        / m.a_short
+        / m.k_short
+        for m in solutions2
     ]
 )
 
@@ -69,10 +91,18 @@ ax.plot(
 )
 ax.plot(
     ak_range,
-    a_modulation_numerical,
+    a_modulation_numerical_linear,
     linestyle="-",
     lw=2,
     color="tab:orange",
[…]

### 2024-08-12 paper/paper.tex
Updates on amplitude modulation
Update intro paragraph

@@
   University of Miami,
-  Miami, FL
+  Coral Gables, FL
 }
@@
 \begin{keypoints}
-\item We revisit analytical and numerical solutions for hydrodynamic modulation of short waves by longer waves from linear wave theory.
-\item We extend the original solution by \citeA{longuet1960changes} to a higher-order solution.
-\item In contrast to earlier studies, we find that there exist stable numerical solutions for modulation by realistic long-wave groups.
+\item We revisit analytical and numerical solutions for hydrodynamic modulation of short waves by longer waves.
+\item We present a higher-order analytical solution based on linear wave theory.
+\item In contrast to an earlier study, we find stable numerical solutions for modulation by realistic long-wave groups.
 \end{keypoints}
@@
 \begin{equation}
-\label{eq:tau}
-\widetilde{k} = k (1 + a_L k_L)
+\label{eq:lhs1960wavenumber}
+\widetilde{k} = k (1 + \varepsilon_L)
 \end{equation}
-where $\widetilde{k}$ and $k$ are the modulated and unmodulated wavenumbers of
-short waves, respectively, and $a_L$ and $k_L$ are amplitude and wavenumber
-of the long wave, respectively.
+where $\widetilde{k}$ and $k$ are modulated and unmodulated short-wave wavenumbers,
+respectively, and $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
+wavenumber $k_L$ and amplitude $a_L$.
 This result has been influential on the development of remote sensing of ocean
-surface waves and currents \cite{keller1975microwave,hara1994hydrodynamic}
-and has been used even recently \cite{peureux2021unsteady} to simulate
-unsteady growth of short-waves due to longer waves.
+surface waves and currents \cite{keller1975microwave,hara1994hydrodynamic}.
+\citeA{longuet1987propagation} revisited this problem by considering variation
+in the gravitational acceleration and fully non-linear wave numerical solutions
+to find significantly stronger modulation than that predicted by
+Eq. \ref{eq:lhs1960wavenumber}.
+\citeA{zhang1990evolution} evaluated the effects of nonlinearity of short waves
+using a non-linear Schr\" odinger equation and found higher wavenumber
+modulation but lesser amplitude modulation than those predicted by
+\citeA{longuet1987propagation}.
+More recently, \cite{peureux2021unsteady} simulated the full wave conservation
+equations and found that the short waves grow unsteadily due to the propagation of
+longer waves, suggesting that the steady-state solutions may only be valid for
+a few long-wave periods before the short waves are steep enough to break.
 
-Working through the theory of \citeA{longuet1960changes,longuet1987propagation}
-and the numerical simulation of \citeA{peureux2021unsteady}, I found several
-gaps in the current understanding of the hydrodynamic modulation of short waves
-based on linear wave theory that I thought were worth revisiting.
-
-In this study, we extend the \citeA{longuet1960changes} theory (hereafter L-HS)
-to a higher-order solution following the linear wave theory, and present revised
-phase-averaged estimates of short-wave properties in presence of long waves.
+In this paper, we revisit this problem from a linear wave theory perspective and
+derive new non-linear solutions for the modulation of short waves, evaluated
+at the long-wave surface.
+We also perform numerical simulations of the full wave crest and action
+conservation laws and evaluate the effect of long-wave groups on the short waves.
+The next section presents the governing equations, followed by the analytical
+solutions in section \ref{section:analytical_solutions}.
+In section \ref{section:numerical_solutions} we present the numerical solutions
+and compare them to the analytical solutions.
+In section \ref{section:scale_analysis} we perform a scale analysis of the
+analytical solutions to quantify the relative importance of the terms in the
+full equation set.
+Finally, we summarize the results and discuss their implications in section
+\ref{section:conclusions}.
 
 \section{Governing equations}
+\label{section:governing_equations}
 
@@
 
+\begin{figure}[h]
+\label{fig:amplitude_modulation}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_amplitude_modulation_2panel.png}
+\caption{
+}
+\end{figure}
+
 \section{Numerical solutions}
@@
 \section{Conclusions}
+\label{section:conclusions}
 

### 2024-08-13 paper/scripts/plot_amplitude_modulation.py
6-panel analytical modulation figures

@@ -1,99 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import matplotlib
-
-matplotlib.rc("font", size=16)
-
-
-def g_eulerian(g, psi, a, k, order=3):
-    assert order in [1, 2, 3]
-    eta = a * np.cos(psi)
-    if order == 1:
-        res = g * (1 - a * k * np.cos(psi))
-    elif order == 2:
-        res = g * (1 - a * k * np.cos(psi) * np.exp(k * eta))
-    else:
-        res = g * (
-            1 - a * k * np.exp(k * eta) * (np.cos(psi) - a * k * np.sin(psi) ** 2)
-        )
-    return res
-
-
-def plot_amplitude_modulation():
-    """Plot the apparent gravity figure."""
-
-    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
-    a_L = 0.1
-    k_L = 1
-    ak = a_L * k_L
-
-    g = 9.8
-
-    g_modulation = g_eulerian(g, phase, a_L, k_L, order=3) / g
-    k_modulation = np.exp(ak * np.cos(phase) * np.exp(ak * np.cos(phase)))
-    N_modulation = k_modulation
-    a_modulation = g_modulation ** (-0.25) * k_modulation**0.25 * N_modulation**0.5
-
-    a_modulation_lhs = 1 + ak * np.cos(phase)
-
-    fig = plt.figure(figsize=(14, 6))
-    ax1 = fig.add_subplot(121)
-    ax2 = fig.add_subplot(122)
-
-    ax1.text(
-        0, 1.02, "(a)", ha="left", va="bottom", transform=ax1.transAxes, fontsize=20
-    )
-    ax2.text(
-        0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
-    )
-
-    ax1.plot(phase, a_modulation_lhs, "k-", label=r"L-HS 1960", lw=2)
-    ax1.plot(phase, a_modulation, "r-", label=r"This paper", lw=2)
-    ax1.legend()
-    ax1.set_xlabel("Phase (rad.)")
-    ax1.set_ylabel(r"$a'/a$")
-    ax1.grid()
-    ax1.plot([0, 2 * np.pi], [1, 1], "k--")
-    ax1.set_xlim(0, 2 * np.pi)
[…]

### 2024-08-13 paper/scripts/plot_analytical_solutions_with_phase.py

@@ -1,0 +1,134 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib
+
+matplotlib.rc("font", size=12)
+
+
+def elevation_linear(a: float, phase: float) -> float:
+    return a * np.cos(phase)
+
+
+def elevation_stokes(a: float, k: float, phase: float) -> float:
+    term1 = np.cos(phase)
+    term2 = 0.5 * a * k * np.cos(2 * phase)
+    term3 = (a * k) ** 2 * (3 / 8 * np.cos(3 * phase) - 1 / 16 * np.cos(phase))
+    return a * (term1 + term2 + term3)
+
+
+def gravity_linear(
+    a: float,
+    k: float,
+    phase: float,
+    g0: float,
+) -> float:
+    """Gravitational acceleration at the surface of a linear wave.
+    """
+    eta = elevation_linear(a, phase)
+    return g0 * (1 - a * k * np.exp(k * eta) * (np.cos(phase) - a * k * np.sin(phase) ** 2))
+
+
+def gravity_stokes(
+    a: float, k: float, phase: float, g0: float
+) -> float:
+    """Gravitational acceleration at the surface of a Stokes wave.
+    """
+    eta = elevation_stokes(a, k, phase)
+    res = g0 * (1 - a * k
+        * (np.cos(phase) - a * k * np.sin(phase)
+            * (
+                (1 - 1 / 16 * (a * k) ** 2) * np.sin(phase)
+                + a * k * np.sin(2 * phase)
+                + 9 / 8 * (a * k) ** 2 * np.sin(3 * phase)
+            )
+        ) * np.exp(k * eta)
+    )
+    return res
+
+
+
+def plot_analytical_solutions(a_L: float):
+
+    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
+    k_L = 1
+    ak_L = a_L * k_L
+
+    k_lhs = 1 + ak_L * np.cos(phase)
+    k_modulation = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
+
+    g_modulation = gravity_linear(a_L, k_L, phase, 1)
+    #g_modulation_stokes = gravity_stokes(a_L, k_L, phase, 1)
+
+    N_modulation = k_modulation
[…]

### 2024-08-13 paper/scripts/plot_frequency_modulation.py

@@ -1,214 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import matplotlib
-
-
-matplotlib.rc("font", size=16)
-
-
-def plot_fig_frequency_modulation():
-    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
-    g = 9.8
-
-    ak_L = 0.2
-
-    g1 = 1 - ak_L * np.cos(phase)
-    k1 = 1 + ak_L * np.cos(phase)
-    omega_rel1 = np.sqrt(g1 * k1)
-
-    term1 = -ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase))
-    term2 = ((ak_L * np.sin(phase)) ** 2) * np.exp(ak_L * np.cos(phase))
-    g2 = 1 + term1 + term2
-    k2 = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
-    omega_rel2 = np.sqrt(g2 * k2)
-
-    fig = plt.figure(figsize=(14, 6))
-    ax1 = fig.add_subplot(121)
-    ax2 = fig.add_subplot(122)
-
-    ax1.text(
-        0, 1.02, "(a)", ha="left", va="bottom", transform=ax1.transAxes, fontsize=20
-    )
-    ax2.text(
-        0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
-    )
-
-    ax1.plot(phase, omega_rel1, "k-", label=r"1$^{st}$ order", lw=2)
-    ax1.plot(phase, omega_rel2, "r-", label=r"3$^{rd}$ order", lw=2)
-    ax1.legend()
-    ax1.set_xlabel("Phase [rad]")
-    ax1.set_ylabel(r"$\widetilde \omega/\omega$")
-    ax1.grid()
-    ax1.plot([0, 2 * np.pi], [1, 1], "k--")
-    ax1.set_xlim(0, 2 * np.pi)
-    ax1.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
-    ax1.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
-
-    # Evaluate frequency modulation as function of ak
-    ak_range = np.arange(0, 0.41, 0.01)
-    omega1_mean, omega1_max, omega1_min = [], [], []
[…]

### 2024-08-13 paper/scripts/plot_gravity_modulation.py

@@ -1,108 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import matplotlib
-
-matplotlib.rc("font", size=16)
-
-
-def g_eulerian(g, psi, a, k, order=3):
-    assert order in [1, 2, 3]
-    eta = a * np.cos(psi)
-    if order == 1:
-        res = g * (1 - a * k * np.cos(psi))
-    elif order == 2:
-        res = g * (1 - a * k * np.cos(psi) * np.exp(k * eta))
-    else:
-        res = g * (
-            1 - a * k * np.exp(k * eta) * (np.cos(psi) - a * k * np.sin(psi) ** 2)
-        )
-    return res
-
-
-def g_lagrangian(g, psi, a, k, order=4):
-    assert order in [3, 4]
-    eta = a * np.cos(psi)
-    g_E = g_eulerian(g, psi, a, k, order=3)
-    if order == 3:
-        res = g_E - a**2 * k**2 * np.exp(2 * k * eta)
-    if order == 4:
-        res = (
-            g_E
-            - a**2 * k**2 * np.exp(2 * k * eta)
-            - a**3
-            * k**3
-            * np.exp(2 * k * eta)
-            * (np.sin(psi) ** 2 * np.cos(psi) - np.sin(2 * psi) * np.sin(psi))
-        )
-    return res
-
-
-def plot_fig_apparent_gravity():
-    """Plot the apparent gravity figure."""
-
-    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
-    a_L = 0.2
-    k_L = 1
-
-    g = 9.8
-
-    g_E1 = g_eulerian(g, phase, a_L, k_L, order=1)
-    g_E3 = g_eulerian(g, phase, a_L, k_L, order=3)
-
-    fig = plt.figure(figsize=(14, 6))
-    ax1 = fig.add_subplot(121)
-    ax2 = fig.add_subplot(122)
-
-    ax1.text(
-        0, 1.02, "(a)", ha="left", va="bottom", transform=ax1.transAxes, fontsize=20
-    )
-    ax2.text(
-        0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
-    )
-
[…]

### 2024-08-13 paper/scripts/plot_numerical_simulations_ramp_properties.py

@@ -23,8 +23,9 @@ ak_init = 0.1
 
 fig = plt.figure(figsize=(8, 6))
 ax = fig.add_subplot(111)
-ax.plot(periods, np.max(ak1, 1) / ak_init, color="tab:blue", lw=2)
-ax.plot(periods, np.max(ak2, 1) / ak_init, color="tab:orange", lw=2)
+ax.plot(periods, np.max(ak1, 1) / ak_init, color="tab:blue", lw=2, label="Sudden onset of long waves")
+ax.plot(periods, np.max(ak2, 1) / ak_init, color="tab:orange", lw=2, label="Linear ramp")
+ax.legend()
 ax.set_xlabel("Time (periods)")
 ax.set_ylabel(r"$\widetilde{ak}/(ak)$")
 ax.grid()

### 2024-08-13 paper/scripts/plot_wavenumber_modulation.py

@@ -1,74 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import matplotlib
-
-matplotlib.rc("font", size=16)
-
-
-def plot_fig_wavenumber_modulation():
-    """Plot the wavenumber modulation comparison figure."""
-
-    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
-    k_L = 1
-    a_L = 0.2
-    ak_L = a_L * k_L
-
-    solution1 = 1 + ak_L * np.cos(phase)
-    solution3 = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
-
-    fig = plt.figure(figsize=(14, 6))
-    ax1 = fig.add_subplot(121)
-    ax2 = fig.add_subplot(122)
-
-    ax1.text(
-        0, 1.02, "(a)", ha="left", va="bottom", transform=ax1.transAxes, fontsize=20
-    )
-    ax2.text(
-        0, 1.02, "(b)", ha="left", va="bottom", transform=ax2.transAxes, fontsize=20
-    )
-
-    ax1.plot(phase, solution1, "k-", lw=2, label="L-H&S")
-    ax1.plot(phase, solution3, "r-", lw=2, label="This paper")
-    ax1.legend()
-    ax1.set_xlabel("Phase (rad.)")
-    ax1.set_ylabel(r"$\widetilde k/k$")
-    ax1.grid()
-    ax1.plot([0, 2 * np.pi], [1, 1], "k--")
-    ax1.set_xlim(0, 2 * np.pi)
-    ax1.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
-    ax1.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
-
-    # Evaluate wa enumber modulation as function of ak
-    ak_range = np.arange(0, 0.41, 0.01)
-    k1_mean, k1_max, k1_min = [], [], []
-    k2_mean, k2_max, k2_min = [], [], []
-    for ak in ak_range:
-        k1 = 1 + ak * np.cos(phase)
-        k2 = np.exp(ak * np.cos(phase) * np.exp(ak * np.cos(phase)))
-        k1_mean.append(np.mean(k1))
-        k1_max.append(np.max(k1))
-        k1_min.append(np.min(k1))
[…]

### 2024-08-13 paper/paper.tex
Incorporate new analytical solution figures in the paper text

@@
 
-\begin{figure}[h]
-\label{fig:wavenumber_modulation}
+\begin{figure}[ht]
+\label{fig:analytical_solutions_ak0.1}
 \centering
-\includegraphics[width=\textwidth]{figures/fig_wavenumber_modulation_2panel.png}
+\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.1.png}
 \caption{
-  Comparison of relative short-wave modulation as function of the
-  long-wave phase, for a moderately steep long wave of $\varepsilon = 0.1$.
-  The dotted line shows the 1st-order approximated solution by L-HS,
-  the dashed line shows the exact solution by L-HS,
-  and the solid line shows the extended solution introduced in this paper.
 }
@@
 faces of the long wave (Fig. \ref{fig:gravity_modulation}).
-
-\begin{figure}[h]
-\label{fig:gravity_modulation}
-\centering
-\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_2panel.png}
-\caption{
-  Relative change in gravity of short waves due to a longer wave with steepness $a_L k_L = 0.2$,
-  as a function of the long-wave phase and evaluated to the first (dashed line) and second (solid line) order in $a_L k_L$.
-}
-\end{figure}
 
@@
 
-\begin{figure}[h]
-\label{fig:frequency_modulation}
-\centering
-\includegraphics[width=\textwidth]{figures/fig_frequency_modulation_2panel.png}
-\caption{
-  (a) Relative change in the short-wave intrinsic frequency on the surface of a
-  longer linear (black) and Stokes (red) wave of steepness $\varepsilon_L = 0.2$
-  as function of long-wave phase;
-  (b) Long-wave phase maximum (solid), mean (dashed), and minimum (dotted)
-  relative change in the short-wave intrinsic frequency as function of long-wave
-  steepness in the case of a linear (black) and Stokes (red) long wave.
-}
-\end{figure}
-
 \subsection{Amplitude and steepness modulation}
@@
 
-\begin{figure}[h]
-\label{fig:amplitude_modulation}
+\begin{figure}[ht]
+\label{fig:analytical_solutions_ak0.2}
 \centering
-\includegraphics[width=\textwidth]{figures/fig_amplitude_modulation_2panel.png}
-\caption{
-}
+\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.2.png}
+\caption{Same a Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.2$.}
+\end{figure}
+
+\begin{figure}[ht]
+\label{fig:analytical_solutions_ak0.4}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.4.png}
+\caption{Same a Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.4$.}
 \end{figure}

### 2024-08-13 paper/scripts/plot_analytical_solutions_with_phase.py

@@ -123,7 +123,7 @@ def plot_analytical_solutions(a_L: float):
     fig.suptitle(r"$\varepsilon_L$" + f" = {a_L * k_L}", fontsize=18)
 
     plt.subplots_adjust(left=0.1, right=0.95, top=0.9, bottom=0.05)
-    plt.savefig(f"../figures/fig_analytical_solutions_ak{a_L * k_L}.png", dpi=150)
+    plt.savefig(f"../figures/fig_analytical_solutions_ak{a_L * k_L}.png", dpi=200)
     plt.close(fig)
 
 

### 2024-08-13 paper/scripts/plot_numerical_vs_analytical_solutions.py
Compare numerical and analytical solutions with LH1987

@@ -1,9 +1,10 @@
 import matplotlib.pyplot as plt
 from model import WaveModulationModel
 import numpy as np
+import pandas as pd
 import matplotlib
 
-matplotlib.rc("font", size=16)
+matplotlib.rc("font", size=14)
 
 
 def g_modulation(psi, a, k):
@@ -14,7 +15,7 @@ def g_modulation(psi, a, k):
 
 k_long = 1
 
-ak_range = np.arange(0, 0.45, 0.05)
+ak_range = np.arange(0, 0.41, 0.01)
 solutions1 = []
 solutions2 = []
 for ak in ak_range:
@@ -77,68 +78,54 @@ k_modulation_analytical = np.array(k_modulation_analytical)
 ak_modulation_analytical = np.array(ak_modulation_analytical)
 
 
+df_k_lh1987 = pd.read_csv("../data/lh1987_k_modulation_r10.csv")
+df_ak_lh1987 = pd.read_csv("../data/lh1987_ak_modulation_r10.csv")
+
 fig = plt.figure(figsize=(8, 6))
 ax = fig.add_subplot(111)
-
-ax.plot(ak_range, (1 + ak_range), "k-", lw=2, label="Amplitude, L-HS 1960")
-ax.plot(
-    ak_range,
-    a_modulation_analytical,
-    linestyle="-",
-    lw=2,
-    color="tab:blue",
-    label="Amplitude, analytical",
-)
-ax.plot(
-    ak_range,
-    a_modulation_numerical_linear,
-    linestyle="-",
-    lw=2,
-    color="tab:orange",
-    label="Amplitude, numerical",
-)
-ax.plot(
-    ak_range,
-    a_modulation_numerical_stokes,
-    linestyle="-",
-    lw=2,
-    color="tab:green",
-    label="Amplitude, numerical",
-)
-
-ax.plot(ak_range, (1 + ak_range) ** 2, "k--", lw=2, label="Steepness, L-HS 1960")
-ax.plot(
-    ak_range,
-    ak_modulation_analytical,
-    linestyle="--",
-    lw=2,
-    color="tab:blue",
-    label="Steepness, analytical",
-)
-ax.plot(
-    ak_range,
-    ak_modulation_numerical_linear,
-    linestyle="--",
-    lw=2,
-    color="tab:orange",
-    label="Steepness, numerical",
-)
-ax.plot(
-    ak_range,
[…]

### 2024-08-14 paper/scripts/plot_analytical_solutions_with_phase.py
Fix plot

@@ -80,8 +80,9 @@ def plot_analytical_solutions(a_L: float):
         plt.subplot2grid((3, 2), (2, 1)),
     ]
 
-    axes[0].plot(phase, k_lhs, "k-", lw=2, label="L-H&S")
+    axes[0].plot(phase, k_lhs, "k-", lw=2, label="L-HS 1960")
     axes[0].plot(phase, k_modulation, "r-", lw=2, label="This paper")
+    axes[0].legend()
 
     axes[1].plot(phase, np.ones(phase.shape), "k-", lw=2)
     axes[1].plot(phase, g_modulation, "r-", lw=2)

### 2024-08-15 paper/scripts/plot_scale_analysis.py
Scale analysis plotting update

@@ -1,0 +1,153 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib
+
+import model
+
+matplotlib.rc("font", size=12)
+
+
+def elevation_linear(a: float, phase: float) -> float:
+    return a * np.cos(phase)
+
+
+def elevation_stokes(a: float, k: float, phase: float) -> float:
+    term1 = np.cos(phase)
+    term2 = 0.5 * a * k * np.cos(2 * phase)
+    term3 = (a * k) ** 2 * (3 / 8 * np.cos(3 * phase) - 1 / 16 * np.cos(phase))
+    return a * (term1 + term2 + term3)
+
+
+def gravity_linear(
+    a: float,
+    k: float,
+    phase: float,
+    g0: float,
+) -> float:
+    """Gravitational acceleration at the surface of a linear wave."""
+    eta = elevation_linear(a, phase)
+    return g0 * (
+        1 - a * k * np.exp(k * eta) * (np.cos(phase) - a * k * np.sin(phase) ** 2)
+    )
+
+
+def gravity_stokes(a: float, k: float, phase: float, g0: float) -> float:
+    """Gravitational acceleration at the surface of a Stokes wave."""
+    eta = elevation_stokes(a, k, phase)
+    res = g0 * (
+        1
+        - a
+        * k
+        * (
+            np.cos(phase)
+            - a
+            * k
+            * np.sin(phase)
+            * (
+                (1 - 1 / 16 * (a * k) ** 2) * np.sin(phase)
+                + a * k * np.sin(2 * phase)
+                + 9 / 8 * (a * k) ** 2 * np.sin(3 * phase)
+            )
+        )
+        * np.exp(k * eta)
+    )
+    return res
+
+
+def calculate_tendencies(a_L: float):
+    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
+    g0 = 9.8
+    k_L = 1
+    ak_L = a_L * k_L
+    omega_L = np.sqrt(g0 * k_L)
+
+    L = 2 * np.pi / k_L
+    dx = L / phase.size
+
+    k0 = 10
+    N0 = 1
+
+    k = k0 * np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
[…]

### 2024-08-15 paper/scripts/scale_analysis.py

@@ -1,39 +1,0 @@
-import matplotlib.pyplot as plt
-from model import diff, WaveModulationModel
-import numpy as np
-
-phase = np.linspace(0, 2 * np.pi, 100, endpoint=False)
-a_long = 0.1
-k_long = 1
-ak = a_long * k_long
-x = phase / k_long
-dx = x[1] - x[0]
-
-k0 = 10
-g0 = 9.8
-
-omega_long = np.sqrt(g0 * k_long)
-U = a_long * omega_long * np.cos(phase) * np.exp(ak * np.cos(phase))
-
-k = k0 * np.exp(ak * np.cos(phase) * np.exp(ak * np.cos(phase)))
-g = g0 * (
-    1 - ak * np.exp(ak * np.cos(phase)) * (np.cos(phase) + ak * np.sin(phase) ** 2)
-)
-omega = np.sqrt(g * k)
-Cg = 0.5 * omega / k
-
-propagation = -Cg * diff(k) / dx
-advection = -U * diff(k) / dx
-convergence = -k * diff(U) / dx
-inhomogeneity = -0.5 * np.sqrt(k / g) * diff(g) / dx
-
-m = WaveModulationModel(a_long=a_long, k_long=k_long)
-m.run(ramp_type="groups")
-ds = m.to_xarray()
-
-fig = plt.figure(figsize=(8, 6))
-ax = fig.add_subplot(111)
-ax.plot(phase, k0 * (1 + ak * np.cos(phase)))
-ax.plot(phase, k)
-ax.plot(phase, ds.wavenumber[1004, :])
-ax.grid()

### 2024-08-15 paper/paper.tex
Compare numerical and analytical solutions with LH1987
Comparing numerical and analytical results with phase

@@
 a few long-wave periods before the short waves are steep enough to break.
+
+\begin{table}[h]
+\label{table:modulation_literature_summary}
+\caption{
+  Summary of prior \cite{longuet1960changes,longuet1987propagation,zhang1990evolution} and present (this paper)
+  approaches to calculating the modulation of short waves by long waves.
+}
+\begin{tabular}{cccccc}
+\hline
+Reference   & Governing & Gravity & Nonlinear  & Nonlinear   & Solution \\
+            & equations & varies  & long waves & short waves &          \\
+\hline
+\hline
+L-HS (1960) & Velocity potential       & No  & No  & No  & Analytical \\
+\hline
+L-H (1987)  & Velocity potential       & Yes & Yes & No  & Numerical \\
+\hline
+ZM (1990)   & Nonlinear Schr\" odinger & Yes & Yes & Yes & Numerical \\
+            & equation                 &     &     &     &           \\
+\hline
+This paper  & Crest and action         & Yes & Yes & No  & Analytical \\
+            & conservation             &     &     &     & and numerical \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-\subsection{Unstable growth in infinite long-wave trains}
+\subsection{Comparison with analytical solutions}
+\label{subsection:comparison_with_analytical_solutions}
+
+\begin{figure}[h]
+\label{fig:numerical_solutions}
+\centering
+\includegraphics[width=\textwidth]{figures/numerical_vs_analytical_solutions_with_phase.png}
+\caption{
+  Comparison of numerical solutions of wavenumber, amplitude, and steepness modulation
+  with their analytical solutions, for $\varepsilon_L$ = 0.1, 0.2, and 0.4.
+}
+\end{figure}
+
+\begin{figure}[h]
+\label{fig:numerical_vs_analytical_k}
+\centering
+\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
+\caption{}
+\end{figure}
+
+\begin{figure}[h]
+\label{fig:numerical_vs_analytical_a}
+\centering
+\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
+\caption{}
+\end{figure}
+
+\begin{figure}[h]
+\label{fig:numerical_vs_analytical_ak}
+\centering
+\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
+\caption{}
+\end{figure}
+
+\subsection{Unstable growth in Pereux et al. (2021)}
 \label{subsection:unstable_growth}

### 2024-08-15 paper/scripts/plot_analytical_solutions_with_phase.py

@@ -22,33 +22,37 @@ def gravity_linear(
     phase: float,
     g0: float,
 ) -> float:
-    """Gravitational acceleration at the surface of a linear wave.
-    """
+    """Gravitational acceleration at the surface of a linear wave."""
     eta = elevation_linear(a, phase)
-    return g0 * (1 - a * k * np.exp(k * eta) * (np.cos(phase) - a * k * np.sin(phase) ** 2))
+    return g0 * (
+        1 - a * k * np.exp(k * eta) * (np.cos(phase) - a * k * np.sin(phase) ** 2)
+    )
 
 
-def gravity_stokes(
-    a: float, k: float, phase: float, g0: float
-) -> float:
-    """Gravitational acceleration at the surface of a Stokes wave.
-    """
+def gravity_stokes(a: float, k: float, phase: float, g0: float) -> float:
+    """Gravitational acceleration at the surface of a Stokes wave."""
     eta = elevation_stokes(a, k, phase)
-    res = g0 * (1 - a * k
-        * (np.cos(phase) - a * k * np.sin(phase)
+    res = g0 * (
+        1
+        - a
+        * k
+        * (
+            np.cos(phase)
+            - a
+            * k
+            * np.sin(phase)
             * (
                 (1 - 1 / 16 * (a * k) ** 2) * np.sin(phase)
                 + a * k * np.sin(2 * phase)
                 + 9 / 8 * (a * k) ** 2 * np.sin(3 * phase)
             )
-        ) * np.exp(k * eta)
+        )
+        * np.exp(k * eta)
     )
     return res
 
 
-
 def plot_analytical_solutions(a_L: float):
-
     phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
     k_L = 1
     ak_L = a_L * k_L
@@ -57,10 +61,10 @@ def plot_analytical_solutions(a_L: float):
     k_modulation = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
 
     g_modulation = gravity_linear(a_L, k_L, phase, 1)
-    #g_modulation_stokes = gravity_stokes(a_L, k_L, phase, 1)
+    # g_modulation_stokes = gravity_stokes(a_L, k_L, phase, 1)
 
     N_modulation = k_modulation
[…]

### 2024-08-15 paper/scripts/plot_numerical_simulations_ramp_properties.py

@@ -23,8 +23,16 @@ ak_init = 0.1
 
 fig = plt.figure(figsize=(8, 6))
 ax = fig.add_subplot(111)
-ax.plot(periods, np.max(ak1, 1) / ak_init, color="tab:blue", lw=2, label="Sudden onset of long waves")
-ax.plot(periods, np.max(ak2, 1) / ak_init, color="tab:orange", lw=2, label="Linear ramp")
+ax.plot(
+    periods,
+    np.max(ak1, 1) / ak_init,
+    color="tab:blue",
+    lw=2,
+    label="Sudden onset of long waves",
+)
+ax.plot(
+    periods, np.max(ak2, 1) / ak_init, color="tab:orange", lw=2, label="Linear ramp"
+)
 ax.legend()
 ax.set_xlabel("Time (periods)")
 ax.set_ylabel(r"$\widetilde{ak}/(ak)$")

### 2024-08-15 paper/scripts/plot_numerical_vs_analytical_solutions.py

@@ -84,10 +84,38 @@ df_ak_lh1987 = pd.read_csv("../data/lh1987_ak_modulation_r10.csv")
 fig = plt.figure(figsize=(8, 6))
 ax = fig.add_subplot(111)
 ax.plot(ak_range, (1 + ak_range), "k-", lw=2, label="L-HS 1960, analytical")
-ax.plot(df_k_lh1987.epsilon, df_k_lh1987.k_modulation, "ko", ms=8, label="L-H 1987, numerical", zorder=10)
-ax.plot(ak_range, k_modulation_analytical, linestyle="-", lw=2, color="tab:blue", label="This paper, analytical")
-ax.plot(ak_range, k_modulation_numerical_linear, linestyle="-", lw=2, color="tab:green", label="This paper, numerical, linear")
-ax.plot(ak_range, k_modulation_numerical_stokes, linestyle="-", lw=2, color="tab:orange", label="This paper, numerical, Stokes")
+ax.plot(
+    df_k_lh1987.epsilon,
+    df_k_lh1987.k_modulation,
+    "ko",
+    ms=8,
+    label="L-H 1987, numerical",
+    zorder=10,
+)
+ax.plot(
+    ak_range,
+    k_modulation_analytical,
+    linestyle="-",
+    lw=2,
+    color="tab:blue",
+    label="This paper, analytical",
+)
+ax.plot(
+    ak_range,
+    k_modulation_numerical_linear,
+    linestyle="-",
+    lw=2,
+    color="tab:green",
+    label="This paper, numerical, linear",
+)
+ax.plot(
+    ak_range,
+    k_modulation_numerical_stokes,
+    linestyle="-",
+    lw=2,
+    color="tab:orange",
+    label="This paper, numerical, Stokes",
+)
 ax.legend(loc="upper left")
 ax.set_xlabel(r"$\varepsilon_L$")
 ax.set_ylabel(r"$\widetilde{k}/k$")
@@ -101,10 +129,38 @@ plt.close()
 fig = plt.figure(figsize=(8, 6))
 ax = fig.add_subplot(111)
 ax.plot(ak_range, (1 + ak_range), "k-", lw=2, label="L-HS 1960, analytical")
-ax.plot(df_ak_lh1987.epsilon, df_ak_lh1987.ak_modulation / df_k_lh1987.k_modulation, "ko", ms=8, label="L-H 1987, numerical", zorder=10)
[…]

### 2024-08-15 paper/scripts/plot_numerical_vs_analytical_solutions_with_phase.py

@@ -1,0 +1,130 @@
+import matplotlib.pyplot as plt
+from model import WaveModulationModel
+import numpy as np
+import pandas as pd
+import matplotlib
+
+matplotlib.rc("font", size=14)
+
+
+def g_modulation(psi, a, k):
+    eta = a * np.cos(psi)
+    res = 1 - a * k * np.exp(k * eta) * (np.cos(psi) - a * k * np.sin(psi) ** 2)
+    return res
+
+
+def calculate_modulation(ak_L: float):
+    k_long = 1
+    a_long = ak_L / k_long
+
+    m = WaveModulationModel(a_long=a_long, k_long=k_long)
+    m.run(ramp_type="groups")
+    ds = m.to_xarray()
+
+    ind = 1002  # Exactly 5 periods
+
+    phase = ds.space * k_long
+
+    k_mod = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
+    N_mod = k_mod
+    g_mod = g_modulation(phase, a_long, k_long)
+    a_mod = g_mod ** (-0.25) * k_mod**0.25 * N_mod**0.5
+
+    k_mod_numerical = ds.wavenumber[ind] / m.k_short
+    a_mod_numerical = ds.amplitude[ind] / m.a_short
+
+    return phase, k_mod, a_mod, k_mod_numerical, a_mod_numerical
+
+
+ak = [0.1, 0.2, 0.4]
+
+fig, axes = plt.subplots(3, len(ak), figsize=(12, 8))
+
+for n, ax in enumerate(axes.flatten()):
+    ax.grid()
+    ax.set_xlim(0, 2 * np.pi)
+    ax.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
+    ax.set_xticklabels([])
+    ax.plot([0, 2 * np.pi], [1, 1], "k--")
+    ax.text(
+        0,
+        1.02,
+        f"({chr(97 + n)})",
+        ha="left",
+        va="bottom",
+        transform=ax.transAxes,
+        fontsize=18,
+    )
+
+for n in range(3):
+    (
+        phase,
+        k_modulation,
+        a_modulation,
+        k_modulation_numerical,
+        a_modulation_numerical,
+    ) = calculate_modulation(ak[n])
+
+    axes[0, n].plot(
+        phase,
+        k_modulation,
+        linestyle="-",
+        lw=2,
+        color="tab:blue",
[…]

### 2024-08-16 paper/paper.tex
Update abstract; captions etc.

@@
 \item We present a higher-order analytical solution based on linear wave theory.
-\item In contrast to an earlier study, we find stable numerical solutions for modulation by realistic long-wave groups.
+\item In contrast to an earlier study, we find stable numerical solutions for modulation by long-wave groups.
 \end{keypoints}
@@
 Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
-a well-known ocean surface process that pertains to remote sensing, the
-interpretation of in situ wave measurements, and numerical wave forecasting.
-In this paper we revisit the early analytical work by \citeA{longuet1960changes},
-as well as the more recent numerical simulations by \citeA{peureux2021unsteady},
-to re-evaluate the changes in short-wave properties due to the presence of longer waves.
-We present two distinct improvements to the current state of the art:
-First, we derive a higher-order, non-linear solution for the short-wave wavenumber
-and action density as function of the long-wave phase and steepness.
-This solution suggests an increase in short-wave wavenumber by 18 and 38\% for
-long-wave steepenesses of 0.1 and 0.2, respectively, compared to the original
-linear solution.
-It is also in close agreement with the full numerical solutions up to the
-long-wave steepness of 0.25.
-Second, we find that the full numerical solutions, which were previously found
-to be unstable for homogeneous initial conditions, are in fact stable for more
-realistic long-wave groups.
-The key physical interpretation from these results is that the short waves can
-indefinitely grow and steepen only in the presence of inifinite long-wave trains,
-an unlikely occurence in the ocean.
+a well-known ocean surface process that affects remote sensing, the
+interpretation of \textit{in situ} wave measurements, and numerical wave
+forecasting.
+In this paper, we revisit the linear wave theory and derive higher-order
+solutions for the change of short-wave wavenumber, action density, and
+gravitational acceleration due to the presence of longer waves.
+We validate the analytical solutions with numerical simulations of the full wave
+crest and action conservation equations.
+The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
+steepness modulation significantly deviate from the linear analytical solutions
+of \citeA{longuet1960changes}, and are similar to the nonlinear numerical
+solutions by \citeA{zhang1990evolution}.
+We also discuss the result of \citeA{peureux2021unsteady} who found through
+numerical simulations that the short-wave modulation grows unsteadily with
+each long-wave phase.
+Here we show that this unsteady growth only occurs for homogeneous initial
+conditions and does not occur for more realistic long-wave groups.
+Finally, the combined analytical and numerical framework allows us to quantify
+the contribution of various terms in the full wave conservation equations.
+The new analytical solutions are a good approximation of the full numerical
+solution in long-wave steepness up to $\approx 0.2$.
 \end{abstract}
@@
 Although there have been analytical and numerical solutions for how the short
-waves should change in the presence of longer waves, they have been simplified
-and incomplete.
+waves should change in the presence of longer waves, they have either been
+simplistic or 
 This paper revisits and extends these solutions to propose a more accurate
@@
 \caption{
-  Summary of prior \cite{longuet1960changes,longuet1987propagation,zhang1990evolution} and present (this paper)
-  approaches to calculating the modulation of short waves by long waves.
+  Summary of prior
+  \cite{longuet1960changes,phillips1981dispersion,longuet1987propagation,zhang1990evolution,peureux2021unsteady}
+  and present (this paper) approaches to calculating the modulation of short
+  waves by long waves.
+  *: long 3rd order Stokes wave is considered only in numerical simulations.
 }
@@
 \hline
-L-HS (1960) & Velocity potential       & No  & No  & No  & Analytical \\
+L-HS 1960   & Velocity potential       & No  & No  & No  & Analytical \\
 \hline
-L-H (1987)  & Velocity potential       & Yes & Yes & No  & Numerical \\
+P 1981      & Crest and energy         & Yes & Yes & No  & Numerical \\
+            & conservation             &     &     &     &           \\
 \hline
-ZM (1990)   & Nonlinear Schr\" odinger & Yes & Yes & Yes & Numerical \\
+L-H 1987    & Velocity potential       & Yes & Yes & No  & Numerical \\
+\hline
+ZM 1990     & Nonlinear Schr\" odinger & Yes & Yes & Yes & Numerical \\
             & equation                 &     &     &     &           \\
 \hline
-This paper  & Crest and action         & Yes & Yes & No  & Analytical \\
+PAG 2021    & Crest and action         & Yes & No & No  & Numerical \\
+            & conservation             &     &     &     &           \\
+\hline
+This paper  & Crest and action         & Yes & Yes* & No  & Analytical \\
             & conservation             &     &      &     & and numerical \\
@@
 
-The second important relation is the conservation of wave crests:
+The second important relation is the conservation of wave crests
+\cite{phillips1981dispersion}:
 
@@
 \caption{
+  Modulation of short-wave (a) wavenumber, (b) gravitational acceleration,
+  (c) amplitude, (d) steepness, (e) intrinsic frequency, and (d) intrinsic phase
+  speed as function of long-wave phase for $\varepsilon_L = 0.1$, based on
+  analytical solutions by \citeA{longuet1960changes} (L-HS 1960, black)
+  and this paper (blue).
+  Long-wave crest and trough are located at $\psi = 0$ and $\psi = \pi$,
+  respectively.
 }
@@
 \includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.2.png}
-\caption{Same a Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.2$.}
+\caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.2$.}
 \end{figure}
@@
 \includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.4.png}
-\caption{Same a Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.4$.}
+\caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.4$.}
 \end{figure}
@@
 \begin{figure}[h]
-  \label{fig:modulation_3panel_ramp}
+  \label{fig:unsteady_growth_timeseries}
   \centering
-  \includegraphics[width=\textwidth]{figures/fig_modulation_ramp_timeseries.png}
+  \includegraphics[width=0.6\textwidth]{figures/fig_modulation_ramp_timeseries.png}
   \caption{
@@
 \acknowledgments
-Thanks to Peisen Tan, Nathan Laxague, and Fabrice Ardhuin for providing helpful
-feedback on the early drafts of this work.
+Thanks to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan for the
+helpful discussions.
 

### 2024-08-20 paper/scripts/effective_gravity_analytical.py
Analytical solution for effective gravity

@@ -1,0 +1,35 @@
+from sympy import (
+    diff,
+    exp,
+    cos,
+    sin,
+    atan,
+    symbols,
+    simplify,
+    init_printing,
+)
+
+init_printing()
+
+x, t, a, k, ω, ψ, U, W, g, g0, η, α = symbols(
+    "x t a k ω ψ U W g g_0 η α", real=True, positive=True
+)
+
+# Long-wave phase
+ψ = k * x - ω * t
+
+# Long-wave elevation
+η = a * cos(ψ)
+
+# Local elevation slope
+α = atan(diff(η, x))
+
+# Orbital velocities
+U = a * ω * exp(k * η) * cos(ψ)
+W = a * ω * exp(k * η) * sin(ψ)
+
+# Orbital accelerations
+dU_dt = simplify(diff(U, t))
+dW_dt = simplify(diff(W, t))
+
+g = simplify(g0 * cos(α) + dW_dt * cos(α) + dU_dt * sin(α))

### 2024-08-20 paper/paper.tex
Incorporate curvilinear effects on g

@@
 higher in the troughs \cite{longuet1986eulerian,longuet1987propagation}.
-A linear wave causes a modulation in gravity acceleration at its surface:
+A long wave modulates gravitational acceleration at its surface:
+
+\begin{equation}
+\label{eq:gravity_modulation_general}
+\widetilde{g}(\psi)
+  = g \cos{\alpha} 
+  + \dfrac{\partial W}{\partial t} \cos{\alpha}
+  + \frac{\partial U}{\partial t} \sin{\alpha}
+\end{equation}
+where $\alpha = \arctan{\frac{\partial \eta}{\partial x}}$ is the local slope
+of the long-wave surface.
+Curvilinear effects on effective gravity can be removed by setting $\alpha = 0$.
+
+For a linear wave, the effective gravity is:
 
@@
 \label{eq:gravity_modulation_linear}
-\widetilde{g}(\psi) = g + \dfrac{\partial W}{\partial t} = g \left[
-  1 - \varepsilon_L e^{\varepsilon_L \cos{\psi}} \left(
-    \cos{\psi} - \varepsilon_L \sin^2{\psi}
-  \right)
-\right]
+\widetilde{g}
+  = g 
+  \frac{
+    1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
+    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 \right]
+  }
+  {\sqrt{\left(\varepsilon_L \sin{\psi}\right)^2 + 1}}
 \end{equation}
+
 The first-order term in $a_L k_L$ is in phase with the long-wave elevation,

### 2024-08-20 paper/scripts/plot_analytical_solutions_with_phase.py

@@ -1,6 +1,7 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import matplotlib
+from model import diff
 
 matplotlib.rc("font", size=12)
 
@@ -20,13 +21,50 @@ def gravity_linear(
     a: float,
     k: float,
     phase: float,
-    g0: float,
+    g0: float = 9.8,
 ) -> float:
     """Gravitational acceleration at the surface of a linear wave."""
     eta = elevation_linear(a, phase)
     return g0 * (
         1 - a * k * np.exp(k * eta) * (np.cos(phase) - a * k * np.sin(phase) ** 2)
     )
+
+
+def gravity_curvilinear(
+    a: float,
+    k: float,
+    phase: float,
+    g0: float = 9.8,
+) -> float:
+    """Gravitational acceleration at the surface of a curvilinear wave."""
+    eta = elevation_linear(a, phase)
+    eps = a * k
+    denominator = np.sqrt((eps * np.sin(phase)) ** 2 + 1)
+    enumerator = g0 * (
+        1
+        - eps * np.exp(eps * np.cos(phase)) * np.cos(phase)
+        - eps**3 * np.exp(eps * np.cos(phase)) * np.cos(phase) * np.sin(phase) ** 2
+    )
+    return enumerator / denominator
+
+
+def gravity_curvilinear_numerical(
+    a: float,
+    k: float,
+    phase: float,
+    g0: float = 9.8,
+) -> float:
+    eta = elevation_linear(a, phase)
+    x = phase / k
+    dx = np.diff(x)[0]
+    omega = np.sqrt(g0 * k)
+    dU_dt = a * omega**2 * np.exp(k * eta) * np.sin(phase) * (a * k * np.cos(phase) + 1)
+    dW_dt = (
+        a * omega**2 * np.exp(k * eta) * (a * k * np.sin(phase) ** 2 - np.cos(phase))
+    )
+    alpha = np.arctan(diff(eta) / dx)
+    g = g0 * np.cos(alpha) + dW_dt * np.cos(alpha) + dU_dt * np.sin(alpha)
+    return g
 
 
 def gravity_stokes(a: float, k: float, phase: float, g0: float) -> float:
@@ -53,14 +91,15 @@ def gravity_stokes(a: float, k: float, phase: float, g0: float) -> float:
 
 
[…]

### 2024-08-23 paper/paper.tex
Comparing effective gravities

@@
 
+We now describe the analytical solutions for the short-wave wavenumber,
+effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
+and intrinsic phase speed in presence of long waves.
+
 \subsection{Wavenumber modulation}
@@
 
-Long waves induce vertical orbital accelerations, so the short
-waves that ride on their surface move in an accelerated reference frame and
-experience effective gravitational acceleration that is lower at the crests and
-higher in the troughs \cite{longuet1986eulerian,longuet1987propagation}.
-A long wave modulates gravitational acceleration at its surface:
+Long waves induce orbital accelerations, so the short waves that ride on their
+surface move in an accelerated reference frame and experience effective
+gravitational acceleration that is lower at the crests and higher in the troughs
+\cite{longuet1986eulerian,longuet1987propagation}.
+Further, a smaller but non-negligible effect is that of the centripetal force
+due to the curvature of the long wave
+\cite{phillips1981dispersion,zhang1990evolution}.
+These effects are completely described by projecting the gravitational
+acceleration from the coordinate that is perpendicular to the mean water surface
+($z=0$) to that which is perpendicular to the long-wave surface ($z=\eta$),
+and likewise for the long-wave orbital accelerations:
 
@@
 \label{eq:gravity_modulation_general}
-\widetilde{g}(\psi)
+\widetilde{g}
   = g \cos{\alpha} 
-  + \dfrac{\partial W}{\partial t} \cos{\alpha}
-  + \frac{\partial U}{\partial t} \sin{\alpha}
+  + \dfrac{\partial W_{z=\eta}}{\partial t} \cos{\alpha}
+  + \dfrac{\partial U_{z=\eta}}{\partial t} \sin{\alpha}
 \end{equation}
-where $\alpha = \arctan{\frac{\partial \eta}{\partial x}}$ is the local slope
-of the long-wave surface.
-Curvilinear effects on effective gravity can be removed by setting $\alpha = 0$.
 
+\begin{equation}
+\label{eq:local_slope}
+\alpha = \tan^{-1}{\dfrac{\partial \eta}{\partial x}}
+\end{equation}
+where $\alpha$ is the local slope of the long-wave surface.
 For a linear wave, the effective gravity is:
@@
 \end{equation}
+
+Curvilinear effects on effective gravity can be neglected by setting $\alpha = 0$.
 
@@
 \end{equation}
-The corresponding gravitational acceleration at the surface of the Stokes wave
-is, to the 4$^{th}$ order:
+Without considering the curvilinear effects in Eq. \ref{eq:gravity_modulation_general},,
+the gravitational acceleration at the surface of a Stokes wave is, to the
+4$^{th}$ order:
 
@@
 \end{equation}
+The equivalent expression that includes the curvilinear effects can be derived
+by evaluating orbital accelerations and local slope $\alpha$ at $z = \eta_{St}$,
+and inserting them into Eq. \ref{eq:gravity_modulation_general}:
 
-With the new long-wave modulated expressions for the short-wave wavenumber
-(Eq. \ref{eq:k_short_exact}) and gravity
-(Eqs. \ref{eq:gravity_modulation_linear}-\ref{eq:gravity_modulation_stokes}),
-we can evaluate the long-wave modulated frequency as
-$\widetilde{\omega}(\psi) = \sqrt{\widetilde{g} \widetilde{k}}$.
-We compare the modulated intrinsic frequency for short waves on the surface of
-linear and Stokes waves with $\varepsilon_L=2$ as function of phase, and their
-phase-integral statistics in Fig. \ref{fig:frequency_modulation}.
+\begin{equation}
+\label{eq:gravity_modulation_general_stokes}
+\widetilde{g}
+  = g \cos{\alpha} 
+  + \dfrac{\partial W_{z=\eta_{St}}}{\partial t} \cos{\alpha_{St}}
+  + \dfrac{\partial U_{z=\eta_{St}}}{\partial t} \sin{\alpha_{St}}
+\end{equation}
+
+\begin{equation}
+\label{eq:local_slope_stokes}
+\alpha_{St} = \tan^{-1}{\dfrac{\partial \eta_{St}}{\partial x}}
+\end{equation}
+
+\begin{figure}[h]
+\label{fig:effective_gravities}
+\centering
+\includegraphics[width=0.6\textwidth]{figures/fig_effective_gravities.png}
+\caption{
+  (a) Analytical solutions for the effective gravitational acceleration
+  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
+  Black is for linear wave evaluated at $z=0$,
+  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
+  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
+  (b) As in (a) but normalized by the modulation of the linear wave evaluated at $z=0$.
+}
+\end{figure}
+
+\begin{figure}[h]
+\label{fig:effective_gravities_mean_max_min}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities_mean_max_min.png}
+\caption{
+  As Fig. \ref{fig:effective_gravities} but showing the mean (solid),
+  maximum (dashed), and minimum (dotted) short-wave gravity modulation as
+  function of long-wave steepness $\varepsilon_L$.
+}
+\end{figure}
 

### 2024-08-23 paper/scripts/compare_effective_gravities.py

@@ -1,0 +1,241 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib
+from model import gravity_linear, gravity_stokes, gravity_curvilinear
+
+matplotlib.rc("font", size=16)
+
+
+def plot_effective_gravities(a: float):
+    a = 0.2
+    phase = np.linspace(0, 2 * np.pi, 10000, endpoint=False)
+    k = 1
+    ak = a * k
+    x = phase / k
+    g0 = 9.8
+    omega = np.sqrt(g0 * k)
+    t = 0
+
+    g1 = g0 * (1 - ak * np.cos(phase))
+    g2 = gravity_linear(x, t, a, k, omega, g0)
+    g3 = gravity_stokes(x, t, a, k, omega, g0)
+    g4 = gravity_curvilinear(x, t, a, k, omega, g0, "linear")
+    g5 = gravity_curvilinear(x, t, a, k, omega, g0, "stokes")
+
+    fig = plt.figure(figsize=(8, 12))
+    axes = fig.subplots(2, 1)
+    axes[0].plot(
+        phase, g1 / g0, label=r"Linear wave, $z=0$", color="black", lw=2, alpha=0.8
+    )
+    axes[0].plot(
+        phase,
+        g2 / g0,
+        label=r"Linear wave, $z=\eta$, $\alpha=0$",
+        color="tab:blue",
+        lw=2,
+        alpha=0.8,
+    )
+    axes[0].plot(
+        phase,
+        g3 / g0,
+        label=r"Stokes wave, $z=\eta$, $\alpha=0$",
+        color="tab:green",
+        lw=2,
+        alpha=0.8,
+    )
+    axes[0].plot(
+        phase,
+        g4 / g0,
+        label=r"Linear wave, $z=\eta$, $\alpha=\frac{\partial \eta}{\partial x}$",
+        color="tab:orange",
+        lw=2,
+        alpha=0.8,
+    )
+    axes[0].plot(
+        phase,
+        g5 / g0,
+        label=r"Stokes wave, $z=\eta$, $\alpha=\frac{\partial \eta}{\partial x}$",
+        color="tab:red",
+        lw=2,
+        alpha=0.8,
+    )
+    axes[0].set_ylabel(r"$\widetilde{g}/g$")
+    axes[0].legend(ncol=1, prop={"size": 13})
+    axes[0].set_title(r"$\varepsilon_L=$" + f"{ak}")
+
[…]

### 2024-08-23 paper/scripts/effective_gravity_analytical.py

@@ -11,25 +11,39 @@ from sympy import (
 
 init_printing()
 
-x, t, a, k, ω, ψ, U, W, g, g0, η, α = symbols(
-    "x t a k ω ψ U W g g_0 η α", real=True, positive=True
+x, t, a, k, ak, ω, ψ, U, W, U_St, W_St, g, g0, η, η_St, α, α_St = symbols(
+    "x t a k ak ω ψ U W U_St W_St g g_0 η η_St α α_St", real=True, positive=True
 )
 
 # Long-wave phase
 ψ = k * x - ω * t
 
-# Long-wave elevation
+ak = a * k
+
+# Long-wave elevation (linear wave)
 η = a * cos(ψ)
+η_St = a * (
+    cos(ψ) + 0.5 * ak * cos(2 * ψ) + ak**2 * (3 / 8.0 * cos(3 * ψ) - 1 / 8.0 * cos(ψ))
+)
 
 # Local elevation slope
 α = atan(diff(η, x))
+α_St = atan(diff(η_St, x))
 
 # Orbital velocities
 U = a * ω * exp(k * η) * cos(ψ)
 W = a * ω * exp(k * η) * sin(ψ)
 
+U_St = a * ω * exp(k * η_St) * cos(ψ)
+W_St = a * ω * exp(k * η_St) * sin(ψ)
+
 # Orbital accelerations
 dU_dt = simplify(diff(U, t))
 dW_dt = simplify(diff(W, t))
 
+dU_dt_St = simplify(diff(U_St, t))
+dW_dt_St = simplify(diff(W_St, t))
+
+# Effective gravities
 g = simplify(g0 * cos(α) + dW_dt * cos(α) + dU_dt * sin(α))
+g_St = simplify(g0 * cos(α_St) + dW_dt_St * cos(α_St) + dU_dt_St * sin(α_St))

### 2024-08-23 paper/scripts/plot_analytical_solutions_with_phase.py

@@ -48,20 +48,75 @@ def gravity_curvilinear(
     return enumerator / denominator
 
 
+def dU_dt_linear(a: float, k: float, phase: float, g0: float = 9.8) -> float:
+    eta = elevation_linear(a, phase)
+    omega = np.sqrt(g0 * k)
+    return a * omega**2 * np.exp(k * eta) * np.sin(phase) * (a * k * np.cos(phase) + 1)
+
+
+def dW_dt_linear(a: float, k: float, phase: float, g0: float = 9.8) -> float:
+    eta = elevation_linear(a, phase)
+    omega = np.sqrt(g0 * k)
+    return a * omega**2 * np.exp(k * eta) * (a * k * np.sin(phase) ** 2 - np.cos(phase))
+
+
+def dU_dt_stokes(a: float, k: float, phase: float, g0: float = 9.8) -> float:
+    eta = elevation_stokes(a, k, phase)
+    omega = np.sqrt(g0 * k)
+    term1 = a * omega**2 * np.exp(k * eta) * np.sin(phase)
+    term2 = (
+        a**2
+        * omega**2
+        * k
+        * np.exp(k * eta)
+        * np.cos(phase)
+        * (
+            np.sin(phase)
+            + a * k * np.sin(2 * phase)
+            + (a * k) ** 2 * (9 / 8.0 * np.sin(3 * phase) - 1 / 16.0 * np.sin(phase))
+        )
+    )
+    return term1 + term2
+
+
+def dW_dt_stokes(a: float, k: float, phase: float, g0: float = 9.8) -> float:
+    eta = elevation_stokes(a, k, phase)
+    omega = np.sqrt(g0 * k)
+    term1 = -a * omega**2 * np.exp(k * eta) * np.cos(phase)
+    term2 = (
+        a**2
+        * omega**2
+        * k
+        * np.exp(k * eta)
+        * np.sin(phase)
+        * (
+            np.sin(phase)
+            + a * k * np.sin(2 * phase)
+            + (a * k) ** 2 * (9 / 8.0 * np.sin(3 * phase) - 1 / 16.0 * np.sin(phase))
+        )
+    )
+    return term1 + term2
+
+
 def gravity_curvilinear_numerical(
     a: float,
     k: float,
     phase: float,
     g0: float = 9.8,
+    long_wave: str = "stokes",
 ) -> float:
-    eta = elevation_linear(a, phase)
[…]

### 2024-08-26 paper/paper.tex
Switch to JFM template

@@
-\documentclass[draft]{agujournal2019}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage{soul}
+\documentclass[lineno]{jfm}
+
+\usepackage{graphicx}
+%\usepackage{epstopdf,epsfig}
+\usepackage{newtxtext}
+\usepackage{newtxmath}
+\usepackage{natbib}
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks = true,
+    urlcolor   = blue,
+    citecolor  = black,
+}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+\newcommand{\RomanNumeralCaps}[1]
 \linenumbers
 
-\draftfalse
+% {\MakeUppercase{\romannumeral #1}}
 
-\journalname{JGR: Oceans}
+\title{Revisiting the hydrodynamic wave modulation from linear wave theory}
+
+\author{
+  Milan Curcic\aff{1}
+  \corresp{\email{mcurcic@miami.edu}}
+}
+
+\affiliation{
+  \aff{1}Rosenstiel School of Marine, Atmospheric, and Earth Science, University of Miami, Miami, FL
+  \aff{2}Frost Institute for Data Science and Computing, University of Miami, Coral Gables, FL
+}
 
 \begin{document}
-
-%\title{Dispersion of short waves riding on longer waves}
-\title{Revisiting the hydrodynamic wave modulation from linear wave theory}
-
-\authors{Milan Curcic\affil{1,2}}
-
-\affiliation{1}{
-  Rosenstiel School of Marine, Atmospheric, and Earth Science,
-  University of Miami,
-  Miami, FL
-}
-\affiliation{2}{
-  Frost Institute for Data Science and Computing,
-  University of Miami,
-  Coral Gables, FL
-}
-
-\correspondingauthor{Milan Curcic}{mcurcic@miami.edu}
-
-\begin{keypoints}
-\item We revisit analytical and numerical solutions for hydrodynamic modulation of short waves by longer waves.
-\item We present a higher-order analytical solution based on linear wave theory.
-\item In contrast to an earlier study, we find stable numerical solutions for modulation by long-wave groups.
-\end{keypoints}
+\maketitle
 
@@
 steepness modulation significantly deviate from the linear analytical solutions
-of \citeA{longuet1960changes}, and are similar to the nonlinear numerical
-solutions by \citeA{zhang1990evolution}.
-We also discuss the result of \citeA{peureux2021unsteady} who found through
+of \citet{longuet1960changes}, and are similar to the nonlinear numerical
+solutions by \citet{zhang1990evolution}.
+We also discuss the result of \citet{peureux2021unsteady} who found through
 numerical simulations that the short-wave modulation grows unsteadily with
@@
 
-\section*{Plain Language Summary}
-The ocean surface is typically covered by waves of many wavelengths.
-Longer waves bunch up and dilate the shorter waves, and this process is
-important to consider for both the measurements and the forecasting of ocean
-waves.
-Although there have been analytical and numerical solutions for how the short
-waves should change in the presence of longer waves, they have either been
-simplistic or 
-This paper revisits and extends these solutions to propose a more accurate
-description of how the short waves should change in the presence of longer
-waves.
+\begin{keywords}
+%Authors should not enter keywords on the manuscript, as these must be chosen by the author during the online submission process and will then be added during the typesetting process (see \href{https://www.cambridge.org/core/journals/journal-of-fluid-mechanics/information/list-of-keywords}{Keyword PDF} for the full list).
+Other classifications will be added at the same time.
+\end{keywords}
+
+%{\bf MSC Codes }  {\it(Optional)} Please enter your MSC Codes here
 
@@
 Short ocean surface waves are modulated by longer swell.
-Following the result by \citeA{unna1941white,unna1942waves,unna1947sea},
-\citeA{longuet1960changes} introduced an analytical solution for hydrodynamic
+Following the result by \citet{unna1941white,unna1942waves,unna1947sea},
+\citet{longuet1960changes} introduced an analytical solution for hydrodynamic
 modulation of short waves at the crest of an underlying longer wave:
@@
 This result has been influential on the development of remote sensing of ocean
-surface waves and currents \cite{keller1975microwave,hara1994hydrodynamic}.
-\citeA{longuet1987propagation} revisited this problem by considering variation
+surface waves and currents \citep{keller1975microwave,hara1994hydrodynamic}.
+\citet{longuet1987propagation} revisited this problem by considering variation
 in the gravitational acceleration and fully non-linear wave numerical solutions
@@
 Eq. \ref{eq:lhs1960wavenumber}.
-\citeA{zhang1990evolution} evaluated the effects of nonlinearity of short waves
+\citet{zhang1990evolution} evaluated the effects of nonlinearity of short waves
 using a non-linear Schr\" odinger equation and found higher wavenumber
 modulation but lesser amplitude modulation than those predicted by
-\citeA{longuet1987propagation}.
-More recently, \cite{peureux2021unsteady} simulated the full wave conservation
+\citet{longuet1987propagation}.
+More recently, \citep{peureux2021unsteady} simulated the full wave conservation
 equations and found that the short waves grow unsteadily due to the propagation of
@@
 
-\begin{table}[h]
-\label{table:modulation_literature_summary}
-\caption{
-  Summary of prior
-  \cite{longuet1960changes,phillips1981dispersion,longuet1987propagation,zhang1990evolution,peureux2021unsteady}
-  and present (this paper) approaches to calculating the modulation of short
-  waves by long waves.
-  *: long 3rd order Stokes wave is considered only in numerical simulations.
-}
+\begin{table}
+\begin{center}
+\def~{\hphantom{0}}
 \begin{tabular}{cccccc}
-\hline
 Reference   & Governing & Gravity & Nonlinear  & Nonlinear   & Solution \\
             & equations & varies  & long waves & short waves &          \\
-\hline
 \hline
@@
 L-H 1987    & Velocity potential       & Yes & Yes & No  & Numerical \\
+\hline
+H 1988      & Hamiltonian              & Yes & Yes & No  & Analytical \\
 \hline
@@
 \end{tabular}
+\caption{
+  Summary of prior
+  \citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,zhang1990evolution,peureux2021unsteady}
+  and present (this paper) approaches to calculating the modulation of short
+  waves by long waves.
+  *: long 3rd order Stokes wave is considered only in numerical simulations.
+}
+\label{table:modulation_literature_summary}
+\end{center}
 \end{table}
@@
 An important limitation here is that $U$ must be slow-varying on the scales of
-the wave period \cite{bretherton1968wavetrains}.
+the wave period \citep{bretherton1968wavetrains}.
 In the context of short waves riding on long waves, the mean advective current
@@
 The second important relation is the conservation of wave crests
-\cite{phillips1981dispersion}:
+\citep{phillips1981dispersion}:
 
@@
 they move in an accelerated reference frame and will thus experience effective
-gravitational acceleration that changes in both space and time \cite{longuet1986eulerian,longuet1987propagation}.
+gravitational acceleration that changes in both space and time
+\citep{longuet1986eulerian,longuet1987propagation}.
 Inserting Eq. (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:
@@
 which here represents the inhomogeneity of the group speed.
-As \citeA{peureux2021unsteady} explained, this term is responsible for unstable
+As \citet{peureux2021unsteady} explained, this term is responsible for unstable
 growth of wave action in infinite long-wave trains.
@@
 Note that here we derive the solutions from a different starting set of
-equations than that of \citeA{longuet1960changes}.
+equations than that of \citet{longuet1960changes}.
 Rather than starting from the conservation of crests and wave action, as we do
@@
 \end{equation}
-Although \citeA{peureux2021unsteady} pointed out that the solution by L-HS
+Although \citet{peureux2021unsteady} pointed out that the solution by L-HS
 requires assuming homogeneous group speed of the short-wave field, in fact,
@@
 
-\begin{figure}[ht]
-\label{fig:analytical_solutions_ak0.1}
+\begin{figure}
 \centering
@@
   speed as function of long-wave phase for $\varepsilon_L = 0.1$, based on
-  analytical solutions by \citeA{longuet1960changes} (L-HS 1960, black)
+  analytical solutions by \citet{longuet1960changes} (L-HS 1960, black)
   and this paper (blue).
@@
 }
+\label{fig:analytical_solutions_ak0.1}
 \end{figure}
@@
 gravitational acceleration that is lower at the crests and higher in the troughs
-\cite{longuet1986eulerian,longuet1987propagation}.
+\citep{longuet1986eulerian,longuet1987propagation}.
 Further, a smaller but non-negligible effect is that of the centripetal force
 due to the curvature of the long wave
-\cite{phillips1981dispersion,zhang1990evolution}.
+\citep{phillips1981dispersion,zhang1990evolution}.
 These effects are completely described by projecting the gravitational
@@
 
-\begin{figure}[h]
-\label{fig:effective_gravities}
+\begin{figure}
 \centering
@@
 }
+\label{fig:effective_gravities}
 \end{figure}
 
-\begin{figure}[h]
-\label{fig:effective_gravities_mean_max_min}
+\begin{figure}
 \centering
@@
 }
+\label{fig:effective_gravities_mean_max_min}
 \end{figure}
@@
 
-\begin{figure}[ht]
-\label{fig:analytical_solutions_ak0.2}
+\begin{figure}
 \centering
@@
 \caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.2$.}
+\label{fig:analytical_solutions_ak0.2}
 \end{figure}
 
-\begin{figure}[ht]
-\label{fig:analytical_solutions_ak0.4}
+\begin{figure}
 \centering
@@
 \caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.4$.}
+\label{fig:analytical_solutions_ak0.4}
 \end{figure}
@@
 2$^{nd}$  order central finite difference in space and integrated in time using the
-4$^{th}$ order Runge-Kutta method \cite{butcher1996runge}.
+4$^{th}$ order Runge-Kutta method \citep{butcher1996runge}.
 The space is divided into 100 grid points.
 This is effectively the same equation set and numerical configuration as that of
-\citeA{peureux2021unsteady}, except that, like \citeA{longuet1960changes}, their
+\citet{peureux2021unsteady}, except that, like \citet{longuet1960changes}, their
 long-wave orbital velocities are evaluated at $z=0$ instead of $z=\eta_L$.
@@
 We show in the next section that our numerical results in infinite long-wave
-trains are qualitatively equivalent to those of \citeA{peureux2021unsteady}.
+trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.
 
@@
 However, this difference is important to keep in mind when comparing the results
-of \citeA{peureux2021unsteady} and those presented here--in their Fig. 1 the
+of \citet{peureux2021unsteady} and those presented here--in their Fig. 1 the
 long-wave phase is fixed and the short waves are moving leftward
@@
 
-\begin{figure}[h]
-\label{fig:numerical_solutions}
+\begin{figure}
 \centering
@@
 }
+\label{fig:numerical_solutions}
 \end{figure}
 
-\begin{figure}[h]
-\label{fig:numerical_vs_analytical_k}
+\begin{figure}
 \centering
@@
 \caption{}
+\label{fig:numerical_vs_analytical_k}
 \end{figure}
 
-\begin{figure}[h]
-\label{fig:numerical_vs_analytical_a}
+\begin{figure}
 \centering
@@
 \caption{}
+\label{fig:numerical_vs_analytical_a}
 \end{figure}
 
-\begin{figure}[h]
-\label{fig:numerical_vs_analytical_ak}
+\begin{figure}
 \centering
@@
 \caption{}
+\label{fig:numerical_vs_analytical_ak}
 \end{figure}
@@
 
-The key result of \citeA{peureux2021unsteady} is that of unsteady steepening of
+The key result of \citet{peureux2021unsteady} is that of unsteady steepening of
 short waves when a homogeneous field of short waves is forced by an infinite
@@
 is approximately doubled, which almost exactly corresponds to the unsteady growth
-reported by \citeA{peureux2021unsteady} in their Fig. 1.
+reported by \citet{peureux2021unsteady} in their Fig. 1.
 Another important property of this solution is that the short-wave group that
@@
 
-\begin{figure}[h]
-\label{fig:modulation_3panel_infinite}
+\begin{figure}
 \centering
@@
 }
+\label{fig:modulation_3panel_infinite}
 \end{figure}
@@
 Why does this unsteady growth of short-wave action occur?
-As \citeA{peureux2021unsteady} explain, "the additional change in advection
+As \citet{peureux2021unsteady} explain, "the additional change in advection
 velocity due to the modulation of the short wave group speed gives an additional
@@
 
-\begin{figure}[h]
-\label{fig:modulation_3panel_ramp}
+\begin{figure}
 \centering
@@
 }
+\label{fig:modulation_3panel_ramp}
 \end{figure}
@@
 initial conditions?
-\citeA{peureux2021unsteady} did not provide an explanation for this behavior.
+\citet{peureux2021unsteady} did not provide an explanation for this behavior.
 At first, it may seem that there is something special about the uniform initial
 conditions that makes its solutions grow unsteadily and indefinitely.
-\citeA{peureux2021unsteady} describe this scenario as "...the sudden appearance
+\citet{peureux2021unsteady} describe this scenario as "...the sudden appearance
 of a long wave perturbation in the middle of a homogeneous short wave field".
@@
 except for a linear ramp applied to $a_L$ (and consequently, $U$ and $W$ as well)
-for the first five periods reveals that the unsteady growh no longer occurs
+for the first five periods reveals that the unsteady growth no longer occurs
 (Fig. \ref{fig:modulation_3panel_ramp}).
 
-\begin{figure}[h]
-  \label{fig:unsteady_growth_timeseries}
+\begin{figure}
   \centering
@@
   }
+  \label{fig:unsteady_growth_timeseries}
 \end{figure}
@@
 
-\begin{figure}[h]
-\label{fig:modulation_3panel_groups}
+\begin{figure}
 \centering
@@
 }
+\label{fig:modulation_3panel_groups}
 \end{figure}
@@
 
-\acknowledgments
+In this paper we revisited the problem of short-wave modulation by long waves
+from a linear wave theory perspective.
+Using the linearized wave crest and action conservation equations, coupled with
+the nonlinear gravity modulation for short waves that includes the the curvature
+effects of the long wave, we derived new analytical solutions for the modulation
+of short wave wavenumber, amplitude, steepness, intrinsic frequency, and phase
+speed.
+The latter two have not been previously examined in the context of short-wave
+modulation, but are important for correctly calculating the wind input into
+short waves, which is proportional to the wind speed relative to the wave
+celerity.
+
+\backsection[Acknowledgements]{
 Thanks to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan for the
 helpful discussions.
+}
 
-\bibliography{references.bib}
+\backsection[Funding]{
+  Milan Curcic has been partially supported by the National Science Foundation
+  under grant number AGS-1745384.
+}
+
+\backsection[Declaration of interests]{The author reports no conflict of interest.}
+
+\backsection[Data availability statement]{
+  The data and code to generate the figures and run the numerical simulations
+  are available at [TODO].
+  The numerical model is available as a standalone package at [TODO].
+  %See JFM's \href{https://www.cambridge.org/core/journals/journal-of-fluid-mechanics/information/journal-policies/research-transparency}{research transparency policy} for more information
+}
+
+\backsection[Author ORCIDs]{M. Curcic, https://orcid.org/0000-0002-8822-7749}
+
+%\backsection[Author contributions]{Authors may include details of the contributions made by each author to the manuscript}
+
+\bibliographystyle{jfm}
+\bibliography{references}
 

### 2024-08-27 paper/paper.tex
Conclusion paragraph

@@
 \hline
-This paper  & Crest and action         & Yes & Yes* & No  & Analytical \\
+This paper  & Crest and action         & Yes & Yes & No  & Analytical \\
             & conservation             &     &     &     & and numerical \\
@@
   Summary of prior
-  \citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,zhang1990evolution,peureux2021unsteady}
+  \citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,henyey1988energy,zhang1990evolution,peureux2021unsteady}
   and present (this paper) approaches to calculating the modulation of short
   waves by long waves.
-  *: long 3rd order Stokes wave is considered only in numerical simulations.
 }
@@
 celerity.
+The approximate steady solutions yield higher wavenumber and amplitude
+modulation than that of \citet{longuet1960changes}, and somewhat lower than 
+the numerical solutions of \citet{longuet1987propagation} and
+\citet{zhang1990evolution}.
+This is mainly due to the analytical solution requiring the crest and action
+conservation equations to be linearized, and the nonlinearity arising only due
+to the evaluation of short waves on the surface of the long wave (as opposed to
+the mean water level) and due to nonlinear effective gravity.
+
+The numerical solutions of the full wave crest and action conservation equations
+are consistent with the analytical solutions for $\varepsilon_L < 0.2$.
+The simulations also reveal that the unsteady growth of short-wave steepness
+reported by \citet{peureux2021unsteady} is due to the sudden appearance of
+long waves in a homogeneous short-wave field, a scenario that is unlikely to
+occur in nature.
+A more likely conditions, such as that of a long-wave groups, are only mildly
+destabilizing for the short waves.
+These results suggest that for mildly-to-moderately steep short waves,
+the analytical steady solutions presented here are sufficient to describe the
+modulation of short waves by long waves.
 

### 2024-08-27 paper/scripts/compare_effective_gravities.py
Update scripts

@@ -1,7 +1,7 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import matplotlib
-from model import gravity_linear, gravity_stokes, gravity_curvilinear
+from twowave import gravity, gravity_curvilinear
 
 matplotlib.rc("font", size=16)
 
@@ -17,8 +17,8 @@ def plot_effective_gravities(a: float):
     t = 0
 
     g1 = g0 * (1 - ak * np.cos(phase))
-    g2 = gravity_linear(x, t, a, k, omega, g0)
-    g3 = gravity_stokes(x, t, a, k, omega, g0)
+    g2 = gravity(x, t, a, k, omega, g0, "linear")
+    g3 = gravity(x, t, a, k, omega, g0, "stokes")
     g4 = gravity_curvilinear(x, t, a, k, omega, g0, "linear")
     g5 = gravity_curvilinear(x, t, a, k, omega, g0, "stokes")
 
@@ -129,8 +129,8 @@ def plot_effective_gravities(a: float):
     for a in np.arange(0, 0.41, 0.01):
         ak = a * k
         g1 = g0 * (1 - ak * np.cos(phase))
-        g2 = gravity_linear(x, t, a, k, omega, g0)
-        g3 = gravity_stokes(x, t, a, k, omega, g0)
+        g2 = gravity(x, t, a, k, omega, g0, "linear")
+        g3 = gravity(x, t, a, k, omega, g0, "stokes")
         g4 = gravity_curvilinear(x, t, a, k, omega, g0, "linear")
         g5 = gravity_curvilinear(x, t, a, k, omega, g0, "stokes")
         g1_max.append(np.max(g1))

### 2024-08-27 paper/scripts/modulation.py

@@ -1,56 +1,0 @@
-import numpy as np
-
-
-def gravity(
-    x: float, t: float, a: float, k: float, omega: float, g0: float, order: int = 3
-) -> float:
-    assert order in [0, 1, 2, 3]
-    phi = phase(x, t, k, omega)
-    term1 = -a * k * np.cos(phi)
-    term2 = -a * k * np.cos(phi) * np.exp(a * k * np.cos(phi))
-    term3 = ((a * k * np.sin(phi)) ** 2) * np.exp(a * k * np.cos(phi))
-    if order == 0:
-        return g0 * np.ones_like(x)
-    elif order == 1:
-        return g0 * (1 + term1)
-    elif order == 2:
-        return g0 * (1 + term2)
-    elif order == 3:
-        return g0 * (1 + term2 + term3)
-    else:
-        raise ValueError("Invalid order")
-
-
-def phase(
-    x: float | np.ndarray,
-    t: float | np.ndarray,
-    k: float | np.ndarray,
-    omega: float | np.ndarray,
-) -> float | np.ndarray:
-    return k * x - omega * t
-
-
-def wavenumber(
-    frequency: float | np.ndarray,
-    water_depth: float,
-    grav: float = 9.8,
-    surface_tension: float = 0.074,
-    water_density: float = 1e3,
-    num_iterations: int = 100,
-) -> float | np.ndarray:
-    """Solve the dispersion relationship for wavenumber using the
-    Newton-Raphson method."""
-    frequency_nondim = 2 * np.pi * np.sqrt(water_depth / grav) * frequency
-    k = frequency_nondim**2
-    surface_tension_nondim = surface_tension / (grav * water_density * water_depth**2)
-    count = 0
-    while count < num_iterations:
-        t = np.tanh(k)
-        dk = -(frequency_nondim**2 - k * t * (1 + surface_tension_nondim * k**2)) / (
-            3 * surface_tension_nondim * k**2 * t
-            + t
-            + k * (1 + surface_tension_nondim * k**2) * (1 - t**2)
-        )
-        k -= dk
-        count += 1
-    return k / water_depth

### 2024-08-27 paper/scripts/plot_analytical_solutions_with_phase.py

@@ -1,163 +1,23 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import matplotlib
-from model import diff
+from twowave import gravity_curvilinear
 
 matplotlib.rc("font", size=12)
-
-
-def elevation_linear(a: float, phase: float) -> float:
-    return a * np.cos(phase)
-
-
-def elevation_stokes(a: float, k: float, phase: float) -> float:
-    term1 = np.cos(phase)
-    term2 = 0.5 * a * k * np.cos(2 * phase)
-    term3 = (a * k) ** 2 * (3 / 8 * np.cos(3 * phase) - 1 / 16 * np.cos(phase))
-    return a * (term1 + term2 + term3)
-
-
-def gravity_linear(
-    a: float,
-    k: float,
-    phase: float,
-    g0: float = 9.8,
-) -> float:
-    """Gravitational acceleration at the surface of a linear wave."""
-    eta = elevation_linear(a, phase)
-    return g0 * (
-        1 - a * k * np.exp(k * eta) * (np.cos(phase) - a * k * np.sin(phase) ** 2)
-    )
-
-
-def gravity_curvilinear(
-    a: float,
-    k: float,
-    phase: float,
-    g0: float = 9.8,
-) -> float:
-    """Gravitational acceleration at the surface of a curvilinear wave."""
-    eta = elevation_linear(a, phase)
-    eps = a * k
-    denominator = np.sqrt((eps * np.sin(phase)) ** 2 + 1)
-    enumerator = g0 * (
-        1
-        - eps * np.exp(eps * np.cos(phase)) * np.cos(phase)
-        - eps**3 * np.exp(eps * np.cos(phase)) * np.cos(phase) * np.sin(phase) ** 2
-    )
-    return enumerator / denominator
-
-
-def dU_dt_linear(a: float, k: float, phase: float, g0: float = 9.8) -> float:
-    eta = elevation_linear(a, phase)
-    omega = np.sqrt(g0 * k)
-    return a * omega**2 * np.exp(k * eta) * np.sin(phase) * (a * k * np.cos(phase) + 1)
-
-
-def dW_dt_linear(a: float, k: float, phase: float, g0: float = 9.8) -> float:
-    eta = elevation_linear(a, phase)
-    omega = np.sqrt(g0 * k)
-    return a * omega**2 * np.exp(k * eta) * (a * k * np.sin(phase) ** 2 - np.cos(phase))
-
-
[…]

### 2024-08-27 paper/scripts/plot_numerical_simulations.py

@@ -1,5 +1,5 @@
 import matplotlib.pyplot as plt
-from model import WaveModulationModel
+from twowave import WaveModulationModel
 import numpy as np
 
 

### 2024-08-27 paper/scripts/plot_numerical_simulations_ramp_properties.py

@@ -1,5 +1,5 @@
 import matplotlib.pyplot as plt
-from model import WaveModulationModel
+from twowave import WaveModulationModel
 import numpy as np
 import matplotlib
 

### 2024-08-27 paper/scripts/plot_numerical_vs_analytical_solutions.py

@@ -1,5 +1,5 @@
 import matplotlib.pyplot as plt
-from model import WaveModulationModel
+from twowave import WaveModulationModel
 import numpy as np
 import pandas as pd
 import matplotlib

### 2024-08-27 paper/scripts/plot_scale_analysis.py

@@ -2,7 +2,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import matplotlib
 
-import model
+import twowave
 
 matplotlib.rc("font", size=12)
 
@@ -76,11 +76,11 @@ def calculate_tendencies(a_L: float):
     Cp = omega / k
     Cg = Cp / 2
 
-    dkdx = model.diff(k) / dx
-    dgdx = model.diff(g) / dx
-    dUdx = model.diff(U) / dx
-    dCgdx = model.diff(Cg) / dx
-    dNdx = model.diff(N) / dx
+    dkdx = twowave.diff(k) / dx
+    dgdx = twowave.diff(g) / dx
+    dUdx = twowave.diff(U) / dx
+    dCgdx = twowave.diff(Cg) / dx
+    dNdx = twowave.diff(N) / dx
 
     k_propagation = -Cg * dkdx
     k_advection = -U * dkdx

### 2024-08-27 paper/paper.tex
WIP paper

@@
 Short ocean surface waves are modulated by longer swell.
-Following the result by \citet{unna1941white,unna1942waves,unna1947sea},
-\citet{longuet1960changes} introduced an analytical solution for hydrodynamic
-modulation of short waves at the crest of an underlying longer wave:
+Following \citet{unna1941white,unna1942waves,unna1947sea},
+\citet{longuet1960changes} introduced a steady, approximate solution for
+hydrodynamic modulation of short waves by longer waves:
 
@@
 \end{equation}
-where $\widetilde{k}$ and $k$ are modulated and unmodulated short-wave wavenumbers,
+where $\widetilde{k}$ and $k$ are the modulated and unmodulated short-wave wavenumbers,
 respectively, and $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
 wavenumber $k_L$ and amplitude $a_L$.
-This result has been influential on the development of remote sensing of ocean
+This result influenced the development of remote sensing of ocean
 surface waves and currents \citep{keller1975microwave,hara1994hydrodynamic}.
-\citet{longuet1987propagation} revisited this problem by considering variation
-in the gravitational acceleration and fully non-linear wave numerical solutions
-to find significantly stronger modulation than that predicted by
-Eq. \ref{eq:lhs1960wavenumber}.
-\citet{zhang1990evolution} evaluated the effects of nonlinearity of short waves
-using a non-linear Schr\" odinger equation and found higher wavenumber
-modulation but lesser amplitude modulation than those predicted by
+\citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited this
+problem by nonlinear long waves and variations in the effective gravity of short
+waves, yielding significantly stronger modulation than previously predicted.
+\citet{henyey1988energy} derived an analytical solution for the modulation of
+short waves by long waves using Hamiltonian mechanics and reported the modulation
+of similar magnitude to that of \citet{longuet1987propagation}.
+\citet{zhang1990evolution} considered weakly nonlinear short waves
+using a non-linear Schr\" odinger equation and found stronger wavenumber
+modulation but weaker amplitude modulation than those predicted by
 \citet{longuet1987propagation}.
-More recently, \citep{peureux2021unsteady} simulated the full wave conservation
-equations and found that the short waves grow unsteadily due to the propagation of
+The variety of analytical and numerical frameworks (Eulerian, Hamiltonian,
+Schr\" odinger) that relatively closely reproduce steady solutions for short-wave
+modulation suggests that the solutions are robust.
+Nevertheless, alternative analytical and numerical approaches to the modulation
+problem are still of interest, especially if higher degrees of nonlinearity
+can be considered using simpler approaches.
+
+More recently, \citep{peureux2021unsteady} asked whether the steady solutions
+of short-wave modulation are realistic by simulating the full wave conservation
+equations
+They found that the short waves grow unsteadily due to the propagation of
 longer waves, suggesting that the steady-state solutions may only be valid for
-a few long-wave periods before the short waves are steep enough to break.
+a few long-wave periods before the short waves begin to break due to excessive
+steepness.
+However, the unsteadiness of their solutions occurs in a specific scenario of
+a uniform short-wave field that suddenly appears in a uniform long-wave train,
+and the numerical simulations stabilize if the short-wave field is initialized
+from existing steady solutions.
 
@@
 \hline
-P 1981      & Crest and energy         & Yes & Yes & No  & Numerical \\
+P 1981      & Crest and action    & Yes & Yes & No  & Numerical \\
             & conservation        &     &     &     &           \\
@@
 \hline
-H 1988      & Hamiltonian              & Yes & Yes & No  & Analytical \\
+HCDSW 1988  & Hamiltonian         & Yes & Yes & No  & Analytical \\
 \hline
-ZM 1990     & Nonlinear Schr\" odinger & Yes & Yes & Yes & Numerical \\
-            & equation                 &     &     &     &           \\
+ZM 1990     & Schr\" odinger      & Yes & Yes & Yes & Numerical \\
 \hline

### 2024-08-28 paper/paper.tex
Working on the text

@@
 modulation suggests that the solutions are robust.
-Nevertheless, alternative analytical and numerical approaches to the modulation
-problem are still of interest, especially if higher degrees of nonlinearity
-can be considered using simpler approaches.
+Alternative analytical and numerical approaches to the modulation problem remain
+of interest, especially if higher degrees of nonlinearity can be considered
+using simpler approaches.
 
@@
 from existing steady solutions.
+Nevertheless, their results do put into question whether the short-wave
+modulation predicted by theory is generally steady over many long-wave periods
+or not.
+
+In this paper, we revisit this problem from linear wave theory and
+derive alternative non-linear solutions for the modulation of short waves,
+evaluated at the long-wave surface.
+The analytical solutions yield modulations that are stronger than the original solution
+by \citet{longuet1960changes}, but similar to the nonlinear numerical solutions
+by \citet{longuet1987propagation} and \citet{zhang1990evolution} in long-wave
+steepness up to $\approx 0.2$.
+We also perform numerical simulations of the full wave crest and action
+conservation laws and evaluate the effect of long-wave groups on the short waves.
+The numerical simulations are used to determine the validity of the approximate
+steady solution and to investigate the unsteadiness of the short-wave modulation
+found by \citet{peureux2021unsteady}.
+The main prior studies on hydrodynamic modulation, and their assumptions and
+approaches to the solution, are summarized in Table
+\ref{table:modulation_literature_summary}.
 
@@
 
-In this paper, we revisit this problem from a linear wave theory perspective and
-derive new non-linear solutions for the modulation of short waves, evaluated
-at the long-wave surface.
-We also perform numerical simulations of the full wave crest and action
-conservation laws and evaluate the effect of long-wave groups on the short waves.
-The next section presents the governing equations, followed by the analytical
-solutions in section \ref{section:analytical_solutions}.
-In section \ref{section:numerical_solutions} we present the numerical solutions
-and compare them to the analytical solutions.
+This paper first presents the governing equations and then the analytical
+solutions for the modulation of short waves by long waves and compares them to
+the original solution by \citet{longuet1960changes}.
+Then, we compare the analytical solutions to numerical solutions of the full
+wave conservation equations in section \ref{section:numerical_solutions}.
 In section \ref{section:scale_analysis} we perform a scale analysis of the
@@
 
-In this paper we consider inviscid, irrotational, and incompressible flow.
+We consider inviscid, irrotational, and incompressible flow.
 In these conditions the linear (sine-like), deep-water, surface gravity waves
@@
 of the wave propagation.
-A current in the direction of the waves thus increases its absolution frequency,
-wavenumber remaining unchanged.
-An important limitation here is that $U$ must be slow-varying on the scales of
+A current in the direction of the waves increases its absolute frequency,
+while the wavenumber remains unchanged.
+An important limitation here is that $U$ must be slowly varying on the scales of
 the wave period \citep{bretherton1968wavetrains}.
@@
 is the horizontal near-surface orbital velocity of the long wave.
-
-The second important relation is the conservation of wave crests
-\citep{phillips1981dispersion}:
+The evolution of the short-wave wavenumber is described by the conservation of
+wave crests \citep{phillips1981dispersion}:
 
@@
 In deep water, the group speed is half the phase speed, so $C_g = 1/2\sqrt{g/k}$.
+Although the absence of wave growth and dissipation is not generally realistic
+in the open ocean, it is a helpful approximation that allows us to isolate the
+hydrodynamic modulation effects solely due to the motion of the long waves.
 
-As the short waves ride on the surface of the long wave,
-they move in an accelerated reference frame and will thus experience effective
-gravitational acceleration that changes in both space and time
-\citep{longuet1986eulerian,longuet1987propagation}.
-Inserting Eq. (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:
+Short waves that ride on the surface of the longer waves move in an accelerated
+reference frame and thus experience effective gravitational acceleration that
+changes in space and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
+From Eqs. (\ref{eq:dispersion}) and (\ref{eq:wave_action}) we obtain:
 
@@
 In a non-accelerated reference frame (i.e. in absence of longer waves), the last
-term is zero.
-However, if gravitational acceleration varies in space, this term is necessary
-to conserve the wave crests.
+term vanishes.
+Otherwise, this term is necessary to conserve the wave crests.
 
@@
 growth of wave action in infinite long-wave trains.
-Eqs. \ref{eq:wavenumber} and \ref{eq:wave_action2} will be the governing equations
-that we solve numerically in section \ref{section:numerical_solutions}.
+Eqs. \ref{eq:wavenumber} and \ref{eq:wave_action2} are the governing equations
+that we approximate and solve analytically in section \ref{section:analytical_solutions},
+and solve numerically in their full form in section \ref{section:numerical_solutions}.
 
@@
 long-wave phase, and $\omega_L$ its angular frequency.
-
 At this time we define the long-wave steepness $\varepsilon_L = a_L k_L$ which
-we will use thoughout the rest of the paper as a scaling parameter.
+we will use throughout the rest of the paper as a scaling parameter, as is common
+in the literature.
 The long wave induces horizontal and vertical orbital velocities whose values
@@
 Eqs. \ref{eq:dispersion}-\ref{eq:w_L} is the full system of equations that we
-exploit in the remainder of this paper.
+use in this paper.
 Note that here we derive the solutions from a different starting set of
@@
 potential of two waves.
-It appears the starting from the conservation of crests and wave action allows
+It appears that starting from the conservation of crests and wave action allows
 for a simpler and more intuitive derivation.

### 2024-08-29 paper/paper.tex
Wavenumber modulation edits

@@
 of \citet{longuet1960changes}, and are similar to the nonlinear numerical
-solutions by \citet{zhang1990evolution}.
-We also discuss the result of \citet{peureux2021unsteady} who found through
+solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
+We also examine the result of \citet{peureux2021unsteady} who found through
 numerical simulations that the short-wave modulation grows unsteadily with
 each long-wave phase.
-Here we show that this unsteady growth only occurs for homogeneous initial
+We show that this unsteady growth only occurs for homogeneous initial
 conditions and does not occur for more realistic long-wave groups.
-Finally, the combined analytical and numerical framework allows us to quantify
-the contribution of various terms in the full wave conservation equations.
+The combined analytical and numerical framework presented here allows us to
+quantify the importance of various terms in the full wave conservation
+equations.
 The new analytical solutions are a good approximation of the full numerical
@@
 defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
-where $a_L$ is the long-wave amplitude, $\psi = k_L x - \omega_L t$ is the
-long-wave phase, and $\omega_L$ its angular frequency.
-At this time we define the long-wave steepness $\varepsilon_L = a_L k_L$ which
-we will use throughout the rest of the paper as a scaling parameter, as is common
-in the literature.
+where $\psi = k_L x - \omega_L t$ is the long-wave phase and $\omega_L$ its
+angular frequency.
 The long wave induces horizontal and vertical orbital velocities whose values
@@
 and intrinsic phase speed in presence of long waves.
+The wavenumber is derived by linearizing the conservation of wave crests
+(Eq. \ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
+The effective gravity is derived by projecting the long-wave induced surface
+accelerations onto the axis that is normal to the long-wave surface.
+(Eq. \ref{eq:wave_action2}) and using the expression for the long-wave orbital
+velocities.
+The amplitude is derived by first evaluation the short-wave action from the
+action conservation equation (Eq. \ref{eq:wave_action}) and then applying the
+modulated effective gravity to obtain the amplitude.
+From the modulated wavenumber, amplitude, and effective gravity, the steepness,
+frequency, and phase speed readily follow from the usual wave relationships.
+Going forward we will use the tilde over the variables to denote the modulated
+quantities (e.g. $\widetilde{k}$ for wavenumber).
 
@@
 \label{eq:k_short_exact}
-\widetilde{k}(\psi) = k e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
+\widetilde{k} = k e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
 \end{equation}
@@
 Notice that the Taylor expansion of Eq. \ref{eq:k_short_exact} to the first order
-recovers the original solution by L-HS:
+recovers the original solution by \citet{longuet1960changes}:
 
@@
 \label{eq:k_short_lhs}
-\widetilde{k}(\psi) = k (1 + \varepsilon_L \cos{\psi})
+\widetilde{k} = k (1 + \varepsilon_L \cos{\psi})
 \end{equation}
-
-The L-HS solution requires two approximations in their solution:
-First, evaluating the long-wave velocity at $z = 0$
-rather than $z = \eta_L$; and second, evaluating the solution to the first order
-of the Taylor expansion series.
-Both of these approximations cause an underestimate of the short-wave
-modulation magnitude (Fig. \ref{fig:wavenumber_modulation}).
-The original L-HS solution is a cosine wave with an amplitude equal
-to the long-wave steepness.
-The other two solutions both introduce an increase in wave modulation
-on the crest and a decrease in the trough.
-At the crest, the relative modulation predicted by the exact L-HS solution
-is 10.7\% larger than that predicted by the 1st order approximate L-HS
-solution, while the solution that evaluates the long-wave orbital velocity
-at the water surface rather than $z = 0$ is 38.4\% larger than that of the
-1$^{st}$ order approximate L-HS solution.
+Coming from the conservation of wave crests, as opposed to the velocity potential
+superposition approach, the L-HS solution requires two approximations in their
+solution:
+First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
+second, further truncating the Taylor expansion series beyond the first order term.
+These approximations cause an underestimate of the short-wave modulation
+magnitude (Fig. \ref{fig:analytical_solutions_ak0.1}a).
+At the wave crest, Eq. \ref{eq:k_short_exact} yields the wavenumber modulation
+that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of \citet{longuet1960changes}
+for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
 

### 2024-09-06 paper/paper.tex
Gravity modulation section

@@
 
-\title{Revisiting the hydrodynamic wave modulation from linear wave theory}
+\title{Revisiting the hydrodynamic modulation of short waves by longer waves}
 
@@
 of short-wave modulation are realistic by simulating the full wave conservation
-equations
+equations.
 They found that the short waves grow unsteadily due to the propagation of
@@
 for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
+Although not immediately obvious from the expressions, the modulated wavenumber
+in Eq. \ref{eq:k_short_lhs} is conservative across the long-wave phase, whereas
+the higher-order solution in Eq. \ref{eq:k_short_exact} is not.
+This is due to the fact that combining the orbital velocities evaluated at
+$z=\eta$ and the linearized conservation of wave crests in Eq. \ref{eq:wavenumber}
+results in excess short-wave wavenumber being created.
+The solution of the full Eq. \ref{eq:wavenumber} of course must be conservative
+across the long-wave phase, as we will see later from the numerical simulations.
 
@@
 \citep{longuet1986eulerian,longuet1987propagation}.
-Further, a smaller but non-negligible effect is that of the centripetal force
-due to the curvature of the long wave
+Further, centripetal force due to the curvature of the long wave induces small
+but non-negligible effects on the short-wave effective gravity
 \citep{phillips1981dispersion,zhang1990evolution}.
@@
 where $\alpha$ is the local slope of the long-wave surface.
-For a linear wave, the effective gravity is:
+For a linear long wave, the effective gravity is:
 
@@
 
-Curvilinear effects on effective gravity can be neglected by setting $\alpha = 0$.
+Curvilinear effects on effective gravity can be removed altogether by setting
+$\alpha = 0$.
+In that case, Eq. \ref{eq:gravity_modulation_general} simplifies to:
 
-The first-order term in $a_L k_L$ is in phase with the long-wave elevation,
-so the apparent gravity of the short waves is increased at the long-wave crests
-and decreased in the troughs.
-The second-order term in $a_L k_L$ is positive-definite and phase-shifted by
-$\pi/2$ relative to the crests.
-The second-order term thus increases the apparent gravity on the front and rear
-faces of the long wave (Fig. \ref{fig:gravity_modulation}).
+\begin{equation}
+\label{eq:gravity_modulation_linear_no_curvature}
+\widetilde{g} = g \left( 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}} \right)
+\end{equation}
+Further, if we evaluate the orbital accelerations at the mean water level $z=0$,
+this further simplifies to:
 
-We can also derive the short-wave gravitational acceleration on the surface of
-a non-linear Stokes wave with the elevation (to the third order in $\varepsilon_L$):
+\begin{equation}
+\label{eq:gravity_modulation_linear_no_curvature_mean_level}
+\widetilde{g} = g \left( 1 - \varepsilon_L \cos{\psi} \right)
+\end{equation}
+which is the form used by \citet{peureux2021unsteady}, for example.
+
+The general form of the gravity modulation in Eq. \ref{eq:gravity_modulation_general}
+is convenient because it allows us to easily attribute the effects of the different
+processes on the gravity modulation.
+It also makes it straightforward to evaluate the gravity modulation for a different
+form of the long-wave elevation, by evaluating the orbital accelerations and the
+local elevation slope at the appropriate level.
+For a third-order Stokes wave, for example, the elevation is:
 
@@
 
+Let us now examine the contributions of the different terms to the short-wave
+effective gravity modulation.
+Fig. \ref{fig:effective_gravities}a shows the effective gravity modulation of
+short-wave as a function of the long-wave phase for $\varepsilon_L = 0.2$, for
+a variety of the gravity modulation forms.
+This value of $\varepsilon_L$ allows easier visualization of the modulation
+differences, while still being relatively moderate.
+To the first order in $\varepsilon_L$, the gravity modulation is out of phase
+with the long-wave elevation, with the short waves experiencing a lower effective
+gravity on the long-wave crests and a higher effective gravity in the troughs.
+Evaluating the orbital accelerations at $z=\eta$ causes a gravity reduction
+around the crests ($\approx6\%$) and the troughs ($\approx 3\%$), and an
+increase on the front and rear faces ($\approx 4\%$) of the long wave, due to
+the exponential change in the orbital accelerations away from the mean water level
+(Fig. \ref{fig:effective_gravities}b).
+The difference between the linear and Stokes long waves at this steepness is
+negligible.
+The long-wave elevation curvature has a second-order effect in $\varepsilon_L$.
+Naturally, there is no curvature exactly at the crests and troughs, and its
+effects are largest at the front and rear faces of the long wave, where the
+covariance between the local slope and the horizontal accelerations are greatest.
+
 \begin{figure}
 \centering
-\includegraphics[width=0.6\textwidth]{figures/fig_effective_gravities.png}
+\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities.png}
 \caption{
@@
 
+It is instructive to also examine the mean, maximum, and minimum values of the
+effective gravity as function of long-wave steepness $\varepsilon_L$
+(Fig. \ref{fig:effective_gravities_mean_max_min}).
+The long-wave averaged effective gravity of short waves remains equal to the
+resting value of $g$ for all values of $\varepsilon_L$, except if the effects
+of curvature are considered, in which case the average effective gravity is
+reduced.
+This reduction, in case of a long Stokes wave, is $\approx 3\%$ for
+$\varepsilon_L = 0.2$ and $\approx 14\%$ for $\varepsilon_L = 0.4$.
+As the effective gravity minima and maxima occur at crests and troughs,
+respectively, the curvature effects only affect the phase-averaged values but
+not the extremes.
+The effective gravity at the crest of the steepest ($\varepsilon_L = 0.4$) long
+Stokes wave is reduced by $\approx 64\%$.
+
 \begin{figure}
@@
   function of long-wave steepness $\varepsilon_L$.
+  The minimum and maximum values for the cases with curvature effects
+  ($\alpha=\frac{\partial \eta}{\partial x}$) are equivalent to those without
+  curvature effects ($\alpha=0$) and obscure them in the plot.
 }
@@
 \backsection[Funding]{
-  Milan Curcic has been partially supported by the National Science Foundation
+  Milan Curcic was partially supported by the National Science Foundation
   under grant number AGS-1745384.

### 2024-09-06 paper/scripts/compare_effective_gravities.py

@@ -107,6 +107,23 @@ def plot_effective_gravities(a: float):
         ax.set_xticks(np.linspace(0, 2 * np.pi, 5))
         ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
 
+    axes[0].text(
+        0.,
+        1.05,
+        "(a)",
+        transform=axes[0].transAxes,
+        fontsize=16,
+        va="top",
+    )
+    axes[1].text(
+        0.,
+        1.05,
+        "(b)",
+        transform=axes[1].transAxes,
+        fontsize=16,
+        va="top",
+    )
+
     plt.subplots_adjust(left=0.12, right=0.95, bottom=0.05, top=0.95)
     plt.savefig("../figures/fig_effective_gravities.png", dpi=150)
     plt.close()

### 2024-09-06 paper/paper.tex
Analytical solutions done

@@
 \label{eq:k_short_exact}
-\widetilde{k} = k e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
+\frac{\widetilde{k}}{k} = e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
 \end{equation}
@@
 \label{eq:k_short_lhs}
-\widetilde{k} = k (1 + \varepsilon_L \cos{\psi})
+\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
 \end{equation}
@@
 \caption{
-  Modulation of short-wave (a) wavenumber, (b) gravitational acceleration,
-  (c) amplitude, (d) steepness, (e) intrinsic frequency, and (d) intrinsic phase
-  speed as function of long-wave phase for $\varepsilon_L = 0.1$, based on
-  analytical solutions by \citet{longuet1960changes} (L-HS 1960, black)
-  and this paper (blue).
+  Steady solutions for the modulation of short-wave (a) wavenumber, (b)
+  gravitational acceleration, (c) amplitude, (d) steepness, (e) intrinsic
+  frequency, and (f) intrinsic phase speed as function of long-wave phase for
+  $\varepsilon_L = 0.1$, based on
+  \citet{longuet1960changes} (L-HS 1960, black) and this paper (blue).
   Long-wave crest and trough are located at $\psi = 0$ and $\psi = \pi$,
@@
 \label{eq:gravity_modulation_general}
-\widetilde{g}
-  = g \cos{\alpha} 
+\frac{\widetilde{g}}{g}
+  = \cos{\alpha} 
   + \dfrac{\partial W_{z=\eta}}{\partial t} \cos{\alpha}
@@
 \label{eq:gravity_modulation_linear}
-\widetilde{g}
-  = g 
+\frac{\widetilde{g}}{g}
+  = 
   \frac{
@@
 \label{eq:gravity_modulation_linear_no_curvature}
-\widetilde{g} = g \left( 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}} \right)
+\frac{\widetilde{g}}{g} = \left( 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}} \right)
 \end{equation}
@@
 \label{eq:gravity_modulation_linear_no_curvature_mean_level}
-\widetilde{g} = g \left( 1 - \varepsilon_L \cos{\psi} \right)
+\widetilde{g} = \left( 1 - \varepsilon_L \cos{\psi} \right)
 \end{equation}
@@
 \label{eq:gravity_modulation_stokes}
-\widetilde{g}(\psi) =
-g \left\{
+\frac{\widetilde{g}}{g} =
+\left\{
   1 - \varepsilon_L e^{k \eta_{St}}
@@
 \label{eq:gravity_modulation_general_stokes}
-\widetilde{g}
-  = g \cos{\alpha} 
+\frac{\widetilde{g}}{g}
+  = \cos{\alpha} 
   + \dfrac{\partial W_{z=\eta_{St}}}{\partial t} \cos{\alpha_{St}}
@@
   and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
-  (b) As in (a) but normalized by the modulation of the linear wave evaluated at $z=0$.
+  (b) As in (a) but normalized by the modulation of the linear wave evaluated at $z=0$, $g_1$.
 }
@@
 As the effective gravity minima and maxima occur at crests and troughs,
-respectively, the curvature effects only affect the phase-averaged values but
-not the extremes.
-The effective gravity at the crest of the steepest ($\varepsilon_L = 0.4$) long
-Stokes wave is reduced by $\approx 64\%$.
+respectively, the curvature affects the phase-averaged values but not the extremes.
+The effective gravity reduction at the crests is $\approx 11\%$, $25\%$, and $66\%$
+for $\varepsilon_L = 0.1$, $0.2$, and $0.4$, respectively.
+These are similar to the fully-nonlinear numerical estimates of
+\citet{longuet1986eulerian,longuet1987propagation}.
+Note that these are strictly Eulerian values, evaluated at fixed positions in
+space.
+As \citet{longuet1986eulerian} pointed out, the effective gravity in the
+Lagrangian frame is smoother and less amplified because the moving particles
+spend a longer time riding near the crests and less time returning through the
+troughs.
+However, when we solve the short-wave crest and action conservation equations
+in an Eulerian frame, it is the Eulerian effective gravity that must be applied
+to conserve the short-wave properties across the long-wave phase.
 
@@
 
-The modulation of short-wave amplitude can be derived in the same way by
-dropping the non-linear terms in the wave action balance equation
-(Eq. \ref{eq:wave_action2}) and integrating wave action in time:
+The modulation of short-wave amplitude can be derived in a similar way as we did
+for the wavenumber, except that here we linearize the wave action balance and
+integrate wave action in time:
 
@@
 \label{eq:wave_action_modulation}
-\widetilde{N}(\psi) = N e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
+\frac{\widetilde{N}}{N} = e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
 \end{equation}
@@
 Since wave action is energy divided by the intrinsic frequency $\sigma$,
-and energy scales with $ga^2$, the modulation of the amplitude is readily
-related to the modulations of gravity, wavenumber, and action:
+and energy scales with $ga^2$, the modulation of the amplitude is related to the
+modulations of gravity, wavenumber, and action:
 
@@
 \end{equation}
-To the 1$^{st}$ order in $\varepsilon_L$, and using Eqs. \ref{eq:k_short_exact}
-and \ref{eq:wave_action_modulation}, we get:
+To the first order in $\varepsilon_L$, and using Eqs. \ref{eq:k_short_exact}
+and \ref{eq:wave_action_modulation}, the above simplifies to:
 
@@
 \end{equation}
+
+The amplitude and steepness modulation of the short waves as a function of the
+long-wave phase for $\varepsilon_L = 0.1$, $0.2$, and $0.4$ are shown and compared
+to the \citet{longuet1960changes} solutions in Figs.
+\ref{fig:analytical_solutions_ak0.1}, \ref{fig:analytical_solutions_ak0.2},
+and \ref{fig:analytical_solutions_ak0.4}, respectively.
+Also shown are the modulation of the short-wave intrinsic frequency and phase
+speed in panels (e) and (f), respectively.
+As the wavenumber and gravity modulations are out of phase, with the latter
+being somewhat larger in magnitude, the overall intrinsic frequency of the
+short waves is reduced across the entire long-wave phase, with highest reduction
+at the long-wave crests.
+For the short-wave phase speed modulation, the wavenumber and gravity modulations
+work in tandem and cause the short-wave phase speed to be reduced on the long-wave
+crests and increased in the troughs.
+We will discuss the implications of this effect later in the context of wind
+input into short waves.
 
@@
 
-We now proceed to integrate the full numerical solutions for the short-wave
-wavenumber and action density in presence of long waves, and in absence of
-sources or sinks of wave energy.
+To quantify the contribution of the nonlinear terms in Eqs. \ref{eq:wavenumber}
+and \ref{eq:wave_action2}, and to evaluate the steadiness assumption of the
+analytical solutions, we now proceed to numerically integrate the full set of
+crest and action conservation equations.
 Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
-2$^{nd}$  order central finite difference in space and integrated in time using the
-4$^{th}$ order Runge-Kutta method \citep{butcher1996runge}.
+second-order central finite difference in space and integrated in time using the
+forth-order Runge-Kutta method \citep{butcher1996runge}.
 The space is divided into 100 grid points.
 This is effectively the same equation set and numerical configuration as that of
-\citet{peureux2021unsteady}, except that, like \citet{longuet1960changes}, their
-long-wave orbital velocities are evaluated at $z=0$ instead of $z=\eta_L$.
-The key difference is that evaluating orbital velocities at $z=0$ neglects the
-Stokes drift of short-wave groups.
-Although not significantly consequential for the modulation results here, there
-is no good reason to exclude it given that its implementation is trivial.
-
+\citet{peureux2021unsteady}, except that their long-wave orbital velocities are
+evaluated at $z=0$ instead of $z=\eta_L$, which neglects the Stokes drift of
+short-wave groups.
+Their gravity modulation also does not consider the nonlinear or the curvature
+effects, however, this difference does not qualitatively affect the results.
 Another difference is the choice of the numerical scheme for spatial differences,
-which in their case is the MUSCL4 scheme by Kurganov and Tadmor (2000 TODO).
-Although the 2$^{nd}$ order central finite difference is not appropriate for many
+which in their case is the MUSCL4 scheme by \citet{kurganov2000new}.
+Although the second-order central finite difference is not appropriate for many
 numerical problems, in our case it is sufficient because it is conservative and
 the fields that we compute derivatives of are smooth and continuous.
-We show in the next section that our numerical results in infinite long-wave
+We show in the next section that our numerical solutions for infinite long-wave
 trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.
@@
 Finally, the numerical equations here are integrated in a fixed reference frame,
-rather than one that moves with the long wave phase speed.
-Either approach produces qualitatively equivalent results, but we find that
-a fixed reference frame allows for more intuitive interpretation of the results.
-However, this difference is important to keep in mind when comparing the results
-of \citet{peureux2021unsteady} and those presented here--in their Fig. 1 the
+rather than that moving with the long-wave phase speed.
+Either approach produces qualitatively equivalent results, however a fixed
+reference frame may allow for a more intuitive interpretation of the results.
+This difference is important to keep in mind when comparing the results
+of \citet{peureux2021unsteady} and those presented here; in their Fig. 1 the
 long-wave phase is fixed and the short waves are moving leftward
-with the speed of $C_{pL} - C_g - U$ (neglecting any group speed inhomogeneity),
+with the speed of $C_{pL} - C_g - U$ (neglecting group speed inhomogeneity),
 whereas in the figures that follow, the long wave is moving rightward with its
 phase speed $C_{pL}$ and the short waves (that is, their energy) are moving
-rightward with the speed of $C_g + U$ (neglecting any group wave inhomogeneity).
+rightward with the speed of $C_g + U$ (neglecting group speed inhomogeneity).
 
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
 \caption{}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
 \caption{}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
 \caption{}

### 2024-09-13 paper/scripts/calculate_numerical_diff_error.py
Specify numerical error of centered finite difference

@@ -1,0 +1,21 @@
+import numpy as np
+from twowave import elevation, diff
+
+phase = np.linspace(0, 2 * np.pi, 100, endpoint=False)
+k = 1
+a = 0.4
+omega = np.sqrt(9.8 * k)
+
+x = phase / k
+t = 0
+
+eta = elevation(x, t, a, k, omega)
+deta_dx_exact = (-a * k * np.sin(phase))[1:-1]
+deta_dx = (diff(eta) / diff(x))[1:-1]
+
+max_abs_error = np.max(np.abs(deta_dx - deta_dx_exact))
+rel_error = np.abs(deta_dx - deta_dx_exact) / np.abs(deta_dx_exact)
+max_rel_error = np.max(rel_error[rel_error < 1])
+
+print("Max. abs. error: ", max_abs_error)
+print("Max. rel. error: ", max_rel_error * 100, "%")

### 2024-09-13 paper/scripts/compare_effective_gravities.py

@@ -108,7 +108,7 @@ def plot_effective_gravities(a: float):
         ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
 
     axes[0].text(
-        0.,
+        0.0,
         1.05,
         "(a)",
         transform=axes[0].transAxes,
@@ -116,7 +116,7 @@ def plot_effective_gravities(a: float):
         va="top",
     )
     axes[1].text(
-        0.,
+        0.0,
         1.05,
         "(b)",
         transform=axes[1].transAxes,

### 2024-09-19 paper/paper.tex
Specify numerical error of centered finite difference
Numerical results

@@
 
-More recently, \citep{peureux2021unsteady} asked whether the steady solutions
+More recently, \citet{peureux2021unsteady} asked whether the steady solutions
 of short-wave modulation are realistic by simulating the full wave conservation
@@
 the gravitational acceleration at the surface of a Stokes wave is, to the
-4$^{th}$ order:
+fourth order:
 
@@
 the fields that we compute derivatives of are smooth and continuous.
+The maximum relative error of the centered finite difference in this model
+is $\approx 0.066\%$.
 We show in the next section that our numerical solutions for infinite long-wave
@@
 
+The first set of simulations is performed for the long-wave steepness of
+$\varepsilon_L = 0.1$, $0.2$, and $0.4$ (Fig. \ref{fig:numerical_solutions}).
+The long-waves are initialized from rest ($a_L = 0$) and are ramped up
+to their target steepness after 5 long-wave periods, thus mimicking the
+passage of a long-wave group
+(we explain in more detail the importance of ramping up the long-wave forcing
+in the next subsection).
+At $\varepsilon = 0.1$, the numerical solutions for the wavenumber, amplitude,
+and steepness modulation are all remarkably similar to the analytical solutions
+(Fig. \ref{fig:numerical_solutions}a,d,g).
+This suggests that at low long-wave steepness that steady analytical solution
+is a reasonable proxy for the fully-nonlinear crest-action conservation
+equation solutions.
+The numerical solutions diverge more notably from the analytical ones at
+moderate steepness of 0.2 (Fig. \ref{fig:numerical_solutions}b,e,h), and are
+considerably different at very high steepness of 0.4
+(Fig. \ref{fig:numerical_solutions}c,f,i).
+These differences suggest that at high steepness the inhomogeneity of
+gravitational acceleration and group speed become important.
+
 \begin{figure}
@@
 
+The maximum modulations of the short-wave wavenumber, amplitude, and steepness
+from the analytical and numerical solutions, are shown on Figs.
+\ref{fig:numerical_vs_analytical_k},
+\ref{fig:numerical_vs_analytical_a}, and
+\ref{fig:numerical_vs_analytical_ak}, respectively, as function of the long-wave
+steepness $\varepsilon_L$.
+Overall, the numerical solutions, based on the linear and third-order Stokes long wave
+alike, begin to diverge from the analytical solution at $\varepsilon_L \approx 0.2$.
+The numerical solution using the Stokes long wave is quantitatively very similar
+to the fully-nonlinear solutions of \citet{longuet1987propagation} in amplitude
+modulation, however with somewhat lower wavenumber and steepness modulations.
+For example, at $\varepsilon_L = 0.4$, the steepness modulation in our solutions
+is close to 6, whereas \citet{longuet1987propagation} finds it close to 8.
+The underestimate in the wavenumber modulation at high long-wave steepness
+relative to the results of \citet{longuet1987propagation}
+is likely due to his long wave velocities being evaluated to a higher order
+than they are here.
+Nevertheless, the overall steepness modulation between the two numerical
+approaches are remarkably similar, and either solution would cause the
+short-waves of even low steepness to break \citep{banner1993wave}.
+
 \begin{figure}
@@
 \includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
-\caption{}
+\caption{
+  Maximum wavenumber modulation as function of long-wave steepness $\varepsilon_L$.
+  Black line and circles are for the analytical and numerical solutions from
+  \citet{longuet1960changes} and \citet{longuet1987propagation}, respectively.
+  Blue line is based on the analytical solutions from this paper.
+  Green and orange lines are for the numerical solutions from this paper
+  using linear and third-order Stokes long waves, respectively.
+}
 \label{fig:numerical_vs_analytical_k}
@@
 \includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
-\caption{}
+\caption{
+  As Fig. \ref{fig:numerical_vs_analytical_k} but for the amplitude modulation.
+}
 \label{fig:numerical_vs_analytical_a}
@@
 \includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
-\caption{}
+\caption{
+  As Fig. \ref{fig:numerical_vs_analytical_k} but for the steepness modulation.
+}
 \label{fig:numerical_vs_analytical_ak}

### 2024-09-19 paper/paper.tex
Plotting inhomogeneity tendencies

@@
 conditions and does not occur for more realistic long-wave groups.
-The combined analytical and numerical framework presented here allows us to
-quantify the importance of various terms in the full wave conservation
+The combined analytical and numerical framework presented here allows
+quantifying the importance of various terms in the full wave conservation
 equations.
-The new analytical solutions are a good approximation of the full numerical
-solution in long-wave steepness up to $\approx 0.2$.
+The proposed analytical solutions are a good approximation of the fully
+nonlinear numerical solutions in long-wave steepness up to $\approx 0.2$.
 \end{abstract}
@@
 \citet{peureux2021unsteady}, except that their long-wave orbital velocities are
-evaluated at $z=0$ instead of $z=\eta_L$, which neglects the Stokes drift of
-short-wave groups.
+evaluated at $z=0$ instead of $z=\eta_L$, thus neglecting the Stokes drift
+induced by long waves on short-wave groups
+\citep{stokes1846,van2018stokes,monismith2020stokes}.
 Their gravity modulation also does not consider the nonlinear or the curvature
@@
 waves that begins its journey in the trough of the long wave.
-This is because the unmodulated short waves that begin in the trough enter into
-the area where the surface velocities are converging.
-Inversely, the short waves that begin around the crest of the long-wave will
-enter into the area of surface divergence and will thus have their wave action
-and wavenumber, decreased.
+This is because the unmodulated short waves that begin their journey in the
+trough first enter the convergence region before reaching the divergence region.
+Inversely, the short waves that begin around the crest of the long-wave first
+enter the divergence region and will thus have their action and wavenumber
+decreased.
 We reproduce this behavior in Fig. \ref{fig:modulation_3panel_infinite}, which
@@
 modulation (here defined as the change of a quantity relative to its initial value).
+The long waves have $\varepsilon_L = 0.1$ and $k_L = 1$, and the short waves
+have $k = 10$.
 After 10 long-wave periods, the short-wave action modulation $\widetilde{N}/N$
-is approximately doubled, which almost exactly corresponds to the unsteady growth
-reported by \citet{peureux2021unsteady} in their Fig. 1.
+is approximately doubled, consistent with the unsteady growth reported by
+\citet{peureux2021unsteady}.
 Another important property of this solution is that the short-wave group that
-experiences peak amplification (dampening) are not locked-in to the long-wave
-crests (troughs), as they are in the L-HS solution.
+experiences peak amplification (dampening) are not locked in to the long-wave
+crests (troughs), as they are in the solution by \citet{longuet1960changes}.
 
@@
 such that their wavenumber (and optionally, action) is higher on the crests
-and lower in the troughs, akin to the L-HS solution, the instability does not
-occur.
+and lower in the troughs, akin to the prior steady solutions, the instability
+does not occur.
 
@@
 
-Why is it, then, that the solution is so different depending on the short-wave
+Why is it, then, that the solution for short-wave modulation is not only
+quantitatively, but also qualitatively, different depending on the short-wave
 initial conditions?
-\citet{peureux2021unsteady} did not provide an explanation for this behavior.
-At first, it may seem that there is something special about the uniform initial
+It may at first seem that there is something special about the uniform initial
 conditions that makes its solutions grow unsteadily and indefinitely.
-\citet{peureux2021unsteady} describe this scenario as "...the sudden appearance
+\citet{peureux2021unsteady} described this scenario as "...the sudden appearance
 of a long wave perturbation in the middle of a homogeneous short wave field".
-The sudden appearance of long waves may be a clue.
-The exact numerical simulation as shown in Fig. \ref{fig:modulation_3panel_infinite},
-except for a linear ramp applied to $a_L$ (and consequently, $U$ and $W$ as well)
-for the first five periods reveals that the unsteady growth no longer occurs
-(Fig. \ref{fig:modulation_3panel_ramp}).
+However, such a scenario is not realistic in the ocean where waves tend to
+organize into groups due to their dispersive nature.
+To test whether it is the sudden onset of long-wave forcing that causes the
+unstable growth, we repeat the numerical simulation of Fig.
+\ref{fig:modulation_3panel_infinite}, except for a linear ramp applied to $a_L$
+(and consequently, $U$ and $W$ as well) for the first five periods:
+$a_L(t) = min(a_L, a_L t / \left(5 T_L\right))$, where $T_L$ is the long-wave period.
+This result is shown in Fig. \ref{fig:modulation_3panel_ramp}.
+The key distinction that arises with the introduction of the ramp is that the
+modulation pattern of short waves is now locked in with the long-wave crests,
+and their indefinite steepening vanishes.
+The stabilization effect of the long-wave ramp is more readily assessed in
+Fig. \ref{fig:unsteady_growth_timeseries}, which shows the maximum short-wave
+steepness modulation as function of time over 30 long-wave periods, for the
+infinite long-wave train case and that when applying the linear ramp to the
+long-wave forcing.
+Comparing the wave action propagation ($C_g \frac{\partial N}{\partial x}$) and
+inhomogeneity ($N \frac{\partial C_g}{\partial x}$) tendencies reveals that
+after only one long-wave period, the propagation and inhomogeneity tendencies
+are unbalanced and amplify the steepening of short waves
+(Fig. \ref{fig:inhomogeneity_tendencies}).
+Although \citet{peureux2021unsteady} correctly pointed out that the unstable
+growth of short-wave steepness occurs due to the resonance of the inhomogeneity
+tendency with the short-wave action perturbation, in case of a gentle ramp
+up of the long-wave forcing, the propagation and inhomogeneity tendencies
+mostly balance each other out, thus allowing the short-wave steepening to
+lock in to the long-wave crests.
 
@@
   \label{fig:unsteady_growth_timeseries}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/inhomogeneity_tendencies.png}
+  \caption{
+    Propagation ($C_g \frac{\partial N}{\partial x}$) (blue) and inhomogeneity
+    ($N \frac{\partial C_g}{\partial x}$) (orange) tendencies and their sum
+    (green) for the short-wave action after one long-wave period, in the case of
+    (a) infinite long-wave train and (b) linear ramp of long-wave amplitude.
+    Dashed line is the long-wave elevation and the dotted line is the short-wave
+    action modulation minus 1.
+  }
+  \label{fig:inhomogeneity_tendencies}
 \end{figure}
@@
 \end{figure}
+
+\section{Sensitivity to wavenumber ratios}
+\label{section:sensitivity_to_wavenumber_ratios}
 

### 2024-09-19 paper/scripts/plot_inhomogeneity_tendencies.py

@@ -1,0 +1,100 @@
+import matplotlib.pyplot as plt
+from twowave import WaveModulationModel
+import numpy as np
+import matplotlib
+
+matplotlib.rc("font", size=16)
+
+m1 = WaveModulationModel(num_periods=10)
+m2 = WaveModulationModel(num_periods=10)
+
+m1.run(ramp_type=None, wave_type="linear", save_tendencies=True)
+m2.run(ramp_type="linear", wave_type="linear", save_tendencies=True)
+
+ds1 = m1.to_xarray()
+ds2 = m2.to_xarray()
+
+T_L = 2 * np.pi / m1.omega_long
+periods = ds1.time / T_L
+
+ak1 = ds1.amplitude * ds1.wavenumber
+ak2 = ds2.amplitude * ds2.wavenumber
+ak_init = 0.1
+
+t = 1 * T_L
+n0 = np.argmin((ds1.time.values - t) ** 2)
+
+eta = m1.a_long * np.cos(ds1.space)
+
+fig = plt.figure(figsize=(12, 8))
+axes = [
+    plt.subplot2grid((2, 1), (0, 0)),
+    plt.subplot2grid((2, 1), (1, 0)),
+]
+axes[0].plot(ds1.space, eta, "k--", lw=3, label=r"$\eta_L$")
+axes[0].plot(ds1.space, ds1.wave_action[n0] - 1, "k:", lw=3, label=r"$N - 1$")
+axes[0].plot(
+    ds1.space,
+    ds1.N_propagation_tendency[n0],
+    lw=3,
+    label=r"$C_g \frac{\partial N}{\partial x}$",
+)
+axes[0].plot(
+    ds1.space,
+    ds1.N_inhomogeneity_tendency[n0],
+    lw=3,
+    label=r"$N \frac{\partial C_g}{\partial x}$",
+)
+axes[0].plot(
+    ds1.space,
+    ds1.N_propagation_tendency.values[n0] + ds1.N_inhomogeneity_tendency.values[n0],
+    lw=3,
+    label=r"$C_g \frac{\partial N}{\partial x} + N \frac{\partial C_g}{\partial x}$",
+)
+axes[1].plot(ds2.space, eta, "k--", lw=3, label=r"$\eta_L$")
+axes[1].plot(ds1.space, ds2.wave_action[n0] - 1, "k:", lw=3, label=r"$N - 1$")
+axes[1].plot(
+    ds2.space,
[…]

### 2024-09-19 paper/paper.tex
Begin work on wave groups
Update paper.tex

@@
 
+We now proceed to examine the numerical solutions with the long-wave forcing
+applied as a long-wave group.
+The wave group is simulated as a non-dispersive train of long waves whose
+amplitude scales with $\cos{2 \pi t / (N T_L)}$, where $N$ is the number of
+long waves in the group.
+The elevation of the long waves thus gradually ramps up from rest to $a_L$ at $t = N/2 T_L$, after which it gradually decreases toward zero.
+In this simulation, $N = 10$ and $\varepsilon = 0.1$ as in the previous two simulations.
+The modulations of short-wave wavenumber, amplitude, and steepness in response to such long-wave group is shown in Fig. \ref{fig:modulation_3panel_groups}.
+Like in the case of a linear ramp, the modulation pattern here is locked in to the long-wave crest and peaks at approximately 1.2 for short-wave steepness.
+After the peak of the long-wave group passes, the short-wave modulation recedes back toward the resting state.
+
 \begin{figure}

### 2024-09-19 paper/paper.tex
Update paper.tex

@@
 from a linear wave theory perspective.
-Using the linearized wave crest and action conservation equations, coupled with
-the nonlinear gravity modulation for short waves that includes the the curvature
-effects of the long wave, we derived new analytical solutions for the modulation
+Using the linearized wave crest and action conservation equations coupled with nonlinear solutions for the effective gravity, we derived steady analytical solutions for the modulation
 of short wave wavenumber, amplitude, steepness, intrinsic frequency, and phase
@@
 This is mainly due to the analytical solution requiring the crest and action
-conservation equations to be linearized, and the nonlinearity arising only due
+conservation equations to be linearized, and the nonlinearity arising due
 to the evaluation of short waves on the surface of the long wave (as opposed to
@@
 long waves in a homogeneous short-wave field, a scenario that is unlikely to
-occur in nature.
-A more likely conditions, such as that of a long-wave groups, are only mildly
+occur in the open ocean.
+A more likely condition such as that of a long-wave groups, are only mildly
 destabilizing for the short waves.

### 2024-09-20 paper/paper.tex
Updating the model

@@
 
-Short ocean surface waves are modulated by longer swell.
+Short ocean surface waves are hydrodynamically modulated by longer swell.
 Following \citet{unna1941white,unna1942waves,unna1947sea},
@@
 wavenumber $k_L$ and amplitude $a_L$.
-This result influenced the development of remote sensing of ocean
-surface waves and currents \citep{keller1975microwave,hara1994hydrodynamic}.
-\citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited this
-problem by nonlinear long waves and variations in the effective gravity of short
-waves, yielding significantly stronger modulation than previously predicted.
+This hydrodynamic modulation process was instrumental for the development and
+interpretation of remote sensing of ocean surface waves and currents
+\citep{keller1975microwave,hara1994hydrodynamic}, and is distinguished
+from the aerodynamic modulation process \citep{donelan1987effect,belcher1999wave,chen2000effects}.
+\citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited the
+hydrodynamic modulation problem by considering nonlinear long waves and
+variations in the effective gravity of short waves.
+Their results yielded significantly stronger modulation than previously predicted.
 \citet{henyey1988energy} derived an analytical solution for the modulation of
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figures/fig_modulation_ramp_timeseries.png}
+  \includegraphics[width=0.8\textwidth]{figures/fig_modulation_ramp_timeseries.png}
   \caption{
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{figures/inhomogeneity_tendencies.png}
+  \includegraphics[width=\textwidth]{figures/inhomogeneity_tendencies.png}
   \caption{
@@
 
-\section{Sensitivity to wavenumber ratios}
-\label{section:sensitivity_to_wavenumber_ratios}
-
-\section{Scale analysis}
-\label{section:scale_analysis}
-
-Now that we showed that the analytical solution is consistent with full numerical
-solutions for $\varepsilon_L < 0.25$, we can use analytical solutions for scale
-analysis to quantify the relative contribution of tendencies in the full equation set.
-
-TODO
+\section{Discussion}
+\label{section:discussion}
 

### 2024-09-20 paper/scripts/plot_numerical_simulations_ramp_properties.py

@@ -8,8 +8,8 @@ matplotlib.rc("font", size=16)
 m1 = WaveModulationModel(num_periods=30)
 m2 = WaveModulationModel(num_periods=30)
 
-m1.run(ramp_type=None, wave_type="stokes")
-m2.run(ramp_type="linear", wave_type="stokes")
+m1.run(ramp_type=None, wave_type="linear")
+m2.run(ramp_type="linear", wave_type="linear")
 
 ds1 = m1.to_xarray()
 ds2 = m2.to_xarray()

### 2024-09-21 paper/paper.tex
WIP discussion

@@
 
-Solving analytically for the wavenumber requires dropping spatial gradients of
-$k$ in Eq. \ref{eq:wavenumber} and assuming homogeneous $g$:
+Solving analytically for the wavenumber requires dropping the spatial derivatives
+of $k$ in Eq. \ref{eq:wavenumber} and assuming homogeneous $g$:
 
@@
 long waves in the group.
-The elevation of the long waves thus gradually ramps up from rest to $a_L$ at $t = N/2 T_L$, after which it gradually decreases toward zero.
-In this simulation, $N = 10$ and $\varepsilon = 0.1$ as in the previous two simulations.
-The modulations of short-wave wavenumber, amplitude, and steepness in response to such long-wave group is shown in Fig. \ref{fig:modulation_3panel_groups}.
-Like in the case of a linear ramp, the modulation pattern here is locked in to the long-wave crest and peaks at approximately 1.2 for short-wave steepness.
-After the peak of the long-wave group passes, the short-wave modulation recedes back toward the resting state.
+The elevation of the long waves thus gradually ramps up from rest to $a_L$ at
+$t = N/2 T_L$, after which it gradually decreases toward zero.
+In this simulation, $N = 10$ and $\varepsilon = 0.1$ as in the previous two
+simulations.
+The modulations of short-wave wavenumber, amplitude, and steepness in response
+to such long-wave group is shown in Fig. \ref{fig:modulation_3panel_groups}.
+Like in the case of a linear ramp, the modulation pattern here is locked in to
+the long-wave crest and peaks at approximately 1.2 for short-wave steepness.
+After the peak of the long-wave group passes, the short-wave modulation recedes
+back toward the resting state.
 
@@
 
+Limitations of this study; no wind or dissipation. Only considering deep water
+gravity waves, but the results could be extended to gravity-capillary waves
+in arbitrary depth.
+Although it is impossible to isolate the hydrodynamic modulation effects
+from other modulation effects in the field and laboratory data, it is
+nevertheless useful to study this process in isolation from other processes
+in a theoretical and numerical framework.
+
+Discuss that we examined r = 10, 100 and find very similar results,
+consistent with \citet{longuet1987propagation}
+
+Results are very similar to prior work, but provide an easy to use analytical
+solution to calculate modulations without numerical iteration.
+
+Discuss the group velocity results of \citet{Willebrand1975energy}.
+What do we get for group velocity modulation?
+Should the short-wave modulations be considered for inclusion in spectral wave
+models?
+
+Implications for spectral wave models and wind input to short waves.
+
 \section{Conclusions}
@@
 from a linear wave theory perspective.
-Using the linearized wave crest and action conservation equations coupled with nonlinear solutions for the effective gravity, we derived steady analytical solutions for the modulation
-of short wave wavenumber, amplitude, steepness, intrinsic frequency, and phase
-speed.
+Using the linearized wave crest and action conservation equations coupled with
+nonlinear solutions for the effective gravity, we derived steady analytical
+solutions for the modulation of short wave wavenumber, amplitude, steepness,
+intrinsic frequency, and phase speed.
 The latter two have not been previously examined in the context of short-wave
@@
 \backsection[Data availability statement]{
-  The data and code to generate the figures and run the numerical simulations
-  are available at [TODO].
-  The numerical model is available as a standalone package at [TODO].
+  The code to run the numerical simulations and generate the figures in this
+  paper is available at \url{https://github.com/wavesgroup/wave-modulation-paper}.
+  The numerical model is available as a standalone package at
+  \url{https://github.com/wavesgroup/2wave}.
   %See JFM's \href{https://www.cambridge.org/core/journals/journal-of-fluid-mechanics/information/journal-policies/research-transparency}{research transparency policy} for more information

### 2024-09-24 paper/paper.tex
Discussion

--- 2024-09-20 paper/paper.tex
@@
 
+The solutions presented here, analytical and numerical alike, are valid only
+for certain scales of short and long waves.
+Specifically, the wave action conservation equation, introduced by
+\citet{bretherton1968wavetrains}, requires that the ambient velocity is slowly
+varying on the time scales of short waves, \textit{i.e.}
+$|\partial U / \partial x| \ll k |U|$
+and $|\partial U / \partial t| \ll \sigma |U|$.
+This implies $k_L \ll k$ and $|\sigma_L| \ll \sigma|$, respectively.
+Another requirement is for the short-waves to remain linear, which implies
+$a k \ll 1$.
+Furthermore, the short waves examined here do not include the effects of
+surface tension (gravity-capillary waves) or intermediate or shallow water.
+Within these limits, the crest-action conservation equations provide an
+elegant and relatively simple framework to study the hydrodynamic modulation
+of short waves by longer waves.
+This approach, of course, isolates this particular process from others
+dominant modulation and nonlinear transfer processes, such as wave growth
+by wind, wave dissipation, and aerodynamic sheltering by the longer waves.
+
+Although in this paper we only explored the wavenumbers 1 and 10 for the long
+and short waves, respectively ($k/k_L = 10$), the results are not sensitive
+to different values of $k_L$, provided that the amplitude $a_L$ is set such that
+the long-wave steepness $\varepsilon_L = a_L / k_L$ is the same.
+Numerical simulations with that compare $k/k_L = 10$ and $k/k_L = 100$ show
+very little sensitivity $k/k_L$ (not shown), consistent with the results of
+\citet{longuet1987propagation}.
+Specifically, the short-wave steepness modulation is 1\% larger at
+$\varepsilon_L = 0.27$ for $k/k_L = 10$ compared to $k/k_L = 100$, and 4\%
+larger at $\varepsilon_L = 0.4$.
+This suggests that as long as $k/k_L \gg 1$, the modulation amplitude is
+largely independent of the ratio $k/k_L$.
+
+Despite several studies that provided qualitatively similar steady solutions
+for the short-wave modulation by long waves
+\citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,henyey1988energy,zhang1990evolution},
+it was not obvious from the governing equations that the fully-nonlinear
+steady solutions should exist.
+\citet{peureux2021unsteady} examined this problem numerically and found that
+the solutions are stable only if the short-wave field is initialized from
+a prior steady solution of the form of that from \citet{longuet1960changes}.
+If the short-wave field is instead initialized from a uniform field, they
+found that the short-wave amplitude, and thus steepness as well, grow
+unsteadily.
+However, in this study we demonstrate that the unsteady steepening of short
+waves only occurs in the case of a sudden appearance of long waves, where
+the velocity fields induced by the long waves act in their full capacity
+on the short waves that have not yet had time to "lock in" to the long-wave
+crests.
+In such a hypothetical scenario, sudden forcing on short waves causes the
+modulation pattern to detach itself from the long-wave crests and to propagate
+at the short-wave group speed.
+When this occurs, the modulated short-waves are no longer in a stabilizing
+resonance with the surface convergence and divergence induced by the long-wave
+orbital velocity, and each long-wave passage further destabilizes the short-wave
+field, causing it to steepen indefinitely.
+
+Another important consideration of the hydrodynamic modulation process is
+whether it should somehow be accounted for in phase-averaged spectral wave
+models used for regional and global wave forecasting. 
+In the early days of spectral wave modeling, \citet{willebrand1975energy}
+examined the effect of a wave spectrum on the short-wave group velocity,
+and found the effect of up to several percent modulation.
+For a practical implementation in a spectral wave model, this would have to
+incur an additional loop over all frequencies for each short-wave frequency.
+It was thus not deemed practical to include this effect in wave models at the time.
+However, with the recent advances in machine learning and phase-resolving modeling
+of short waves, the effect of wave-modulation deserves re-examination.
+
+TODO Implications for spectral wave models and wind input to short waves.
+
 \section{Conclusions}

### 2024-09-24 paper/paper.tex
Edit in intro
Incorporate diagram

@@
 In this paper, we revisit the linear wave theory and derive higher-order
-solutions for the change of short-wave wavenumber, action density, and
+steady solutions for the change of short-wave wavenumber, action density, and
 gravitational acceleration due to the presence of longer waves.
@@
 solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
-We also examine the result of \citet{peureux2021unsteady} who found through
+We further examine the result of \citet{peureux2021unsteady} who found through
 numerical simulations that the short-wave modulation grows unsteadily with
-each long-wave phase.
-We show that this unsteady growth only occurs for homogeneous initial
-conditions and does not occur for more realistic long-wave groups.
+each long-wave passage.
+We show that this unsteady growth only occurs only for homogeneous initial
+conditions as a special case and does not generally occur, for example in more
+realistic long-wave groups.
 The combined analytical and numerical framework presented here allows
@@
 nonlinear numerical solutions in long-wave steepness up to $\approx 0.2$.
+As a corollary, except for a subset of initial conditions, the solutions to
+fully nonlinear crest-action conservation equations are mostly steady in the
+reference frame of the long waves.
 \end{abstract}
@@
 Short ocean surface waves are hydrodynamically modulated by longer swell.
+As long waves propagate through a field of shorter waves, their orbital
+velocities cause near-surface convergence and divergence on the front and rear
+faces of the long waves, respectively.
 Following \citet{unna1941white,unna1942waves,unna1947sea},
@@
 wavenumber $k_L$ and amplitude $a_L$.
-This hydrodynamic modulation process was instrumental for the development and
+Hydrodynamic modulation is a key process for the development and
 interpretation of remote sensing of ocean surface waves and currents
 \citep{keller1975microwave,hara1994hydrodynamic}, and is distinguished
-from the aerodynamic modulation process \citep{donelan1987effect,belcher1999wave,chen2000effects}.
+from aerodynamic modulation \citep{donelan1987effect,belcher1999wave,chen2000effects}.
 \citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited the
@@
 for a simpler and more intuitive derivation.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_hydrodynamic_modulation_diagram.pdf}
+\caption{
+  Diagram of the hydrodynamic modulation process.
+}
+\label{fig:hydrodynamic_modulation_diagram}
+\end{figure}
 

### 2024-09-24 paper/scripts/hydrodynamic_modulation_diagram.py

@@ -1,0 +1,147 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Set up the figure
+fig, ax = plt.subplots(figsize=(12, 4))  # Increased width to accommodate two periods
+
+# Generate x values
+x = np.linspace(0, 3.5 * np.pi, 2000)  # Extended to 4π, increased number of points
+
+# Generate long wave
+long_wave = np.sin(x)
+
+# Generate short wave (shorter and more subtle)
+short_wave_base = 0.05 * np.sin(20 * x)
+
+# Modulate short wave amplitude based on long wave phase
+modulation = 0.5 * (1 + long_wave)  # This will be 1 at peaks and 0 at troughs
+short_wave = short_wave_base * modulation
+
+# Combine waves
+combined_wave = long_wave + short_wave
+
+# Plot the combined wave
+ax.plot(x, combined_wave, "k-", lw=3, zorder=1)
+
+# Add surface velocity arrows
+arrow_positions = np.arange(np.pi, 4 * np.pi, np.pi)
+for pos in arrow_positions:
+    # Calculate wave slope at this position
+    slope = np.cos(pos)  # Derivative of sin(x)
+    angle = np.arctan(slope)
+
+    # Calculate arrow components
+    arrow_length = 0.6
+    dx = arrow_length * np.cos(angle)
+    dy = arrow_length * np.sin(angle)
+
+    # Position arrow slightly below the wave
+    y_offset = -0.25
+
+    ax.arrow(
+        pos - 1.6 * dx,
+        np.sin(pos - 1.6 * dx) + y_offset,
+        dx,
+        dy * 0.9,
+        head_width=0.2,
+        head_length=0.2,
+        fc="tab:blue",
+        ec="tab:blue",
+        width=0.1,
+    )
+    ax.arrow(
+        pos + 1.6 * dx,
+        np.sin(pos + 1.6 * dx) + y_offset,
+        -dx,
+        -dy,
+        head_width=0.2,
+        head_length=0.2,
+        fc="tab:blue",
+        ec="tab:blue",
+        width=0.1,
+    )
+
+    ax.arrow(
+        pos,
+        np.sin(pos) - y_offset,
+        -1.5 * dx,
+        -1.5 * dy,
+        head_width=0.12,
+        head_length=0.2,
[…]

### 2024-09-26 paper/scripts/hydrodynamic_modulation_diagram.py
Reverse the Cg arrows in the diagram

@@ -62,10 +62,10 @@ for pos in arrow_positions:
     )
 
     ax.arrow(
-        pos,
-        np.sin(pos) - y_offset,
-        -1.5 * dx,
-        -1.5 * dy,
+        pos - 0.25 * np.pi,
+        np.sin(pos - 0.25 * np.pi) - y_offset,
+        1.5 * dx,
+        1.5 * dy,
         head_width=0.12,
         head_length=0.2,
         fc="k",

### 2024-09-26 paper/paper.tex
Near final edits

@@
 wavenumber $k_L$ and amplitude $a_L$.
+This process is illustrated in Fig. \ref{fig:hydrodynamic_modulation_diagram}.
 Hydrodynamic modulation is a key process for the development and
@@
 \citet{henyey1988energy} derived an analytical solution for the modulation of
-short waves by long waves using Hamiltonian mechanics and reported the modulation
-of similar magnitude to that of \citet{longuet1987propagation}.
+short waves using Hamiltonian mechanics and reported similar modulation magnitudes
+to that of \citet{longuet1987propagation}.
 \citet{zhang1990evolution} considered weakly nonlinear short waves
@@
 \citet{longuet1987propagation}.
-The variety of analytical and numerical frameworks (Eulerian, Hamiltonian,
-Schr\" odinger) that relatively closely reproduce steady solutions for short-wave
-modulation suggests that the solutions are robust.
-Alternative analytical and numerical approaches to the modulation problem remain
-of interest, especially if higher degrees of nonlinearity can be considered
-using simpler approaches.
-
-More recently, \citet{peureux2021unsteady} asked whether the steady solutions
-of short-wave modulation are realistic by simulating the full wave conservation
+The variety of analytical and numerical frameworks (Eulerian crest and action
+conservation, Hamiltonian, Schr\" odinger) that relatively closely reproduce
+steady solutions for short-wave modulation suggests that the solutions are robust.
+Nevertheless, alternative analytical and numerical approaches to the modulation
+problem remain of interest, especially if higher degrees of nonlinearity can be
+considered using simpler approaches.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_hydrodynamic_modulation_diagram.pdf}
+\caption{
+  A diagram of short waves riding on longer waves that propagate with their
+  phase-speed $C_{pL}$ from left to right.
+  The short waves propagate on the long-wave surface with their group velocity
+  $C_g$.
+  Long waves induce orbital velocities $U$ that cause near-surface convergence
+  and divergence on their front and rear faces, respectively.
+  They also induce downward and upward centripetal accelerations
+  ($\partial W/\partial t$) in the crests and troughs, respectively, that
+  modulate the effective gravity at the surface.
+  As a consequence of surface convergence and divergence and the effective-gravity
+  modulation, the short waves become shorter and higher preferentially
+  on the long-wave crests.
+  The diagram is not to scale.
+}
+\label{fig:hydrodynamic_modulation_diagram}
+\end{figure}
+
+All modulation solutions mentioned thus far are steady in the reference frame
+of the long wave---short waves become shorter and higher (and thus, steeper)
+preferentially on the long-wave crests.
+Conversely, they elongate and flatten preferentially in the long-wave troughs.
+Recently, \citet{peureux2021unsteady} asked whether the steady solutions
+of short-wave modulation are appropriate by simulating the full wave conservation
 equations.
@@
 steepness.
-However, the unsteadiness of their solutions occurs in a specific scenario of
-a uniform short-wave field that suddenly appears in a uniform long-wave train,
-and the numerical simulations stabilize if the short-wave field is initialized
-from existing steady solutions.
+However, the unsteadiness of their solutions occurs in a specific scenario in
+which a long-wave train suddenly appears to modulate a uniform field of short
+waves.
+The numerical simulations stabilize if the short-wave field is initialized
+from existing steady solutions, like those by \citet{longuet1960changes}.
 Nevertheless, their results do put into question whether the short-wave
@@
 or not.
-
-In this paper, we revisit this problem from linear wave theory and
-derive alternative non-linear solutions for the modulation of short waves,
-evaluated at the long-wave surface.
-The analytical solutions yield modulations that are stronger than the original solution
-by \citet{longuet1960changes}, but similar to the nonlinear numerical solutions
-by \citet{longuet1987propagation} and \citet{zhang1990evolution} in long-wave
-steepness up to $\approx 0.2$.
-We also perform numerical simulations of the full wave crest and action
-conservation laws and evaluate the effect of long-wave groups on the short waves.
+Looking at the conservation equations alone, it is not obvious that it is.
+
+In this paper, we revisit this problem from linear wave theory but consider
+the long-wave slope to be significant enough to require evaluating the
+short-waves at the long-wave surface rather than at the mean water level.
+This allows us to derive alternative, steady, non-linear solutions for the
+modulation of short waves.
+The analytical solutions yield modulations that are stronger than the original
+solution by \citet{longuet1960changes}, but similar to the nonlinear numerical
+solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution} in
+long-wave steepness up to $\approx 0.2$.
+To validate the steadiness of the analytical solutions, we perform numerical
+simulations of the full wave crest and action conservation laws, and compare
+the results to the analytical solutions presented here, as well as the prior
+numerical results.
 The numerical simulations are used to determine the validity of the approximate
@@
 \ref{table:modulation_literature_summary}.
+We first present the governing equations in \S\ref{section:governing_equations},
+and then the analytical solutions for the modulation of short waves by long waves
+in \S\ref{section:analytical_solutions}.
+Then, we describe the numerical simulations and their results in
+\S\ref{section:numerical_solutions}.
+Finally, we discuss the results and their implications in
+\S\ref{section:discussion} and conclude the paper in \S\ref{section:conclusions}.
 
@@
 \hline
-L-HS 1960   & Velocity potential  & No  & No  & No  & Analytical \\
+\citet{longuet1960changes} & Velocity potential  & No  & No  & No  & Analytical \\
 \hline
-P 1981      & Crest and action    & Yes & Yes & No  & Numerical \\
+\citet{phillips1981dispersion} & Crest and action & Yes & Yes & No  & Numerical \\
                                & conservation     &     &     &     &           \\
 \hline
-L-H 1987    & Velocity potential  & Yes & Yes & No  & Numerical \\
+\citet{longuet1987propagation} & Velocity potential & Yes & Yes & No  & Numerical \\
 \hline
-HCDSW 1988  & Hamiltonian         & Yes & Yes & No  & Analytical \\
+\citet{henyey1988energy} & Hamiltonian & Yes & Yes & No  & Analytical \\
 \hline
-ZM 1990     & Schr\" odinger      & Yes & Yes & Yes & Numerical \\
+\citet{zhang1990evolution} & Schr\" odinger & Yes & Yes & Yes & Numerical \\
 \hline
-PAG 2021    & Crest and action    & Yes & No  & No  & Numerical \\
+\citet{peureux2021unsteady} & Crest and action & Yes & No  & No  & Numerical \\
                             & conservation     &     &     &     &           \\
@@
 \caption{
-  Summary of prior
-  \citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,henyey1988energy,zhang1990evolution,peureux2021unsteady}
-  and present (this paper) approaches to calculating the modulation of short
-  waves by long waves.
+  Summary of prior and present (this paper) approaches to calculating the
+  modulation of short waves by long waves.
 }
@@
 
-This paper first presents the governing equations and then the analytical
-solutions for the modulation of short waves by long waves and compares them to
-the original solution by \citet{longuet1960changes}.
-Then, we compare the analytical solutions to numerical solutions of the full
-wave conservation equations in section \ref{section:numerical_solutions}.
-In section \ref{section:scale_analysis} we perform a scale analysis of the
-analytical solutions to quantify the relative importance of the terms in the
-full equation set.
-Finally, we summarize the results and discuss their implications in section
-\ref{section:conclusions}.
-
 \section{Governing equations}
@@
 
-We consider inviscid, irrotational, and incompressible flow.
-In these conditions the linear (sine-like), deep-water, surface gravity waves
-obey the dispersion relation:
+The flow is inviscid, irrotational, and incompressible, and without sources or
+sinks.
+In these conditions, linear, deep-water, surface gravity waves obey the
+dispersion relation:
 
@@
 \end{equation}
-where $\omega$ is the angular frequency, $g$ is the gravitational acceleration,
-$k$ is the wavenumber, and $U$ is the mean advective current in the direction
-of the wave propagation.
-A current in the direction of the waves increases its absolute frequency,
-while the wavenumber remains unchanged.
+where $\omega$ is the angular frequency in a fixed reference frame, $g$ is the
+gravitational acceleration, $k$ is the wavenumber, and $U$ is the mean advective
+current in the direction of the wave propagation.
+A current in the direction of the waves increases their absolute frequency
+without changing their wavenumber.
 An important limitation here is that $U$ must be slowly varying on the scales of
 the wave period \citep{bretherton1968wavetrains}.
-In the context of short waves riding on long waves, the mean advective current
+From the perspective of short waves riding on long waves, the advective current
 is the horizontal near-surface orbital velocity of the long wave.
+Strictly, a velocity profile must be considered when determining the
+advective velocity for a wave of any given wavenumber \citep{stewart1974hf},
+however for brevity here we will assume a surface velocity to be the advective
+one for short waves.
 The evolution of the short-wave wavenumber is described by the conservation of
@@
 where $t$ and $x$ are time and space, respectively.
-This relation states that the wavenumber must change when the frequency varies
-in space.
-
-Equations (\ref{eq:dispersion}) and (\ref{eq:wave_crests}) only describe the
-evolution of the short-wave wavenumber.
-To describe the evolution of the wave amplitude, we use the conservation of wave
-action in absence of sources (e.g. due to wind) and sinks (e.g. due to whitecapping):
+This relation states that the wavenumber must change locally when the frequency
+varies in space.
+
+(\ref{eq:dispersion}) and (\ref{eq:wave_crests}) only describe the evolution of
+the short-wave wavenumber.
+To describe the evolution of the wave amplitude, use the wave action balance
+\citep{bretherton1968wavetrains} in absence of sources (e.g. due to wind) and
+sinks (e.g. due to whitecapping):
 
@@
 \end{equation}
-where $N$ is the action density, and $C_g = \partial \omega / \partial k$ is the
-group speed.
-In deep water, the group speed is half the phase speed, so $C_g = 1/2\sqrt{g/k}$.
+where $N$ is the action density, and $C_g = 1/2\sqrt{g/k}$ is the group speed
+in deep water.
 Although the absence of wave growth and dissipation is not generally realistic
-in the open ocean, it is a helpful approximation that allows us to isolate the
-hydrodynamic modulation effects solely due to the motion of the long waves.
+in the open ocean, it is a helpful approximation here to isolate the effects
+of hydrodynamic modulation solely due to the motion of the long waves.
 
 Short waves that ride on the surface of the longer waves move in an accelerated
-reference frame and thus experience effective gravitational acceleration that
-changes in space and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
-From Eqs. (\ref{eq:dispersion}) and (\ref{eq:wave_action}) we obtain:
+reference frame (due to the centripetal acceleration at the long-wave surface)
+and thus experience effective gravitational acceleration that varies in space
+and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
+Combine (\ref{eq:dispersion}) and (\ref{eq:wave_crests}) to get:
 
@@
 term vanishes.
-Otherwise, this term is necessary to conserve the wave crests.
-
-Similar to Eq. \ref{eq:wavenumber}, we expand Eq. \ref{eq:wave_action} to obtain:
+It is otherwise necessary to conserve the wave crests.
+
+For the wave action, expand (\ref{eq:wave_action}) to get:
 
@@
 \end{equation}
-This equation is similar to Eq. \ref{eq:wavenumber}, except for the last term
-which here represents the inhomogeneity of the group speed.
-As \citet{peureux2021unsteady} explained, this term is responsible for unstable
-growth of wave action in infinite long-wave trains.
-Eqs. \ref{eq:wavenumber} and \ref{eq:wave_action2} are the governing equations
-that we approximate and solve analytically in section \ref{section:analytical_solutions},
-and solve numerically in their full form in section \ref{section:numerical_solutions}.
+This equation is similar to (\ref{eq:wavenumber}), except for the last term
+which represents the inhomogeneity of the group speed.
+As \citet{peureux2021unsteady} explained, the presence of this term causes
+unstable growth of wave action in infinite long-wave trains.
+(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are the governing equations
+that we approximate and solve analytically in \S\ref{section:analytical_solutions},
+and numerically in their full form in \S\ref{section:numerical_solutions}.
 
@@
 defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
-where $\psi = k_L x - \omega_L t$ is the long-wave phase and $\omega_L$ its
+where $\psi = k_L x - \sigma_L t$ is the long-wave phase and $\sigma_L$ its
 angular frequency.
@@
 
-Eqs. \ref{eq:dispersion}-\ref{eq:w_L} is the full system of equations that we
-use in this paper.
-Note that here we derive the solutions from a different starting set of
-equations than that of \citet{longuet1960changes}.
-Rather than starting from the conservation of crests and wave action, as we do
-here, they derived their solution from the linear superposition of the velocity
-potential of two waves.
-It appears that starting from the conservation of crests and wave action allows
-for a simpler and more intuitive derivation.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figures/fig_hydrodynamic_modulation_diagram.pdf}
-\caption{
-  Diagram of the hydrodynamic modulation process.
-}
-\label{fig:hydrodynamic_modulation_diagram}
-\end{figure}
+(\ref{eq:dispersion})-(\ref{eq:w_L}) are the governing equations used in this paper.
+Note that rather than deriving the modulation solution from the velocity
+potential of two waves, as \citet{longuet1960changes} did, starting from the
+crest-action conservation balances allows for a simpler derivation.
 
@@
 The wavenumber is derived by linearizing the conservation of wave crests
-(Eq. \ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
+(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
 The effective gravity is derived by projecting the long-wave induced surface
-accelerations onto the axis that is normal to the long-wave surface.
-(Eq. \ref{eq:wave_action2}) and using the expression for the long-wave orbital
-velocities.
-The amplitude is derived by first evaluation the short-wave action from the
-action conservation equation (Eq. \ref{eq:wave_action}) and then applying the
-modulated effective gravity to obtain the amplitude.
+accelerations onto the local vertical axis (that is, the axis normal to the
+long-wave surface), akin to \citet{zhang1990evolution}.
+The amplitude is derived by first evaluating the short-wave action from the
+action conservation (\ref{eq:wave_action}) and then applying the modulated
+effective gravity to obtain the amplitude.
 From the modulated wavenumber, amplitude, and effective gravity, the steepness,
@@
 \end{equation}
-Although \citet{peureux2021unsteady} pointed out that the solution by L-HS
-requires assuming homogeneous group speed of the short-wave field, in fact,
-it also requires assuming no horizontal propagation and advection of the short
-waves.
-Combining Eqs. \ref{eq:U_L} and \ref{eq:wavenumber2}, and integrating it in time
-yields:
+Although \citet{peureux2021unsteady} pointed out that the solution by
+\citet{longuet1960changes} requires assuming homogeneous group speed of the
+short-wave field, in fact, it also requires assuming no horizontal propagation
+and advection of the short waves.
+Combine (\ref{eq:U_L})-(\ref{eq:wavenumber2}), and integrate in time to get:
 
@@
 
-Notice that the Taylor expansion of Eq. \ref{eq:k_short_exact} to the first order
+Notice that the Taylor expansion of (\ref{eq:k_short_exact}) to the first order
 recovers the original solution by \citet{longuet1960changes}:
@@
 \end{equation}
-Coming from the conservation of wave crests, as opposed to the velocity potential
-superposition approach, the L-HS solution requires two approximations in their
-solution:
+Coming from the conservation of crests, as opposed to the velocity potential
+superposition, the solution by \citet{longuet1960changes} requires two
+approximations:
 First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
-second, further truncating the Taylor expansion series beyond the first order term.
-These approximations cause an underestimate of the short-wave modulation
-magnitude (Fig. \ref{fig:analytical_solutions_ak0.1}a).
-At the wave crest, Eq. \ref{eq:k_short_exact} yields the wavenumber modulation
-that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of \citet{longuet1960changes}
-for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
+second, further truncating the Taylor expansion series beyond the first-order
+term.
+These approximations underestimate the short-wave modulation magnitude
+(Fig. \ref{fig:analytical_solutions_ak0.1}a).
+At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
+that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of
+\citet{longuet1960changes} for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
 Although not immediately obvious from the expressions, the modulated wavenumber
-in Eq. \ref{eq:k_short_lhs} is conservative across the long-wave phase, whereas
-the higher-order solution in Eq. \ref{eq:k_short_exact} is not.
-This is due to the fact that combining the orbital velocities evaluated at
-$z=\eta$ and the linearized conservation of wave crests in Eq. \ref{eq:wavenumber}
-results in excess short-wave wavenumber being created.
-The solution of the full Eq. \ref{eq:wavenumber} of course must be conservative
-across the long-wave phase, as we will see later from the numerical simulations.
+(\ref{eq:k_short_lhs}) is conserved across the long-wave phase, whereas
+the higher-order solution (\ref{eq:k_short_exact}) is not.
+This is because combining the orbital velocities evaluated at $z=\eta$ and the
+linearized conservation of wave crests in (\ref{eq:wavenumber}) results in
+excess short-wave wavenumber being created.
+The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
+phase, as we will see later from the numerical simulations.
 
@@
 
-Long waves induce orbital accelerations, so the short waves that ride on their
-surface move in an accelerated reference frame and experience effective
-gravitational acceleration that is lower at the crests and higher in the troughs
+Long waves induce orbital (centripetal) accelerations on their surface, so the
+short waves ride in an accelerated reference frame and experience effective
+gravity that is lower at the crests and higher in the troughs
 \citep{longuet1986eulerian,longuet1987propagation}.
-Further, centripetal force due to the curvature of the long wave induces small
-but non-negligible effects on the short-wave effective gravity
+Further, the projection of the long-wave induced horizontal acceleration
+contributes to small but non-negligible effects on the short-wave effective
+gravity in the curvilinear coordinate system
 \citep{phillips1981dispersion,zhang1990evolution}.
-These effects are completely described by projecting the gravitational
-acceleration from the coordinate that is perpendicular to the mean water surface
-($z=0$) to that which is perpendicular to the long-wave surface ($z=\eta$),
-and likewise for the long-wave orbital accelerations:
+These two effects are completely described by projecting the gravitational
+acceleration vector from the coordinate that is perpendicular to the mean water
+surface ($z=0$) to that which is perpendicular to the long-wave surface
+($z=\eta_L$), and likewise for the long-wave orbital accelerations:
 
@@
 where $\alpha$ is the local slope of the long-wave surface.
-For a linear long wave, the effective gravity is:
+For short waves on a linear long wave, the gravity modulation is:
 
@@
 \end{equation}
-
 Curvilinear effects on effective gravity can be removed altogether by setting
 $\alpha = 0$.
-In that case, Eq. \ref{eq:gravity_modulation_general} simplifies to:
+In that case, (\ref{eq:gravity_modulation_general}) simplifies to:
 
@@
 
-The general form of the gravity modulation in Eq. \ref{eq:gravity_modulation_general}
-is convenient because it allows us to easily attribute the effects of the different
-processes on the gravity modulation.
-It also makes it straightforward to evaluate the gravity modulation for a different
-form of the long-wave elevation, by evaluating the orbital accelerations and the
-local elevation slope at the appropriate level.
+The general form of the gravity modulation (\ref{eq:gravity_modulation_general})
+is convenient because it allows us to easily attribute the modulation to
+different processes.
+It also makes it straightforward to evaluate the gravity modulation for a
+different long-wave form, by evaluating the orbital accelerations and the local
+slope for that wave.
 For a third-order Stokes wave, for example, the elevation is:
@@
 \end{equation}
-Without considering the curvilinear effects in Eq. \ref{eq:gravity_modulation_general},,
-the gravitational acceleration at the surface of a Stokes wave is, to the
-fourth order:
+Without considering the curvilinear effects in (\ref{eq:gravity_modulation_general}),
+the gravitational acceleration at the surface of a Stokes wave is, to the third
+order:
 
@@
 The equivalent expression that includes the curvilinear effects can be derived
-by evaluating orbital accelerations and local slope $\alpha$ at $z = \eta_{St}$,
-and inserting them into Eq. \ref{eq:gravity_modulation_general}:
+by evaluating the velocities and the local slope $\alpha$ at $z = \eta_{St}$,
+and inserting them into (\ref{eq:gravity_modulation_general}):
 
@@
 gravity on the long-wave crests and a higher effective gravity in the troughs.
-Evaluating the orbital accelerations at $z=\eta$ causes a gravity reduction
+Evaluating the orbital accelerations at $z=\eta_L$ causes a gravity reduction
 around the crests ($\approx6\%$) and the troughs ($\approx 3\%$), and an
@@
 The long-wave elevation curvature has a second-order effect in $\varepsilon_L$.
-Naturally, there is no curvature exactly at the crests and troughs, and its
+Naturally, there is no curvature exactly at the crests and the troughs, and its
 effects are largest at the front and rear faces of the long wave, where the
-covariance between the local slope and the horizontal accelerations are greatest.
+covariance between the local slope and the horizontal accelerations are largest.
 
@@
 As \citet{longuet1986eulerian} pointed out, the effective gravity in the
-Lagrangian frame is smoother and less amplified because the moving particles
-spend a longer time riding near the crests and less time returning through the
-troughs.
+Lagrangian frame (\textit{i.e.}, following the surface particles) is smoother
+and less amplified because the moving particles spend a longer time riding near
+the crests and less time returning through the troughs.
 However, when we solve the short-wave crest and action conservation equations
@@
 The modulation of short-wave amplitude can be derived in a similar way as we did
-for the wavenumber, except that here we linearize the wave action balance and
-integrate wave action in time:
+for the wavenumber, except that here we linearize the wave action balance
+(\ref{eq:wave_action2}) and integrate it in time:
 
@@
 \end{equation}
-To the first order in $\varepsilon_L$, and using Eqs. \ref{eq:k_short_exact}
-and \ref{eq:wave_action_modulation}, the above simplifies to:
+To the first order in $\varepsilon_L$, and using (\ref{eq:k_short_exact})
+and (\ref{eq:wave_action_modulation}), the above simplifies to:
 
@@
 crests and increased in the troughs.
-We will discuss the implications of this effect later in the context of wind
-input into short waves.
+We discuss the implications of this effect later in \S\ref{section:discussion}
+in the context of wind input to short waves.
 
@@
 
-To quantify the contribution of the nonlinear terms in Eqs. \ref{eq:wavenumber}
-and \ref{eq:wave_action2}, and to evaluate the steadiness assumption of the
+\subsection{Model description}
+
+To quantify the contribution of the nonlinear terms in (\ref{eq:wavenumber})
+and (\ref{eq:wave_action2}), and to evaluate the steadiness assumption of the
 analytical solutions, we now proceed to numerically integrate the full set of
 crest and action conservation equations.
-Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
+(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are discretized using
 second-order central finite difference in space and integrated in time using the
-forth-order Runge-Kutta method \citep{butcher1996runge}.
+fourth-order Runge-Kutta method \citep{butcher1996runge}.
 The space is divided into 100 grid points.
@@
 induced by long waves on short-wave groups
-\citep{stokes1846,van2018stokes,monismith2020stokes}.
+\citep{stokes1847,van2018stokes,monismith2020stokes}.
 Their gravity modulation also does not consider the nonlinear or the curvature
@@
 Another difference is the choice of the numerical scheme for spatial differences,
-which in their case is the MUSCL4 scheme by \citet{kurganov2000new}.
+which in their case is the more sophisticated MUSCL4 scheme \citep{kurganov2000new}.
 Although the second-order central finite difference is not appropriate for many
@@
 rather than that moving with the long-wave phase speed.
-Either approach produces qualitatively equivalent results, however a fixed
+Either approach produces equivalent modulation results, however a fixed
 reference frame may allow for a more intuitive interpretation of the results.
@@
 whereas in the figures that follow, the long wave is moving rightward with its
-phase speed $C_{pL}$ and the short waves (that is, their energy) are moving
+phase speed $C_{pL}$ and the short waves (that is, their action) are moving
 rightward with the speed of $C_g + U$ (neglecting group speed inhomogeneity).
@@
 whether it should somehow be accounted for in phase-averaged spectral wave
-models used for regional and global wave forecasting. 
+models \citep{group1988wam,tolman1991third,donelan2012modeling} used for global
+and regional wave forecasting.
 In the early days of spectral wave modeling, \citet{willebrand1975energy}

### 2024-09-27 paper/scripts/plot_numerical_vs_analytical_solutions_with_phase.py
Fix indexing issue with new time step scheme

@@ -21,7 +21,9 @@ def calculate_modulation(ak_L: float):
     m.run(ramp_type="groups")
     ds = m.to_xarray()
 
-    ind = 1002  # Exactly 5 periods
+    T = 2 * np.pi / np.sqrt(m.grav0 * k_long)
+
+    ind = np.argmin((m.time - 5 * T) ** 2)
 
     phase = ds.space * k_long
 

### 2024-09-28 paper/paper.tex
Tiny edits

@@
 Although the absence of wave growth and dissipation is not generally realistic
-in the open ocean, it is a helpful approximation here to isolate the effects
+in the open ocean, it is a useful approximation here to isolate the effects
 of hydrodynamic modulation solely due to the motion of the long waves.
@@
   paper is available at \url{https://github.com/wavesgroup/wave-modulation-paper}.
-  The numerical model is available as a standalone package at
+  The numerical model is available as a standalone Python package at
   \url{https://github.com/wavesgroup/2wave}.

### 2024-09-30 paper/paper.tex
Near-final edits to the text

@@
 
-\title{Revisiting the hydrodynamic modulation of short waves by longer waves}
+\title{Revisiting the hydrodynamic modulation of short surface waves by longer waves}
 
@@
 and numerically in their full form in \S\ref{section:numerical_solutions}.
+It is worth noting that this system of equations is semi-coupled, meaning that
+it is possible to solve (\ref{eq:wavenumber}) on its own, but solving
+(\ref{eq:wave_action2}) requires also the solution of (\ref{eq:wavenumber}).
+In other words, the wave kinematics (the wavenumber distribution) are not
+concerned with the wave action.
+In contrast, the wave action is governed, among others, by the
+group velocity and thus depends on the wavenumber.
 
@@
 $\varepsilon_L = 0.1$, $0.2$, and $0.4$ (Fig. \ref{fig:numerical_solutions}).
-The long-waves are initialized from rest ($a_L = 0$) and are ramped up
-to their target steepness after 5 long-wave periods, thus mimicking the
-passage of a long-wave group
-(we explain in more detail the importance of ramping up the long-wave forcing
-in the next subsection).
+The long-waves are initialized from rest ($a_L = 0$) and gradually ramped up
+to their target steepness over 5 long-wave periods, thus emulating the passage
+of a long-wave group
+(the importance of which we discuss in more detail in the next subsection).
 At $\varepsilon = 0.1$, the numerical solutions for the wavenumber, amplitude,
@@
 (Fig. \ref{fig:numerical_solutions}a,d,g).
-This suggests that at low long-wave steepness that steady analytical solution
+This suggests that at low long-wave steepness that the steady analytical solution
 is a reasonable proxy for the fully-nonlinear crest-action conservation
@@
 These differences suggest that at high steepness the inhomogeneity of
-gravitational acceleration and group speed become important.
+gravitational acceleration and group speed, otherwise ignored in the steady
+solutions, become important.
 
@@
 \caption{
-  Comparison of numerical solutions of wavenumber, amplitude, and steepness modulation
-  with their analytical solutions, for $\varepsilon_L$ = 0.1, 0.2, and 0.4.
+  Comparison of numerical solutions (orange) of wavenumber (top), amplitude
+  (middle), and steepness (bottom) modulation with their analytical solutions
+  (blue), for $\varepsilon_L$ = 0.1, 0.2, and 0.4.
 }
@@
 steepness $\varepsilon_L$.
-Overall, the numerical solutions, based on the linear and third-order Stokes long wave
-alike, begin to diverge from the analytical solution at $\varepsilon_L \approx 0.2$.
+Overall, the numerical solutions, based on the linear and the third-order Stokes
+long waves alike, begin to diverge from the analytical solution at
+$\varepsilon_L \approx 0.2$.
 The numerical solution using the Stokes long wave is quantitatively very similar
@@
 relative to the results of \citet{longuet1987propagation}
-is likely due to his long wave velocities being evaluated to a higher order
-than they are here.
+is likely due to his long wave velocities being evaluated to a higher order in
+$\varepsilon_L$ than they are here.
 Nevertheless, the overall steepness modulation between the two numerical
@@
 short-waves of even low steepness to break \citep{banner1993wave}.
+Where the two modulation solutions differ (at $\varepsilon_L > 0.2$),
+either solution would lead to short-wave breaking after one long-wave period.
+This suggests that in the cases where the short-waves would survive the
+modulation and persist over multiple long-wave periods, the two solutions are
+effectively equivalent.
 
@@
 The steepening is driven by the progressive increase in the action of short
-waves (as opposed to their wavenumber), and is centered on the group of short
-waves that begins its journey in the trough of the long wave.
+waves (as opposed to their wavenumber), and is centered on the short-wave group
+that begins its journey in the trough of the long wave.
 This is because the unmodulated short waves that begin their journey in the
@@
 have $k = 10$.
-After 10 long-wave periods, the short-wave action modulation $\widetilde{N}/N$
-is approximately doubled, consistent with the unsteady growth reported by
-\citet{peureux2021unsteady}.
+After 10 long-wave periods, the short-wave action is approximately doubled,
+consistent with \citet{peureux2021unsteady}.
 Another important property of this solution is that the short-wave group that
-experiences peak amplification (dampening) are not locked in to the long-wave
-crests (troughs), as they are in the solution by \citet{longuet1960changes}.
+experiences peak amplification (dampening) are not locked-in to the long-wave
+crests (troughs), as they are in the prior steady solutions
+\citep{longuet1960changes,longuet1987propagation,zhang1990evolution}.
 
@@
 Why does this unsteady growth of short-wave action occur?
-As \citet{peureux2021unsteady} explain, "the additional change in advection
-velocity due to the modulation of the short wave group speed gives an additional
-amplification."
-That is, it is the inhomogeneity term $N \dfrac{\partial C_g}{\partial x}$ that
-is responsible for the instability.
+\citet{peureux2021unsteady} explain that the change in advection velocity due to
+the modulation of the short-wave group speed yields an additional amplification
+of the short-wave action.
+That is, the inhomogeneity term $N \dfrac{\partial C_g}{\partial x}$ is
+responsible for the instability.
 However, the instability only occurs if the short waves are initialized as a
@@
 and lower in the troughs, akin to the prior steady solutions, the instability
-does not occur.
+vanishes and the short-wave modulation pattern locks-in to the long-wave
+crests.
 
@@
 However, such a scenario is not realistic in the ocean where waves tend to
-organize into groups due to their dispersive nature.
-To test whether it is the sudden onset of long-wave forcing that causes the
-unstable growth, we repeat the numerical simulation of Fig.
+organize into groups due to dispersion.
+To test whether the sudden onset of long-wave forcing causes the unstable growth,
+we repeat the numerical simulation of Fig.
 \ref{fig:modulation_3panel_infinite}, except for a linear ramp applied to $a_L$
-(and consequently, $U$ and $W$ as well) for the first five periods:
+(and consequently, the other long-wave properties including the orbital
+velocities and the local steepness) for the first five periods:
 $a_L(t) = min(a_L, a_L t / \left(5 T_L\right))$, where $T_L$ is the long-wave period.
@@
 The key distinction that arises with the introduction of the ramp is that the
-modulation pattern of short waves is now locked in with the long-wave crests,
+modulation pattern of short waves is now locked-in to the long-wave crests,
 and their indefinite steepening vanishes.
@@
 steepness modulation as function of time over 30 long-wave periods, for the
-infinite long-wave train case and that when applying the linear ramp to the
-long-wave forcing.
+infinite long-wave train case and the case of applying a linear ramp.
 Comparing the wave action propagation ($C_g \frac{\partial N}{\partial x}$) and
@@
 (Fig. \ref{fig:inhomogeneity_tendencies}).
-Although \citet{peureux2021unsteady} correctly pointed out that the unstable
+\citet{peureux2021unsteady} correctly pointed out that the unstable
 growth of short-wave steepness occurs due to the resonance of the inhomogeneity
-tendency with the short-wave action perturbation, in case of a gentle ramp
-up of the long-wave forcing, the propagation and inhomogeneity tendencies
-mostly balance each other out, thus allowing the short-wave steepening to
-lock in to the long-wave crests.
+tendency with the short-wave action perturbation.
+However, in the case of a gentle ramp-up of the long-wave forcing, the
+propagation and inhomogeneity tendencies mostly balance each other and allow
+the short-wave steepening to lock-in to the long-wave crests.
 
@@
 The wave group is simulated as a non-dispersive train of long waves whose
-amplitude scales with $\cos{2 \pi t / (N T_L)}$, where $N$ is the number of
+amplitude scales with $\sin[\pi t / (n T_L)]$, where $n$ is the number of
 long waves in the group.
@@
 $t = N/2 T_L$, after which it gradually decreases toward zero.
-In this simulation, $N = 10$ and $\varepsilon = 0.1$ as in the previous two
+In this simulation, $n = 10$ and $\varepsilon_L = 0.1$ as in the previous two
 simulations.
@@
 Furthermore, the short waves examined here do not include the effects of
-surface tension (gravity-capillary waves) or intermediate or shallow water.
+surface tension (gravity-capillary waves) or intermediate or shallow water
+(although the solutions can be extended to those conditions, as
+\citet{phillips1981dispersion} did, for example).
 Within these limits, the crest-action conservation equations provide an
@@
 of short waves by longer waves.
-This approach, of course, isolates this particular process from others
+This approach, of course, isolates this particular process from other
 dominant modulation and nonlinear transfer processes, such as wave growth
@@
 
-Although in this paper we only explored the wavenumbers 1 and 10 for the long
-and short waves, respectively ($k/k_L = 10$), the results are not sensitive
-to different values of $k_L$, provided that the amplitude $a_L$ is set such that
-the long-wave steepness $\varepsilon_L = a_L / k_L$ is the same.
-Numerical simulations with that compare $k/k_L = 10$ and $k/k_L = 100$ show
-very little sensitivity $k/k_L$ (not shown), consistent with the results of
+Although in this paper we only explored the solutions for $k_L = 1$ and $k=10$
+the results are not sensitive to different values of $k_L$, provided that the
+amplitude $a_L$ is set such that the long-wave steepness $\varepsilon_L = a_L / k_L$
+is the same.
+Numerical simulations with $k/k_L = 10$ and $k/k_L = 100$ show little
+sensitivity to the ratio $k/k_L$ (not shown), consistent with the results of
 \citet{longuet1987propagation}.
@@
 largely independent of the ratio $k/k_L$.
+These results can be reproduced with the companion numerical model.
 
@@
 unsteadily.
-However, in this study we demonstrate that the unsteady steepening of short
+However, here we demonstrate that the unsteady steepening of short
 waves only occurs in the case of a sudden appearance of long waves, where
 the velocity fields induced by the long waves act in their full capacity
-on the short waves that have not yet had time to "lock in" to the long-wave
+on the short waves that have not yet had time to lock-in to the long-wave
 crests.
@@
 and regional wave forecasting.
+Currently, most spectral wave models used for research or forecasting
+do not consider the preferential steepening of short waves in the context
+of their dispersion properties or the wind input
+(\textit{i.e.} via the short-wave phase speed).
+Some dissipation source functions use the cumulative mean squared slope
+to quantify preferential dissipation of short waves due to this effect
+\citep{donelan2012modeling,romero2019distribution}.
 In the early days of spectral wave modeling, \citet{willebrand1975energy}
@@
 and found the effect of up to several percent modulation.
-For a practical implementation in a spectral wave model, this would have to
-incur an additional loop over all frequencies for each short-wave frequency.
+For a practical implementation, this would incur an additional loop over all
+frequencies longer than the considered short-wave frequency.
 It was thus not deemed practical to include this effect in wave models at the time.
-However, with the recent advances in machine learning and phase-resolving modeling
-of short waves, the effect of wave-modulation deserves re-examination.
-
-TODO Implications for spectral wave models and wind input to short waves.
+However, with the recent advances in computing power, machine learning for
+acceleration of complex functions, and phase-resolved modeling of short waves,
+considering the effect of wave-modulation beyond the simple cumulative mean
+squared slope deserves re-examination.
 
@@
 The latter two have not been previously examined in the context of short-wave
-modulation, but are important for correctly calculating the wind input into
+modulation, but may be important for correctly calculating the wind input into
 short waves, which is proportional to the wind speed relative to the wave
@@
 The numerical solutions of the full wave crest and action conservation equations
-are consistent with the analytical solutions for $\varepsilon_L < 0.2$.
+are consistent with the analytical solutions for $\varepsilon_L \lesssim 0.2$.
 The simulations also reveal that the unsteady growth of short-wave steepness
@@
 modulation of short waves by long waves.
+It may be worthwhile to revisit a more systematic inclusion of the hydrodynamic
+modulation effects in phase-averaged spectral wave models, beyond the simple
+cumulative mean squared slope for wave dissipation, but also for modulating
+the dispersion and wind input into short waves.
 

### 2024-09-30 paper/scripts/plot_analytical_solutions_with_phase.py

@@ -83,6 +83,7 @@ def plot_analytical_solutions(a_L: float):
 
     for ax in axes[-2:]:
         ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
+        ax.set_xlabel(r"$\psi$")
 
     axes[0].set_title(r"$\widetilde{k}/k$")
     axes[1].set_title(r"$\widetilde{g}/g$")
@@ -90,7 +91,7 @@ def plot_analytical_solutions(a_L: float):
     axes[3].set_title(r"$\widetilde{ak}/(ak)$")
     axes[4].set_title(r"$\widetilde{\sigma}/\sigma$")
     axes[5].set_title(r"$\widetilde{C_p}/C_p$")
-
+    
     for n, ax in enumerate(axes):
         ax.text(
             0,

### 2024-09-30 paper/scripts/plot_numerical_vs_analytical_solutions_with_phase.py

@@ -126,7 +126,7 @@ axes[2, 0].set_ylabel(r"$\widetilde{ak}/(ak)$")
 
 for ax in axes[-1]:
     ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
-    ax.set_xlabel("Phase")
+    ax.set_xlabel(r"$\psi$")
 
 plt.savefig("../figures/numerical_vs_analytical_solutions_with_phase.png", dpi=200)
 plt.close()

### 2024-10-01 paper/paper.tex
A few more references + spelling

@@
 waves.
-The numerical simulations stabilize if the short-wave field is initialized
+The numerical simulations stabilise if the short-wave field is initialised
 from existing steady solutions, like those by \citet{longuet1960changes}.
@@
 The main prior studies on hydrodynamic modulation, and their assumptions and
-approaches to the solution, are summarized in Table
+approaches to the solution, are summarised in Table
 \ref{table:modulation_literature_summary}.
@@
 For the ambient forcing, we consider a train of monochromatic long wave
-defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
+defined, to the first order in steepness, by the elevation $\eta_L = a_L \cos{\psi}$,
 where $\psi = k_L x - \sigma_L t$ is the long-wave phase and $\sigma_L$ its
@@
 and intrinsic phase speed in presence of long waves.
-The wavenumber is derived by linearizing the conservation of wave crests
+The wavenumber is derived by linearising the conservation of wave crests
 (\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
@@
 This is because combining the orbital velocities evaluated at $z=\eta$ and the
-linearized conservation of wave crests in (\ref{eq:wavenumber}) results in
+linearised conservation of wave crests in (\ref{eq:wavenumber}) results in
 excess short-wave wavenumber being created.
@@
 a variety of the gravity modulation forms.
-This value of $\varepsilon_L$ allows easier visualization of the modulation
+This value of $\varepsilon_L$ allows easier visualisation of the modulation
 differences, while still being relatively moderate.
@@
   and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
-  (b) As in (a) but normalized by the modulation of the linear wave evaluated at $z=0$, $g_1$.
+  (b) As in (a) but normalised by the modulation of the linear wave evaluated at $z=0$, $g_1$.
 }
@@
 The modulation of short-wave amplitude can be derived in a similar way as we did
-for the wavenumber, except that here we linearize the wave action balance
+for the wavenumber, except that here we linearise the wave action balance
 (\ref{eq:wave_action2}) and integrate it in time:
@@
 crest and action conservation equations.
-(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are discretized using
+(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are discretised using
 second-order central finite difference in space and integrated in time using the
@@
 the fields that we compute derivatives of are smooth and continuous.
-The maximum relative error of the centered finite difference in this model
+The maximum relative error of the centred finite difference in this model
 is $\approx 0.066\%$.
@@
 
-Finally, the numerical equations here are integrated in a fixed reference frame,
+The numerical equations here are integrated in a fixed reference frame,
 rather than that moving with the long-wave phase speed.
@@
 rightward with the speed of $C_g + U$ (neglecting group speed inhomogeneity).
+The crest and action conservation equations are integrated in the curvilinear
+coordinate system in which the horizontal coordinate follows the long-wave
+surface, akin to \citet{zhang1990evolution}.
 
@@
 $\varepsilon_L = 0.1$, $0.2$, and $0.4$ (Fig. \ref{fig:numerical_solutions}).
-The long-waves are initialized from rest ($a_L = 0$) and gradually ramped up
+The long-waves are initialised from rest ($a_L = 0$) and gradually ramped up
 to their target steepness over 5 long-wave periods, thus emulating the passage
@@
 The steepening is driven by the progressive increase in the action of short
-waves (as opposed to their wavenumber), and is centered on the short-wave group
+waves (as opposed to their wavenumber), and is centred on the short-wave group
 that begins its journey in the trough of the long wave.
@@
 decreased.
-We reproduce this behavior in Fig. \ref{fig:modulation_3panel_infinite}, which
+We reproduce this behaviour in Fig. \ref{fig:modulation_3panel_infinite}, which
 shows the evolution of the short-wave action, wavenumber, and steepness
@@
   $k_L = 1$, $k = 10$, $\varepsilon_L = 0.1$.
+  The black contour follows the value of 1, which corresponds to no change
+  from the initial values.
 }
@@
 responsible for the instability.
-However, the instability only occurs if the short waves are initialized as a
+However, the instability only occurs if the short waves are initialised as a
 uniform field.
-If they are initialized instead as a periodic function of the long-wave phase
+If they are initialised instead as a periodic function of the long-wave phase
 such that their wavenumber (and optionally, action) is higher on the crests
@@
 However, such a scenario is not realistic in the ocean where waves tend to
-organize into groups due to dispersion.
+organise into groups due to dispersion.
 To test whether the sudden onset of long-wave forcing causes the unstable growth,
@@
 and their indefinite steepening vanishes.
-The stabilization effect of the long-wave ramp is more readily assessed in
+The stabilisation effect of the long-wave ramp is more readily assessed in
 Fig. \ref{fig:unsteady_growth_timeseries}, which shows the maximum short-wave
@@
 dominant modulation and nonlinear transfer processes, such as wave growth
-by wind, wave dissipation, and aerodynamic sheltering by the longer waves.
+by wind, wave dissipation, and aerodynamic sheltering by the longer waves,
+that are otherwise present in most measurements
+\citep{plant1986two,laxague2017gravity}.
 
@@
 \citet{peureux2021unsteady} examined this problem numerically and found that
-the solutions are stable only if the short-wave field is initialized from
+the solutions are stable only if the short-wave field is initialised from
 a prior steady solution of the form of that from \citet{longuet1960changes}.
-If the short-wave field is instead initialized from a uniform field, they
+If the short-wave field is instead initialised from a uniform field, they
 found that the short-wave amplitude, and thus steepness as well, grow
@@
 at the short-wave group speed.
-When this occurs, the modulated short-waves are no longer in a stabilizing
+When this occurs, the modulated short-waves are no longer in a stabilising
 resonance with the surface convergence and divergence induced by the long-wave
-orbital velocity, and each long-wave passage further destabilizes the short-wave
+orbital velocity, and each long-wave passage further destabilises the short-wave
 field, causing it to steepen indefinitely.
@@
 \citep{donelan2012modeling,romero2019distribution}.
-In the early days of spectral wave modeling, \citet{willebrand1975energy}
+In the early days of spectral wave modelling, \citet{willebrand1975energy}
 examined the effect of a wave spectrum on the short-wave group velocity,
@@
 However, with the recent advances in computing power, machine learning for
-acceleration of complex functions, and phase-resolved modeling of short waves,
+acceleration of complex functions, and phase-resolved modelling of short waves,
 considering the effect of wave-modulation beyond the simple cumulative mean
@@
 from a linear wave theory perspective.
-Using the linearized wave crest and action conservation equations coupled with
+Using the linearised wave crest and action conservation equations coupled with
 nonlinear solutions for the effective gravity, we derived steady analytical
@@
 This is mainly due to the analytical solution requiring the crest and action
-conservation equations to be linearized, and the nonlinearity arising due
+conservation equations to be linearised, and the nonlinearity arising due
 to the evaluation of short waves on the surface of the long wave (as opposed to
@@
 A more likely condition such as that of a long-wave groups, are only mildly
-destabilizing for the short waves.
+destabilising for the short waves.
 These results suggest that for mildly-to-moderately steep short waves,
@@
   Milan Curcic was partially supported by the National Science Foundation
-  under grant number AGS-1745384.
+  under the grant number AGS-1745384.
 }

### 2024-10-01 paper/scripts/plot_scale_analysis.py
Reorganize

@@ -1,153 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import matplotlib
-
-import twowave
-
-matplotlib.rc("font", size=12)
-
-
-def elevation_linear(a: float, phase: float) -> float:
-    return a * np.cos(phase)
-
-
-def elevation_stokes(a: float, k: float, phase: float) -> float:
-    term1 = np.cos(phase)
-    term2 = 0.5 * a * k * np.cos(2 * phase)
-    term3 = (a * k) ** 2 * (3 / 8 * np.cos(3 * phase) - 1 / 16 * np.cos(phase))
-    return a * (term1 + term2 + term3)
-
-
-def gravity_linear(
-    a: float,
-    k: float,
-    phase: float,
-    g0: float,
-) -> float:
-    """Gravitational acceleration at the surface of a linear wave."""
-    eta = elevation_linear(a, phase)
-    return g0 * (
-        1 - a * k * np.exp(k * eta) * (np.cos(phase) - a * k * np.sin(phase) ** 2)
-    )
-
-
-def gravity_stokes(a: float, k: float, phase: float, g0: float) -> float:
-    """Gravitational acceleration at the surface of a Stokes wave."""
-    eta = elevation_stokes(a, k, phase)
-    res = g0 * (
-        1
-        - a
-        * k
-        * (
-            np.cos(phase)
-            - a
-            * k
-            * np.sin(phase)
-            * (
-                (1 - 1 / 16 * (a * k) ** 2) * np.sin(phase)
-                + a * k * np.sin(2 * phase)
-                + 9 / 8 * (a * k) ** 2 * np.sin(3 * phase)
-            )
-        )
-        * np.exp(k * eta)
-    )
-    return res
-
-
-def calculate_tendencies(a_L: float):
-    phase = np.arange(0, 2 * np.pi + 1e-4, 1e-4)
-    g0 = 9.8
-    k_L = 1
-    ak_L = a_L * k_L
-    omega_L = np.sqrt(g0 * k_L)
-
-    L = 2 * np.pi / k_L
-    dx = L / phase.size
-
-    k0 = 10
-    N0 = 1
-
-    k = k0 * np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
[…]

### 2024-10-01 scripts/plot_analytical_solutions_with_phase.py

@@ -91,7 +91,7 @@ def plot_analytical_solutions(a_L: float):
     axes[3].set_title(r"$\widetilde{ak}/(ak)$")
     axes[4].set_title(r"$\widetilde{\sigma}/\sigma$")
     axes[5].set_title(r"$\widetilde{C_p}/C_p$")
-    
+
     for n, ax in enumerate(axes):
         ax.text(
             0,

### 2024-10-04 paper.tex
Fix typo
Minor edits and clarifications

--- 2024-10-01 paper/paper.tex
@@
 equations.
-The proposed analytical solutions are a good approximation of the fully
+The proposed steady solutions are a good approximation of the fully
 nonlinear numerical solutions in long-wave steepness up to $\approx 0.2$.
 As a corollary, except for a subset of initial conditions, the solutions to
-fully nonlinear crest-action conservation equations are mostly steady in the
-reference frame of the long waves.
+the fully nonlinear crest-action conservation equations are mostly steady in
+the reference frame of the long waves.
 \end{abstract}
@@
 As \citet{peureux2021unsteady} explained, the presence of this term causes
-unstable growth of wave action in infinite long-wave trains.
+unsteady growth of wave action in infinite long-wave trains.
 (\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are the governing equations
@@
 \label{eq:gravity_modulation_linear_no_curvature_mean_level}
-\widetilde{g} = \left( 1 - \varepsilon_L \cos{\psi} \right)
+\frac{\widetilde{g}}{g} = \left( 1 - \varepsilon_L \cos{\psi} \right)
 \end{equation}
@@
 
-\subsection{Unstable growth in Pereux et al. (2021)}
-\label{subsection:unstable_growth}
+\subsection{Unsteady growth in Pereux et al. (2021)}
+\label{subsection:unsteady_growth}
 
@@
 of a long wave perturbation in the middle of a homogeneous short wave field".
-However, such a scenario is not realistic in the ocean where waves tend to
-organise into groups due to dispersion.
-To test whether the sudden onset of long-wave forcing causes the unstable growth,
+However, it is not obvious whether such a scenario is realistic as waves tend to
+disperse into groups, and the time scales of incoming swell fronts are much
+longer than those of short-wave generation by wind.
+To test whether the sudden onset of long-wave forcing causes the unsteady growth,
 we repeat the numerical simulation of Fig.
@@
 (Fig. \ref{fig:inhomogeneity_tendencies}).
-\citet{peureux2021unsteady} correctly pointed out that the unstable
-growth of short-wave steepness occurs due to the resonance of the inhomogeneity
+\citet{peureux2021unsteady} correctly pointed out that the unsteady growth of
+short-wave steepness occurs due to the resonance of the inhomogeneity
 tendency with the short-wave action perturbation.
@@
 The elevation of the long waves thus gradually ramps up from rest to $a_L$ at
-$t = N/2 T_L$, after which it gradually decreases toward zero.
+$t = n T_L / 2$, after which it gradually decreases toward zero.
 In this simulation, $n = 10$ and $\varepsilon_L = 0.1$ as in the previous two
@@
 and $|\partial U / \partial t| \ll \sigma |U|$.
-This implies $k_L \ll k$ and $|\sigma_L| \ll \sigma|$, respectively.
+This implies $k_L \ll k$ and $\sigma_L \ll \sigma$, respectively.
 Another requirement is for the short-waves to remain linear, which implies
@@
 Furthermore, the short waves examined here do not include the effects of
-surface tension (gravity-capillary waves) or intermediate or shallow water
-(although the solutions can be extended to those conditions, as
-\citet{phillips1981dispersion} did, for example).
+surface tension (gravity-capillary waves) or intermediate or shallow water,
+although the solutions can be extended to those conditions, as
+\citet{phillips1981dispersion} did, for example.
 Within these limits, the crest-action conservation equations provide an
@@
 crests.
-In such a hypothetical scenario, sudden forcing on short waves causes the
-modulation pattern to detach itself from the long-wave crests and to propagate
-at the short-wave group speed.
+In such a scenario, sudden forcing on short waves causes the modulation
+pattern to detach itself from the long-wave crests and to propagate at the
+short-wave group speed.
 When this occurs, the modulated short-waves are no longer in a stabilising
@@
 do not consider the preferential steepening of short waves in the context
-of their dispersion properties or the wind input
-(\textit{i.e.} via the short-wave phase speed).
+of their dispersion properties or the wind input via the short-wave phase speed.
 Some dissipation source functions use the cumulative mean squared slope
@@
 In the early days of spectral wave modelling, \citet{willebrand1975energy}
-examined the effect of a wave spectrum on the short-wave group velocity,
-and found the effect of up to several percent modulation.
+examined the effect of a wave spectrum on the short-wave group velocity
+and found the modulation effect to be up to several percent.
 For a practical implementation, this would incur an additional loop over all
 frequencies longer than the considered short-wave frequency.
-It was thus not deemed practical to include this effect in wave models at the time.
+It was thus not deemed practical to include this effect in wave models at the
+time.
 However, with the recent advances in computing power, machine learning for

### 2024-10-10 scripts/hydrodynamic_modulation_diagram.py
Fix caption

@@ -25,7 +25,7 @@ ax.plot(x, combined_wave, "k-", lw=3, zorder=1)
 
 # Add surface velocity arrows
 arrow_positions = np.arange(np.pi, 4 * np.pi, np.pi)
-for pos in arrow_positions:
+for i, pos in enumerate(arrow_positions):
     # Calculate wave slope at this position
     slope = np.cos(pos)  # Derivative of sin(x)
     angle = np.arctan(slope)
@@ -38,28 +38,52 @@ for pos in arrow_positions:
     # Position arrow slightly below the wave
     y_offset = -0.25
 
-    ax.arrow(
-        pos - 1.6 * dx,
-        np.sin(pos - 1.6 * dx) + y_offset,
-        dx,
-        dy * 0.9,
-        head_width=0.2,
-        head_length=0.2,
-        fc="tab:blue",
-        ec="tab:blue",
-        width=0.1,
-    )
-    ax.arrow(
-        pos + 1.6 * dx,
-        np.sin(pos + 1.6 * dx) + y_offset,
-        -dx,
-        -dy,
-        head_width=0.2,
-        head_length=0.2,
-        fc="tab:blue",
-        ec="tab:blue",
-        width=0.1,
-    )
+    if i % 2 == 0:  # Even positions (arrows pointing inward)
+        ax.arrow(
+            pos - 1.6 * dx,
+            np.sin(pos - 1.6 * dx) + y_offset,
+            dx,
+            dy * 0.9,
+            head_width=0.2,
+            head_length=0.2,
+            fc="tab:blue",
+            ec="tab:blue",
+            width=0.1,
+        )
+        ax.arrow(
+            pos + 1.6 * dx,
+            np.sin(pos + 1.6 * dx) + y_offset,
+            -dx,
+            -dy,
+            head_width=0.2,
+            head_length=0.2,
+            fc="tab:blue",
+            ec="tab:blue",
+            width=0.1,
+        )
+    else:  # Odd positions (arrows pointing outward)
+        ax.arrow(
+            pos - 0.4 * dx,
+            np.sin(pos - 0.4 * dx) + y_offset,
+            -dx,
+            -dy * 0.9,
+            head_width=0.2,
+            head_length=0.2,
+            fc="tab:blue",
+            ec="tab:blue",
+            width=0.1,
+        )
+        ax.arrow(
+            pos + 0.4 * dx,
[…]

### 2024-10-12 paper.tex
Fix caption
Expand on limitations

@@
 \label{eq:lhs1960wavenumber}
-\widetilde{k} = k (1 + \varepsilon_L)
+\widetilde{k} = k (1 + \varepsilon_L \cos{\psi})
 \end{equation}
@@
 respectively, and $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
-wavenumber $k_L$ and amplitude $a_L$.
+wavenumber $k_L$ and amplitude $a_L$, and $\psi$ is the long-wave phase.
 This process is illustrated in Fig. \ref{fig:hydrodynamic_modulation_diagram}.
@@
 \citet{zhang1990evolution} considered weakly nonlinear short waves
-using a non-linear Schr\" odinger equation and found stronger wavenumber
+using a nonlinear Schr\" odinger equation and found stronger wavenumber
 modulation but weaker amplitude modulation than those predicted by
@@
   phase-speed $C_{pL}$ from left to right.
-  The short waves propagate on the long-wave surface with their group velocity
-  $C_g$.
+  The short waves propagate on the long-wave surface, their action density
+  moving with the group velocity $C_g$.
   Long waves induce orbital velocities $U$ that cause near-surface convergence
@@
 short-waves at the long-wave surface rather than at the mean water level.
-This allows us to derive alternative, steady, non-linear solutions for the
+This allows us to derive alternative, steady, nonlinear solutions for the
 modulation of short waves.
@@
 \begin{tabular}{cccccc}
-Reference   & Governing & Gravity & Nonlinear  & Nonlinear   & Solution \\
-            & equations & varies  & long waves & short waves &          \\
+Reference   & Analysis   & Gravity & Nonlinear  & Nonlinear   & Solution \\
+            & frameworks & varies  & long waves & short waves &          \\
 \hline
@@
 \caption{
-  Summary of prior and present (this paper) approaches to calculating the
-  modulation of short waves by long waves.
+  Non-exhaustive summary of prior and present (this paper) approaches to
+  calculating the modulation of short waves by long waves.
 }
@@
 
-(\ref{eq:dispersion}) and (\ref{eq:wave_crests}) only describe the evolution of
+(\ref{eq:dispersion}) and (\ref{eq:wave_crests}) describe the evolution of
 the short-wave wavenumber.
@@
 \end{equation}
-where $N$ is the action density, and $C_g = 1/2\sqrt{g/k}$ is the group speed
-in deep water.
-Although the absence of wave growth and dissipation is not generally realistic
-in the open ocean, it is a useful approximation here to isolate the effects
+where $N$ is the action density of short waves defined as the ratio of their
+energy density to their intrinsic frequency, $E/\sigma$,
+and $C_g = 1/2\sqrt{g/k}$ is their group speed in deep water.
+The wave action density expressed as $E/\sigma$ is commonly taken as
+conservative in wave modulation theory
+\citep{phillips1981dispersion,longuet1987propagation}, however as
+\citet{dysthe1988orbiting} show, it is conservative only for gently sloped waves.
+Although the absence of wave growth and dissipation is not generally realistic in
+the open ocean, it is a useful approximation here to isolate the effects
 of hydrodynamic modulation solely due to the motion of the long waves.
@@
 
-For the ambient forcing, we consider a train of monochromatic long wave
-defined, to the first order in steepness, by the elevation $\eta_L = a_L \cos{\psi}$,
-where $\psi = k_L x - \sigma_L t$ is the long-wave phase and $\sigma_L$ its
-angular frequency.
-The long wave induces horizontal and vertical orbital velocities whose values
-at the surface $\eta_L$ are:
+For the ambient forcing, we consider a train of monochromatic long waves
+defined, to the first order in $\varepsilon_L$, by the elevation
+$\eta_L = a_L \cos{\psi}$, where $\psi = k_L x - \sigma_L t$ is the long-wave
+phase and $\sigma_L$ its angular frequency.
+The velocity potential of the long wave is:
+
+\begin{equation}
+\label{eq:phi_L}
+\phi_L = \dfrac{a_L \omega_L}{k_L} e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
+\end{equation}
+so, up to the third order in $\varepsilon_L$, the long-wave induced horizontal
+and vertical orbital velocities, evaluated at the wave surface $z = \eta_L$, are:
 
@@
 \end{equation}
-
-(\ref{eq:dispersion})-(\ref{eq:w_L}) are the governing equations used in this paper.
-Note that rather than deriving the modulation solution from the velocity
+Evaluating these velocities at the wave surface $z = \eta_L$ rather than at
+the mean water level $z=0$ is a departure from the linear theory, making the
+treatment of the long waves here as weakly nonlinear.
+
+(\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are the governing equations used
+in this paper, with (\ref{eq:U_L})-(\ref{eq:w_L}) describing the long-wave
+induced ambient velocity forcing for the short waves. 
+Rather than deriving the modulation solution from the velocity
 potential of two waves, as \citet{longuet1960changes} did, starting from the
@@
 These results can be reproduced with the companion numerical model.
+Also not considered here are the short-wave trains that are oblique to the
+longer waves, which is relatively straightforward to include
+\citep{peureux2021unsteady}.
 

### 2024-10-14 paper.tex
Updates from reviews

@@
 forecasting.
-In this paper, we revisit the linear wave theory and derive higher-order
+In this paper, we revisit the linear (Airy) wave theory and derive higher-order
 steady solutions for the change of short-wave wavenumber, action density, and
@@
 The variety of analytical and numerical frameworks (Eulerian crest and action
-conservation, Hamiltonian, Schr\" odinger) that relatively closely reproduce
-steady solutions for short-wave modulation suggests that the solutions are robust.
+conservation, Hamiltonian, Schr\" odinger) that tackle the same fundamental
+physics and that relatively closely reproduce steady solutions for short-wave
+modulation suggests that the solutions are robust.
 Nevertheless, alternative analytical and numerical approaches to the modulation
@@
 
-In this paper, we revisit this problem from linear wave theory but consider
-the long-wave slope to be significant enough to require evaluating the
-short-waves at the long-wave surface rather than at the mean water level.
+In this paper, we revisit this problem from the perspective of linear wave theory
+but consider the long-wave slope to be significant enough to require evaluating
+the short-waves at the long-wave surface rather than at the mean water level.
 This allows us to derive alternative, steady, nonlinear solutions for the
-modulation of short waves.
+modulation of short waves, within the limits of small $\varepsilon_L$.
 The analytical solutions yield modulations that are stronger than the original
@@
 and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
-Combine (\ref{eq:dispersion}) and (\ref{eq:wave_crests}) to get:
+Inserting (\ref{eq:dispersion}) into (\ref{eq:wave_crests}) and noting that
+$\partial \sqrt{gk} / \partial x = 1/2 \sqrt{g/k} \partial k / \partial x + 1/2 \sqrt{k/g} \partial g / \partial x$,
+we get:
 
@@
 \end{equation}
-so, up to the third order in $\varepsilon_L$, the long-wave induced horizontal
-and vertical orbital velocities, evaluated at the wave surface $z = \eta_L$, are:
+where $z$ is the vertical distance from the mean water level that is positive
+upwards.
+(\ref{eq:phi_L}) is accurate to the third order in $\varepsilon_L$.
+The long-wave induced horizontal and vertical orbital velocities, evaluated at
+the wave surface $z = \eta_L$, are:
 
@@
 \label{eq:U_L}
-U = a_L \omega_L e^{\varepsilon_L \cos{\psi}} \cos{\psi}
+U = a_L \omega_L e^{k_L z} \cos{\psi} + \mathcal{O}(\varepsilon_L^4)
 \end{equation}
@@
 \label{eq:w_L}
-W = a_L \omega_L e^{\varepsilon_L \cos{\psi}} \sin{\psi}
+W = a_L \omega_L e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
 \end{equation}
@@
 treatment of the long waves here as weakly nonlinear.
+As \citet{zhang1990evolution} point out, this is appropriate to do when the
+amplitude of the long waves is of the similar magnitude or large than the
+wavelength of the short waves.
 
@@
   \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
-\right]
+\right] + \mathcal{O}(\varepsilon_L^4)
 \end{equation}
@@
 To the first order in $\varepsilon_L$, and using (\ref{eq:k_short_exact})
-and (\ref{eq:wave_action_modulation}), the above simplifies to:
+and (\ref{eq:wave_action_modulation}), the above simplifies to the result of
+\citet{longuet1960changes}:
 
@@
   \left( 1 + \varepsilon_L \right)^{0.75}
+  \approx \frac{(1 + \varepsilon_L)^{0.75}}{(1 - \varepsilon_L)^{0.25}}
+  \approx 1 + \varepsilon_L
 \end{equation}
@@
 is $\approx 0.066\%$.
+The wavenumber and wave action are conservative over time within
+$\mathcal{O}(10^{-7})$ and $\mathcal{O}(10^{-5})$ relative error, respectively.
 We show in the next section that our numerical solutions for infinite long-wave
@@
 Thanks to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan for the
-helpful discussions.
+helpful discussions that improved this paper.
 }

### 2024-10-15 paper.tex
Near-final edits

@@
 solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
+The short-wave steepness modulation is attributed to be 5/8 due to the
+wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
 We further examine the result of \citet{peureux2021unsteady} who found through
@@
 \begin{equation}
-\label{eq:phi_L}
-\phi_L = \dfrac{a_L \omega_L}{k_L} e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
+\label{eq:phi}
+\phi = \dfrac{a_L \omega_L}{k_L} e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
 \end{equation}
@@
 upwards.
-(\ref{eq:phi_L}) is accurate to the third order in $\varepsilon_L$.
-The long-wave induced horizontal and vertical orbital velocities, evaluated at
-the wave surface $z = \eta_L$, are:
+(\ref{eq:phi}) is accurate to the third order in $\varepsilon_L$ in deep water
+because the second- and third-order terms in the Stokes expansion series are
+exactly zero.
+The long-wave induced horizontal and vertical orbital velocities are:
 
@@
 \label{eq:U_L}
-U = a_L \omega_L e^{k_L z} \cos{\psi} + \mathcal{O}(\varepsilon_L^4)
+U =
+\frac{\partial \phi}{\partial x} =
+a_L \omega_L e^{k_L z} \cos{\psi} + \mathcal{O}(\varepsilon_L^4)
 \end{equation}
@@
 \label{eq:w_L}
-W = a_L \omega_L e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
+W =
+\frac{\partial \phi}{\partial z} =
+a_L \omega_L e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
 \end{equation}
@@
 the higher-order solution (\ref{eq:k_short_exact}) is not.
-This is because combining the orbital velocities evaluated at $z=\eta$ and the
-linearised conservation of wave crests in (\ref{eq:wavenumber}) results in
-excess short-wave wavenumber being created.
+This is because combining the orbital velocities evaluated at $z=\eta_L$ and the
+linearised conservation of wave crests in (\ref{eq:wavenumber}) creates excess
+short-wave wavenumber.
 The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
@@
 \label{eq:gravity_modulation_linear_no_curvature}
-\frac{\widetilde{g}}{g} = \left( 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}} \right)
-\end{equation}
-Further, if we evaluate the orbital accelerations at the mean water level $z=0$,
+\frac{\widetilde{g}}{g} = 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
+\end{equation}
+If we evaluate the orbital accelerations at the mean water level $z=0$,
 this further simplifies to:
@@
 \label{eq:gravity_modulation_linear_no_curvature_mean_level}
-\frac{\widetilde{g}}{g} = \left( 1 - \varepsilon_L \cos{\psi} \right)
+\frac{\widetilde{g}}{g} = 1 - \varepsilon_L \cos{\psi}
 \end{equation}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities.png}
+\includegraphics[width=0.7\textwidth]{figures/fig_effective_gravities.png}
 \caption{
@@
 respectively, the curvature affects the phase-averaged values but not the extremes.
+As the slope effects on the effective gravity exist only away from crests and
+troughs, they may be neglected in the steady solutions where modulation is
+typically evaluated at the long-wave crests. 
 The effective gravity reduction at the crests is $\approx 11\%$, $25\%$, and $66\%$
@@
 =
-  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
-  \left( \dfrac{\widetilde{k}}{k} \right)^{0.25}
-  \left( \dfrac{\widetilde{N}}{N} \right)^{0.5}
+  \left( \dfrac{\widetilde{k}}{k} \right)^{\frac{1}{4}}
+  \left( \dfrac{\widetilde{N}}{N} \right)^{\frac{1}{2}}
+  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
 \end{equation}
@@
 \dfrac{\widetilde{a}}{a} = 
-  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
-  \left( 1 + \varepsilon_L \right)^{0.75}
-  \approx \frac{(1 + \varepsilon_L)^{0.75}}{(1 - \varepsilon_L)^{0.25}}
+  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
+  \left( 1 + \varepsilon_L \right)^{\frac{3}{4}}
+  \approx \frac{(1 + \varepsilon_L)^{\frac{3}{4}}}{(1 - \varepsilon_L)^{\frac{1}{4}}}
   \approx 1 + \varepsilon_L
@@
 The amplitude and steepness modulation of the short waves as a function of the
-long-wave phase for $\varepsilon_L = 0.1$, $0.2$, and $0.4$ are shown and compared
-to the \citet{longuet1960changes} solutions in Figs.
-\ref{fig:analytical_solutions_ak0.1}, \ref{fig:analytical_solutions_ak0.2},
-and \ref{fig:analytical_solutions_ak0.4}, respectively.
+long-wave phase for $\varepsilon_L = 0.1$ are compared to the
+\citet{longuet1960changes} solutions in Fig.
+\ref{fig:analytical_solutions_ak0.1} panels (c) and (d), respectively.
 Also shown are the modulation of the short-wave intrinsic frequency and phase
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.2.png}
-\caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.2$.}
-\label{fig:analytical_solutions_ak0.2}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.4.png}
-\caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.4$.}
-\label{fig:analytical_solutions_ak0.4}
+Finally, the short-wave steepness modulation follows from (\ref{eq:k_short_exact})
+and (\ref{eq:wave_amplitude_modulation}):
+
+\begin{equation}
+\label{eq:ak_modulation}
+\dfrac{\widetilde{ak}}{ak} = 
+  \left( \dfrac{\widetilde{k}}{k} \right)^{\frac{5}{4}}
+  \left( \dfrac{\widetilde{N}}{N} \right)^{\frac{1}{2}}
+  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
+\end{equation}
+The short-wave steepness modulation at the long-wave crests is thus largely
+controlled by the wavenumber modulation (5/8, or, 62.5\%), followed by the
+convergence of wave action (1/4, or, 25\%), and by least amount by the
+effective gravity reduction (1/8, or, 12.5\%) (Fig. \ref{fig:modulation_contributors}).
+(\ref{eq:ak_modulation}) and Fig. \ref{fig:modulation_contributors} demonstrate
+that more than half of the steepness modulation can be captured by the
+wavenumber modulation alone.
+Further, considering the wavenumber and wave action modulations, while
+neglecting the gravity modulation, captures $\approx$88\% of the steepness
+modulation at the long-wave crests.
+This result may be useful for the interpretation of remote sensing products
+that resolve some but not all of the short-wave modulation effects.
+Finally, the relative contributions of different modulation factors remain
+mostly constant as $\varepsilon_L$ increases, with only a mild increase of the
+effective gravity modulation contribution, and mild increases of the
+wavenumber and wave action modulation contributions.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_modulation_contributors.png}
+\caption{
+  Contributions of short-wave wavenumber, action, and effective gravity
+  modulations to the steepness modulation, as function of the long-wave
+  steepness $\varepsilon_L$.
+  Panel (a) shows each modulation factor (colour) and their product (black),
+  and panel (b) shows their relative contributions in percentage, calculated as
+  the ratio of the logarithm of each modulation factor to the logarithm of the
+  product of all modulations.
+}
+\label{fig:modulation_contributors}
 \end{figure}
@@
 of the short-wave action.
-That is, the inhomogeneity term $N \dfrac{\partial C_g}{\partial x}$ is
+That is, the inhomogeneity term $N \partial C_g / \partial x$ is
 responsible for the instability.
@@
 infinite long-wave train case and the case of applying a linear ramp.
-Comparing the wave action propagation ($C_g \frac{\partial N}{\partial x}$) and
-inhomogeneity ($N \frac{\partial C_g}{\partial x}$) tendencies reveals that
+Comparing the wave action propagation ($C_g \partial N / \partial x$) and
+inhomogeneity ($N \partial C_g / \partial x$) tendencies reveals that
 after only one long-wave period, the propagation and inhomogeneity tendencies
@@
 the mean water level) and due to nonlinear effective gravity.
+In the steady solutions, the short-wave steepness modulation is dominated by
+the wavenumber modulation (5/8), followed by the wave action modulation (1/4),
+and is least affected by the effective gravity modulation (1/8).
 
 The numerical solutions of the full wave crest and action conservation equations
-are consistent with the analytical solutions for $\varepsilon_L \lesssim 0.2$.
+are consistent with the analytical solutions for $\varepsilon_L \lesssim 0.2$,
+and are similar in amplitude to the numerical solutions of
+\citet{longuet1987propagation} and \citet{zhang1990evolution}.
 The simulations also reveal that the unsteady growth of short-wave steepness
@@
 occur in the open ocean.
-A more likely condition such as that of a long-wave groups, are only mildly
+A more likely condition such as that of long-wave groups, are only mildly
 destabilising for the short waves.
@@
 \backsection[Acknowledgements]{
-Thanks to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan for the
-helpful discussions that improved this paper.
+I am thankful to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan
+for the discussions that helped improve this paper.
 }
@@
 \backsection[Funding]{
-  Milan Curcic was partially supported by the National Science Foundation
-  under the grant number AGS-1745384.
+  Milan Curcic was partly supported by the National Science Foundation
+  grant AGS-1745384, Transatlantic Research Partnership, and the Office of Naval
+  Research Coastal Land Air-Sea Interaction DRI.
 }

### 2024-10-15 paper.tex
Switch to PDF for figures

@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.1.png}
+\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.1.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.7\textwidth]{figures/fig_effective_gravities.png}
+\includegraphics[width=0.7\textwidth]{figures/fig_effective_gravities.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities_mean_max_min.png}
+\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities_mean_max_min.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=\textwidth]{figures/fig_modulation_contributors.png}
+\includegraphics[width=\textwidth]{figures/fig_modulation_contributors.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=\textwidth]{figures/numerical_vs_analytical_solutions_with_phase.png}
+\includegraphics[width=\textwidth]{figures/numerical_vs_analytical_solutions_with_phase.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_infinite_wave_train.png}
+\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_infinite_wave_train.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_linear_ramp.png}
+\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_linear_ramp.pdf}
 \caption{
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{figures/fig_modulation_ramp_timeseries.png}
+  \includegraphics[width=0.8\textwidth]{figures/fig_modulation_ramp_timeseries.pdf}
   \caption{
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/inhomogeneity_tendencies.png}
+  \includegraphics[width=\textwidth]{figures/inhomogeneity_tendencies.pdf}
   \caption{
@@
 \centering
-\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_wave_group.png}
+\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_wave_group.pdf}
 \caption{

### 2024-10-15 scripts/plot_analytical_solutions_with_phase.py

@@ -106,12 +106,12 @@ def plot_analytical_solutions(a_L: float):
     fig.suptitle(r"$\varepsilon_L$" + f" = {a_L * k_L}", fontsize=18)
 
     plt.subplots_adjust(left=0.1, right=0.95, top=0.9, bottom=0.05)
-    plt.savefig(f"../figures/fig_analytical_solutions_ak{a_L * k_L}.png", dpi=200)
+    plt.savefig(f"../figures/fig_analytical_solutions_ak{a_L * k_L}.pdf")
     plt.close(fig)
 
 
 if __name__ == "__main__":
     plot_analytical_solutions(0.1)
-    plot_analytical_solutions(0.2)
+    # plot_analytical_solutions(0.2)
     # plot_analytical_solutions(0.3)
-    plot_analytical_solutions(0.4)
+    # plot_analytical_solutions(0.4)

### 2024-10-15 scripts/plot_effective_gravity.py

@@ -125,7 +125,7 @@ def plot_effective_gravities(a: float):
     )
 
     plt.subplots_adjust(left=0.12, right=0.95, bottom=0.05, top=0.95)
-    plt.savefig("../figures/fig_effective_gravities.png", dpi=150)
+    plt.savefig("../figures/fig_effective_gravities.pdf")
     plt.close()
 
     g1_mean = []
@@ -250,7 +250,7 @@ def plot_effective_gravities(a: float):
     ax.text(0.15, 1.3, "Max. (troughs)", color="black", fontsize=16)
     ax.text(0.25, 0.4, "Min. (crests)", color="black", fontsize=16)
 
-    plt.savefig("../figures/fig_effective_gravities_mean_max_min.png", dpi=150)
+    plt.savefig("../figures/fig_effective_gravities_mean_max_min.pdf")
     plt.close()
 
 

### 2024-10-15 scripts/plot_inhomogeneity_tendencies.py

@@ -96,5 +96,5 @@ axes[1].text(
 
 plt.subplots_adjust(left=0.1, right=0.95, bottom=0.1, top=0.95)
 
-plt.savefig("../figures/inhomogeneity_tendencies.png", dpi=200)
+plt.savefig("../figures/inhomogeneity_tendencies.pdf")
 plt.close()

### 2024-10-15 scripts/plot_modulation_contributors.py

@@ -1,0 +1,57 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+import matplotlib
+
+matplotlib.rcParams["font.size"] = 14
+
+ak = np.arange(0, 0.41, 0.01)
+
+k_mod = np.exp(ak * np.exp(ak)) ** 1.25
+N_mod = np.exp(ak * np.exp(ak)) ** 0.5
+g_mod = (1 - ak * np.exp(ak)) ** (-0.25)
+
+ak_mod = g_mod * k_mod * N_mod
+
+k_mod_log = np.log(k_mod)
+N_mod_log = np.log(N_mod)
+g_mod_log = np.log(g_mod)
+ak_mod_log = k_mod_log + N_mod_log + g_mod_log
+
+# Create a figure with two subplots side by side
+fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(12, 5))
+
+# First subplot
+ax1.plot(ak, k_mod, label=r"$(\widetilde{k}/k)^{\frac{5}{4}}$", lw=2)
+ax1.plot(ak, N_mod, label=r"$(\widetilde{N}/N)^{\frac{1}{2}}$", lw=2)
+ax1.plot(ak, g_mod, label=r"$(\widetilde{g}/g)^{-\frac{1}{4}}$", lw=2)
+ax1.plot(ak, ak_mod, "k-", label=r"$(\widetilde{ak}/(ak))$", lw=2)
+ax1.legend()
+ax1.set_title("Modulation Factors")
+ax1.set_xlabel(r"$\varepsilon_L$")
+ax1.set_ylabel("Modulation")
+ax1.set_xlim(0, 0.4)
+ax1.set_ylim(1, 4)
+ax1.text(0, 1.02, "(a)", transform=ax1.transAxes, va="bottom", ha="left")
+ax1.grid(True, linestyle="--", alpha=0.7)
+
+# Second subplot (modified to show percentages)
+k_contrib_percent = 100 * k_mod_log / ak_mod_log
+N_contrib_percent = 100 * N_mod_log / ak_mod_log
+g_contrib_percent = 100 * g_mod_log / ak_mod_log
+
+ax2.plot(ak, k_contrib_percent, label="Wavenumber", lw=2)
+ax2.plot(ak, N_contrib_percent, label="Wave action", lw=2)
+ax2.plot(ak, g_contrib_percent, label="Effective gravity", lw=2)
+ax2.legend()
[…]

### 2024-10-15 scripts/plot_numerical_simulations.py

@@ -13,20 +13,28 @@ def plot_modulation_3panel(model, contour_levels, title):
     fig = plt.figure(figsize=(8, 6))
     axes = fig.subplots(1, 3, sharey=True)
 
+    # contourf 3 times to alleviate the antialiasing issue with matplotlib emitting to PDF
+    axes[0].contourf(ds.space, time, ds.wave_action, contour_levels)
+    axes[0].contourf(ds.space, time, ds.wave_action, contour_levels)
     axes[0].contourf(ds.space, time, ds.wave_action, contour_levels)
     axes[0].contour(ds.space, time, ds.wave_action, [1], colors="k", linewidths=0.5)
     axes[0].set_ylabel("Time (long-wave periods)")
     axes[0].set_title(r"$N$")
 
     axes[1].contourf(ds.space, time, ds.wavenumber / model.k_short, contour_levels)
+    axes[1].contourf(ds.space, time, ds.wavenumber / model.k_short, contour_levels)
+    axes[1].contourf(ds.space, time, ds.wavenumber / model.k_short, contour_levels)
     axes[1].contour(
         ds.space, time, ds.wavenumber / model.k_short, [1], colors="k", linewidths=0.5
     )
     axes[1].set_title(r"$k$")
 
-    cm = axes[2].contourf(
-        ds.space, time, ak / model.k_short / model.a_short, contour_levels
-    )
+    ak_modulation = ak / model.k_short / model.a_short
+    ak_modulation = np.clip(ak_modulation, contour_levels[0], contour_levels[-1])
+
+    cm = axes[2].contourf(ds.space, time, ak_modulation, contour_levels)
+    cm = axes[2].contourf(ds.space, time, ak_modulation, contour_levels)
+    cm = axes[2].contourf(ds.space, time, ak_modulation, contour_levels)
     axes[2].contour(
         ds.space,
         time,
@@ -41,7 +49,7 @@ def plot_modulation_3panel(model, contour_levels, title):
         ax.set_yticks(range(0, num_periods + 1, 1))
         ax.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
         ax.set_xticklabels([r"$0$", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
-        ax.set_xlabel("Phase")
+        ax.set_xlabel(r"$\Psi$")
 
[…]

### 2024-10-15 scripts/plot_numerical_simulations_ramp_properties.py

@@ -40,5 +40,5 @@ ax.grid()
 ax.set_xlim(0, 30)
 ax.set_ylim(1, 4.5)
 ax.plot([0, 30], [4, 4], "k--")
-plt.savefig("../figures/fig_modulation_ramp_timeseries.png", dpi=200)
+plt.savefig("../figures/fig_modulation_ramp_timeseries.pdf")
 plt.close()

### 2024-10-15 scripts/plot_numerical_vs_analytical_solutions.py

@@ -122,7 +122,7 @@ ax.set_ylabel(r"$\widetilde{k}/k$")
 ax.grid()
 ax.set_xlim(0, 0.405)
 ax.set_ylim(1, 3.5)
-plt.savefig("../figures/numerical_vs_analytical_solutions_k.png", dpi=200)
+plt.savefig("../figures/numerical_vs_analytical_solutions_k.pdf", dpi=200)
 plt.close()
 
 
@@ -167,7 +167,7 @@ ax.set_ylabel(r"$\widetilde{a}/a$")
 ax.grid()
 ax.set_xlim(0, 0.405)
 ax.set_ylim(1, 3)
-plt.savefig("../figures/numerical_vs_analytical_solutions_a.png", dpi=200)
+plt.savefig("../figures/numerical_vs_analytical_solutions_a.pdf")
 plt.close()
 
 fig = plt.figure(figsize=(8, 6))
@@ -211,5 +211,5 @@ ax.set_ylabel(r"$\widetilde{ak}/(ak)$")
 ax.grid()
 ax.set_xlim(0, 0.405)
 ax.set_ylim(1, 8)
-plt.savefig("../figures/numerical_vs_analytical_solutions_ak.png", dpi=200)
+plt.savefig("../figures/numerical_vs_analytical_solutions_ak.pdf")
 plt.close()

### 2024-10-15 scripts/plot_numerical_vs_analytical_solutions_with_phase.py

@@ -128,5 +128,5 @@ for ax in axes[-1]:
     ax.set_xticklabels(["0", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
     ax.set_xlabel(r"$\psi$")
 
-plt.savefig("../figures/numerical_vs_analytical_solutions_with_phase.png", dpi=200)
+plt.savefig("../figures/numerical_vs_analytical_solutions_with_phase.pdf")
 plt.close()

### 2024-10-15 paper.tex
Final edits for submission on 10/15/2024

@@
 forecasting.
-In this paper, we revisit the linear (Airy) wave theory and derive higher-order
+In this paper, we revisit the linear wave theory and derive higher-order
 steady solutions for the change of short-wave wavenumber, action density, and
@@
 realistic long-wave groups.
-The combined analytical and numerical framework presented here allows
-quantifying the importance of various terms in the full wave conservation
-equations.
 The proposed steady solutions are a good approximation of the fully
-nonlinear numerical solutions in long-wave steepness up to $\approx 0.2$.
-As a corollary, except for a subset of initial conditions, the solutions to
-the fully nonlinear crest-action conservation equations are mostly steady in
-the reference frame of the long waves.
+nonlinear numerical solutions in long-wave steepness up to $\approx$0.2.
+Except for a subset of initial conditions, the solutions to the fully nonlinear
+crest-action conservation equations are mostly steady in the reference frame of
+the long waves.
 \end{abstract}
@@
 
+\begin{table}
+\begin{center}
+\def~{\hphantom{0}}
+\begin{tabular}{cccccc}
+Reference   & Analysis   & Gravity & Nonlinear  & Nonlinear   & Solution \\
+            & frameworks & varies  & long waves & short waves &          \\
+\hline
+\citet{longuet1960changes} & Velocity potential  & No  & No  & No  & Analytical \\
+\hline
+\citet{phillips1981dispersion} & Crest and action & Yes & Yes & No  & Numerical \\
+                               & conservation     &     &     &     &           \\
+\hline
+\citet{longuet1987propagation} & Velocity potential & Yes & Yes & No  & Numerical \\
+\hline
+\citet{henyey1988energy} & Hamiltonian & Yes & Yes & No  & Analytical \\
+\hline
+\citet{zhang1990evolution} & Schr\" odinger & Yes & Yes & Yes & Numerical \\
+\hline
+\citet{peureux2021unsteady} & Crest and action & Yes & No  & No  & Numerical \\
+                            & conservation     &     &     &     &           \\
+\hline
+This paper  & Crest and action & Yes & Yes & No  & Analytical \\
+            & conservation     &     &     &     & and numerical \\
+\hline
+\end{tabular}
+\caption{
+  Non-exhaustive summary of prior and present (this paper) approaches to
+  calculating the modulation of short waves by long waves.
+}
+\label{table:modulation_literature_summary}
+\end{center}
+\end{table}
+
 In this paper, we revisit this problem from the perspective of linear wave theory
@@
 
-\begin{table}
-\begin{center}
-\def~{\hphantom{0}}
-\begin{tabular}{cccccc}
-Reference   & Analysis   & Gravity & Nonlinear  & Nonlinear   & Solution \\
-            & frameworks & varies  & long waves & short waves &          \\
-\hline
-\citet{longuet1960changes} & Velocity potential  & No  & No  & No  & Analytical \\
-\hline
-\citet{phillips1981dispersion} & Crest and action & Yes & Yes & No  & Numerical \\
-                               & conservation     &     &     &     &           \\
-\hline
-\citet{longuet1987propagation} & Velocity potential & Yes & Yes & No  & Numerical \\
-\hline
-\citet{henyey1988energy} & Hamiltonian & Yes & Yes & No  & Analytical \\
-\hline
-\citet{zhang1990evolution} & Schr\" odinger & Yes & Yes & Yes & Numerical \\
-\hline
-\citet{peureux2021unsteady} & Crest and action & Yes & No  & No  & Numerical \\
-                            & conservation     &     &     &     &           \\
-\hline
-This paper  & Crest and action & Yes & Yes & No  & Analytical \\
-            & conservation     &     &     &     & and numerical \\
-\hline
-\end{tabular}
-\caption{
-  Non-exhaustive summary of prior and present (this paper) approaches to
-  calculating the modulation of short waves by long waves.
-}
-\label{table:modulation_literature_summary}
-\end{center}
-\end{table}
-
 \section{Governing equations}
@@
 and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
-Inserting (\ref{eq:dispersion}) into (\ref{eq:wave_crests}) and noting that
-$\partial \sqrt{gk} / \partial x = 1/2 \sqrt{g/k} \partial k / \partial x + 1/2 \sqrt{k/g} \partial g / \partial x$,
+Inserting (\ref{eq:dispersion}) into (\ref{eq:wave_crests}) and noting that:
+
+\begin{equation}
+\label{eq:dispersion_derivative}
+\dfrac{\partial \sqrt{gk}}{\partial x} =
+\dfrac{1}{2} \sqrt{\dfrac{g}{k}} \dfrac{\partial k}{\partial x} + \dfrac{1}{2} \sqrt{\dfrac{k}{g}} \dfrac{\partial g}{\partial x}
+\end{equation}
 we get:
@@
 The wavenumber is derived by linearising the conservation of wave crests
-(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
+(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$ and $g$.
 The effective gravity is derived by projecting the long-wave induced surface
@@
 Going forward we will use the tilde over the variables to denote the modulated
-quantities (e.g. $\widetilde{k}$ for wavenumber).
+quantities (\textit{e.g.} $\widetilde{k}$ for wavenumber).
 
@@
   Milan Curcic was partly supported by the National Science Foundation
-  grant AGS-1745384, Transatlantic Research Partnership, and the Office of Naval
-  Research Coastal Land Air-Sea Interaction DRI.
+  grant AGS-1745384, Transatlantic Research Partnership by the French Embassy
+  in the United States and the FACE Foundation, and the Office of Naval Research
+  Coastal Land Air-Sea Interaction DRI.
 }

### 2024-10-15 scripts/plot_modulation_diagram.py
Emit diagram to PNG as well for the graphical abstract

@@ -167,4 +167,5 @@ ax.set_ylim(-1.5, 2)
 
 plt.tight_layout()
 plt.savefig("../figures/fig_hydrodynamic_modulation_diagram.pdf")
+plt.savefig("../figures/fig_hydrodynamic_modulation_diagram.png", dpi=300)
 plt.close()

### 2024-10-16 paper.tex
Fix typo

@@
 As \citet{zhang1990evolution} point out, this is appropriate to do when the
-amplitude of the long waves is of the similar magnitude or large than the
+amplitude of the long waves is of the similar magnitude or larger than the
 wavelength of the short waves.

### 2024-10-16 paper.tex
Style fix
Fix typo

@@
 Solving analytically for the wavenumber requires dropping the spatial derivatives
-of $k$ in Eq. \ref{eq:wavenumber} and assuming homogeneous $g$:
+of $k$ in (\ref{eq:wavenumber}) and assuming homogeneous $g$:
 
@@
 Although in this paper we only explored the solutions for $k_L = 1$ and $k=10$
-the results are not sensitive to different values of $k_L$, provided that the
-amplitude $a_L$ is set such that the long-wave steepness $\varepsilon_L = a_L / k_L$
-is the same.
+the results are not sensitive to different values of $k_L$, provided the same
+long-wave steepness $\varepsilon_L = a_L k_L$.
 Numerical simulations with $k/k_L = 10$ and $k/k_L = 100$ show little

### 2024-10-19 paper.tex
Fix typo
Fix typos

@@
 each long-wave passage.
-We show that this unsteady growth only occurs only for homogeneous initial
+We show that this unsteady growth only occurs for homogeneous initial
 conditions as a special case and does not generally occur, for example in more
@@
 Reference   & Analysis   & Gravity & Nonlinear  & Nonlinear   & Solution \\
-            & frameworks & varies  & long waves & short waves &          \\
+            & framework & varies  & long waves & short waves &          \\
 \hline
@@
 \frac{\widetilde{g}}{g}
-  = \cos{\alpha} 
+  = \cos{\alpha_{St}}
   + \dfrac{\partial W_{z=\eta_{St}}}{\partial t} \cos{\alpha_{St}}

### 2024-10-20 paper.tex
Fix typos

@@
 (the importance of which we discuss in more detail in the next subsection).
-At $\varepsilon = 0.1$, the numerical solutions for the wavenumber, amplitude,
+At $\varepsilon_L = 0.1$, the numerical solutions for the wavenumber, amplitude,
 and steepness modulation are all remarkably similar to the analytical solutions
 (Fig. \ref{fig:numerical_solutions}a,d,g).
-This suggests that at low long-wave steepness that the steady analytical solution
+This suggests that for low long-wave steepness the steady analytical solution
 is a reasonable proxy for the fully-nonlinear crest-action conservation
@@
 Like in the case of a linear ramp, the modulation pattern here is locked in to
-the long-wave crest and peaks at approximately 1.2 for short-wave steepness.
+the long-wave crest and peaks at approximately 1.2 for short-wave steepness of 0.1.
 After the peak of the long-wave group passes, the short-wave modulation recedes

### 2025-02-09 scripts/plot_velocity_error.py
Figures to quantify surface velocity error

@@ -1,0 +1,134 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from ssgw import SSGW
+from twowave import elevation, orbital_horizontal_velocity
+
+import matplotlib
+
+matplotlib.rcParams.update({"font.size": 14})
+
+g = 9.8
+k = 1
+a = 0.4
+ak = a * k
+d = np.inf
+omega = np.sqrt(g * k * np.tanh(k * d))
+t = 0
+
+# Fully nonlinear solution
+wave = SSGW(d, ak)
+
+phase, eta = np.real(wave.zs), np.imag(wave.zs)
+x = phase / k
+U = (np.real(wave.ws) + wave.ce) * np.sqrt(g)
+
+eta_linear = elevation(x, t, a, k, omega, wave_type="linear")
+eta_stokes = elevation(x, t, a, k, omega, wave_type="stokes")
+
+U_linear = orbital_horizontal_velocity(x, eta_linear, t, a, k, omega)
+U_stokes = orbital_horizontal_velocity(x, eta_stokes, t, a, k, omega)
+
+
+fig, (ax1, ax2, ax3) = plt.subplots(3, 1, figsize=(8, 9), sharex=True)
+
+# Top panel: Surface elevation
+ax1.plot(phase, eta_linear, label="Linear")
+ax1.plot(phase, eta_stokes, label="3$^{rd}$ order Stokes")
+ax1.plot(phase, eta, color="k", label="Fully nonlinear")
+ax1.axhline(y=0, color="k", linestyle="--", alpha=0.5)
+ax1.set_ylabel("Surface elevation (m)")
+ax1.legend()
+ax1.grid(True)
+
+# Middle panel: Velocities
+ax2.plot(phase, U_linear, label="Linear")
+ax2.plot(phase, U_stokes, label="3$^{rd}$ order Stokes")
+ax2.plot(phase, U, color="k", label="Fully nonlinear")
+ax2.axhline(y=0, color="k", linestyle="--", alpha=0.5)
+ax2.set_ylabel("Horizontal velocity (m/s)")
+ax2.grid(True)
+
+# Middle panel: Absolute errors
+error_linear = U_linear - U
+error_stokes = U_stokes - U
+
+ax3.plot(phase, error_linear, label="Linear")
+ax3.plot(phase, error_stokes, label="3$^{rd}$ order Stokes")
+ax3.axhline(y=0, color="k", linestyle="--", alpha=0.5)
[…]

### 2025-02-11 scripts/plot_action_conservation_criteria.py
Compute wave action conservation criteria

@@ -1,0 +1,87 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from twowave import WaveModulationModel, diff
+
+import matplotlib
+
+matplotlib.rcParams["font.size"] = 14
+
+
+def get_homogeneity_and_stationarity(a_long):
+    m = WaveModulationModel(a_long=a_long)
+    m.run(ramp_type="groups")
+    ds = m.to_xarray()
+
+    x, t = ds.space.values, ds.time.values
+    dx, dt = np.diff(x)[0], np.diff(t)[0]
+
+    # Homogeneity criteria
+    n, _ = np.unravel_index(np.argmax(ds.wave_action.values), ds.wave_action.shape)
+    k_h = ds.wavenumber.values[n]
+    g_h = ds.gravitational_acceleration.values[n]
+    N_h = ds.wave_action.values[n]
+
+    H_k = 1 - np.abs(diff(k_h)) / dx / k_h / k_h
+    H_g = 1 - np.abs(diff(g_h)) / dx / g_h / k_h
+    H_N = 1 - np.abs(diff(N_h)) / dx / N_h / k_h
+
+    # Stationarity criteria
+    k_s = ds.wavenumber.values[:, 0]
+    g_s = ds.gravitational_acceleration.values[:, 0]
+    N_s = ds.wave_action.values[:, 0]
+    sigma = np.sqrt(g_s * k_s)
+
+    S_k = 1 - np.abs(diff(k_s)) / dt / k_s / sigma
+    S_g = 1 - np.abs(diff(g_s)) / dt / g_s / sigma
+    S_N = 1 - np.abs(diff(N_s)) / dt / N_s / sigma
+
+    return (
+        np.min(H_k),
+        np.min(H_g),
+        np.min(H_N),
+        np.min(S_k),
+        np.min(S_g),
+        np.min(S_N),
+    )
+
+
+ak_values = np.arange(0, 0.41, 0.01)
+H_k_values = np.zeros_like(ak_values)
+H_g_values = np.zeros_like(ak_values)
+H_N_values = np.zeros_like(ak_values)
+
+S_k_values = np.zeros_like(ak_values)
+S_g_values = np.zeros_like(ak_values)
+S_N_values = np.zeros_like(ak_values)
+
+for i, ak in enumerate(ak_values):
+    (
+        H_k_values[i],
+        H_g_values[i],
+        H_N_values[i],
[…]

### 2025-02-15 scripts/plot_action_conservation_criteria.py
Homogeneity and stationarity from analytical and numerical solutions

@@ -7,9 +7,9 @@ import matplotlib
 matplotlib.rcParams["font.size"] = 14
 
 
-def get_homogeneity_and_stationarity(a_long):
-    m = WaveModulationModel(a_long=a_long)
-    m.run(ramp_type="groups")
+def get_homogeneity_and_stationarity(a_long, k_long, k_short):
+    m = WaveModulationModel(a_long=a_long, k_long=k_long, k_short=k_short)
+    m.run(wave_type="stokes", ramp_type="groups")
     ds = m.to_xarray()
 
     x, t = ds.space.values, ds.time.values
@@ -45,43 +45,413 @@ def get_homogeneity_and_stationarity(a_long):
     )
 
 
+def homogeneity_k(k0, k_L, ak, phase):
+    return 1 - np.abs(k_L / k0 * ak * np.sin(phase) / (1 + ak * np.cos(phase)) ** 2)
+
+
+def homogeneity_g(k0, k_L, ak, phase):
+    return 1 - np.abs(k_L / k0 * ak * np.sin(phase) / (1 - ak**2 * np.cos(phase) ** 2))
+
+
+def stationarity_k(sigma0, sigma_L, ak, phase):
+    return 1 - np.abs(
+        sigma_L
+        / sigma0
+        * ak
+        * np.sin(phase)
+        / (1 + ak * np.cos(phase))
+        / np.sqrt(1 - ak**2 * np.cos(phase) ** 2)
+    )
+
+
+def stationarity_g(sigma0, sigma_L, ak, phase):
+    return 1 - np.abs(
+        sigma_L
+        / sigma0
+        * ak
+        * np.sin(phase)
+        / (1 - ak * np.cos(phase))
+        / np.sqrt(1 - ak**2 * np.cos(phase) ** 2)
+    )
+
+
+# Numerical model results
+ak_values = np.arange(0, 0.42, 0.02)
+r_values = np.logspace(1, 2, 10)
+k_L = 1
+H_k_values = np.zeros((len(r_values), len(ak_values)))
+H_g_values = np.zeros((len(r_values), len(ak_values)))
+H_N_values = np.zeros((len(r_values), len(ak_values)))
+
+S_k_values = np.zeros((len(r_values), len(ak_values)))
+S_g_values = np.zeros((len(r_values), len(ak_values)))
+S_N_values = np.zeros((len(r_values), len(ak_values)))
+
+for i, r in enumerate(r_values):
[…]

### 2025-02-15 paper.tex
Compute wave action conservation criteria
WIP wave action conservation limits

@@
 for the discussions that helped improve this paper.
+I also thank three anonymous reviewers for their helpful comments.
 }
@@
 
+\appendix
+\section{Wave action conservation requirements}
+\label{appendix:wave_action_conservation}
+
+Following the variational approach by \citet{whitham1965general},
+\citet{bretherton1968wavetrains} showed that the wave action is conserved for
+small amplitude (linear) waves propagating in slowly-varying media.
+Although \citet{longuet1987propagation} correctly pointed out both requirements
+(that the medium is slowly-varying and that the short-wave properties are
+approximately homogeneous and stationary) for the wave action conservation to
+hold, he incorrectly stated that there is no explicit restriction on the
+steepness of the long wave, and that it appears necessary to only assume
+small-amplitude short waves and significant separation between $k$ and $k_L$.
+We will show that in the general case, both the homogeneity and the stationarity
+of a short wave quantity $q$ are dependent on the long-wave steepness
+$\varepsilon_L$.
+
+The homogeneity and stationarity of a scalar quantify $q$ can be expressed as
+conditions on the instantaneous fractional rate of change of $q$ being much
+smaller than the short-wave inverse scales:
+
+\begin{equation}
+  \frac{1}{q} \frac{\partial q}{\partial x} \ll k
+  \label{eq:homogeneity_condition}
+\end{equation}
+
+\begin{equation}
+  \frac{1}{q} \frac{\partial q}{\partial t} \ll \sigma
+  \label{eq:stationarity_condition}
+\end{equation}
+The slowly-varying conditions on the medium (in our case, the long-wave orbital
+velocity $U$) are then
+\begin{equation}
+  \frac{1}{U} \frac{\partial U}{\partial x} \ll \widetilde{k},
+\end{equation}
+and
+
+\begin{equation}
+  \frac{1}{U} \frac{\partial U}{\partial t} \ll \widetilde{\sigma},
+\end{equation}
+respectively.
+For a linear long wave, these conditions are equivalent to the scale separation
+between the long and short waves:
+
+\begin{equation}
+  k_L \ll \widetilde{k}
+  \label{eq:k_scale_separation}
+\end{equation}
+
+\begin{equation}
+  \sigma_L \ll \widetilde{\sigma}
+\end{equation}
+\citet{bretherton1968wavetrains} also require that the medium is nearly uniform
+in the vertical, over the vertical scales of motion of the short waves.
+For the ambient velocity being the orbital velocity of the long wave, this is
+trivially satisfied with (\ref{eq:k_scale_separation}).
+
+Now, to address the variation of short-wave properties,
+we apply (\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}) to
+the long-wave phase dependent wavenumber, gravitational acceleration, and
+wave action, and define the homogeneity and stationarity quantities as:
+
+\begin{equation}
+  H_q = 1 - \left| \frac{1}{qk} \frac{\partial q}{\partial x} \right|
+\end{equation}
+
+\begin{equation}
+  S_q = 1 - \left| \frac{1}{q\sigma} \frac{\partial q}{\partial t} \right|
+\end{equation}
+With the above definitions, we consider $q$ to be homogeneous and stationary as
+$H_q \rightarrow 1$ and $S_q \rightarrow 1$, respectively.
+
+For the linearized modulation solutions (\ref{eq:k_short_exact}),
+(\ref{eq:gravity_modulation_linear_no_curvature_mean_level}), and
+(\ref{eq:wave_action_modulation}), the respective homogeneity expressions are:
+
+\begin{equation}
+  H_{k,N} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right)^2} \right|
+\end{equation}
+
+\begin{equation}
+  H_{g} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L^2 \cos^2{\psi}\right)} \right|
+\end{equation}
+
+\begin{equation}
+  S_{k,N} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
+\end{equation}
+
+\begin{equation}
+  S_{g} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
+\end{equation}
+
+\section{Surface velocity errors}
+\label{appendix:surface_velocity_errors}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/fig_velocity_error_by_phase.pdf}
+\caption{
+}
+\label{fig:velocity_error_by_phase}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/fig_relative_velocity_error.pdf}
+\caption{
+}
+\label{fig:relative_velocity_error}
+\end{figure}
+
 \bibliographystyle{jfm}

### 2025-02-16 scripts/plot_action_conservation_criteria.py
WIP homogeneity and stationarity figures

@@ -9,7 +9,7 @@ matplotlib.rcParams["font.size"] = 14
 
 def get_homogeneity_and_stationarity(a_long, k_long, k_short):
     m = WaveModulationModel(a_long=a_long, k_long=k_long, k_short=k_short)
-    m.run(wave_type="stokes", ramp_type="groups")
+    m.run(wave_type="linear", ramp_type="groups")
     ds = m.to_xarray()
 
     x, t = ds.space.values, ds.time.values
@@ -101,32 +101,6 @@ for i, r in enumerate(r_values):
             S_N_values[i, j],
         ) = get_homogeneity_and_stationarity(ak, k_long=k_L, k_short=k_L * r)
 
-
-ak_values = np.arange(0, 0.42, 0.02)
-r_values = np.logspace(1, 2, 10)
-k_L = 1
-H_k_values = np.zeros((len(r_values), len(ak_values)))
-H_g_values = np.zeros((len(r_values), len(ak_values)))
-H_N_values = np.zeros((len(r_values), len(ak_values)))
-
-S_k_values = np.zeros((len(r_values), len(ak_values)))
-S_g_values = np.zeros((len(r_values), len(ak_values)))
-S_N_values = np.zeros((len(r_values), len(ak_values)))
-
-for i, r in enumerate(r_values):
-    for j, ak in enumerate(ak_values):
-        print(
-            f"Processing r={r:.2f}, ak={ak:.2f} ({i + 1}/{len(r_values)}, {j + 1}/{len(ak_values)})"
-        )
-        (
-            H_k_values[i, j],
-            H_g_values[i, j],
-            H_N_values[i, j],
-            S_k_values[i, j],
-            S_g_values[i, j],
-            S_N_values[i, j],
-        ) = get_homogeneity_and_stationarity(ak, k_long=k_L, k_short=k_L * r)
-
 # Clip stationarity values to minimum of 0.5
 S_k_values = np.clip(S_k_values, 0.5, None)
 S_g_values = np.clip(S_g_values, 0.5, None)
@@ -134,175 +108,114 @@ S_N_values = np.clip(S_N_values, 0.5, None)
 
 
 # Create 6-panel figure for numerical results
-fig, ((ax1, ax2, ax3), (ax4, ax5, ax6)) = plt.subplots(2, 3, figsize=(15, 10))
[…]

### 2025-02-17 paper.tex
WIP appendices

@@
   H_{k,N} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right)^2} \right|
+  \label{eq:homogeneity_k}
 \end{equation}
@@
   H_{g} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L^2 \cos^2{\psi}\right)} \right|
+  \label{eq:homogeneity_g}
 \end{equation}
@@
   S_{k,N} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
+  \label{eq:stationarity_k}
 \end{equation}
@@
   S_{g} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
-\end{equation}
+  \label{eq:stationarity_g}
+\end{equation}
+Notice that the steady, linearised solutions depend on both the the wavenumber
+ratio $k_L/k$ and the long-wave steepness $\varepsilon_L$.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_homogeneity_stationarity_analytical.pdf}
+\caption{
+Homogeneity (left) and stationarity (right) of the short-wave wavenumber (top)
+and effective gravity (bottom) as a function of the long-wave steepness
+$\varepsilon_L$ and the wavenumber ratio $k/k_L$, based on the linearised
+solutions (\ref{eq:homogeneity_k}-\ref{eq:stationarity_g}).
+}
+\label{fig:homogeneity_stationarity_analytical}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_homogeneity_stationarity_numerical.pdf}
+\caption{
+Homogeneity (left) and stationarity (right) of the short-wave wavenumber (top),
+action (middle), and effective gravity (bottom) as a function of the long-wave
+steepness $\varepsilon_L$ and the wavenumber ratio $k/k_L$, based on the
+numerical solutions of the full wave crest and action conservation equations.
+Note that the color ranges are different than those in Fig. \ref{fig:homogeneity_stationarity_analytical}.
+}
+\label{fig:homogeneity_stationarity_numerical}
+\end{figure}
 

### 2025-02-17 scripts/plot_action_conservation_criteria.py

@@ -108,8 +108,8 @@ S_N_values = np.clip(S_N_values, 0.5, None)
 
 
 # Create 6-panel figure for numerical results
-fig, ((ax1, ax2), (ax3, ax4), (ax5, ax6)) = plt.subplots(3, 2, figsize=(10, 15))
-fig.subplots_adjust(hspace=0.3, wspace=0.3)
+fig, ((ax1, ax2), (ax3, ax4), (ax5, ax6)) = plt.subplots(3, 2, figsize=(12, 15))
+fig.subplots_adjust(hspace=0.2, wspace=0.2)
 axes = [ax1, ax2, ax3, ax4, ax5, ax6]
 
 # Common plotting parameters
@@ -262,7 +262,7 @@ fig.subplots_adjust(hspace=0.2, wspace=0.2)
 axes = [ax1, ax2, ax3, ax4]
 
 # Common plotting parameters
-levels = np.arange(0.85, 1.0, 0.01)
+levels = np.linspace(0.85, 1.0, 16)
 cmap = plt.cm.viridis
 
 # Plot homogeneity and stationarity contours

### 2025-02-20 paper.tex
Add ref to Clamond & Dutykh

@@
 \begin{equation}
-\label{eq:w_L}
+\label{eq:W_L}
 W =
@@
 (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are the governing equations used
-in this paper, with (\ref{eq:U_L})-(\ref{eq:w_L}) describing the long-wave
+in this paper, with (\ref{eq:U_L})-(\ref{eq:W_L}) describing the long-wave
 induced ambient velocity forcing for the short waves. 
@@
 
+The long-wave induced velocities (\ref{eq:phi})-(\ref{eq:W_L}) are accurate to
+$\mathcal{O}(\varepsilon_L^3)$ when evaluated at the mean water level.
+However, in this paper we depart from the mean water level and evaluate the
+velocities at the surface of the long wave.
+This introduces an additional error at the order of $\mathcal{O}(\varepsilon_L^2)$.
+To quantify this error exactly, we compare the velocities evaluated at the
+surface for a linear (small-amplitude) wave and a third-order Stokes wave,
+with the surface velocity of a fully nonlinear wave.
+The nonlinear wave surface velocity is calculated numerically using the method
+by \citet{clamond2018accurate}.
+
 \begin{figure}

### 2025-02-20 scripts/plot_velocity_error.py
Update velocity error plot

@@ -79,6 +79,8 @@ a_omega = []
 ak_values = np.arange(0, 0.41, 0.01)
 max_error_linear = []
 max_error_stokes = []
+mean_error_linear = []
+mean_error_stokes = []
 
 for ak in ak_values:
     a = ak / k
@@ -106,29 +108,49 @@ for ak in ak_values:
     max_error_linear.append(error_linear)
     max_error_stokes.append(error_stokes)
 
+    mean_error_linear.append(np.mean(np.abs(U_linear - U)))
+    mean_error_stokes.append(np.mean(np.abs(U_stokes - U)))
 
 # Plot errors vs steepness
 fig = plt.figure(figsize=(8, 6))
 ax = fig.add_subplot(111)
-ax.semilogy(
+ax.plot(
     ak_values,
-    np.array(max_error_linear) / a_omega,
+    np.array(max_error_linear),
+    lw=2,
     color="tab:blue",
-    lw=2,
-    label="Linear wave",
+    label="1$^{st}$ order, max.",
 )
-ax.semilogy(
+ax.plot(
     ak_values,
-    np.array(max_error_stokes) / a_omega,
-    color="tab:orange",
+    np.array(mean_error_linear),
     lw=2,
-    label=r"3$^{rd}$ order Stokes wave",
+    color="tab:blue",
+    linestyle="--",
+    label="1$^{st}$ order, mean",
 )
-ax.set_xlabel("Wave steepness (ak)")
-ax.set_ylabel("Surface velocity relative error")
+ax.plot(
+    ak_values,
+    np.array(max_error_stokes),
+    lw=2,
+    color="tab:orange",
+    label="3$^{rd}$ order, max.",
+)
+ax.plot(
+    ak_values,
+    np.array(mean_error_stokes),
+    lw=2,
+    color="tab:orange",
+    linestyle="--",
+    label="3$^{rd}$ order, mean",
+)
+ax.plot(ak_values, ak_values**2, "k--", lw=2, label=r"$\varepsilon_L^2$")
+ax.plot(ak_values, ak_values**3, "k:", lw=2, label=r"$\varepsilon_L^3$")
+ax.set_xlabel(r"$\varepsilon_L$")
+ax.set_ylabel("Surface velocity error")
 ax.legend()
 ax.set_xlim(0, 0.4)
+ax.set_ylim(0, 0.3)
 ax.grid()
-
[…]

### 2025-02-20 paper.tex
Add section on wave action limits

@@
 upwards.
-(\ref{eq:phi}) is accurate to the third order in $\varepsilon_L$ in deep water
-because the second- and third-order terms in the Stokes expansion series are
-exactly zero.
+When evaluated at the mean water level ($z=0$), (\ref{eq:phi}) is accurate to
+the third order in $\varepsilon_L$ in deep water because the second- and
+third-order terms in the Stokes expansion series are exactly zero.
 The long-wave induced horizontal and vertical orbital velocities are:
@@
 wavelength of the short waves.
+However, evaluating these velocities at the free surface $z = \eta_L$ introduces
+an additional error of $\mathcal{O}(\varepsilon_L^2)$.
+The surface velocity errors are described in more detail in Appendix
+\ref{appendix:surface_velocity_errors}.
 
@@
 
+\subsection{Limits to the wave action conservation}
+\label{appendix:wave_action_conservation}
+
+Following the variational approach by \citet{whitham1965general},
+\citet{bretherton1968wavetrains} showed that the wave action is conserved for
+small-amplitude (linear) waves that propagate in slowly-varying media.
+\citet{longuet1987propagation} correctly pointed out both
+requirements--that both the medium and the short wavetrain must be
+slowly-varying--for the wave action conservation to hold.
+However, he incorrectly stated that there is no explicit restriction on the
+steepness of the long wave, and that it appears necessary to only assume
+small-amplitude short waves and significant scale separation ($k/k_L$).
+The restriction on $\varepsilon_L$ becomes readily apparent when we recognize
+that the long wave-induced orbital velocity scales with $\varepsilon_L$, and
+that the divergence of this velocity is the dominant term in (\ref{eq:wavenumber})
+and (\ref{eq:wave_action2}).
+Here we show that in the general case both the homogeneity and the stationarity
+of a short wave quantity $q$ depend on the long-wave steepness $\varepsilon_L$.
+
+The homogeneity and stationarity of a scalar quantify $q$ can be expressed as
+conditions on the instantaneous fractional rate of change of $q$ being much
+smaller than the short-wave inverse scales:
+
+\begin{equation}
+  \frac{1}{q} \frac{\partial q}{\partial x} \ll k
+  \label{eq:homogeneity_condition}
+\end{equation}
+
+\begin{equation}
+  \frac{1}{q} \frac{\partial q}{\partial t} \ll \sigma
+  \label{eq:stationarity_condition}
+\end{equation}
+The slowly-varying conditions on the medium (in our case, the long-wave orbital
+velocity $U$) are then
+\begin{equation}
+  \frac{1}{U} \frac{\partial U}{\partial x} \ll \widetilde{k},
+\end{equation}
+and
+
+\begin{equation}
+  \frac{1}{U} \frac{\partial U}{\partial t} \ll \widetilde{\sigma},
+\end{equation}
+respectively.
+For a linear long wave, these conditions are equivalent to the scale separation
+between the long and short waves:
+
+\begin{equation}
+  k_L \ll \widetilde{k}
+  \label{eq:k_scale_separation}
+\end{equation}
+
+\begin{equation}
+  \sigma_L \ll \widetilde{\sigma}
+\end{equation}
+\citet{bretherton1968wavetrains} also require that the medium is nearly uniform
+over the vertical scales of motion of the short waves.
+For the ambient velocity being the orbital velocity of the long wave, this is
+trivially satisfied with (\ref{eq:k_scale_separation}).
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_homogeneity_stationarity_analytical.pdf}
+\caption{
+Homogeneity (left) and stationarity (right) of the short-wave wavenumber (top)
+and effective gravity (bottom) as a function of the long-wave steepness
+$\varepsilon_L$ and the wavenumber ratio $k/k_L$, based on the linearised
+solutions (\ref{eq:homogeneity_k}-\ref{eq:stationarity_g}).
+Solid and dashed lines highlight the 0.99 and 0.90 values, respectively.
+}
+\label{fig:homogeneity_stationarity_analytical}
+\end{figure}
+
+Now, to address the variation of short-wave properties,
+we apply (\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}) to
+the long-wave phase dependent wavenumber, gravitational acceleration, and
+wave action, and define the homogeneity and stationarity quantities as:
+
+\begin{equation}
+  H_q = 1 - \left| \frac{1}{qk} \frac{\partial q}{\partial x} \right|
+\end{equation}
+
+\begin{equation}
+  S_q = 1 - \left| \frac{1}{q\sigma} \frac{\partial q}{\partial t} \right|
+\end{equation}
+With the above definitions, we consider $q$ to be homogeneous and stationary as
+$H_q \rightarrow 1$ and $S_q \rightarrow 1$, respectively.
+
+For the linearized modulation solutions (\ref{eq:k_short_exact}),
+(\ref{eq:gravity_modulation_linear_no_curvature_mean_level}), and
+(\ref{eq:wave_action_modulation}), the respective homogeneity expressions are:
+
+\begin{equation}
+  H_{k,N} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right)^2} \right|
+  \label{eq:homogeneity_k}
+\end{equation}
+
+\begin{equation}
+  H_{g} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L^2 \cos^2{\psi}\right)} \right|
+  \label{eq:homogeneity_g}
+\end{equation}
+
+\begin{equation}
+  S_{k,N} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
+  \label{eq:stationarity_k}
+\end{equation}
+
+\begin{equation}
+  S_{g} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
+  \label{eq:stationarity_g}
+\end{equation}
+Notice that the steady, linearised solutions depend on both the the wavenumber
+ratio $k_L/k$ and the long-wave steepness $\varepsilon_L$.
+
+Fig. \ref{fig:homogeneity_stationarity_analytical} shows the homogeneities and
+stationarities of the short-wave wavenumber and effective gravity based on
+the steady solutions and their derived criteria as minima of
+(\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}).
+Homogeneity is a weaker requirement than stationarity.
+At the lowest scale separation considered ($k/k_L = 10$), short-wave wavenumber,
+action, and effective gravity are all strongly homogeneous ($H_{k,N,g} > 0.99$).
+However, the solutions are only strongly stationary ($S_{k,N,g} > 0.99$) at a
+high scale separation of $k/k_L = 100$ for $\varepsilon_L < 0.1$.
+If we consider weakly stationary conditions as $S_{k,N,g} > 0.9$, then the
+solutions satisfy the requirements for the wave action conservation for
+$\varepsilon_L < 0.3$ at $k/k_L = 10$, and for $\varepsilon_L < 0.4$ at
+$k/k_L > 20$.
+Although \citet{bretherton1968wavetrains} state that the wave action balance
+is valid to the first order in $k_L/k$, applying the general criteria
+(\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}) yields an
+error of $\mathcal{O}(\varepsilon_L k/k_L)$.
+We will revisit these criteria based on the numerical solutions of nonlinear
+wave action and crest conservation equations in the next section.
+
 \section{Numerical solutions}
@@
 \label{fig:numerical_vs_analytical_ak}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_homogeneity_stationarity_numerical.pdf}
+\caption{
+Homogeneity (left) and stationarity (right) of the short-wave wavenumber (top),
+action (middle), and effective gravity (bottom) as a function of the long-wave
+steepness $\varepsilon_L$ and the wavenumber ratio $k/k_L$, based on the
+numerical solutions of the full wave crest and action conservation equations.
+Note that the color ranges are different than those in Fig. \ref{fig:homogeneity_stationarity_analytical}.
+}
+\label{fig:homogeneity_stationarity_numerical}
 \end{figure}
@@
 I am thankful to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan
-for the discussions that helped improve this paper.
-I also thank three anonymous reviewers for their helpful comments.
+for fruitful discussions.
+I also thank three anonymous reviewers for their expertise that helped improve
+this paper.
 }
@@
   \url{https://github.com/wavesgroup/2wave}.
+  A Python implementation of SSGW by \citet{clamond2018accurate} is available
+  at \url{https://github.com/wavesgroup/ssgw}.
   %See JFM's \href{https://www.cambridge.org/core/journals/journal-of-fluid-mechanics/information/journal-policies/research-transparency}{research transparency policy} for more information
@@
 \appendix
-\section{Wave action conservation requirements}
-\label{appendix:wave_action_conservation}
-
-Following the variational approach by \citet{whitham1965general},
-\citet{bretherton1968wavetrains} showed that the wave action is conserved for
-small amplitude (linear) waves propagating in slowly-varying media.
-Although \citet{longuet1987propagation} correctly pointed out both requirements
-(that the medium is slowly-varying and that the short-wave properties are
-approximately homogeneous and stationary) for the wave action conservation to
-hold, he incorrectly stated that there is no explicit restriction on the
-steepness of the long wave, and that it appears necessary to only assume
-small-amplitude short waves and significant separation between $k$ and $k_L$.
-We will show that in the general case, both the homogeneity and the stationarity
-of a short wave quantity $q$ are dependent on the long-wave steepness
-$\varepsilon_L$.
-
-The homogeneity and stationarity of a scalar quantify $q$ can be expressed as
-conditions on the instantaneous fractional rate of change of $q$ being much
-smaller than the short-wave inverse scales:
-
-\begin{equation}
-  \frac{1}{q} \frac{\partial q}{\partial x} \ll k
-  \label{eq:homogeneity_condition}
-\end{equation}
-
-\begin{equation}
-  \frac{1}{q} \frac{\partial q}{\partial t} \ll \sigma
-  \label{eq:stationarity_condition}
-\end{equation}
-The slowly-varying conditions on the medium (in our case, the long-wave orbital
-velocity $U$) are then
-\begin{equation}
-  \frac{1}{U} \frac{\partial U}{\partial x} \ll \widetilde{k},
-\end{equation}
-and
-
-\begin{equation}
-  \frac{1}{U} \frac{\partial U}{\partial t} \ll \widetilde{\sigma},
-\end{equation}
-respectively.
-For a linear long wave, these conditions are equivalent to the scale separation
-between the long and short waves:
-
-\begin{equation}
-  k_L \ll \widetilde{k}
-  \label{eq:k_scale_separation}
-\end{equation}
-
-\begin{equation}
-  \sigma_L \ll \widetilde{\sigma}
-\end{equation}
-\citet{bretherton1968wavetrains} also require that the medium is nearly uniform
-in the vertical, over the vertical scales of motion of the short waves.
-For the ambient velocity being the orbital velocity of the long wave, this is
-trivially satisfied with (\ref{eq:k_scale_separation}).
-
-Now, to address the variation of short-wave properties,
-we apply (\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}) to
-the long-wave phase dependent wavenumber, gravitational acceleration, and
-wave action, and define the homogeneity and stationarity quantities as:
-
-\begin{equation}
-  H_q = 1 - \left| \frac{1}{qk} \frac{\partial q}{\partial x} \right|
-\end{equation}
-
-\begin{equation}
-  S_q = 1 - \left| \frac{1}{q\sigma} \frac{\partial q}{\partial t} \right|
-\end{equation}
-With the above definitions, we consider $q$ to be homogeneous and stationary as
-$H_q \rightarrow 1$ and $S_q \rightarrow 1$, respectively.
-
-For the linearized modulation solutions (\ref{eq:k_short_exact}),
-(\ref{eq:gravity_modulation_linear_no_curvature_mean_level}), and
-(\ref{eq:wave_action_modulation}), the respective homogeneity expressions are:
-
-\begin{equation}
-  H_{k,N} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right)^2} \right|
-  \label{eq:homogeneity_k}
-\end{equation}
-
-\begin{equation}
-  H_{g} = 1 - \frac{k_L}{k} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L^2 \cos^2{\psi}\right)} \right|
-  \label{eq:homogeneity_g}
-\end{equation}
-
-\begin{equation}
-  S_{k,N} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 + \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
-  \label{eq:stationarity_k}
-\end{equation}
-
-\begin{equation}
-  S_{g} = 1 - \frac{\sigma_L}{\sigma} \left| \frac{\varepsilon_L \sin{\psi}}{\left(1 - \varepsilon_L \cos{\psi}\right) \sqrt{1 - \varepsilon_L^2 \cos^2{\psi}}} \right|
-  \label{eq:stationarity_g}
-\end{equation}
-Notice that the steady, linearised solutions depend on both the the wavenumber
-ratio $k_L/k$ and the long-wave steepness $\varepsilon_L$.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figures/fig_homogeneity_stationarity_analytical.pdf}
-\caption{
-Homogeneity (left) and stationarity (right) of the short-wave wavenumber (top)
-and effective gravity (bottom) as a function of the long-wave steepness
-$\varepsilon_L$ and the wavenumber ratio $k/k_L$, based on the linearised
-solutions (\ref{eq:homogeneity_k}-\ref{eq:stationarity_g}).
-}
-\label{fig:homogeneity_stationarity_analytical}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figures/fig_homogeneity_stationarity_numerical.pdf}
-\caption{
-Homogeneity (left) and stationarity (right) of the short-wave wavenumber (top),
-action (middle), and effective gravity (bottom) as a function of the long-wave
-steepness $\varepsilon_L$ and the wavenumber ratio $k/k_L$, based on the
-numerical solutions of the full wave crest and action conservation equations.
-Note that the color ranges are different than those in Fig. \ref{fig:homogeneity_stationarity_analytical}.
-}
-\label{fig:homogeneity_stationarity_numerical}
-\end{figure}
 
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/fig_relative_velocity_error.pdf}
-\caption{
-}
-\label{fig:relative_velocity_error}
+\includegraphics[width=0.8\textwidth]{figures/fig_velocity_error_by_ak.pdf}
+\caption{
+}
+\label{fig:velocity_error_by_ak}
 \end{figure}

### 2025-02-21 paper.tex
Appendix A

@@
 In this paper, we revisit the linear wave theory and derive higher-order
-steady solutions for the change of short-wave wavenumber, action density, and
+steady solutions for the change of short-wave wavenumber, action, and
 gravitational acceleration due to the presence of longer waves.
@@
   phase-speed $C_{pL}$ from left to right.
-  The short waves propagate on the long-wave surface, their action density
+  The short waves propagate on the long-wave surface, their action
   moving with the group velocity $C_g$.
@@
 \end{equation}
-where $N$ is the action density of short waves defined as the ratio of their
+where $N$ is the action of short waves defined as the ratio of their
 energy density to their intrinsic frequency, $E/\sigma$,
 and $C_g = 1/2\sqrt{g/k}$ is their group speed in deep water.
-The wave action density expressed as $E/\sigma$ is commonly taken as
+The wave action expressed as $E/\sigma$ is commonly taken as
 conservative in wave modulation theory
@@
 \citet{dysthe1988orbiting} show, it is conservative only for gently sloped waves.
+We describe this requirement on gentle wave slope in more detail further below.
 Although the absence of wave growth and dissipation is not generally realistic in
@@
 \subsection{Limits to the wave action conservation}
-\label{appendix:wave_action_conservation}
+\label{sec:wave_action_conservation}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.pdf}
+\caption{
+  Maximum wavenumber modulation as function of long-wave steepness $\varepsilon_L$.
+  Black line and circles are for the analytical and numerical solutions from
+  \citet{longuet1960changes} and \citet{longuet1987propagation}, respectively.
+  Blue line is based on the analytical solutions from this paper.
+  Green and orange lines are for the numerical solutions from this paper
+  using linear and third-order Stokes long waves, respectively.
+}
+\label{fig:numerical_vs_analytical_k}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.pdf}
+\caption{
+  As Fig. \ref{fig:numerical_vs_analytical_k} but for the amplitude modulation.
+}
+\label{fig:numerical_vs_analytical_a}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.pdf}
+\caption{
+  As Fig. \ref{fig:numerical_vs_analytical_k} but for the steepness modulation.
+}
+\label{fig:numerical_vs_analytical_ak}
+\end{figure}
+
 The maximum modulations of the short-wave wavenumber, amplitude, and steepness
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.pdf}
-\caption{
-  Maximum wavenumber modulation as function of long-wave steepness $\varepsilon_L$.
-  Black line and circles are for the analytical and numerical solutions from
-  \citet{longuet1960changes} and \citet{longuet1987propagation}, respectively.
-  Blue line is based on the analytical solutions from this paper.
-  Green and orange lines are for the numerical solutions from this paper
-  using linear and third-order Stokes long waves, respectively.
-}
-\label{fig:numerical_vs_analytical_k}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.pdf}
-\caption{
-  As Fig. \ref{fig:numerical_vs_analytical_k} but for the amplitude modulation.
-}
-\label{fig:numerical_vs_analytical_a}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.pdf}
-\caption{
-  As Fig. \ref{fig:numerical_vs_analytical_k} but for the steepness modulation.
-}
-\label{fig:numerical_vs_analytical_ak}
-\end{figure}
+As the numerical solutions are for nonlinear wave action and crest conservation
+equations, it is important to re-evaluate the homogeneity and stationarity
+requirements for the wave action conservation to remain valid.
+Fig. \ref{fig:homogeneity_stationarity_numerical} shows the homogeneities
+and stationarities of the short-wave wavenumber, action, and effective gravity
+as function of the long-wave steepness $\varepsilon_L$ and the wavenumber ratio
+$k/k_L$, based on the numerical simulations.
+Although homogeneity is slightly lower ($H_{k,N,g} \approx 0.92$ for $k/k_L = 10$
+and $\varepsilon_L = 0.4$) compared to the analytical solutions,
+the stationarity is significantly reduced.
+Following the same criteria for strong ($S_{k,N,g} > 0.99$) and weak
+($S_{k,N,g} > 0.9$) stationarity as in subsection \ref{sec:wave_action_conservation},
+the numerical solutions are only weakly stationary for $\varepsilon_L < 0.21$
+at $k/k_L = 10$, and for $\varepsilon_L < 0.36$ at $k/k_L = 100$.
+This provides a quantitative guidance on when the short-wave action can be
+considered to be conserved.
 
@@
 velocities at the surface of the long wave.
-This introduces an additional error at the order of $\mathcal{O}(\varepsilon_L^2)$.
+This introduces a additional errors at $\mathcal{O}(\varepsilon_L^2)$.
 To quantify this error exactly, we compare the velocities evaluated at the
 surface for a linear (small-amplitude) wave and a third-order Stokes wave,
-with the surface velocity of a fully nonlinear wave.
-The nonlinear wave surface velocity is calculated numerically using the method
-by \citet{clamond2018accurate}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/fig_velocity_error_by_phase.pdf}
-\caption{
+with the surface velocity of a fully nonlinear wave
+(Fig. \ref{fig:velocity_error_by_phase}).
+The elevation and the surface velocity for the fully nonlinear wave is
+calculated numerically using the method by \citet{clamond2018accurate},
+which relies on Petviashvili's iterations and in this case uses 2048 Fourier
+modes to represent the wave.
+Both the 1$^{st}$ and the 3$^{rd}$ order waves underestimate the surface velocity
+at the crest and in the trough, and overestimate it on the front and rear faces
+of the crest.
+
+Maximum values of surface velocity errors for both the linear and the
+3$^{rd}$ order Stokes wave asymptote to $\varepsilon_L^3$ for small
+$\varepsilon_L$ (Fig. \ref{fig:velocity_error_by_ak}).
+However, they approach and exceed $\varepsilon_L^2$ at $\varepsilon_L \approx 0.3$
+and beyond.
+Although the maximum error is $\approx 10-20\%$ smaller for the 3$^rd$ order
+Stokes wave relative to the linear wave, its average error is slightly larger.
+The main takeaway from this figure is that evaluating the surface velocity
+of a linear wave carries an error no greater than $\varepsilon_L^2$ for
+$\varepsilon_L < 0.32$.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_velocity_error_by_phase.pdf}
+\caption{
+  Surface elevation (top) and horizontal velocity (middle) for the linear wave
+  (blue), the 3$^{rd}$ order Stokes wave (orange), and the fully nonlinear wave
+  (black) with steepness of $\varepsilon_L = 0.4$.
+  The bottom panel shows the surface velocity error for the linear wave (blue)
+  and the 3$^{rd}$ order Stokes wave (orange), relative to the fully nonlinear
+  wave solution.
 }
@@
 \caption{
+  Surface velocity error as function of wave steepness $\varepsilon_L$,
+  for the linear wave (blue) and the 3$^{rd}$ order Stokes wave (orange).
+  Maximum and mean errors are indicated with solid and dashed lines,
+  respectively.
+  For reference, $\varepsilon_L^2$ and $\varepsilon_L^3$ curves are shown in
+  black dashed and dotted lines, respectively.
 }

### 2025-02-21 scripts/plot_velocity_error.py

@@ -58,7 +58,7 @@ ax3.axhline(y=0, color="k", linestyle="--", alpha=0.5)
 ax3.set_ylabel("Velocity Error (m/s)")
 ax3.legend()
 ax3.grid(True)
-ax3.set_xlabel("Phase")
+ax3.set_xlabel(r"$\Psi$")
 
 # Set x-axis limits and ticks for all subplots
 xticks = np.array([0, np.pi / 2, np.pi, 3 * np.pi / 2, 2 * np.pi])
@@ -119,7 +119,7 @@ ax.plot(
     np.array(max_error_linear),
     lw=2,
     color="tab:blue",
-    label="1$^{st}$ order, max.",
+    label="1$^{st}$ order, maximum error",
 )
 ax.plot(
     ak_values,
@@ -127,14 +127,14 @@ ax.plot(
     lw=2,
     color="tab:blue",
     linestyle="--",
-    label="1$^{st}$ order, mean",
+    label="1$^{st}$ order, mean absolute error",
 )
 ax.plot(
     ak_values,
     np.array(max_error_stokes),
     lw=2,
     color="tab:orange",
-    label="3$^{rd}$ order, max.",
+    label="3$^{rd}$ order, maximum error",
 )
 ax.plot(
     ak_values,
@@ -142,12 +142,12 @@ ax.plot(
     lw=2,
     color="tab:orange",
     linestyle="--",
-    label="3$^{rd}$ order, mean",
+    label="3$^{rd}$ order, mean absolute error",
 )
 ax.plot(ak_values, ak_values**2, "k--", lw=2, label=r"$\varepsilon_L^2$")
 ax.plot(ak_values, ak_values**3, "k:", lw=2, label=r"$\varepsilon_L^3$")
 ax.set_xlabel(r"$\varepsilon_L$")
-ax.set_ylabel("Surface velocity error")
+ax.set_ylabel("Surface velocity error (m/s)")
 ax.legend()
 ax.set_xlim(0, 0.4)
 ax.set_ylim(0, 0.3)

### 2025-02-25 paper.tex
Fix typo

@@
 \end{equation}
-Notice that the steady, linearised solutions depend on both the the wavenumber
+Notice that the steady, linearised solutions depend on both the wavenumber
 ratio $k_L/k$ and the long-wave steepness $\varepsilon_L$.

### 2025-03-05 scripts/plot_action_conservation_criteria.py
Fixes to figures

@@ -9,7 +9,7 @@ matplotlib.rcParams["font.size"] = 14
 
 def get_homogeneity_and_stationarity(a_long, k_long, k_short):
     m = WaveModulationModel(a_long=a_long, k_long=k_long, k_short=k_short)
-    m.run(wave_type="linear", ramp_type="groups")
+    m.run(wave_type="stokes", ramp_type="groups")
     ds = m.to_xarray()
 
     x, t = ds.space.values, ds.time.values
@@ -116,7 +116,7 @@ axes = [ax1, ax2, ax3, ax4, ax5, ax6]
 cmap = plt.cm.viridis
 
 # Plot homogeneity (left column)
-H_levels = np.arange(0.9, 1.01, 0.01)
+H_levels = np.arange(0.88, 1.01, 0.01)
 c1 = ax1.contourf(ak_values, r_values, H_k_values, levels=H_levels, cmap=cmap)
 c3 = ax3.contourf(ak_values, r_values, H_N_values, levels=H_levels, cmap=cmap)
 c5 = ax5.contourf(ak_values, r_values, H_g_values, levels=H_levels, cmap=cmap)
@@ -136,7 +136,7 @@ plt.colorbar(c3, ax=ax3)
 plt.colorbar(c5, ax=ax5)
 
 # Plot stationarity (right column)
-S_levels = np.arange(0.6, 1.05, 0.05)
+S_levels = np.arange(0.5, 1.05, 0.05)
 c2 = ax2.contourf(ak_values, r_values, S_k_values, levels=S_levels, cmap=cmap)
 c4 = ax4.contourf(ak_values, r_values, S_N_values, levels=S_levels, cmap=cmap)
 c6 = ax6.contourf(ak_values, r_values, S_g_values, levels=S_levels, cmap=cmap)

### 2025-03-05 scripts/plot_analytical_solutions_with_phase.py

@@ -14,10 +14,13 @@ def plot_analytical_solutions(a_L: float):
     t = 0
     g0 = 9.8
     omega = np.sqrt(g0 * k_L)
+    k_short = 10
+    omega_short = np.sqrt(g0 * k_short)
+    cg_short = 0.5 * omega_short / k_short
 
     k_lhs = 1 + ak_L * np.cos(phase)
     k_modulation = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
-    g_modulation = gravity(x, t, a_L, k_L, omega, g0, "linear") / g0
+    g_modulation = gravity(x, t, a_L, k_L, omega, g0, cg_short, "linear") / g0
     N_modulation = k_modulation
     a_modulation = g_modulation ** (-0.25) * k_modulation**0.25 * N_modulation**0.5
     a_lhs = k_lhs

### 2025-03-05 scripts/plot_effective_gravity.py

@@ -17,10 +17,10 @@ def plot_effective_gravities(a: float):
     t = 0
 
     g1 = g0 * (1 - ak * np.cos(phase))
-    g2 = gravity(x, t, a, k, omega, g0, "linear", curvilinear=False)
-    g3 = gravity(x, t, a, k, omega, g0, "stokes", curvilinear=False)
-    g4 = gravity(x, t, a, k, omega, g0, "linear", curvilinear=True)
-    g5 = gravity(x, t, a, k, omega, g0, "stokes", curvilinear=True)
+    g2 = gravity(x, t, a, k, omega, g0, wave_type="linear", curvilinear=False)
+    g3 = gravity(x, t, a, k, omega, g0, wave_type="stokes", curvilinear=False)
+    g4 = gravity(x, t, a, k, omega, g0, wave_type="linear", curvilinear=True)
+    g5 = gravity(x, t, a, k, omega, g0, wave_type="stokes", curvilinear=True)
 
     fig = plt.figure(figsize=(8, 12))
     axes = fig.subplots(2, 1)
@@ -146,10 +146,10 @@ def plot_effective_gravities(a: float):
     for a in np.arange(0, 0.41, 0.01):
         ak = a * k
         g1 = g0 * (1 - ak * np.cos(phase))
-        g2 = gravity(x, t, a, k, omega, g0, "linear", curvilinear=False)
-        g3 = gravity(x, t, a, k, omega, g0, "stokes", curvilinear=False)
-        g4 = gravity(x, t, a, k, omega, g0, "linear", curvilinear=True)
-        g5 = gravity(x, t, a, k, omega, g0, "stokes", curvilinear=True)
+        g2 = gravity(x, t, a, k, omega, g0, wave_type="linear", curvilinear=False)
+        g3 = gravity(x, t, a, k, omega, g0, wave_type="stokes", curvilinear=False)
+        g4 = gravity(x, t, a, k, omega, g0, wave_type="linear", curvilinear=True)
+        g5 = gravity(x, t, a, k, omega, g0, wave_type="stokes", curvilinear=True)
         g1_max.append(np.max(g1))
         g2_max.append(np.max(g2))
         g3_max.append(np.max(g3))

### 2025-03-05 scripts/plot_numerical_vs_analytical_solutions.py

@@ -1,73 +1,118 @@
 import matplotlib.pyplot as plt
-from twowave import WaveModulationModel
+from twowave import WaveModulationModel, diff
 import numpy as np
 import pandas as pd
 import matplotlib
+import multiprocessing as mp
+import xarray as xr
+import os
 
 matplotlib.rc("font", size=14)
 
 
-def g_modulation(psi, a, k):
+def process_ak(a_long: float, k_long: float, wave_type: str):
+    filename = f"data/model_{wave_type}_ak_{a_long:.2f}.nc"
+    if os.path.exists(filename):
+        return xr.open_dataset(filename)
+    m = WaveModulationModel(a_long=a_long, k_long=k_long)
+    m.run(wave_type=wave_type, ramp_type="groups")
+    ds = m.to_xarray()
+    ds["a_long"] = a_long
+    ds["k_long"] = k_long
+    ds["a_short"] = m.a_short
+    ds["k_short"] = m.k_short
+    ds.to_netcdf(filename)
+    return ds
+
+
+def g_modulation(psi, a, k, Cg_short):
     eta = a * np.cos(psi)
-    res = 1 - a * k * np.exp(k * eta) * (np.cos(psi) - a * k * np.sin(psi) ** 2)
-    return res
+    g0 = 9.8
+    omega = np.sqrt(g0 * k)
+    Cp = omega / k
+    U = a * omega * np.cos(psi) * np.exp(k * eta)
+    W = a * omega * np.sin(psi) * np.exp(k * eta)
+    dx = np.diff(psi / k)[0]
+    a_z = -diff(W) / dx * (Cp - U - Cg_short)
+    gmod = (g0 + a_z) / g0
+    return gmod
 
 
 k_long = 1
 
-ak_range = np.arange(0, 0.41, 0.01)
-solutions1 = []
-solutions2 = []
-for ak in ak_range:
-    a_long = ak / k_long
+if not os.path.exists("data"):
+    os.makedirs("data")
 
-    m = WaveModulationModel(a_long=a_long, k_long=k_long)
-    m.run(ramp_type="groups")
-    solutions1.append(m)
+# Run in parallel using multiprocessing
+ak_values = np.arange(0.01, 0.41, 0.01)
+print("Running linear solutions")
+with mp.Pool(processes=mp.cpu_count()) as pool:
[…]

### 2025-03-05 paper.tex
Lagrangian gravity

@@
 \hline
-\citet{longuet1987propagation} & Velocity potential & Yes & Yes & No  & Numerical \\
+\citet{longuet1987propagation} & Crest and action & Yes & Yes & No  & Numerical \\
+                               & conservation     &     &     &     &           \\
 \hline
@@
 \caption{
-  Steady solutions for the modulation of short-wave (a) wavenumber, (b)
+  Steady analytical solutions for the modulation of short-wave (a) wavenumber, (b)
   gravitational acceleration, (c) amplitude, (d) steepness, (e) intrinsic
@@
 
-Long waves induce orbital (centripetal) accelerations on their surface, so the
-short waves ride in an accelerated reference frame and experience effective
+Three distinct effects contribute to the modulation of the effective gravity of
+short waves in the presence of long waves.
+First, long waves induce orbital (centripetal) accelerations on their surface,
+so the short waves ride in an accelerated reference frame and experience effective
 gravity that is lower at the crests and higher in the troughs
 \citep{longuet1986eulerian,longuet1987propagation}.
-Further, the projection of the long-wave induced horizontal acceleration
-contributes to small but non-negligible effects on the short-wave effective
+This is an $\mathcal{O}(\varepsilon_L)$ effect.
+Second, it is important to distinguish between the Eulerian and Lagrangian effective
+gravities \citep{longuet1986eulerian}.
+The Eulerian gravity is that which would be measured by a fixed wave staff,
+for example, and features sharp and strong minima at the crests and broad
+and weak maxima in the troughs.
+In contrast, the Lagrangian gravity is that which is experienced by a short wave
+group that propagates with its own group speed and is advected by the ambient,
+long-wave induced, velocity.
+The Lagrangian velocity is overall smoother and more attenuated compared to
+its Eulerian counterpart because the local slope as experienced by the
+travelling short-wave group is gentler.
+The two gravities are related by:
+
+\begin{equation}
+\label{eq:gravity_modulation_lagrangian}
+\widetilde{g_l} =
+g + \frac{dW}{dt} =
+g + \frac{\partial W}{\partial t} + \left(C_g + U\right) \frac{\partial W}{\partial x} =
+\widetilde{g_e} + \left(C_g + U\right) \frac{\partial W}{\partial x}
+\end{equation}
+where $l$ and $e$ subscripts serve to denote "Lagrangian" and "Eulerian",
+respectively.
+Although \citet{longuet1986eulerian} thoroughly described the differences
+between the Eulerian and Lagrangian effective gravities, he did so for
+a fully nonlinear irrotational gravity wave, which can be computed numerically
+but is not readily available analytically.
+Note also that when evaluating the Lagrangian effective gravity,
+\citet{longuet1987propagation} considered the orbital motion of the long wave
+but not the propagation speed of the short-wave group, which can exceed the
+ambient velocity $U$ depending on the properties of the short and long waves
+considered.
+Group speed of the short waves aside, the Lagrangian correction to the gravity
+being an advective term is $\mathcal{O}(\varepsilon_L^2)$.
+Third, the projection of the long-wave induced horizontal acceleration
+contributes small but non-negligible effects on the short-wave effective
 gravity in the curvilinear coordinate system
 \citep{phillips1981dispersion,zhang1990evolution}.
-These two effects are completely described by projecting the gravitational
+These three effects are completely described by projecting the gravitational
 acceleration vector from the coordinate that is perpendicular to the mean water
@@
   = \cos{\alpha} 
-  + \dfrac{\partial W_{z=\eta}}{\partial t} \cos{\alpha}
-  + \dfrac{\partial U_{z=\eta}}{\partial t} \sin{\alpha}
+  + \frac{1}{g} \dfrac{dW_{z=\eta_L}}{dt} \cos{\alpha}
+  + \frac{1}{g} \dfrac{dU_{z=\eta_L}}{dt} \sin{\alpha}
 \end{equation}
@@
 where $\alpha$ is the local slope of the long-wave surface.
-For short waves on a linear long wave, the gravity modulation is:
+For short waves on a linear long wave, the Eulerian gravity modulation is:
 
@@
 \label{eq:gravity_modulation_linear}
-\frac{\widetilde{g}}{g}
+\frac{\widetilde{g_e}}{g}
   = 
@@
 \end{equation}
+The same derivation can be carried out for the Lagrangian gravity
+modulation following (\ref{eq:gravity_modulation_lagrangian})-(\ref{eq:local_slope}),
+and is omitted here for brevity.
 Curvilinear effects on effective gravity can be removed altogether by setting
@@
   = \cos{\alpha_{St}}
-  + \dfrac{\partial W_{z=\eta_{St}}}{\partial t} \cos{\alpha_{St}}
-  + \dfrac{\partial U_{z=\eta_{St}}}{\partial t} \sin{\alpha_{St}}
+  + \dfrac{dW_{z=\eta_{St}}}{dt} \cos{\alpha_{St}}
+  + \dfrac{dU_{z=\eta_{St}}}{dt} \sin{\alpha_{St}}
 \end{equation}
@@
 Fig. \ref{fig:effective_gravities}a shows the effective gravity modulation of
-short-wave as a function of the long-wave phase for $\varepsilon_L = 0.2$, for
+short waves as a function of the long-wave phase for $\varepsilon_L = 0.2$, for
 a variety of the gravity modulation forms.
@@
 \citet{longuet1986eulerian,longuet1987propagation}.
-Note that these are strictly Eulerian values, evaluated at fixed positions in
-space.
-As \citet{longuet1986eulerian} pointed out, the effective gravity in the
-Lagrangian frame (\textit{i.e.}, following the surface particles) is smoother
-and less amplified because the moving particles spend a longer time riding near
-the crests and less time returning through the troughs.
-However, when we solve the short-wave crest and action conservation equations
-in an Eulerian frame, it is the Eulerian effective gravity that must be applied
-to conserve the short-wave properties across the long-wave phase.
 
@@
 small-amplitude short waves and significant scale separation ($k/k_L$).
-The restriction on $\varepsilon_L$ becomes readily apparent when we recognize
+The restriction on $\varepsilon_L$ becomes apparent when we recognize
 that the long wave-induced orbital velocity scales with $\varepsilon_L$, and
@@
   in the United States and the FACE Foundation, and the Office of Naval Research
-  Coastal Land Air-Sea Interaction DRI.
+  grants N000142012102 (Coastal Land Air-Sea Interaction DRI) and
+  N000142412598 (in collaboration with SASCWATCH: Study on Air-Sea Coupling with
+  WAves, Turbulence, and Clouds at High winds MURI).
 }
@@
 velocities at the surface of the long wave.
-This introduces a additional errors at $\mathcal{O}(\varepsilon_L^2)$.
+This introduces additional errors at $\mathcal{O}(\varepsilon_L^2)$.
 To quantify this error exactly, we compare the velocities evaluated at the

### 2025-03-06 scripts/plot_effective_gravity.py
Update dependencies and gravity script

@@ -1,14 +1,49 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import matplotlib
-from twowave import gravity
+from twowave import (
+    diff,
+    elevation,
+    orbital_horizontal_velocity,
+    orbital_vertical_velocity,
+    surface_slope,
+    nonlinear_wave_properties,
+)
 
-matplotlib.rc("font", size=16)
+
+def gravity(
+    x: float,
+    t: float,
+    a: float,
+    k: float,
+    omega: float,
+    g0: float = 9.8,
+    cg: float = 0.0,
+    wave_type: str = "linear",
+    nonlinear_props: tuple = None,
+    curvilinear: bool = True,
+    lagrangian: bool = True,
+) -> float:
+    z = elevation(x, t, a, k, omega, wave_type, nonlinear_props)
+    U = orbital_horizontal_velocity(x, z, t, a, k, omega, wave_type, nonlinear_props)
+    W = orbital_vertical_velocity(x, z, t, a, k, omega, wave_type, nonlinear_props)
+    Cp = omega / k
+    dx = np.diff(x)[0]
+    vel = Cp
+    if lagrangian:
+        vel = vel - U - cg
+    a_z = -diff(W) / dx * vel
+    if curvilinear:
+        a_x = -diff(U) / dx * vel
+        slope = surface_slope(x, t, a, k, omega, wave_type, nonlinear_props)
+        g = g0 * np.cos(slope) + a_z * np.cos(slope) + a_x * np.sin(slope)
+    else:
+        g = g0 + a_z
+    return g
 
 
 def plot_effective_gravities(a: float):
-    a = 0.2
-    phase = np.linspace(0, 2 * np.pi, 10000, endpoint=False)
+    phase = np.linspace(0, 2 * np.pi, 1000, endpoint=False)
     k = 1
     ak = a * k
     x = phase / k
@@ -16,243 +51,193 @@ def plot_effective_gravities(a: float):
     omega = np.sqrt(g0 * k)
     t = 0
 
-    g1 = g0 * (1 - ak * np.cos(phase))
-    g2 = gravity(x, t, a, k, omega, g0, wave_type="linear", curvilinear=False)
-    g3 = gravity(x, t, a, k, omega, g0, wave_type="stokes", curvilinear=False)
-    g4 = gravity(x, t, a, k, omega, g0, wave_type="linear", curvilinear=True)
[…]

### 2025-03-06 paper.tex
Text revisions

@@
 forecasting.
-In this paper, we revisit the linear wave theory and derive higher-order
-steady solutions for the change of short-wave wavenumber, action, and
-gravitational acceleration due to the presence of longer waves.
+In this paper, we revisit the wave crest and action conservation laws and derive
+higher-order steady solutions for the change of short-wave wavenumber, action,
+and gravitational acceleration due to the presence of longer waves.
 We validate the analytical solutions with numerical simulations of the full wave
@@
 wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
-We further examine the result of \citet{peureux2021unsteady} who found through
+We further examine the results of \citet{peureux2021unsteady} who found through
 numerical simulations that the short-wave modulation grows unsteadily with
@@
   They also induce downward and upward centripetal accelerations
-  ($\partial W/\partial t$) in the crests and troughs, respectively, that
+  ($dW/dt$) in the crests and troughs, respectively, that
   modulate the effective gravity at the surface.
@@
 Conversely, they elongate and flatten preferentially in the long-wave troughs.
-Recently, \citet{peureux2021unsteady} asked whether the steady solutions
+Recently, \citet{peureux2021unsteady} asked whether the steady (stationary) solutions
 of short-wave modulation are appropriate by simulating the full wave conservation
@@
 phase, as we will see later from the numerical simulations.
+Although the analytical solutions by \citet{longuet1960changes} has been updated
+by the numerical simulations of \citet{longuet1987propagation}, they are here
+nevertheless relevant to use as reference for the proposed analytical solutions.
 
@@
 Fig. \ref{fig:effective_gravities}a shows the effective gravity modulation of
-short waves as a function of the long-wave phase for $\varepsilon_L = 0.2$, for
+short waves as a function of the long-wave phase for $\varepsilon_L = 0.3$, for
 a variety of the gravity modulation forms.
-This value of $\varepsilon_L$ allows easier visualisation of the modulation
-differences, while still being relatively moderate.
+Although quite steep, this value of $\varepsilon_L$ allows easier visualisation
+of the modulation differences.
 To the first order in $\varepsilon_L$, the gravity modulation is out of phase
@@
 gravity on the long-wave crests and a higher effective gravity in the troughs.
-Evaluating the orbital accelerations at $z=\eta_L$ causes a gravity reduction
-around the crests ($\approx6\%$) and the troughs ($\approx 3\%$), and an
-increase on the front and rear faces ($\approx 4\%$) of the long wave, due to
-the exponential change in the orbital accelerations away from the mean water level
-(Fig. \ref{fig:effective_gravities}b).
-The difference between the linear and Stokes long waves at this steepness is
-negligible.
-The long-wave elevation curvature has a second-order effect in $\varepsilon_L$.
-Naturally, there is no curvature exactly at the crests and the troughs, and its
+Evaluating the orbital accelerations at $z=\eta$ and the Lagrangian gravity
+correction each have an $\mathcal{O}(\varepsilon_L^2)$ contribution but of
+opposite signs.
+The former amplifies the gravity modulation while the latter attenuates it.
+The curvilinear effects are significant at the front and rear faces of the long
+wave but not near the crests and troughs.
+The long-wave elevation slope has a second-order effect in $\varepsilon_L$.
+Naturally, the slope is exactly zero at the crests and the troughs, and its
 effects are largest at the front and rear faces of the long wave, where the
 covariance between the local slope and the horizontal accelerations are largest.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.7\textwidth]{figures/fig_effective_gravities.pdf}
-\caption{
-  (a) Analytical solutions for the effective gravitational acceleration
-  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
-  Black is for linear wave evaluated at $z=0$,
-  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
-  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
-  (b) As in (a) but normalised by the modulation of the linear wave evaluated at $z=0$, $g_1$.
+Evaluating the velocities and accelerations at the surface of third-order Stokes
+wave causes a correction mostly at the faces of the long wave.
+Finally, numerically computing the surface velocities and accelerations of a
+fully nonlinear wave allows us to exclude uncertainties associated with the
+truncation errors of the Stokes expansion at first and third orders in
+$\varepsilon_L$.
+At this steepness, the fully-nonlinear solution is marginally different from
+that of the third-order Stokes wave
+(see also Appendix \ref{appendix:surface_velocity_errors} for the errors
+in surface velocity estimates of the linear and third-order Stokes
+approximations).
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/fig_effective_gravities.pdf}
+\caption{
+  Analytical solutions for the effective gravitational acceleration
+  modulation by long waves, as function of the long-wave phase, for $\varepsilon_L = 0.3$.
+  Black is for Eulerian gravity of a linear wave evaluated at $z=0$;
+  blue is the same as black but evaluated at $z=\eta$;
+  orange is the same as blue but for Lagrangian gravity;
+  green is the same as orange but in a curvilinear reference frame;
+  red is the same as green but for a third-order Stokes wave;
+  and purple is the same as red but for a fully-nonlinear wave.
+  The elevation and surface velocities for the fully-nonlinear wave are computed
+  following \citet{clamond2018accurate}.
 }
@@
 
-It is instructive to also examine the mean, maximum, and minimum values of the
+It is instructive to also examine the minimum values of the
 effective gravity as function of long-wave steepness $\varepsilon_L$
-(Fig. \ref{fig:effective_gravities_mean_max_min}).
-The long-wave averaged effective gravity of short waves remains equal to the
-resting value of $g$ for all values of $\varepsilon_L$, except if the effects
-of curvature are considered, in which case the average effective gravity is
-reduced.
-This reduction, in case of a long Stokes wave, is $\approx 3\%$ for
-$\varepsilon_L = 0.2$ and $\approx 14\%$ for $\varepsilon_L = 0.4$.
-As the effective gravity minima and maxima occur at crests and troughs,
-respectively, the curvature affects the phase-averaged values but not the extremes.
+(Fig. \ref{fig:effective_gravities_min}), as the minima occur at or
+near the long-wave crests where the short-wave modulation is largest.
+Without the Lagrangian correction to the effective gravity, the gravity
+modulation is overestimated; for accelerations evaluated at the surface of a
+linear wave, the maximum gravity reduction is 60\% at $\varepsilon_L = 0.4$.
+The Lagrangian correction attenuates the gravity modulation, making the
+gravity reduction at the crests $\approx 25\%$.
 As the slope effects on the effective gravity exist only away from crests and
@@
 typically evaluated at the long-wave crests.
-The effective gravity reduction at the crests is $\approx 11\%$, $25\%$, and $66\%$
-for $\varepsilon_L = 0.1$, $0.2$, and $0.4$, respectively.
-These are similar to the fully-nonlinear numerical estimates of
+Here, the slope effect causes a difference in the minimum effective gravity
+for $\varepsilon_L \gtrsim 0.3$.
+The effective gravity reduction at the crests in the case of a fully nonlinear
+wave is $\approx$ 10\%, 19\%, 27\%, and 40\%, 
+for $\varepsilon_L$ of 0.1, 0.2, 0.3, and 0.4, respectively.
+These are similar to the fully nonlinear numerical estimates of
 \citet{longuet1986eulerian,longuet1987propagation}.
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities_mean_max_min.pdf}
-\caption{
-  As Fig. \ref{fig:effective_gravities} but showing the mean (solid),
-  maximum (dashed), and minimum (dotted) short-wave gravity modulation as
-  function of long-wave steepness $\varepsilon_L$.
-  The minimum and maximum values for the cases with curvature effects
-  ($\alpha=\frac{\partial \eta}{\partial x}$) are equivalent to those without
-  curvature effects ($\alpha=0$) and obscure them in the plot.
-}
-\label{fig:effective_gravities_mean_max_min}
+\includegraphics[width=\textwidth]{figures/fig_effective_gravities_min.pdf}
+\caption{
+  As Fig. \ref{fig:effective_gravities} but showing the minimum short-wave
+  gravity modulation as function of long-wave steepness $\varepsilon_L$.
+}
+\label{fig:effective_gravities_min}
 \end{figure}

### 2025-03-07 scripts/plot_analytical_solutions_with_phase.py
Update analytical solution script

@@ -1,7 +1,6 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import matplotlib
-from twowave import gravity
 
 matplotlib.rc("font", size=12)
 
@@ -10,22 +9,19 @@ def plot_analytical_solutions(a_L: float):
     phase = np.linspace(0, 2 * np.pi, 10000, endpoint=False)
     k_L = 1
     ak_L = a_L * k_L
-    x = phase / k_L
-    t = 0
     g0 = 9.8
-    omega = np.sqrt(g0 * k_L)
-    k_short = 10
-    omega_short = np.sqrt(g0 * k_short)
-    cg_short = 0.5 * omega_short / k_short
 
     k_lhs = 1 + ak_L * np.cos(phase)
+    g_lhs = 1 - ak_L * np.cos(phase)
     k_modulation = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
-    g_modulation = gravity(x, t, a_L, k_L, omega, g0, cg_short, "linear") / g0
+    g_modulation = 1 + ak_L * np.exp(ak_L * np.cos(phase)) * (
+        -np.cos(phase) + ak_L * np.sin(phase) ** 2
+    )
     N_modulation = k_modulation
     a_modulation = g_modulation ** (-0.25) * k_modulation**0.25 * N_modulation**0.5
     a_lhs = k_lhs
 
-    omega_lhs = np.sqrt(k_lhs)
+    omega_lhs = np.sqrt(g_lhs * k_lhs)
     omega_modulation = np.sqrt(g_modulation * k_modulation)
 
     Cp_lhs = omega_lhs / k_lhs
@@ -53,7 +49,7 @@ def plot_analytical_solutions(a_L: float):
     )
     axes[0].legend()
 
-    axes[1].plot(phase, np.ones(phase.shape), "k-", lw=2)
+    axes[1].plot(phase, g_lhs, "k-", lw=2)
     axes[1].plot(phase, g_modulation, marker="", linestyle="-", color="tab:blue", lw=2)
 
     axes[2].plot(phase, a_lhs, "k-", lw=2)

### 2025-03-08 paper.tex
Addressing minor reviews

@@
 wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
+Examining the homogeneity and stationarity requirements of the wave action
+conservation equation reveals that stationarity is more easily violated than
+homogeneity, and that the short-wave action is strongly conserved ($>$99\%) only
+for very small amplitude long waves, and weakly conserved ($>$90\%) for steep
+long waves.
 We further examine the results of \citet{peureux2021unsteady} who found through
@@
 interpretation of remote sensing of ocean surface waves and currents
-\citep{keller1975microwave,hara1994hydrodynamic}, and is distinguished
+\citep{keller1975microwave,alpers1978two,hara1994hydrodynamic}, and is distinguished
 from aerodynamic modulation \citep{donelan1987effect,belcher1999wave,chen2000effects}.
+Experimental studies of the general short-wave modulation problem included both
+the laboratory \citep[e.g.,][]{keller1975microwave,donelan2010modulation} and field
+\citep[e.g.,][]{hara2003observation,plant1977studies} measurements.
 \citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited the
@@
 variations in the effective gravity of short waves.
-Their results yielded significantly stronger modulation than previously predicted.
+Their results yielded significantly stronger modulation than previously predicted
+by \citet{longuet1960changes}.
 \citet{henyey1988energy} derived an analytical solution for the modulation of
@@
 \citet{longuet1987propagation}.
-The variety of analytical and numerical frameworks (Eulerian crest and action
-conservation, Hamiltonian, Schr\" odinger) that tackle the same fundamental
-physics and that relatively closely reproduce steady solutions for short-wave
+Aside from the early linear solutions by \citet{longuet1960changes}, the variety
+of analytical and numerical frameworks (Eulerian crest and action conservation,
+Hamiltonian, Schr\" odinger) that tackle the same fundamental
+physics and that relatively closely reproduce the steady solutions for short-wave
 modulation suggests that the solutions are robust.
@@
 
-In this paper, we revisit this problem from the perspective of linear wave theory
-but consider the long-wave slope to be significant enough to require evaluating
-the short-waves at the long-wave surface rather than at the mean water level.
+In this paper, we revisit this problem from the perspective of wave action
+conservation but consider the long-wave slope to be significant enough to
+require evaluating the short-waves at the long-wave surface rather than at the
+mean water level.
 This allows us to derive alternative, steady, nonlinear solutions for the
@@
 These approximations underestimate the short-wave modulation magnitude
-(Fig. \ref{fig:analytical_solutions_ak0.1}a).
+(Fig. \ref{fig:analytical_solutions_ak0.1}).
 At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
@@
 phase, as we will see later from the numerical simulations.
-Although the analytical solutions by \citet{longuet1960changes} has been updated
-by the numerical simulations of \citet{longuet1987propagation}, they are here
+The steady analytical solutions for modulation of short-wave effective gravity,
+amplitude, steepness, intrinsic frequency, and phase speed are also shown in
+Fig. \ref{fig:analytical_solutions_ak0.1}, for reference, however, their
+formulations appear in the following subsections.
+Although the analytical solutions by \citet{longuet1960changes} have been updated
+by the numerical results of \citet{longuet1987propagation}, they are here
 nevertheless relevant to use as reference for the proposed analytical solutions.
+This comparison allows quantifying solely the effect of evaluating the
+long-wave properties at the surface rather than at the mean water level.
 
@@
 a fully nonlinear irrotational gravity wave, which can be computed numerically
-but is not readily available analytically.
+but not analytically.
 Note also that when evaluating the Lagrangian effective gravity,
@@
     1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
-    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 \right]
+    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 e^{\varepsilon_L \cos{\psi}} \right]
   }
@@
 \label{eq:gravity_modulation_linear_no_curvature}
-\frac{\widetilde{g}}{g} = 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
+\frac{\widetilde{g_e}}{g} = 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
 \end{equation}
@@
 \label{eq:gravity_modulation_linear_no_curvature_mean_level}
-\frac{\widetilde{g}}{g} = 1 - \varepsilon_L \cos{\psi}
-\end{equation}
-which is the form used by \citet{peureux2021unsteady}, for example.
+\frac{\widetilde{g_e}}{g} = 1 - \varepsilon_L \cos{\psi}
+\end{equation}
+which is the form used by \citet{longuet1960changes} and
+\citet{peureux2021unsteady}, for example.
 
@@
   following \citet{clamond2018accurate}.
+  Long-wave crest and trough are located at $\psi = 0$ and $\psi = \pi$,
+  respectively.
 }
@@
 fourth-order Runge-Kutta method \citep{butcher1996runge}.
-The space is divided into 100 grid points.
+The space is divided into 128 grid points.
 This is effectively the same equation set and numerical configuration as that of
@@
 
-The numerical equations here are integrated in a fixed reference frame,
-rather than that moving with the long-wave phase speed.
+The numerical equations here are integrated in a fixed reference frame with
+periodic boundary conditions, rather than that moving with the long-wave phase speed.
 Either approach produces equivalent modulation results, however a fixed
@@
 This suggests that for low long-wave steepness the steady analytical solution
-is a reasonable proxy for the fully-nonlinear crest-action conservation
+is a reasonable approximation for the fully-nonlinear crest-action conservation
 equation solutions.
@@
   Green and orange lines are for the numerical solutions from this paper
-  using linear and third-order Stokes long waves, respectively.
+  using linear and third-order Stokes approximations of long waves, respectively.
+  Red line is for the fully-nonlinear gravity wave based on \citet{clamond2018accurate}.
 }
@@
 long waves alike, begin to diverge from the analytical solution at
-$\varepsilon_L \approx 0.2$.
-The numerical solution using the Stokes long wave is quantitatively very similar
-to the fully-nonlinear solutions of \citet{longuet1987propagation} in amplitude
-modulation, however with somewhat lower wavenumber and steepness modulations.
-For example, at $\varepsilon_L = 0.4$, the steepness modulation in our solutions
-is close to 6, whereas \citet{longuet1987propagation} finds it close to 8.
-The underestimate in the wavenumber modulation at high long-wave steepness
-relative to the results of \citet{longuet1987propagation}
-is likely due to his long wave velocities being evaluated to a higher order in
-$\varepsilon_L$ than they are here.
+$\varepsilon_L \approx 0.15$.
+The numerical solution using the third-order Stokes long wave is quantitatively
+very similar to the fully-nonlinear solutions of \citet{longuet1987propagation}
+in wavenumber modulation, however with somewhat lower amplitude and steepness
+modulations.
+Using the velocity and gravity properties of a fully-nonlinear wave, based on
+the approach by \citet{clamond2018accurate}, the numerical solutions diverge
+from those of the third-order Stokes wave only at very large steepnesses
+($\varepsilon_L > 0.3$).
+Using a fully-nonlinear form of the long-wave thus does not collapse the
+difference between the solutions here and those of \citet{longuet1987propagation}.
+For example, at $\varepsilon_L = 0.4$, the steepness modulation in our solution
+is 6.6, whereas \citet{longuet1987propagation} finds it to be 7.8.
+A possible reason for this difference is that the solutions by
+\citet{longuet1987propagation} are stationary (see his \S2), whereas in our
+numerical model we solve the prognostic nonlinear equations for the short-wave
+wavenumber and action.
+As we will see in the next section, and as was suggested by \citet{peureux2021unsteady},
+they are often, but not always, mostly steady, and can be strongly unsteady in
+special cases.
 Nevertheless, the overall steepness modulation between the two numerical
@@
 
-As the numerical solutions are for nonlinear wave action and crest conservation
+As the numerical solutions are for the nonlinear wave action and crest conservation
 equations, it is important to re-evaluate the homogeneity and stationarity
@@
 Following the same criteria for strong ($S_{k,N,g} > 0.99$) and weak
-($S_{k,N,g} > 0.9$) stationarity as in subsection \ref{sec:wave_action_conservation},
+($S_{k,N,g} > 0.9$) stationarity as in \S\ref{sec:wave_action_conservation},
 the numerical solutions are only weakly stationary for $\varepsilon_L < 0.21$
@@
 This provides a quantitative guidance on when the short-wave action can be
-considered to be conserved.
+considered to be conserved in the presence of longer waves.
 
@@
 This implies $k_L \ll k$ and $\sigma_L \ll \sigma$, respectively.
+The same requirement is imposed on the properties of short waves, including
+wavenumber, amplitude, and effective gravity.
 Another requirement is for the short-waves to remain linear, which implies
 $a k \ll 1$.
+These requirements have been evaluated for the solutions presented here by
+quantifying the homogeneity and stationarity of each relevant property of the
+flow in \S\ref{sec:wave_action_conservation}.
 Furthermore, the short waves examined here do not include the effects of
@@
 that are otherwise present in most measurements
-\citep{plant1986two,laxague2017gravity}.
+\citep[e.g.,][]{plant1986two,laxague2017gravity}.
+An alternative approach using a fully nonlinear velocity potential flow solver
+that allows for two-way interaction between short and long waves would provide
+additional insight into the modulation of the long waves by the short waves.
 
@@
 and regional wave forecasting.
+Second-order wave spectra have historically been provided by some wave models
+due to the importance for understanding radar altimetry data, however only in a
+diagnostic capacity \citep[e.g.,][]{janssen2009some}.
 Currently, most spectral wave models used for research or forecasting
@@
 In this paper we revisited the problem of short-wave modulation by long waves
-from a linear wave theory perspective.
+from a wave action conservation perspective.
 Using the linearised wave crest and action conservation equations coupled with

### 2025-03-10 paper.tex
Addressing reviews

@@
 homogeneity, and that the short-wave action is strongly conserved ($>$99\%) only
-for very small amplitude long waves, and weakly conserved ($>$90\%) for steep
+for very small-amplitude long waves, and weakly conserved ($>$90\%) for steep
 long waves.
@@
 \citet{peureux2021unsteady}, for example.
+The Lagrangian correction in (\ref{eq:gravity_modulation_lagrangian}) is:
+
+\begin{equation}
+\label{eq:gravity_modulation_lagrangian_correction}
+U \frac{\partial W}{\partial x} = g \varepsilon_L^2 \cos^2{\psi} e^{2 \varepsilon_L \cos{\psi}}
++ \mathcal{O}(\varepsilon_L^3)
+\end{equation}
+and so it becomes important only for large $\varepsilon_L$.
 
@@
 \dfrac{\widetilde{a}}{a} = 
-  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
+  \left( \dfrac{\widetilde{g}}{g} \right)^{-\frac{1}{4}}
   \left( 1 + \varepsilon_L \right)^{\frac{3}{4}}
@@
 
-Finally, the short-wave steepness modulation follows from (\ref{eq:k_short_exact})
-and (\ref{eq:wave_amplitude_modulation}):
+Finally, the short-wave steepness modulation follows from (\ref{eq:wave_amplitude_modulation}):
 
@@
   \left( \dfrac{\widetilde{N}}{N} \right)^{\frac{1}{2}}
-  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
+  \left( \dfrac{\widetilde{g}}{g} \right)^{-\frac{1}{4}}
 \end{equation}
@@
 Finally, the relative contributions of different modulation factors remain
-mostly constant as $\varepsilon_L$ increases, with only a mild increase of the
+mostly constant as $\varepsilon_L$ increases, with only a mild decrease of the
 effective gravity modulation contribution, and mild increases of the
@@
 approaches are remarkably similar, and either solution would cause the
-short-waves of even low steepness to break \citep{banner1993wave}.
-Where the two modulation solutions differ (at $\varepsilon_L > 0.2$),
-either solution would lead to short-wave breaking after one long-wave period.
-This suggests that in the cases where the short-waves would survive the
+short-waves of moderate steepness to break \citep{banner1993wave}.
+Specifically, where the two modulation solutions differ ($\varepsilon_L \gtrsim 0.3$),
+the steepness modulation approaches and exceeds 3.
+Tripling the short-wave steepness of even $\varepsilon = 0.1$ would bring the
+wave train near the observed breaking thresholds of $\varepsilon \gtrsim 0.32$
+\citep{perlin2013breaking}, although, recent measurements show evidence of
+significantly higher breaking thresholds
+\citep{toffoli2010maximum,mcallister2024three}, and even exceeding the
+geometric Stokes limit of 0.44.
+Nevertheless, in the cases where the short-waves would survive the
 modulation and persist over multiple long-wave periods, the two solutions are
@@
 
-\subsection{Short-wave modulation by long-wave groups}
+\subsection{Modulation by long waves of varying amplitude}
 \label{subsection:wave_groups}
 
-We now proceed to examine the numerical solutions with the long-wave forcing
-applied as a long-wave group.
-The wave group is simulated as a non-dispersive train of long waves whose
+We now proceed to examine the numerical solutions of short-wave modulation by
+long waves whose amplitude varies with time.
+The amplitude variation aims to mimic the behaviour of long-wave groups, a
+phenomenon that is ubiquitous in the ocean due to the dispersive nature of
+surface gravity waves, as well as due to the fact that wind-generated waves
+evolve to have a broad-banded spectra.
+For simplicity of implementation and interpretation, 
+the wave group here is simulated as a non-dispersive train of long waves whose
 amplitude scales with $\sin[\pi t / (n T_L)]$, where $n$ is the number of
@@
 intrinsic frequency, and phase speed.
-The latter two have not been previously examined in the context of short-wave
-modulation, but may be important for correctly calculating the wind input into
-short waves, which is proportional to the wind speed relative to the wave
-celerity.
+The latter two were previously examined in the context of short-wave
+modulation \citep{longuet1962resonant,longuet1962phase}, and may be important
+for correctly calculating the wind input into short waves, which is proportional
+to the wind speed relative to the wave celerity.
 The approximate steady solutions yield higher wavenumber and amplitude

### 2025-03-10 scripts/plot_velocity_error.py

@@ -32,18 +32,18 @@ U_stokes = orbital_horizontal_velocity(x, eta_stokes, t, a, k, omega)
 fig, (ax1, ax2, ax3) = plt.subplots(3, 1, figsize=(8, 9), sharex=True)
 
 # Top panel: Surface elevation
-ax1.plot(phase, eta_linear, label="Linear")
-ax1.plot(phase, eta_stokes, label="3$^{rd}$ order Stokes")
-ax1.plot(phase, eta, color="k", label="Fully nonlinear")
+ax1.plot(phase, eta_linear, label="1$^{st}$ order (linear)", lw=3)
+ax1.plot(phase, eta_stokes, label="3$^{rd}$ order Stokes", lw=3)
+ax1.plot(phase, eta, color="k", label="Fully nonlinear", lw=3)
 ax1.axhline(y=0, color="k", linestyle="--", alpha=0.5)
 ax1.set_ylabel("Surface elevation (m)")
 ax1.legend()
 ax1.grid(True)
 
 # Middle panel: Velocities
-ax2.plot(phase, U_linear, label="Linear")
-ax2.plot(phase, U_stokes, label="3$^{rd}$ order Stokes")
-ax2.plot(phase, U, color="k", label="Fully nonlinear")
+ax2.plot(phase, U_linear, label="1$^{st}$ order (linear)", lw=3)
+ax2.plot(phase, U_stokes, label="3$^{rd}$ order Stokes", lw=3)
+ax2.plot(phase, U, color="k", label="Fully nonlinear", lw=3)
 ax2.axhline(y=0, color="k", linestyle="--", alpha=0.5)
 ax2.set_ylabel("Horizontal velocity (m/s)")
 ax2.grid(True)
@@ -52,8 +52,8 @@ ax2.grid(True)
 error_linear = U_linear - U
 error_stokes = U_stokes - U
 
-ax3.plot(phase, error_linear, label="Linear")
-ax3.plot(phase, error_stokes, label="3$^{rd}$ order Stokes")
+ax3.plot(phase, error_linear, label="1$^{st}$ order (linear)", lw=3)
+ax3.plot(phase, error_stokes, label="3$^{rd}$ order Stokes", lw=3)
 ax3.axhline(y=0, color="k", linestyle="--", alpha=0.5)
 ax3.set_ylabel("Velocity Error (m/s)")
 ax3.legend()
@@ -117,14 +117,14 @@ ax = fig.add_subplot(111)
 ax.plot(
     ak_values,
[…]

### 2025-03-11 paper.tex
Finishing touches

@@
 wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
-Examining the homogeneity and stationarity requirements of the wave action
-conservation equation reveals that stationarity is more easily violated than
-homogeneity, and that the short-wave action is strongly conserved ($>$99\%) only
-for very small-amplitude long waves, and weakly conserved ($>$90\%) for steep
-long waves.
+Examining the homogeneity and stationarity requirements for the conservation of
+wave action reveals that stationarity is a stronger requirement and is
+generally not satisfied for very steep long waves.
 We further examine the results of \citet{peureux2021unsteady} who found through
@@
 nonlinear numerical solutions in long-wave steepness up to $\approx$0.2.
-Except for a subset of initial conditions, the solutions to the fully nonlinear
+Except for a subset of initial conditions, the solutions to the non-linearised
 crest-action conservation equations are mostly steady in the reference frame of
@@
 of a long wave perturbation in the middle of a homogeneous short wave field".
-However, it is not obvious whether such a scenario is realistic as waves tend to
+However, it is not obvious whether such a scenario is common as waves tend to
 disperse into groups, and the time scales of incoming swell fronts are much
 longer than those of short-wave generation by wind.
+Exceptional situations with such a sudden onset of swell may occur could be
+around complex coastlines or in shadows behind islands, where wind-generated
+ripples may propagate from an otherwise calm sea state into the path of a
+crossing swell.
+
 To test whether the sudden onset of long-wave forcing causes the unsteady growth,
@@
 surface gravity waves, as well as due to the fact that wind-generated waves
-evolve to have a broad-banded spectra.
+evolve to have broad-banded spectra.
 For simplicity of implementation and interpretation, 
@@
 These results can be reproduced with the companion numerical model.
+Note, however, according to \S\ref{sec:wave_action_conservation}, that the
+wave action conservation requirements are more easily satisfied for larger
+$k/k_L$.
 Also not considered here are the short-wave trains that are oblique to the
@@
 and is least affected by the effective gravity modulation (1/8).
+The homogeneity and stationarity criteria for wave action conservation are
+evaluated for both the analytical and numerical solutions.
+The stationarity is a stronger criterion than homogeneity, and is found to be
+strongly satisfied (> 99\% stationary) for very gently sloped waves and weakly
+satisfied (> 90\% stationary) for steeper waves.
 
@@
 long waves in a homogeneous short-wave field, a scenario that is unlikely to
-occur in the open ocean.
-A more likely condition such as that of long-wave groups, are only mildly
+be common in the open ocean.
+A more likely condition such as that of long-wave groups, is only mildly
 destabilising for the short waves.
-These results suggest that for mildly-to-moderately steep short waves,
+These results suggest that for the long-wave steepness $\varepsilon_L \lesssim 0.2$,
 the analytical steady solutions presented here are sufficient to describe the
 modulation of short waves by long waves.
+For moderately steep long waves ($0.2 \lesssim \varepsilon_L \lesssim 0.3$),
+numerical solutions of non-linearised crest and action conservation equations
+are necessary to capture the short-wave steepness modulation.
+For steeper long waves ($\varepsilon_L > 0.3$), the short-wave properties can
+no longer be considered stationary \citep{bretherton1968wavetrains}, and the
+wave action conservation equation is not valid in this context.
 It may be worthwhile to revisit a more systematic inclusion of the hydrodynamic

### 2025-03-12 paper.tex
Final edits for revision

@@
 We show that this unsteady growth only occurs for homogeneous initial
-conditions as a special case and does not generally occur, for example in more
+conditions as a special case and does not generally occur, for example, in more
 realistic long-wave groups.
-The proposed steady solutions are a good approximation of the fully
-nonlinear numerical solutions in long-wave steepness up to $\approx$0.2.
+The proposed steady solutions are a good approximation of the nonlinear
+crest-action conservation solutions in long-wave steepness $\lesssim 0.2$.
 Except for a subset of initial conditions, the solutions to the non-linearised
@@
 where $\widetilde{k}$ and $k$ are the modulated and unmodulated short-wave wavenumbers,
-respectively, and $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
+respectively, $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
 wavenumber $k_L$ and amplitude $a_L$, and $\psi$ is the long-wave phase.
@@
 interpretation of remote sensing of ocean surface waves and currents
-\citep{keller1975microwave,alpers1978two,hara1994hydrodynamic}, and is distinguished
+\citep{keller1975microwave,alpers1978two,hara1994hydrodynamic} and is distinguished
 from aerodynamic modulation \citep{donelan1987effect,belcher1999wave,chen2000effects}.
@@
 Conversely, they elongate and flatten preferentially in the long-wave troughs.
-Recently, \citet{peureux2021unsteady} asked whether the steady (stationary) solutions
+Recently, \citet{peureux2021unsteady} asked whether the steady
+(\textit{i.e.}, stationary; not depending on time) solutions
 of short-wave modulation are appropriate by simulating the full wave conservation
@@
 \begin{tabular}{cccccc}
-Reference   & Analysis   & Gravity & Nonlinear  & Nonlinear   & Solution \\
-            & framework & varies  & long waves & short waves &          \\
+Reference   & Analysis   & Solution & Solution    & Nonlinear  & Nonlinear   \\
+            & framework  & type     & steadiness  & long waves & short waves \\
 \hline
-\citet{longuet1960changes} & Velocity potential  & No  & No  & No  & Analytical \\
+\citet{longuet1960changes} & Velocity potential  & Analytical & Steady & No  & No  \\
 \hline
-\citet{phillips1981dispersion} & Crest and action & Yes & Yes & No  & Numerical \\
+\citet{phillips1981dispersion} & Crest and action & Numerical & Steady & Yes & No  \\
                                & conservation     &           &        &     &     \\
 \hline
-\citet{longuet1987propagation} & Crest and action & Yes & Yes & No  & Numerical \\
+\citet{longuet1987propagation} & Crest and action & Numerical & Steady & Yes & No  \\
                                & conservation     &           &        &     &     \\
 \hline
-\citet{henyey1988energy} & Hamiltonian & Yes & Yes & No  & Analytical \\
+\citet{henyey1988energy} & Hamiltonian & Analytical & Steady & Yes & No  \\
 \hline
-\citet{zhang1990evolution} & Schr\" odinger & Yes & Yes & Yes & Numerical \\
+\citet{zhang1990evolution} & Schr\" odinger & Numerical & Steady & Yes & Yes \\
 \hline
-\citet{peureux2021unsteady} & Crest and action & Yes & No  & No  & Numerical \\
+\citet{peureux2021unsteady} & Crest and action & Numerical & Unsteady & No  & No  \\
                             & conservation     &           &          &     &     \\
 \hline
-This paper  & Crest and action & Yes & Yes & No  & Analytical \\
-            & conservation     &     &     &     & and numerical \\
+This paper  & Crest and action & Analytical    & Steady       & Yes & No  \\
+            & conservation     & and numerical & and unsteady &     &     \\
 \hline
@@
   calculating the modulation of short waves by long waves.
+  The solution steadiness refers to whether the solution depends on time
+  (unsteady) or not (steady).
 }
@@
 solution by \citet{longuet1960changes}, but similar to the nonlinear numerical
-solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution} in
-long-wave steepness up to $\approx 0.2$.
+solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution} for
+$\varepsilon_L \lesssim 0.2$.
 To validate the steadiness of the analytical solutions, we perform numerical
@@
 advective velocity for a wave of any given wavenumber \citep{stewart1974hf},
-however for brevity here we will assume a surface velocity to be the advective
+however, for brevity, here we will assume a surface velocity to be the advective
 one for short waves.
 The evolution of the short-wave wavenumber is described by the conservation of
-wave crests \citep{phillips1981dispersion}:
+wave crests \citep[e.g.,][]{whitham1974dispersive,phillips1981dispersion}:
 
@@
 where $N$ is the action of short waves defined as the ratio of their
-energy density to their intrinsic frequency, $E/\sigma$,
+energy to their intrinsic frequency, $E/\sigma$,
 and $C_g = 1/2\sqrt{g/k}$ is their group speed in deep water.
@@
 \citet{dysthe1988orbiting} show, it is conservative only for gently sloped waves.
-We describe this requirement on gentle wave slope in more detail further below.
+We describe this and other requirements in more detail further below and show
+in what long-wave conditions they are satisfied.
+The wave crest conservation (\ref{eq:wave_crests}) follows from the assumption
+of a conserved wave phase, which is limited by the same set of requirements as
+those for the wave action conservation \citep[e.g.,][]{whitham1974dispersive}.
 Although the absence of wave growth and dissipation is not generally realistic in
@@
 and numerically in their full form in \S\ref{section:numerical_solutions}.
-It is worth noting that this system of equations is semi-coupled, meaning that
+This system of equations is semi-coupled, meaning that
 it is possible to solve (\ref{eq:wavenumber}) on its own, but solving
@@
 We now describe the analytical solutions for the short-wave wavenumber,
-effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
-and intrinsic phase speed in presence of long waves.
+effective gravitational acceleration, amplitude, and steepness in the presence
+of long waves.
 The wavenumber is derived by linearising the conservation of wave crests
@@
 and advection of the short waves.
-Combine (\ref{eq:U_L})-(\ref{eq:wavenumber2}), and integrate in time to get:
+Combine (\ref{eq:U_L}) and (\ref{eq:wavenumber2}) and integrate in time to get:
 
@@
 First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
-second, further truncating the Taylor expansion series beyond the first-order
-term.
+second, truncating the Taylor expansion series beyond $\mathcal{O}(\varepsilon_L)$.
 These approximations underestimate the short-wave modulation magnitude
@@
 between the Eulerian and Lagrangian effective gravities, he did so for
-a fully nonlinear irrotational gravity wave, which can be computed numerically
+a fully-nonlinear irrotational gravity wave, which can be computed numerically
 but not analytically.
@@
 being an advective term is $\mathcal{O}(\varepsilon_L^2)$.
-Third, the projection of the long-wave induced horizontal acceleration
-contributes small but non-negligible effects on the short-wave effective
-gravity in the curvilinear coordinate system
-\citep{phillips1981dispersion,zhang1990evolution}.
+Third, the projection of the long-wave induced horizontal acceleration in the
+curvilinear coordinate system contributes small but non-negligible effects on
+the short-wave effective gravity \citep{phillips1981dispersion,zhang1990evolution}.
 These three effects are completely described by projecting the gravitational
@@
 $\alpha = 0$.
-In that case, (\ref{eq:gravity_modulation_general}) simplifies to:
+In that case, (\ref{eq:gravity_modulation_linear}) simplifies to:
 
@@
 different processes.
-It also makes it straightforward to evaluate the gravity modulation for a
-different long-wave form, by evaluating the orbital accelerations and the local
-slope for that wave.
+It also makes it straightforward to evaluate the gravity modulation for
+different long-wave forms, by evaluating the orbital accelerations and the local
+slope for each wave form.
 For a third-order Stokes wave, for example, the elevation is:
@@
 of the modulation differences.
-To the first order in $\varepsilon_L$, the gravity modulation is out of phase
+To $\mathcal{O}(\varepsilon_L)$, the gravity modulation is out of phase
 with the long-wave elevation, with the short waves experiencing a lower effective
@@
 wave but not near the crests and troughs.
-The long-wave elevation slope has a second-order effect in $\varepsilon_L$.
+The long-wave surface slope has a $\mathcal{O}(\varepsilon_L^2)$ effect.
 Naturally, the slope is exactly zero at the crests and the troughs, and its
 effects are largest at the front and rear faces of the long wave, where the
-covariance between the local slope and the horizontal accelerations are largest.
-Evaluating the velocities and accelerations at the surface of third-order Stokes
+covariance between the local slope and the horizontal accelerations is largest.
+Evaluating the velocities and accelerations at the surface of a third-order Stokes
 wave causes a correction mostly at the faces of the long wave.
 Finally, numerically computing the surface velocities and accelerations of a
-fully nonlinear wave allows us to exclude uncertainties associated with the
+fully-nonlinear wave allows us to exclude uncertainties associated with the
 truncation errors of the Stokes expansion at first and third orders in
@@
 for $\varepsilon_L \gtrsim 0.3$.
-The effective gravity reduction at the crests in the case of a fully nonlinear
+The effective gravity reduction at the crests in the case of a fully-nonlinear
 wave is $\approx$ 10\%, 19\%, 27\%, and 40\%, 
 for $\varepsilon_L$ of 0.1, 0.2, 0.3, and 0.4, respectively.
-These are similar to the fully nonlinear numerical estimates of
+These are similar to the fully-nonlinear numerical estimates of
 \citet{longuet1986eulerian,longuet1987propagation}.
@@
   \left( \dfrac{\widetilde{N}}{N} \right)^{\frac{1}{2}}
-  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
-\end{equation}
-To the first order in $\varepsilon_L$, and using (\ref{eq:k_short_exact})
+  \left( \dfrac{\widetilde{g}}{g} \right)^{-\frac{1}{4}}
+\end{equation}
+To $\mathcal{O}(\varepsilon_L)$, and using (\ref{eq:k_short_exact})
 and (\ref{eq:wave_action_modulation}), the above simplifies to the result of
@@
 speed in panels (e) and (f), respectively.
-As the wavenumber and gravity modulations are out of phase, with the latter
-being somewhat larger in magnitude, the overall intrinsic frequency of the
-short waves is reduced across the entire long-wave phase, with highest reduction
-at the long-wave crests.
+As the wavenumber and gravity modulations are both $\mathcal{O}(\varepsilon_L)$
+and out of phase, the intrinsic frequency modulation is
+$\mathcal{O}(\varepsilon_L^2)$, positive on the long-wave faces and negative
+on the crests and troughs.
 For the short-wave phase speed modulation, the wavenumber and gravity modulations
@@
 crests and increased in the troughs.
-We discuss the implications of this effect later in \S\ref{section:discussion}
-in the context of wind input to short waves.
 
@@
 
-The homogeneity and stationarity of a scalar quantify $q$ can be expressed as
+The homogeneity and stationarity of a scalar quantity $q$ can be expressed as
 conditions on the instantaneous fractional rate of change of $q$ being much
@@
 The slowly-varying conditions on the medium (in our case, the long-wave orbital
-velocity $U$) are then
+velocity $U$) are then:
 \begin{equation}
@@
 \end{equation}
-and
 
@@
 the long-wave phase dependent wavenumber, gravitational acceleration, and
-wave action, and define the homogeneity and stationarity quantities as:
+wave action, and define the homogeneity and stationarity as:
 
@@
 
-For the linearized modulation solutions (\ref{eq:k_short_exact}),
+For the linearised modulation solutions (\ref{eq:k_short_exact}),
 (\ref{eq:gravity_modulation_linear_no_curvature_mean_level}), and
@@
 \end{equation}
-Notice that the steady, linearised solutions depend on both the wavenumber
-ratio $k_L/k$ and the long-wave steepness $\varepsilon_L$.
+The steady, linearised solutions thus depend on both the wavenumber ratio
+$k_L/k$ (homogeneity) or the frequency ratio $\sigma_L/\sigma$ (stationarity)
+and the long-wave steepness $\varepsilon_L$.
 
@@
 (\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}).
-Homogeneity is a weaker requirement than stationarity.
+Stationarity is a stronger requirement than homogeneity.
 At the lowest scale separation considered ($k/k_L = 10$), short-wave wavenumber,
-action, and effective gravity are all strongly homogeneous ($H_{k,N,g} > 0.99$).
+action, and effective gravity are all strongly homogeneous ($H_{k,N,g} > 0.99$)
+for $\varepsilon_L < 0.1$.
 However, the solutions are only strongly stationary ($S_{k,N,g} > 0.99$) at a
@@
 Although \citet{bretherton1968wavetrains} state that the wave action balance
-is valid to the first order in $k_L/k$, applying the general criteria
+is valid to $\mathcal{O}(k_L/k)$, applying the general criteria
 (\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}) yields an
-error of $\mathcal{O}(\varepsilon_L k/k_L)$.
+error of $\mathcal{O}(\varepsilon_L k_L/k)$.
 We will revisit these criteria based on the numerical solutions of nonlinear
@@
 \citep{stokes1847,van2018stokes,monismith2020stokes}.
-Their gravity modulation also does not consider the nonlinear or the curvature
-effects, however, this difference does not qualitatively affect the results.
+Their gravity modulation also does not consider the nonlinear, Lagrangian, or
+curvature effects, however, this difference does not qualitatively affect the results.
 Another difference is the choice of the numerical scheme for spatial differences,
@@
 $\mathcal{O}(10^{-7})$ and $\mathcal{O}(10^{-5})$ relative error, respectively.
-We show in the next section that our numerical solutions for infinite long-wave
+We show in the next subsection that our numerical solutions for infinite long-wave
 trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.
@@
 Either approach produces equivalent modulation results, however a fixed
-reference frame may allow for a more intuitive interpretation of the results.
+reference frame allows for a more intuitive interpretation of the results.
 This difference is important to keep in mind when comparing the results
@@
 The long-waves are initialised from rest ($a_L = 0$) and gradually ramped up
-to their target steepness over 5 long-wave periods, thus emulating the passage
-of a long-wave group
+to their target steepness over 5 long-wave periods to allow for a gentle
+increase of the long-wave forcing on the short waves
 (the importance of which we discuss in more detail in the next subsection).
@@
 (Fig. \ref{fig:numerical_solutions}a,d,g).
-This suggests that for low long-wave steepness the steady analytical solution
-is a reasonable approximation for the fully-nonlinear crest-action conservation
-equation solutions.
+This suggests that for low $\varepsilon_L$ the steady analytical solution
+is a reasonable approximation for the nonlinear crest-action solutions.
 The numerical solutions diverge more notably from the analytical ones at
-moderate steepness of 0.2 (Fig. \ref{fig:numerical_solutions}b,e,h), and are
-considerably different at very high steepness of 0.4
+moderate $\varepsilon_L = 0.2$ (Fig. \ref{fig:numerical_solutions}b,e,h),
+and are considerably different at very high $\varepsilon_L = 0.4$
 (Fig. \ref{fig:numerical_solutions}c,f,i).
-These differences suggest that at high steepness the inhomogeneity of
+These differences suggest that at high $\varepsilon_L$ the inhomogeneity of
 gravitational acceleration and group speed, otherwise ignored in the steady
@@
 Another important property of this solution is that the short-wave group that
-experiences peak amplification (dampening) are not locked-in to the long-wave
+experiences peak amplification (attenuation) are not locked-in to the long-wave
 crests (troughs), as they are in the prior steady solutions
@@
   applied to $a_L$ during the initial five long-wave periods.
+  Notice the change in the colour range.
 }
@@
 longer than those of short-wave generation by wind.
-Exceptional situations with such a sudden onset of swell may occur could be
+Exceptional situations with such a sudden onset of swell may occur
 around complex coastlines or in shadows behind islands, where wind-generated
@@
 crossing swell.
+Within the limits of the modulation theory used here, such short waves could
+quickly steepen and break.
 
@@
 This result is shown in Fig. \ref{fig:modulation_3panel_ramp}.
+Note that the linear ramp is strictly a numerical modeling technique to test
+the sensitivity of the solution to the sudden forcing at the beginning of the
+simulation, and is not intended to represent a physical process.
 The key distinction that arises with the introduction of the ramp is that the
@@
 However, in the case of a gentle ramp-up of the long-wave forcing, the
-propagation and inhomogeneity tendencies mostly balance each other and allow
+propagation and inhomogeneity tendencies balance each other and allow
 the short-wave steepening to lock-in to the long-wave crests.
@@
 $t = n T_L / 2$, after which it gradually decreases toward zero.
-In this simulation, $n = 10$ and $\varepsilon_L = 0.1$ as in the previous two
-simulations.
+In this simulation, $k/k_L = 10$ and $\varepsilon_L = 0.1$ as in the previous
+two simulations.
 The modulations of short-wave wavenumber, amplitude, and steepness in response
@@
 Within these limits, the crest-action conservation equations provide an
-elegant and relatively simple framework to study the hydrodynamic modulation
+elegant and simple framework to study the hydrodynamic modulation
 of short waves by longer waves.
@@
 \citep[e.g.,][]{plant1986two,laxague2017gravity}.
-An alternative approach using a fully nonlinear velocity potential flow solver
+An alternative approach using a fully-nonlinear velocity potential flow solver
 that allows for two-way interaction between short and long waves would provide
@@
 \citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,henyey1988energy,zhang1990evolution},
-it was not obvious from the governing equations that the fully-nonlinear
-steady solutions should exist.
+it was not obvious from the governing prognostic equations that the steady
+solutions should exist generally.
 \citet{peureux2021unsteady} examined this problem numerically and found that
@@
 found that the short-wave amplitude, and thus steepness as well, grow
-unsteadily.
-However, here we demonstrate that the unsteady steepening of short
+unsteadily and indefinitely.
+Here we demonstrate that the unsteady steepening of short
 waves only occurs in the case of a sudden appearance of long waves, where
@@
 Another important consideration of the hydrodynamic modulation process is
-whether it should somehow be accounted for in phase-averaged spectral wave
+whether it should be accounted for in phase-averaged spectral wave
 models \citep{group1988wam,tolman1991third,donelan2012modeling} used for global
@@
 diagnostic capacity \citep[e.g.,][]{janssen2009some}.
+The framework by \citet{janssen2009some}, based on the Hamiltonian and
+the action series expansion by \citet{zakharov1968stability}, is an alternative
+and more general approach to nonlinear wave evolution than the one described
+here.
 Currently, most spectral wave models used for research or forecasting
@@
 Using the linearised wave crest and action conservation equations coupled with
-nonlinear solutions for the effective gravity, we derived steady analytical
+nonlinear solutions for the effective gravity, we derived steady, nonlinear, analytical
 solutions for the modulation of short wave wavenumber, amplitude, steepness,
@@
 The stationarity is a stronger criterion than homogeneity, and is found to be
-strongly satisfied (> 99\% stationary) for very gently sloped waves and weakly
-satisfied (> 90\% stationary) for steeper waves.
+strongly satisfied ($> 99\%$ stationary) for very gently sloped waves and weakly
+satisfied ($> 90\%$ stationary) for steeper waves.
 
@@
 are consistent with the analytical solutions for $\varepsilon_L \lesssim 0.2$,
-and are similar in amplitude to the numerical solutions of
+and are similar in magnitude to the numerical solutions of
 \citet{longuet1987propagation} and \citet{zhang1990evolution}.
-The simulations also reveal that the unsteady growth of short-wave steepness
+The simulations also reveal that the unsteady growth of the short-wave steepness
 reported by \citet{peureux2021unsteady} is due to the sudden appearance of
@@
 be common in the open ocean.
-A more likely condition such as that of long-wave groups, is only mildly
+A more likely condition, such as that of long-wave groups, is only mildly
 destabilising for the short waves.
@@
 numerical solutions of non-linearised crest and action conservation equations
-are necessary to capture the short-wave steepness modulation.
-For steeper long waves ($\varepsilon_L > 0.3$), the short-wave properties can
-no longer be considered stationary \citep{bretherton1968wavetrains}, and the
-wave action conservation equation is not valid in this context.
+are necessary to properly capture the short-wave steepness modulation.
+For steeper long waves ($\varepsilon_L \gtrsim 0.3$), the short-wave properties can
+no longer be considered stationary \citep{bretherton1968wavetrains};
+the wave action conservation equation is thus not valid in this context.
 It may be worthwhile to revisit a more systematic inclusion of the hydrodynamic
@@
 surface for a linear (small-amplitude) wave and a third-order Stokes wave,
-with the surface velocity of a fully nonlinear wave
+with the surface velocity of a fully-nonlinear wave
 (Fig. \ref{fig:velocity_error_by_phase}).
-The elevation and the surface velocity for the fully nonlinear wave is
+The elevation and the surface velocity for the fully-nonlinear wave is
 calculated numerically using the method by \citet{clamond2018accurate},
@@
 modes to represent the wave.
-Both the 1$^{st}$ and the 3$^{rd}$ order waves underestimate the surface velocity
+Both the first-order and the third-order waves underestimate the surface velocity
 at the crest and in the trough, and overestimate it on the front and rear faces
@@
 Maximum values of surface velocity errors for both the linear and the
-3$^{rd}$ order Stokes wave asymptote to $\varepsilon_L^3$ for small
+third-order Stokes wave asymptote to $\varepsilon_L^3$ for small
 $\varepsilon_L$ (Fig. \ref{fig:velocity_error_by_ak}).
@@
 and beyond.
-Although the maximum error is $\approx 10-20\%$ smaller for the 3$^rd$ order
+Although the maximum error is $\approx 10-20\%$ smaller for the third-order
 Stokes wave relative to the linear wave, its average error is slightly larger.
@@
 $\varepsilon_L < 0.32$.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figures/fig_velocity_error_by_phase.pdf}
+Field measurements of near-surface orbital velocities by
+\citet{laxague2020observations} qualitatively support this result, although
+they find larger departures from the linear theory in stronger winds, that is,
+steeper waves.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fig_velocity_error_by_phase.pdf}
 \caption{
   Surface elevation (top) and horizontal velocity (middle) for the linear wave
-  (blue), the 3$^{rd}$ order Stokes wave (orange), and the fully nonlinear wave
+  (blue), the third-order Stokes wave (orange), and the fully-nonlinear wave
   (black) with steepness of $\varepsilon_L = 0.4$.
   The bottom panel shows the surface velocity error for the linear wave (blue)
-  and the 3$^{rd}$ order Stokes wave (orange), relative to the fully nonlinear
+  and the third-order Stokes wave (orange), relative to the fully-nonlinear
   wave solution.

### 2025-03-13 paper.tex
Revision 1 (#1)
* Figures to quantify surface velocity error

* Compute wave action conservation criteria

* Homogeneity and stationarity from analytical and numerical solutions

* WIP wave action conservation limits

* WIP homogeneity and stationarity figures

* WIP appendices

* Add ref to Clamond & Dutykh

* Update velocity error plot

* Add section on wave action limits

* Appendix A

* Fix typo

* Update plotting scripts to reflect updated functions in 2wave

* Fix x-label and contour levels

* Fixes to figures

* Total derivative for acceleration in diagram

* Lagrangian gravity

* Updat

@@
 Their gravity modulation also does not consider the nonlinear, Lagrangian, or
-curvature effects, however, this difference does not qualitatively affect the results.
+slope effects, however, this difference does not qualitatively affect the results.
 Another difference is the choice of the numerical scheme for spatial differences,

### 2025-03-13 scripts/plot_action_conservation_criteria.py

@@ -1,0 +1,367 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from twowave import WaveModulationModel, diff
+
+import matplotlib
+
+matplotlib.rcParams["font.size"] = 14
+
+
+def get_homogeneity_and_stationarity(a_long, k_long, k_short):
+    m = WaveModulationModel(a_long=a_long, k_long=k_long, k_short=k_short)
+    m.run(wave_type="stokes", ramp_type="groups")
+    ds = m.to_xarray()
+
+    x, t = ds.space.values, ds.time.values
+    dx, dt = np.diff(x)[0], np.diff(t)[0]
+
+    # Homogeneity criteria
+    n, _ = np.unravel_index(np.argmax(ds.wave_action.values), ds.wave_action.shape)
+    k_h = ds.wavenumber.values[n]
+    g_h = ds.gravitational_acceleration.values[n]
+    N_h = ds.wave_action.values[n]
+
+    H_k = 1 - np.abs(diff(k_h)) / dx / k_h / k_h
+    H_g = 1 - np.abs(diff(g_h)) / dx / g_h / k_h
+    H_N = 1 - np.abs(diff(N_h)) / dx / N_h / k_h
+
+    # Stationarity criteria
+    k_s = ds.wavenumber.values[:, 0]
+    g_s = ds.gravitational_acceleration.values[:, 0]
+    N_s = ds.wave_action.values[:, 0]
+    sigma = np.sqrt(g_s * k_s)
+
+    S_k = 1 - np.abs(diff(k_s)) / dt / k_s / sigma
+    S_g = 1 - np.abs(diff(g_s)) / dt / g_s / sigma
+    S_N = 1 - np.abs(diff(N_s)) / dt / N_s / sigma
+
+    return (
+        np.min(H_k),
+        np.min(H_g),
+        np.min(H_N),
+        np.min(S_k),
+        np.min(S_g),
+        np.min(S_N),
+    )
+
+
+def homogeneity_k(k0, k_L, ak, phase):
+    return 1 - np.abs(k_L / k0 * ak * np.sin(phase) / (1 + ak * np.cos(phase)) ** 2)
+
+
+def homogeneity_g(k0, k_L, ak, phase):
+    return 1 - np.abs(k_L / k0 * ak * np.sin(phase) / (1 - ak**2 * np.cos(phase) ** 2))
+
+
+def stationarity_k(sigma0, sigma_L, ak, phase):
+    return 1 - np.abs(
+        sigma_L
+        / sigma0
[…]

### 2025-03-13 scripts/plot_analytical_solutions_with_phase.py

@@ -1,7 +1,6 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import matplotlib
-from twowave import gravity_curvilinear
 
 matplotlib.rc("font", size=12)
 
@@ -10,19 +9,19 @@ def plot_analytical_solutions(a_L: float):
     phase = np.linspace(0, 2 * np.pi, 10000, endpoint=False)
     k_L = 1
     ak_L = a_L * k_L
-    x = phase / k_L
-    t = 0
     g0 = 9.8
-    omega = np.sqrt(g0 * k_L)
 
     k_lhs = 1 + ak_L * np.cos(phase)
+    g_lhs = 1 - ak_L * np.cos(phase)
     k_modulation = np.exp(ak_L * np.cos(phase) * np.exp(ak_L * np.cos(phase)))
-    g_modulation = gravity_curvilinear(x, t, a_L, k_L, omega, g0, "linear") / g0
+    g_modulation = 1 + ak_L * np.exp(ak_L * np.cos(phase)) * (
+        -np.cos(phase) + ak_L * np.sin(phase) ** 2
+    )
     N_modulation = k_modulation
     a_modulation = g_modulation ** (-0.25) * k_modulation**0.25 * N_modulation**0.5
     a_lhs = k_lhs
 
-    omega_lhs = np.sqrt(k_lhs)
+    omega_lhs = np.sqrt(g_lhs * k_lhs)
     omega_modulation = np.sqrt(g_modulation * k_modulation)
 
     Cp_lhs = omega_lhs / k_lhs
@@ -50,7 +49,7 @@ def plot_analytical_solutions(a_L: float):
     )
     axes[0].legend()
 
-    axes[1].plot(phase, np.ones(phase.shape), "k-", lw=2)
+    axes[1].plot(phase, g_lhs, "k-", lw=2)
     axes[1].plot(phase, g_modulation, marker="", linestyle="-", color="tab:blue", lw=2)
 
     axes[2].plot(phase, a_lhs, "k-", lw=2)

### 2025-03-13 scripts/plot_effective_gravity.py

@@ -1,14 +1,49 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import matplotlib
-from twowave import gravity, gravity_curvilinear
+from twowave import (
+    diff,
+    elevation,
+    orbital_horizontal_velocity,
+    orbital_vertical_velocity,
+    surface_slope,
+    nonlinear_wave_properties,
+)
 
-matplotlib.rc("font", size=16)
+
+def gravity(
+    x: float,
+    t: float,
+    a: float,
+    k: float,
+    omega: float,
+    g0: float = 9.8,
+    cg: float = 0.0,
+    wave_type: str = "linear",
+    nonlinear_props: tuple = None,
+    curvilinear: bool = True,
+    lagrangian: bool = True,
+) -> float:
+    z = elevation(x, t, a, k, omega, wave_type, nonlinear_props)
+    U = orbital_horizontal_velocity(x, z, t, a, k, omega, wave_type, nonlinear_props)
+    W = orbital_vertical_velocity(x, z, t, a, k, omega, wave_type, nonlinear_props)
+    Cp = omega / k
+    dx = np.diff(x)[0]
+    vel = Cp
+    if lagrangian:
+        vel = vel - U - cg
+    a_z = -diff(W) / dx * vel
+    if curvilinear:
+        a_x = -diff(U) / dx * vel
+        slope = surface_slope(x, t, a, k, omega, wave_type, nonlinear_props)
+        g = g0 * np.cos(slope) + a_z * np.cos(slope) + a_x * np.sin(slope)
+    else:
+        g = g0 + a_z
+    return g
 
 
 def plot_effective_gravities(a: float):
-    a = 0.2
-    phase = np.linspace(0, 2 * np.pi, 10000, endpoint=False)
+    phase = np.linspace(0, 2 * np.pi, 1000, endpoint=False)
     k = 1
     ak = a * k
     x = phase / k
@@ -16,243 +51,193 @@ def plot_effective_gravities(a: float):
     omega = np.sqrt(g0 * k)
     t = 0
 
-    g1 = g0 * (1 - ak * np.cos(phase))
-    g2 = gravity(x, t, a, k, omega, g0, "linear")
-    g3 = gravity(x, t, a, k, omega, g0, "stokes")
-    g4 = gravity_curvilinear(x, t, a, k, omega, g0, "linear")
-    g5 = gravity_curvilinear(x, t, a, k, omega, g0, "stokes")
[…]

### 2025-03-13 scripts/plot_modulation_contributors_numerical.py

@@ -1,5 +1,6 @@
 import matplotlib.pyplot as plt
 import numpy as np
+from twowave import WaveModulationModel
 
 import matplotlib
 
@@ -7,21 +8,30 @@ matplotlib.rcParams["font.size"] = 14
 
 ak = np.arange(0, 0.41, 0.01)
 
-k_mod = np.exp(ak * np.exp(ak)) ** 1.25
-N_mod = np.exp(ak * np.exp(ak)) ** 0.5
-g_mod = (1 - ak * np.exp(ak) + ak**2 * np.exp(2 * ak)) ** (-0.25)
-
-ak_mod = g_mod * k_mod * N_mod
+k_mod = []
+N_mod = []
+g_mod = []
+ak_mod = []
+for a in ak:
+    m = WaveModulationModel(a_long=a)
+    m.run(ramp_type="groups", wave_type="linear")
+    ds = m.to_xarray()
+    k_mod.append(np.max(ds.wavenumber) / m.k_short)
+    N_mod.append(np.max(ds.wave_action))
+    g_mod.append(np.min(ds.gravitational_acceleration) / m.grav0)
+    ak_mod.append(np.max(ds.amplitude * ds.wavenumber) / (m.a_short * m.k_short))
+k_mod = np.array(k_mod) ** 1.25
+N_mod = np.array(N_mod) ** 0.5
+g_mod = np.array(g_mod) ** -0.25
+ak_mod = np.array(ak_mod)
 
 k_mod_log = np.log(k_mod)
 N_mod_log = np.log(N_mod)
 g_mod_log = np.log(g_mod)
 ak_mod_log = k_mod_log + N_mod_log + g_mod_log
 
-# Create a figure with two subplots side by side
 fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(12, 5))
 
-# First subplot
 ax1.plot(ak, k_mod, label=r"$(\widetilde{k}/k)^{\frac{5}{4}}$", lw=2)
 ax1.plot(ak, N_mod, label=r"$(\widetilde{N}/N)^{\frac{1}{2}}$", lw=2)
 ax1.plot(ak, g_mod, label=r"$(\widetilde{g}/g)^{-\frac{1}{4}}$", lw=2)
@@ -31,11 +41,10 @@ ax1.set_title("Modulation Factors")
 ax1.set_xlabel(r"$\varepsilon_L$")
[…]

### 2025-03-13 scripts/plot_modulation_diagram.py

@@ -130,7 +130,7 @@ pos = arrow_positions[0]
 ax.text(
     pos - 0.2,
     np.sin(pos) - 0.4,
-    r"$\frac{\partial W}{\partial t}$",
+    r"$\frac{dW}{dt}$",
     fontsize=28,
     ha="right",
     va="top",

### 2025-03-13 scripts/plot_numerical_simulations.py

@@ -49,7 +49,7 @@ def plot_modulation_3panel(model, contour_levels, title):
         ax.set_yticks(range(0, num_periods + 1, 1))
         ax.set_xticks(np.arange(0, 2.5 * np.pi, 0.5 * np.pi))
         ax.set_xticklabels([r"$0$", r"$\pi/2$", r"$\pi$", r"$3\pi/2$", r"$2\pi$"])
-        ax.set_xlabel(r"$\Psi$")
+        ax.set_xlabel(r"$k_L x$")
 
     fig.subplots_adjust(left=0.1, right=0.9, bottom=0.0, top=0.9)
     cbar = fig.colorbar(
@@ -64,7 +64,7 @@ def plot_modulation_3panel(model, contour_levels, title):
 
 m = WaveModulationModel()
 m.run(ramp_type=None)
-plot_modulation_3panel(m, np.arange(0.5, 2.3, 0.1), "Infinite wave train")
+plot_modulation_3panel(m, np.linspace(0.5, 2.4, 20), "Infinite wave train")
 
 m.run(ramp_type="linear")
 plot_modulation_3panel(m, np.arange(0.75, 1.3, 0.05), "Linear ramp")

### 2025-03-13 scripts/plot_numerical_vs_analytical_solutions.py

@@ -1,74 +1,118 @@
 import matplotlib.pyplot as plt
-from twowave import WaveModulationModel
+from twowave import WaveModulationModel, diff
 import numpy as np
 import pandas as pd
 import matplotlib
+import multiprocessing as mp
+import xarray as xr
+import os
 
 matplotlib.rc("font", size=14)
 
 
-def g_modulation(psi, a, k):
+def process_ak(a_long: float, k_long: float, wave_type: str):
+    filename = f"data/model_{wave_type}_ak_{a_long:.2f}.nc"
+    if os.path.exists(filename):
+        return xr.open_dataset(filename)
+    m = WaveModulationModel(a_long=a_long, k_long=k_long)
+    m.run(wave_type=wave_type, ramp_type="groups")
+    ds = m.to_xarray()
+    ds["a_long"] = a_long
+    ds["k_long"] = k_long
+    ds["a_short"] = m.a_short
+    ds["k_short"] = m.k_short
+    ds.to_netcdf(filename)
+    return ds
+
+
+def g_modulation(psi, a, k, Cg_short):
     eta = a * np.cos(psi)
-    res = 1 - a * k * np.exp(k * eta) * (np.cos(psi) - a * k * np.sin(psi) ** 2)
-    return res
+    g0 = 9.8
+    omega = np.sqrt(g0 * k)
+    Cp = omega / k
+    U = a * omega * np.cos(psi) * np.exp(k * eta)
+    W = a * omega * np.sin(psi) * np.exp(k * eta)
+    dx = np.diff(psi / k)[0]
+    a_z = -diff(W) / dx * (Cp - U - Cg_short)
+    gmod = (g0 + a_z) / g0
+    return gmod
 
 
 k_long = 1
 
-ak_range = np.arange(0, 0.41, 0.01)
-solutions1 = []
-solutions2 = []
-for ak in ak_range:
-    a_long = ak / k_long
+if not os.path.exists("data"):
+    os.makedirs("data")
 
-    m = WaveModulationModel(a_long=a_long, k_long=k_long)
-    m.run(ramp_type="groups")
-    solutions1.append(m)
+# Run in parallel using multiprocessing
+ak_values = np.arange(0.01, 0.41, 0.01)
+print("Running linear solutions")
+with mp.Pool(processes=mp.cpu_count()) as pool:
[…]

### 2025-03-13 scripts/plot_velocity_error.py

@@ -1,0 +1,156 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from ssgw import SSGW
+from twowave import elevation, orbital_horizontal_velocity
+
+import matplotlib
+
+matplotlib.rcParams.update({"font.size": 14})
+
+g = 9.8
+k = 1
+a = 0.4
+ak = a * k
+d = np.inf
+omega = np.sqrt(g * k * np.tanh(k * d))
+t = 0
+
+# Fully nonlinear solution
+wave = SSGW(d, ak)
+
+phase, eta = np.real(wave.zs), np.imag(wave.zs)
+x = phase / k
+U = (np.real(wave.ws) + wave.ce) * np.sqrt(g)
+
+eta_linear = elevation(x, t, a, k, omega, wave_type="linear")
+eta_stokes = elevation(x, t, a, k, omega, wave_type="stokes")
+
+U_linear = orbital_horizontal_velocity(x, eta_linear, t, a, k, omega)
+U_stokes = orbital_horizontal_velocity(x, eta_stokes, t, a, k, omega)
+
+
+fig, (ax1, ax2, ax3) = plt.subplots(3, 1, figsize=(8, 9), sharex=True)
+
+# Top panel: Surface elevation
+ax1.plot(phase, eta_linear, label="1$^{st}$ order (linear)", lw=3)
+ax1.plot(phase, eta_stokes, label="3$^{rd}$ order Stokes", lw=3)
+ax1.plot(phase, eta, color="k", label="Fully nonlinear", lw=3)
+ax1.axhline(y=0, color="k", linestyle="--", alpha=0.5)
+ax1.set_ylabel("Surface elevation (m)")
+ax1.legend()
+ax1.grid(True)
+
+# Middle panel: Velocities
+ax2.plot(phase, U_linear, label="1$^{st}$ order (linear)", lw=3)
+ax2.plot(phase, U_stokes, label="3$^{rd}$ order Stokes", lw=3)
+ax2.plot(phase, U, color="k", label="Fully nonlinear", lw=3)
+ax2.axhline(y=0, color="k", linestyle="--", alpha=0.5)
+ax2.set_ylabel("Horizontal velocity (m/s)")
+ax2.grid(True)
+
+# Middle panel: Absolute errors
+error_linear = U_linear - U
+error_stokes = U_stokes - U
+
+ax3.plot(phase, error_linear, label="1$^{st}$ order (linear)", lw=3)
[…]

### 2025-03-13 paper.tex
Shorten abstract to fit 250 word limit

@@
 \begin{abstract}
-Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
-a well-known ocean surface process that affects remote sensing, the
-interpretation of \textit{in situ} wave measurements, and numerical wave
-forecasting.
-In this paper, we revisit the wave crest and action conservation laws and derive
-higher-order steady solutions for the change of short-wave wavenumber, action,
-and gravitational acceleration due to the presence of longer waves.
-We validate the analytical solutions with numerical simulations of the full wave
-crest and action conservation equations.
+Hydrodynamic modulation of short ocean surface waves by longer ambient waves
+significantly influences remote sensing, interpretation of \textit{in situ} wave
+measurements, and numerical wave forecasting.
+This paper revisits the wave crest and action conservation laws and derives
+steady, nonlinear, analytical solutions for the change of short-wave wavenumber,
+action, and gravitational acceleration due to the presence of longer waves.
+We validate the analytical solutions with numerical simulations of the full 
+conservation equations.
 The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
@@
 solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
-The short-wave steepness modulation is attributed to be 5/8 due to the
-wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
+The short-wave steepness modulation is attributed 5/8 to
+wavenumber, 1/4 due to wave action, and 1/8 due to effective gravity.
 Examining the homogeneity and stationarity requirements for the conservation of
@@
 generally not satisfied for very steep long waves.
-We further examine the results of \citet{peureux2021unsteady} who found through
+We examine the results of \citet{peureux2021unsteady} who found through
 numerical simulations that the short-wave modulation grows unsteadily with
@@
 We show that this unsteady growth only occurs for homogeneous initial
-conditions as a special case and does not generally occur, for example, in more
-realistic long-wave groups.
+conditions as a special case and not generally.
 The proposed steady solutions are a good approximation of the nonlinear

### 2025-06-09 paper.tex
Second round of revisions

@@
 action, and gravitational acceleration due to the presence of longer waves.
-We validate the analytical solutions with numerical simulations of the full 
-conservation equations.
+We validate the analytical solutions with numerical solutions of the full 
+crest and action conservation equations.
 The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
@@
 is the horizontal near-surface orbital velocity of the long wave.
-Strictly, a velocity profile must be considered when determining the
-advective velocity for a wave of any given wavenumber \citep{stewart1974hf},
-however, for brevity, here we will assume a surface velocity to be the advective
-one for short waves.
+Strictly, shear must be considered when determining the advective velocity for
+a wave of any given wavenumber \citep{stewart1974hf,ellingsen2014ship}.
+Here, we assume the surface velocity to be the advective one for short waves,
+and demonstrate later in $\S$\ref{sec:wave_action_conservation} that this is
+valid for sufficient scale separation between the short and long waves.
 The evolution of the short-wave wavenumber is described by the conservation of
@@
 The wave crest conservation (\ref{eq:wave_crests}) follows from the assumption
-of a conserved wave phase, which is limited by the same set of requirements as
-those for the wave action conservation \citep[e.g.,][]{whitham1974dispersive}.
+of a conserved wave phase, and is another prerequisite for wave action
+conservation \citep[e.g.,][]{whitham1974dispersive}.
 Although the absence of wave growth and dissipation is not generally realistic in
@@
 of hydrodynamic modulation solely due to the motion of the long waves.
+The governing equations (\ref{eq:dispersion})-(\ref{eq:wave_action}) are
+derived in Appendix \ref{appendix:derivation}.
 
@@
 group velocity and thus depends on the wavenumber.
+This convenient property is only possible if we consider the linear form
+of the dispersion relation, that is, if the series are truncated at the
+second-order solution in the Stokes expansion.
+For a third-order expansion, an additional term that is proportional to the
+square of the wave steepness appears in the dispersion relation.
+In that case, equations (\ref{eq:wavenumber}) and (\ref{eq:wave_action2})
+become fully coupled and the analytical solution for the wavenumber is no longer
+feasible.
 
@@
 \citet{bretherton1968wavetrains} also require that the medium is nearly uniform
-over the vertical scales of motion of the short waves.
-For the ambient velocity being the orbital velocity of the long wave, this is
-trivially satisfied with (\ref{eq:k_scale_separation}).
+over the vertical scales of motion of the short waves:
+
+\begin{equation}
+  \frac{1}{U} \frac{\partial U}{\partial z} \ll \widetilde{k}
+\end{equation}
+which, for a long deep water wave, is equivalent to (\ref{eq:k_scale_separation}).
+Alternatively, if we use the second Froude number criterion following
+\citet{ellingsen2014ship} and quantify the shear length
+under the long wave surface as:
+
+\begin{equation}
+  l_S = \frac{g}{S^2} \approx \frac{g}{\varepsilon_L^2 \omega_L^2}
+\end{equation}
+we arrive to an additional requirement:
+
+\begin{equation}
+  \frac{k}{k_L} \gg \varepsilon_L^2
+\end{equation}
+which relates the scale separation to the long-wave steepness.
 
@@
 wave action and crest conservation equations in the next section.
+For the asymptotic limits of wave action conservation in terms of the wave
+short-wave steepness, see Appendix \ref{appendix:wave_action_conservation}.
 
@@
 intrinsic frequency, and phase speed.
-The latter two were previously examined in the context of short-wave
-modulation \citep{longuet1962resonant,longuet1962phase}, and may be important
-for correctly calculating the wind input into short waves, which is proportional
-to the wind speed relative to the wave celerity.
+The latter two were previously examined in the context of resonant interactions
+\citep{longuet1962resonant,longuet1962phase}, a modulation process distinct from
+the one discussed here.
+Both effects, however, may be important for correctly calculating the wind input
+into short waves, which is proportional to the wind speed relative to the wave
+celerity.
 The approximate steady solutions yield higher wavenumber and amplitude
@@
 
+\section{Derivation of the governing equations}
+\label{appendix:derivation}
+
+Here we derive the governing equations for the short waves
+(\ref{eq:dispersion})-(\ref{eq:wave_action}) and discuss their asymptotic
+consistency.
+
+\subsection{Wave dispersion}
+
+An irrotational flow has a velocity potential such that the continuity can be
+expressed as:
+
+\begin{equation}
+  \nabla^2 \phi = 0
+  \label{eq:laplace_equation}
+\end{equation}
+Boundary conditions that allow solving for $\phi$ are the dynamic and kinematic
+free surface boundary conditions, respectively:
+
+\begin{equation}
+  \frac{\partial \phi}{\partial t} + \frac{1}{2} (\nabla{\phi})^2 + gz = 0 \quad \text{at} \quad z = \eta(x,t)
+  \label{eq:dynamic_free_surface_boundary_condition}
+\end{equation}
+
+\begin{equation}
+  \frac{\partial \eta}{\partial t} + \frac{\partial \phi}{\partial x} \frac{\partial \eta}{\partial x} + \frac{\partial \phi}{\partial z} = 0 \quad \text{at} \quad z = \eta(x,t)
+  \label{eq:kinematic_free_surface_boundary_condition}
+\end{equation}
+Also required is the bottom boundary condition:
+
+\begin{equation}
+  \frac{\partial \phi}{\partial z} = 0 \quad \text{at} \quad z = -h
+\end{equation}
+In the \citet{stokes1847} perturbation approach, the velocity potential and
+elevation are expanded in a power series of the wave steepness
+$\varepsilon = ak$ as:
+
+\begin{equation}
+  \phi = \sum_{n=1}^{\infty} \varepsilon^n \phi_n
+\end{equation}
+\begin{equation}
+  \eta = \sum_{n=1}^{\infty} \varepsilon^n \eta_n
+\end{equation}
+The power series expansion requires $\varepsilon \ll 1$.
+The velocity potential and elevation series are truncated at the desired order
+and inserted into the dynamic free surface boundary condition (\ref{eq:dynamic_free_surface_boundary_condition}).
+Up to the third order, the solutions for the velocity potential and the
+corresponding velocity components are (\ref{eq:phi})-(\ref{eq:W_L}), and the
+surface elevation is (\ref{eq:eta_stokes}).
+Evaluating these quantities to higher orders in $\varepsilon$ is possible at
+the expense of higher complexity and with diminishing returns in accuracy.
+
+Combining the solutions for the potential and the surface elevation with the
+kinematic free surface boundary condition (\ref{eq:kinematic_free_surface_boundary_condition})
+yields the dispersion relation, which in the linear case and in deep water is
+(\ref{eq:dispersion}).
+At the third order in $\varepsilon$, the frequency begins to depend on the wave
+steepness $\varepsilon$ as well:
+
+\begin{equation}
+  \omega = \sqrt{g k} \left(1 + \frac{1}{2} \varepsilon^2 + \mathcal{O}(\varepsilon^4)\right) + U k
+  \label{eq:dispersion_nonlinear}
+\end{equation}
+In this paper we constrain the analysis framework to the linear case for short
+waves, as this permits the analytical solutions in $\S\ref{section:analytical_solutions}$.
+The Doppler shift by the ambient velocity $U$ is unaffected by the nonlinear
+effects in the solution.
+
+\subsection{Conservation of wave crests}
+
+A system with a well defined phase function $\psi(x,t)$ also has the wavenumber
+and frequency defined as its gradients with respect to space and time, respectively:
+
+\begin{equation}
+  k = \frac{\partial \psi}{\partial x}
+  \label{eq:wavenumber_from_phase}
+\end{equation}
+
+\begin{equation}
+  \omega = - \frac{\partial \psi}{\partial t}
+  \label{eq:frequency_from_phase}
+\end{equation}
+implying $\psi = kx - \omega t$.
+Differentiating (\ref{eq:wavenumber_from_phase})-(\ref{eq:frequency_from_phase})
+with respect to time and space, respectively, and then eliminating $\psi$ yields:
+
+\begin{equation}
+  \frac{\partial k}{\partial t} + \frac{\partial \omega}{\partial x} = 0
+  \label{eq:wavenumber_conservation}
+\end{equation}
+which is the crest conservation equation (\ref{eq:wave_crests}).
+That this is the wavenumber conservation equation with the group speed as the
+transport velocity can be recognized by writing:
+
+\begin{equation}
+  \frac{\partial k}{\partial t} + \frac{\partial \omega}{\partial k} \frac{\partial k}{\partial x} = 0
+\end{equation}
+This conservation equation is valid regardless of the order of nonlinearity
+considered, as long as $\psi$ is well defined.
+Nonlinearity in terms of $\varepsilon$ can, however, enter (\ref{eq:wavenumber_conservation})
+through nonlinear forms of $\omega$ such as (\ref{eq:dispersion_nonlinear}).
+
+\subsection{Conservation of wave action}
+\label{appendix:wave_action_conservation}
+
+The conservation of wave action (\ref{eq:wave_action}) is most readily derived
+following the variational approach by \citet{whitham1965general}.
+Begin with a statement that infinitesimal variations of a Lagrangian of a system
+over a spatial domain and time interval must vanish:
+
+\begin{equation}
+  \delta \iint L\, dx\,dt = 0
+\end{equation}
+Here, $L$ is the instantaneous Lagrangian density, a scalar quantity that
+corresponds to the difference between the kinetic and potential energy of the
+system:
+
+\begin{equation}
+  L = \int_{-h}^{\eta} \left[ \frac{1}{2} (\nabla{\phi})^2 + gz \right] dz
+  \label{eq:lagrangian_density_energy_difference}
+\end{equation}
+\citet{luke1967variational} showed that the equations for irrotational
+surface waves follow from the variational approach if the Lagrangian $L$ is
+defined instead as:
+
+\begin{equation}
+  L = \int_{-h}^{\eta} \left[ \frac{\partial \phi}{\partial t} + \frac{1}{2} (\nabla{\phi})^2 + gz \right] dz
+  \label{eq:lagrangian_density_luke}
+\end{equation}
+(\ref{eq:lagrangian_density_energy_difference}) and (\ref{eq:lagrangian_density_luke})
+are equivalent in the sense that they both yield (\ref{eq:laplace_equation}),
+however, varying the latter is necessary to attain the correct boundary conditions
+\citep{whitham1967non}.
+Notice that for water waves the instantaneous Lagrangian being zero is
+equivalent to the Bernoulli's equation, which is also the dynamic free surface
+boundary condition.
+
+Then, Whitham's averaged variation principle is obtained by averaging the
+instantaneous Lagrangian $L$ over one wave period:
+
+\begin{equation}
+  \mathcal{L} = \frac{1}{2\pi} \int_0^{2\pi} L\ d\psi
+  \label{eq:averaged_lagrangian}
+\end{equation}
+where $\mathcal{L}$ is the averaged Lagrangian density.
+The averaged variational principle then states:
+
+\begin{equation}
+  \delta \iint \mathcal{L}\, dx\,dt = 0
+\end{equation}
+The averaged Lagrangian for water waves is obtained by integrating
+(\ref{eq:lagrangian_density_luke}) over depth, and then averaging over the phase
+(\ref{eq:averaged_lagrangian}).
+In the deep water limit, it is:
+
+\begin{equation}
+  \mathcal{L} = \frac{1}{4} \left( 1 - \frac{\omega^2}{gk} \right) g a^2 + \mathcal{O}(\varepsilon^4)
+\end{equation}
+which, after recognizing the wave energy$E = \frac{1}{2} g a^2$, yields:
+
+\begin{equation}
+  \mathcal{L} = \frac{1}{2} \left( 1 - \frac{\omega^2}{gk} \right) E + \mathcal{O}(\varepsilon^4)
+  \label{eq:averaged_lagrangian_deep_water}
+\end{equation}
+
+Whitham's conservation equations for water waves are:
+
+\begin{equation}
+  \frac{\partial \mathcal{L}}{\partial a} = 0
+\end{equation}
+
+\begin{equation}
+  \frac{\partial}{\partial t} \left(\frac{\partial \mathcal{L}}{\partial \omega}\right) - \frac{\partial}{\partial x} \left(\frac{\partial \mathcal{L}}{\partial k}\right) = 0
+  \label{eq:whitham_conservation_equation}
+\end{equation}
+Inserting (\ref{eq:averaged_lagrangian_deep_water}) into (\ref{eq:whitham_conservation_equation})
+and defining:
+
+\begin{equation}
+  \frac{\partial \mathcal{L}}{\partial \omega} = \frac{E}{\sigma} \equiv N
+  \label{eq:adiabatic1}
+\end{equation}
+yields the conservation of wave action:
+
+\begin{equation}
+  \frac{\partial}{\partial t} \left[ N + \mathcal{O}(\varepsilon^4) \right] +
+  \frac{\partial}{\partial x} \left[ (C_g + \mathcal{O}(\varepsilon^2)) N + \mathcal{O}(\varepsilon^4) \right] = 0
+\end{equation}
+Due to the $\mathcal{O}(\varepsilon^2)$ correction to the group speed, the
+overall truncation error for the wave action balance is $\mathcal{O}(\varepsilon^2)$
+as well:
+
+\begin{equation}
+  \frac{\partial N}{\partial t} +
+  \frac{\partial (C_g N)}{\partial x} = \mathcal{O}(\varepsilon^2)
+  \label{eq:wave_action_conservation_nonlinear}
+\end{equation}
+which was introduced as a governing equation in \S\ref{section:governing_equations}.
+Although the truncation error of $\mathcal{O}(\varepsilon^2)$ arising from $C_g$
+seems to have been omitted in the original derivation by \citet{whitham1967non},
+it appeared correctly later in \citet{whitham1974linear}.
+An important nuance about applying the nonlinear dispersion here is that the
+secondary (wave-induced) mean flow and set-up effects can be ignored in the
+deep water limit; this is discussed by \citet{whitham1974dispersive}.
+The terms proportional to $\varepsilon^4$ do not appear in the averaged Lagrangian
+(\ref{eq:averaged_lagrangian_deep_water}) until the third-order Stokes waves
+are considered, for which the dispersion gains a $\mathcal{O}(\varepsilon^2)$
+correction.
+In that case, the conserved quantity is:
+
+\begin{equation}
+  \frac{\partial \mathcal{L}}{\partial \omega} = \frac{E}{\sigma} (1 + \frac{1}{2} \varepsilon^2)
+  \label{eq:adiabatic2}
+\end{equation}
+and the group speed gains a nonlinear factor in $\varepsilon^2$ as well.
+(\ref{eq:wave_action_conservation_nonlinear}), then, maintains the same form,
+but with the truncation error of $\mathcal{O}(\varepsilon^4)$.
+This derivation, as well as the averaged Lagrangian principle applied to the
+wave-induced mean flow over arbitrary water depths, is detailed at greater length
+by \citet{whitham1967non}.
+
+To consider the effect of slowly varying ambient currents,
+\citet{bretherton1968wavetrains} carried out the same derivation except that
+they propagated the ambient current derivatives $\partial U / \partial t$ and
+$\partial U / \partial x$ through the averaged Lagrangian.
+To allow these derivatives to drop out of the averaging integral, the authors
+make an argument that the variation must be sufficiently small, no larger than
+$\mathcal{O}(\frac{1}{\sigma U} \frac{\partial U}{\partial t})$ and
+$\mathcal{O}(\frac{1}{k U} \frac{\partial U}{\partial x})$, respectively.
+Taking into consideration the asymptotic limits in terms of the wave steepness
+$\varepsilon$ as well as the variation of the ambient current, the wave action
+balance can be written as:
+
+\begin{equation}
+  \frac{\partial N}{\partial t} +
+  \frac{\partial}{\partial x} \left[ \left(C_g + U\right) N \right] =
+  \mathcal{O}\left(\varepsilon^4, \frac{1}{\sigma U} \frac{\partial U}{\partial t}, \frac{1}{k U} \frac{\partial U}{\partial x}\right)
+\end{equation}
+\S\ref{sec:wave_action_conservation} deals specifically with quantifying the
+ambient velocity variation in the context of long-wave orbital velocities.
+
+There is an alternative derivation of the wave action conservation that was
+pointed out by \citet{bretherton1968wavetrains}.
+Starting from the wave energy conservation in non-uniform currents derived by
+\citet{longuet1961changes}
+
+\begin{equation}
+  \frac{\partial E}{\partial t} +
+  \frac{\partial}{\partial x} \left[ \left(C_g + U\right) E \right] +
+  S_{x} \frac{\partial U}{\partial x} = 0
+  \label{eq:energy_conservation_nonuniform_currents}
+\end{equation}
+where $S_x$ is the wave radiation stress, and combining it with the dispersion
+relation (\ref{eq:dispersion}) and conservation of crests
+(\ref{eq:wave_crests}), they show that (\ref{eq:energy_conservation_nonuniform_currents})
+is equivalent to the wave action conservation (\ref{eq:wave_action}).
+This derivation is found in the Appendix of \citet{bretherton1968wavetrains},
+and was later recognized by \citet{whitham1974linear} as well.
+
+In summary, asymptotically consistent sets of dispersion and action conservation
+equations vary depending on the level of nonlinearity considered.
+In the case of linear waves, the dispersion relation (\ref{eq:dispersion}),
+the averaged Lagrangian (\ref{eq:averaged_lagrangian_deep_water}), the adiabatic
+quantity (\ref{eq:adiabatic1}) and the wave action conservation
+(\ref{eq:wave_action_conservation_nonlinear}) are asymptotically consistent.
+For third-order Stokes waves, the dispersion relation
+(\ref{eq:dispersion_nonlinear}), the adiabatic quantity (\ref{eq:adiabatic2}),
+together with the group speed that includes a $\mathcal{O}(\varepsilon^2)$ correction,
+are governed by the action conservation law
+(\ref{eq:wave_action_conservation_nonlinear}) and are asymptotically consistent.
+This is summarized in Table \ref{table:asymptotic_consistency}.
+
+\begin{table}
+\begin{center}
+\def~{\hphantom{0}}
+\begin{tabular}{cccc}
+Order & Frequency  & Group speed & Wave action conservation \\
+\hline
+1, 2 & $\sqrt{gk}(1 + \mathcal{O}(\varepsilon^2))$ & $\frac{1}{2}\sqrt{\frac{g}{k}}(1 + \mathcal{O}(\varepsilon^2))$ & $\frac{\partial N}{\partial t} + \frac{\partial (C_g N)}{\partial x} = \mathcal{O}(\varepsilon^2)$ \\
+\hline
+3 & $\sqrt{gk}(1 + \frac{1}{2}\varepsilon^2 + \mathcal{O}(\varepsilon^4))$ & $\frac{1}{2}\sqrt{\frac{g}{k}}(1 + \frac{5}{2}\varepsilon^2 + \mathcal{O}(\varepsilon^4))$ & $\frac{\partial N}{\partial t} + \frac{\partial (C_g N)}{\partial x} = \mathcal{O}(\varepsilon^4)$ \\
+\hline
+\end{tabular}
+\caption{
+  Asymptotically consistent sets of dispersion and wave action relations for
+  deep water and their respective truncation errors in terms of the wave
+  steepness $\varepsilon$, for the first three Stokes expansion orders.
+}
+\label{table:asymptotic_consistency}
+\end{center}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fig_velocity_error_by_phase.pdf}
+\caption{
+  Surface elevation (top) and horizontal velocity (middle) for the linear wave
+  (blue), the third-order Stokes wave (orange), and the fully-nonlinear wave
+  (black) with steepness of $\varepsilon_L = 0.4$.
+  The bottom panel shows the surface velocity error for the linear wave (blue)
+  and the third-order Stokes wave (orange), relative to the fully-nonlinear
+  wave solution.
+}
+\label{fig:velocity_error_by_phase}
+\end{figure}
+
 \section{Surface velocity errors}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/fig_velocity_error_by_phase.pdf}
-\caption{
-  Surface elevation (top) and horizontal velocity (middle) for the linear wave
-  (blue), the third-order Stokes wave (orange), and the fully-nonlinear wave
-  (black) with steepness of $\varepsilon_L = 0.4$.
-  The bottom panel shows the surface velocity error for the linear wave (blue)
-  and the third-order Stokes wave (orange), relative to the fully-nonlinear
-  wave solution.
-}
-\label{fig:velocity_error_by_phase}
-\end{figure}
-
-\begin{figure}
-\centering
 \includegraphics[width=0.8\textwidth]{figures/fig_velocity_error_by_ak.pdf}

### 2025-06-17 paper.tex
Fix affil. numbers
Fix omitted parentheses in (A 22); Clarify truncation order on L798

@@
 \documentclass[lineno]{jfm}
+%\documentclass[]{jfm}
 
@@
 \author{
-  Milan Curcic\aff{1}
+  Milan Curcic\aff{1,2}
   \corresp{\email{mcurcic@miami.edu}}
@@
   \frac{\partial}{\partial t} \left[ N + \mathcal{O}(\varepsilon^4) \right] +
-  \frac{\partial}{\partial x} \left[ (C_g + \mathcal{O}(\varepsilon^2)) N + \mathcal{O}(\varepsilon^4) \right] = 0
+  \frac{\partial}{\partial x} \left[ (C_g + \mathcal{O}(\varepsilon^2)) (N + \mathcal{O}(\varepsilon^4)) \right] = 0
 \end{equation}
@@
 are governed by the action conservation law
-(\ref{eq:wave_action_conservation_nonlinear}) and are asymptotically consistent.
+(\ref{eq:wave_action_conservation_nonlinear}, but with the truncation error of
+$\mathcal{O}(\varepsilon^4)$), and are asymptotically consistent.
 This is summarized in Table \ref{table:asymptotic_consistency}.

### 2025-06-17 scripts/plot_graphical_abstract.py
Update graphical abstract script

@@ -2,7 +2,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 # Set up the figure
-fig, ax = plt.subplots(figsize=(12, 4))  # Increased width to accommodate two periods
+fig, ax = plt.subplots(figsize=(6, 5), dpi=300)  # Increased width to accommodate two periods
 
 # Generate x values
 x = np.linspace(0, 3.5 * np.pi, 2000)  # Extended to 4π, increased number of points
@@ -36,7 +36,7 @@ for i, pos in enumerate(arrow_positions):
     dy = arrow_length * np.sin(angle)
 
     # Position arrow slightly below the wave
-    y_offset = -0.25
+    y_offset = -0.3
 
     if i % 2 == 0:  # Even positions (arrows pointing inward)
         ax.arrow(
@@ -85,37 +85,38 @@ for i, pos in enumerate(arrow_positions):
             width=0.1,
         )
 
-    ax.arrow(
-        pos - 0.25 * np.pi,
-        np.sin(pos - 0.25 * np.pi) - y_offset,
-        1.5 * dx,
-        1.5 * dy,
-        head_width=0.12,
-        head_length=0.2,
-        fc="k",
-        width=0.04,
-    )
+    # Short-wave Cg arrow
+    #ax.arrow(
+    #    pos - 0.25 * np.pi,
+    #    np.sin(pos - 0.25 * np.pi) - y_offset,
+    #    1.5 * dx,
+    #    1.5 * dy,
+    #    head_width=0.12,
+    #    head_length=0.2,
+    #    fc="k",
+    #    width=0.04,
+    #)
 
 pos = arrow_positions[0]
-ax.text(
-    pos, np.sin(pos) - 0.3, r"$U$", fontsize=24, ha="right", va="top", color="tab:blue"
-)
+#ax.text(
+#    pos, np.sin(pos) - 0.3, r"$U$", fontsize=24, ha="right", va="top", color="tab:blue"
+#)
 
-ax.text(
-    pos - 0.05,
-    np.sin(pos) + 1.1,
-    r"$C_g$",
-    fontsize=24,
-    ha="right",
-    va="top",
-    color="k",
-)
+#ax.text(
+#    pos - 0.05,
+#    np.sin(pos) + 1.1,
+#    r"$C_g$",
+#    fontsize=24,
+#    ha="right",
[…]
